### 2017-07-13 eugene_notes.tex
notes from falling sky

\documentclass{article}
\usepackage{fullpage}
\usepackage{amsmath, amssymb}

\newcommand{\mL}{\mathcal{L}}
\newcommand{\ip}[2]{\langle #1, #2 \rangle}

\begin{document}

\paragraph{Outline of paper}

\begin{itemize}

    \item  Introduction:
\begin{itemize}
        \item developmental systems drift
        \item justification of passage to the infinitesimal model
\end{itemize}

    \item Model:
\begin{itemize}
        \item heuristics and pictures from simulation below
        \item justification of passage to the infinitesimal model
        \item comparison to simulation: shape of the distribution; recombination load
    \end{itemize}

    \item Conclusion
\begin{itemize}
        \item limiting distribution: general sizes and how it depends on selection and segregation variance
        \item recombionation load
        \item speed of approach to optimum (?)
        \item influence of dimensionality
    \end{itemize}

\end{itemize}

\paragraph{General picture}
We want to describe the dynamics of a population that is evolving under stabilizing selection around not a single point but a manifold
of optimal phenotypes, using the infinitesimal model.
We think the right picture is that as the population size $N$ grows,
the dynamics of the measure become deterministic,
which flow quickly to within $1/N$ of the manifold,
and then we can follow only the mean as it diffuses as an OU-like process
near the manifold,
with drift and diffusion terms depending on the (deterministic) shape of the stable population measure
around that point.

\paragraph{Deterministic flow}
Let $X$ be a population of size $N$, recorded as the empirical distribution on phenotypes
\begin{align*}
    X = \frac{1}{N} \sum_{i=1}^N \delta_{x_i} .
\end{align*}
This evolves as a Moran process under the infinitesimal model
with death rates depending on state $x$.
Concretely, an individual whose phenotype is $x$
dies at rate $1 + \mu(x)$
and is replaced by a new individual of phenotype
\begin{align*}
  \frac{y + z}{2} + \xi ,
\end{align*}
where $y$ and $z$ are independent draws from $X$
and $\xi$ is an independent $N(0, \eta^2)$.
Let $\nu(X)$ denote the distribution of $\frac{y + z}{2} + \xi$.
This has generator
\begin{align*}
    \mL \Phi(X)
    &=
    \int \int \int \int
    \left( \Phi(X + \frac{1}{N} \delta_{(y+z)/2 + \xi} - \frac{1}{N} \delta_x) - \Phi(X) \right)
    (1 + \mu(x)) X(dx) X(dy) X(dz) e^{-(\xi/\eta)^2/2} d\xi \\
    &=
    \int \int
    \left( \Phi(X + \frac{1}{N} \delta_u - \frac{1}{N} \delta_x) - \Phi(X) \right)
    (1 + \mu(x)) X(dx) \nu(X)(du) .
\end{align*}

For a test function $\phi$, setting $\Phi(X) = \ip{\phi}{X}$
we get the first moment
\begin{align*}
    \mL \ip{\phi}{X}
    &=
    \frac{1}{N} \int \int \int \int
    \left( \phi \left( \frac{y+z}{2} + \xi \right)
         - \phi(x) \right)
    (1 + \mu(x)) X(dx) X(dy) X(dz) e^{-(\xi/\eta)^2/2} d\xi  \\
    &=
    \frac{1}{N} \ip{\phi}{\nu(X)}\ip{1+\mu}{X}
    - \frac{1}{N} \ip{\phi}{(1+\mu) X} .
\end{align*}
Note that $\ip{\phi}{\partial_{\nu(X)} X} = \ip{\phi}{\nu(X)}$,
so that this is a first derivative in the direction of
$\ip{1+\mu}{X} \nu(X) - (1+\mu) X$.

% The second moment is
% \begin{align*}
%     \mL \ip{\phi}{X} \ip{\psi}{X}
%     &=
%     \frac{1}{N} \int \int \int \int
%     \left( \phi \left( \frac{y+z}{2} + \xi \right)
%          - \phi(x) \right)
%     \ip{\psi}{X}
%     (1 + \mu(x)) X(dx) X(dy) X(dz) e^{-(\xi/\eta)^2/2} d\xi \\
%     &\qquad {} +
%     \frac{1}{N} \int \int \int \int
%     \ip{\phi}{X}
%     \left( \psi \left( \frac{y+z}{2} + \xi \right)
%          - \psi(x) \right)
%     (1 + \mu(x)) X(dx) X(dy) X(dz) e^{-(\xi/\eta)^2/2} d\xi 
%     \\ &\qquad {} +
%     O(1/N^2)  \\
%     &=
%     \ip{\phi}{\partial_{\nu(X)} X} \ip{\psi}{X}
%     +
%     \ip{\phi}{X} \ip{\psi}{\nu(X)} 
%     + O(1/N^2)  \\
% \end{align*}

More generally, if $\Phi(X) = \prod_i \ip{\phi_i}{X}$ then
\begin{align*}
    \mL \Phi(X)
    &=
    \frac{1}{N} \sum_i \left( 
        \ip{1+\mu}{X} \ip{\phi_i}{\nu(X)}  - \ip{\phi_i}{(1+\mu)X}
    \right)
    \prod_{j \neq i} \ip{\phi_j}{X} 
    + O(1/N^2) .
\end{align*}
This implies that after rescaling time by $N$
and suitible other caveats,
the limiting process is a deterministic flow in the direction of 
$\nu(X) \ip{1+\mu}{X} - (1+\mu)X$.

If $Z$ is a fixed point of this flow,
then $\nu(Z)\ip{1+\mu}{Z} = (1+\mu)Z$.
We assume that $\mu(x) = 0$ on a nice smooth ``optimal'' manifold,
and that it increases with distance to the manifold.
With $\eta = 0$,
any fixed point is either a point mass,
or possibly uniform on the optimal manifold if it is linear.
We think an argument for this is obtained SOMETHING LIKE THE FOLLOWING: 
if $Z$ is a fixed point
and $\phi$ is a test function that is convex and increasing with distance from the optimal manifold,
then $\ip{\phi}{\nu(Z)} < \ip{\phi}{Z} < \ip{\phi}{(1+\mu)Z}/\ip{1+\mu}{Z}$,
since averaging contracts $Z$ while reweighting by $1+\mu$ expands $Z$.

If $\eta$ is not zero, then we hope that for each point $z$ on the optimal manifold,
there exists a unique measure $Z(z)$ with mean $z$ that is a fixed point
of the flow.

\paragraph{Diffusion of the mean}
Since we work with a Moran model,
the mean evolves simply by jumping at rate $\ip{1+\mu}{X}$,
and at each jump, adding $w/N$ and subtracting $\tilde x/N$,
where $w$ is drawn from $\nu(X)$ and $\tilde x$ is drawn from 
$(1+\mu)X/\ip{1+\mu}{X}$.
This implies that if the mean is currently $\bar X$
and the population distribution is $X$,
then with $\phi_{id}(x) = x$,
if
\begin{align*}
    m(X) = N \left( \ip{\phi_{id}}{\nu(X)}\ip{1+\mu}{X}
    -
    \ip{\phi_{id}}{(1+\mu)X} \right)
\end{align*}
is of order 1,
then
the $\bar X$ run on a time scale of $N^2$
should behave as a diffusion with drift $m(X)$
and variance equal to the variance of $w$ plus the variance of $\tilde x$, which is
\begin{align*}
    \frac{1}{2} \int (x-\bar X)^2 X(dx) + \eta^2
    + \frac{ \int x^2 (1+\mu(x)) X(dx)  
    - \left(\int x (1+\mu(x)) X(dx)\right)^2 }{ \ip{1+\mu}{X} } .
\end{align*}

\paragraph{Joint process}
By writing the joint dynamics of the mean and the measure realtive to the mean,
we hope to show that the picture above makes sense,
that in the limit we can take the measure as deterministic given the mean.

\paragraph{Weak selection}
It may be nice to consider weak selection,
in which case perhaps the limiting measures are all Gaussian.

\paragraph{Rough calculations of variances}
We did some of these,
by solving for $\sigma^2$ that was stable under
average $\to$ mollify $\to$ reweight by $1+\mu$,
assuming that fourth moments are $C$ times second moments, squared.

\end{document}

### 2017-07-14 eugene_notes.tex
moments

@@
 \newcommand{\ip}[2]{\langle #1, #2 \rangle}
+
+\include{macros}
 
@@
 
-% The second moment is
-% \begin{align*}
-%     \mL \ip{\phi}{X} \ip{\psi}{X}
-%     &=
-%     \frac{1}{N} \int \int \int \int
-%     \left( \phi \left( \frac{y+z}{2} + \xi \right)
-%          - \phi(x) \right)
-%     \ip{\psi}{X}
-%     (1 + \mu(x)) X(dx) X(dy) X(dz) e^{-(\xi/\eta)^2/2} d\xi \\
-%     &\qquad {} +
-%     \frac{1}{N} \int \int \int \int
-%     \ip{\phi}{X}
-%     \left( \psi \left( \frac{y+z}{2} + \xi \right)
-%          - \psi(x) \right)
-%     (1 + \mu(x)) X(dx) X(dy) X(dz) e^{-(\xi/\eta)^2/2} d\xi 
-%     \\ &\qquad {} +
-%     O(1/N^2)  \\
-%     &=
-%     \ip{\phi}{\partial_{\nu(X)} X} \ip{\psi}{X}
-%     +
-%     \ip{\phi}{X} \ip{\psi}{\nu(X)} 
-%     + O(1/N^2)  \\
-% \end{align*}
-
 More generally, if $\Phi(X) = \prod_i \ip{\phi_i}{X}$ then
@@
 
+\paragraph{Stable measures}
 If $Z$ is a fixed point of this flow,
@@
 of the flow.
+
+Any measure $Z$ that is a fixed point has the property that
+(the average of two independent draws plus segregation noise)
+is equal in distribution to
+(a $1+\mu$-weighted draw).
+Concretely, with $A$, $B$, and $C$ independent draws from $Z$,
+$$\begin{aligned}
+    \E\left[ f\left( \frac{A+B}{2} + \xi \right) \right]
+    &=
+    \frac{\E\left[ f(C) (1+\mu(C)) \right]}{\E[1+\mu(C)]} .
+\end{aligned}$$
+Let
+$$\begin{aligned}
+    m_k = \E[A^k] = \ip{x^k}{X} ,
+\end{aligned}$$
+and note that since $\E[\xi^w] = \eta^w (w-1)!!$ for even $w$,
+$$\begin{aligned}
+    \ip{x^k}{\nu(X)}
+    &=
+    \sum_{u+v+w=k} \binom{k}{u,v,w} 2^{w-k} m_u m_v \eta^w (w-1)!!.
+\end{aligned}$$
+
+We can calculate some things about this measure
+for the case that $\mu(x) = ax^2$ 
+(or generally, if this is a Taylor expansion).
+Then
+$$\begin{aligned}
+    \frac{\E\left[ C^k (1+\mu(C)) \right]}{\E[1+\mu(C)]} 
+    &=
+    \frac{ m_k + a m_{k+2} }{ 1 + a m_2 } .
+\end{aligned}$$
+This gives us an equation for $m_{k+2}$ in terms of lower-order terms.
+With $k=2$ and assuming that $Z$ is symmetric,
+$$\begin{aligned}
+    \frac{m_2}{2} + \eta^2
+    &=
+    \frac{ m_2 + a m_4 }{ 1 + a m_2 },
+\end{aligned}$$
+and so 
+$$\begin{aligned}
+    m_4
+    &= m_2 \left( \eta^2 + \frac{1}{2} m_2 \right) 
+       + \frac{1}{a} \left( \eta^2 - \frac{1}{2} m_2 \right) .
+\end{aligned}$$
+We believe that as $a \to 0$ this does not blow up,
+implying that $m_2 = 2 \eta^2 + O(a)$,
+and that in fact approaches Gaussian, for which $m_4 = 3m_2^2$,
+and so $\eta^2 - m_2/2 = a \eta^2 + O(a^2)$,
+and hence
+$$\begin{aligned}
+    m_2 &= 2 \eta^2 ( 1 - a \eta^2 ) + O(a^2)  \\
+    m_4 &= 4 \eta^4 + O(a) .
+\end{aligned}$$
+
+If we take $1+\mu(x) = e^{ax^2}$,
+then there is a Gaussian solution,
+as is common in quant gen.
+(Write this down in arbitrary dimensions!)
+
+\paragraph{Fitness and load}
+Note that since reproduction is unweighted,
+fitness is proportional to lifetime,
+so the fitness function is $1/(1+\mu(x)$.
+Since replacement occurs at rate $\ip{1+\mu}{X}$,
+the mean number of offspring per individual of type $x$
+in population $X$ is $\ip{1+\mu}{X} / (1+\mu(x))$.
+
+Take $\mu(x) = ax^2$.
+The mean fitness in the population is then
+$$\begin{aligned}
+    \ip{1/(1+\mu)}{X}
+    &=
+    \ip{1 - ax^2 + a^2 x^4}{X} + O(a^3) \\
+    &=
+    1 - a m_2 + a^2 m_4 + O(a^3) \\
+    &=
+    1 - 2 a \eta^2 + 6 a^2 \eta^4 + O(a^3) ,
+\end{aligned}$$
+and the mean fitness of \emph{offspring} is
+$$\begin{aligned}
+    \ip{1/(1+\mu)}{\nu(X)}
+    &=
+    \ip{1 - ax^2 + a^2 x^4}{\nu(X)} + O(a^3) \\
+    &=
+    1 - a \left( \frac{m_2}{2} + \eta^2 \right) 
+        + a^2 \left( 2 m_4 + 6 m_2^2 + 12 m_2 \eta^2 + 3 \eta^4\right) + O(a^3) \\
+    &=
+    1 - a \left( 2 \eta^2 - a \eta^4 \right)
+        + a^2 \left( 8 \eta^4 + 24 \eta^4 + 24 \eta^4 + 3 \eta^4\right) + O(a^3) \\
+    &=
+    1 - 2 a \eta^2 + 60 a^2 \eta^4 + O(a^3) .
+\end{aligned}$$
+This implies that the difference in mean fitness between parents and offspring
+(maybe this is called recombination load?)
+is $54 a^2 \eta^4 + O(a^3)$.
+(SOMETHING IS WRONG here because mean fitness in offspring should be less than in parents).
 
@@
 
-\paragraph{Rough calculations of variances}
-We did some of these,
-by solving for $\sigma^2$ that was stable under
-average $\to$ mollify $\to$ reweight by $1+\mu$,
-assuming that fourth moments are $C$ times second moments, squared.
-
 \end{document}

### 2017-07-14 eugene_notes.tex
better way of getting load

@@
 \newcommand{\mL}{\mathcal{L}}
-\newcommand{\ip}[2]{\langle #1, #2 \rangle}
+\newcommand{\ip}[2]{\left\langle #1, #2 \right\rangle}
 
@@
 
+Since $\nu(X)$ is the offspring distribution,
+and we know that $\nu(X) = (1+\mu) X / \ip{1+\mu}{X}$,
+we can compute mean fitness of offspring as
+$$\begin{aligned}
+    \ip{\frac{1}{1+\mu}}{\nu(X)}
+    &=
+    \frac{\ip{\frac{1}{1+\mu}}{(1+\mu)X}}{\ip{1+\mu}{X}} \\
+    &=
+    \frac{1}{\ip{1+\mu}{X}} .
+\end{aligned}$$
+
 Take $\mu(x) = ax^2$.
@@
 \end{aligned}$$
-and the mean fitness of \emph{offspring} is
+and the mean fitness among offspring is
+$$\begin{aligned}
+    \frac{1}{1+am_2} 
+    &= 1 - 2 a \eta^2 + 2 a^2 \eta^4 + O(a^3),
+\end{aligned}$$
+so the difference in mean fitness between parents and offspring
+(maybe this is called recombination load?)
+is $4 a^2 \eta^4$.
+
+Here is the same thing calculated the other way,
+but something is wrong:
+The mean fitness of offspring is
 $$\begin{aligned}
@@
 \end{aligned}$$
-This implies that the difference in mean fitness between parents and offspring
-(maybe this is called recombination load?)
+The difference here is
 is $54 a^2 \eta^4 + O(a^3)$.
-(SOMETHING IS WRONG here because mean fitness in offspring should be less than in parents).
+(something is wrong here because mean fitness in offspring should be less than in parents).
 

### 2017-09-20 MeanFlow.tex
added mean flow doc

\documentclass[11pt]{amsart}

\usepackage{amsmath,amssymb,amsthm,bm,enumerate,mathtools}
%\usepackage[all]{xy}
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\geometry{a4paper}

\usepackage{chngcntr}
\usepackage{apptools}
\AtAppendix{\counterwithin{lem}{section}}

\newtheorem{thm}{Theorem}
\newtheorem{lem}{Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{cor}{Corollary}
\theoremstyle{remark} 
	\newtheorem{rem}{Remark}
	\newtheorem{assn}{Assumption}
\theoremstyle{definition} 
	\newtheorem{mydef}{Definition} 
	\newtheorem{exmp}{Example} 
	\newtheorem{cond}{Condition}
	\newtheorem{conj}{Conjecture}

\newtheorem{innercustomthm}{Propostion}
\newenvironment{customthm}[1]
  {\renewcommand\theinnercustomthm{#1}\innercustomthm}
  {\endinnercustomthm}

\newcommand{\abs}[1]{\left| #1 \right|}
\newcommand{\BigO}[1]{\mathcal{O}{\textstyle\left( #1\right)}}

\newcommand{\ie}{\textit{i.e.,}\,}
\newcommand{\eg}{\textit{e.g.,}\,}
\newcommand{\nb}{\textit{n.b.,}\, }
\newcommand{\defn}{:=}

%\input{../../GlobalDefs}

\begin{document}

\title{Fixed Points of the Mean Flow}
%\subtitle{}
\author{Todd L. Parsons}
\address{Laboratoire de Probabilit\'{e}s et Mod\`{e}les Al\'{e}atoires, CNRS UMR 7599, Universit\'{e} Pierre et Marie Curie, Paris, 75005, France.}
%\ead{todd.parsons@upmc.fr}

\date{\today}

%\begin{abstract}

%\end{abstract}

\maketitle

We had that $\nu(X)$ is the distribution of $\frac{y+z}{2}$, where $y$ and $z$ are independent draws from $X$ and $\xi$ is an independent $N(0,\eta^{2})$ random variable, and that mean was a deterministic flow in the direction of $\nu(x) \langle 1+ \mu, X \rangle - (1+\mu) X$.  We were interested in the fixed points of this flow, \ie measures $Z$ such that 
\[
	\nu(Z) = \frac{(1+\mu) Z}{\langle 1+ \mu, Z \rangle}.
\]
Here, I'm going to look at characterising such $Z$ in a few special cases.  I'm going to assume that $Z$ is a finite measure, and without loss of generality normalised to have total mass one.

Throughout, I'll be using the characteristic function of $Z$, which I'll denote by 
\[
	\phi(\theta) = \mathbb{E}\left[e^{i\theta Z}\right].
\]
First, consider the case without selection, so $\mu$ is a constant.  Then, our condition on the characteristic function simplifies to
\begin{equation}\label{CHARACTERISTIC}
	{\textstyle \phi\left(\frac{\theta}{2}\right)^{2}} e^{-\frac{\eta^{2}\theta^{2}}{2}} = \phi(\theta).
\end{equation}
Now, suppose that $Z$ has a finite first moment.  Then, since $\phi$ is a characteristic function, $\phi(0) = 1$, $\phi$ is continuously differentiable at 0, and $\phi$ is non-zero in a neighbourhood of 0.
In particular, $\psi(\theta) = \ln{\phi}(\theta)$ and its derivative are well defined in a neighbourhood 0, and the derivative is continuous at 0.  Then,
\[
	{\textstyle 2\psi\left(\frac{\theta}{2}\right)} -\frac{\eta^{2}\theta^{2}}{2} = \psi(\theta),
\]
and
\[ 
	\psi'\left(\frac{\theta}{2}\right) - \eta^{2} \theta = \psi(\theta)'.
\]
Iterating the latter, we have 
\[
	\psi(\theta)' = \psi'\left(\frac{\theta}{2^{n+1}}\right) - \eta^{2} \theta \sum_{i=0}^{n} 2^{-n},
\]
and taking limits as $n \to \infty$, we have
\[
	\psi(\theta)' = \psi'(0) - 2\eta^{2} \theta.
\]
Since $\phi(0) = 1$, we have $\psi'(0) = \phi'(0) = -i \mathbb{E}[Z] = m_{1}$, so 
\[
	\phi(\theta) = e^{-i m_{1} \theta - \eta^{2} \theta^{2}},
\]
which we recognise as the characteristic function of a point mass at $m_{1}$ if $\eta = 0$, and a Gaussian with mean $m_{1}$ and variance $2\eta^{2}$ otherwise.

We note that this does not exclude the possibility of distributions without a mean; \eg if $\eta =0$, then\eqref{CHARACTERISTIC} is satisfied by the family of Cauchy distributions with characteristic functions $\phi(\theta) = e^{i m_{1} \theta - \gamma |\theta|}$; one can check that these are the only stable laws satisfying \eqref{CHARACTERISTIC}.  One can similarly exclude the stable laws without mean ($\alpha \leq 1$) by inspection in the case of $\eta > 0$.

Now, consider the case when $1+ \mu(x) = e^{a x^{2}}$.  In this case there is a Gaussian solution: suppose $Z$ is Gaussian with mean $m_{1}$ and variance $\sigma^{2}$.  Then,
\[
	\phi(\theta) = e^{i m_{1} \theta - \frac{1}{2} \sigma^{2} \theta^{2}}.
\]
Moreover, provided $a < \frac{1}{2\sigma^{2}}$, we can explicitly compute the characteristic function of $\frac{(1+\mu) Z}{\langle 1+ \mu, Z \rangle}$ to conclude that
\[
	e^{i m_{1} \theta - \frac{1}{2} \left(\frac{\sigma^{2}}{2} + \eta^{2}\right)\theta^{2}}
	= e^{i \frac{m_{1}}{1-2 a \sigma^{2}}\theta 
		- \frac{1}{2} \frac{\sigma^{2}}{1-2 a \sigma^{2}}\theta^{2}},
\]
which, equating real and complex components, has a solution provided $m_{1} = 0$ and
\[
	\sigma^{2} = \frac{\sqrt{16 a^{2} \eta^{4} + 24 a \eta^{2}+1} - 4a \eta^{2} -1}{4a} 
	=  2\eta^{2}(1-8a\eta^{2}) + O(a^{2}).
\]

For uniqueness: suppose $1+\mu$ has a Fourier transform $\widehat{(1+\mu)}(-\theta)$. Then, the characteristic function becomes 
\[
	{\textstyle \phi\left(\frac{\theta}{2}\right)^{2}} e^{-\frac{\eta^{2}\theta^{2}}{2}} = 
	\frac{1}{\langle 1+\mu, Z\rangle} (\widehat{(1+\mu)} * \phi)(\theta).
\]
Can this be used to construct other examples or exclude those with a transform?

\bibliography{../../Global}
\bibliographystyle{plain}

\end{document}

### 2023-02-21 MeanFlow.tex
Pushed my local version of MeanFlow.tex

@@
 \author{Todd L. Parsons}
-\address{Laboratoire de Probabilit\'{e}s et Mod\`{e}les Al\'{e}atoires, CNRS UMR 7599, Universit\'{e} Pierre et Marie Curie, Paris, 75005, France.}
+\address{Laboratoire de Probabilit\'es, Statistique et Mod\'elisation, UMR8001, Sorbonne Universit\'e, Paris, 75005, France.}
 %\ead{todd.parsons@upmc.fr}
@@
 
-We had that $\nu(X)$ is the distribution of $\frac{y+z}{2}$, where $y$ and $z$ are independent draws from $X$ and $\xi$ is an independent $N(0,\eta^{2})$ random variable, and that mean was a deterministic flow in the direction of $\nu(x) \langle 1+ \mu, X \rangle - (1+\mu) X$.
+We had that $\nu(X)$ is the distribution of $\frac{y+z}{2}$, where $y$ and $z$ are independent draws from $X$ and $\xi$ is an independent  random variable with a fixed distribution (we will discuss suitable distributions below), and that mean was a deterministic flow in the direction of $\nu(X) \langle 1+ \mu, X \rangle - (1+\mu) X$.
 We were interested in the fixed points of this flow, \ie measures $Z$ such that
@@
 \]
-Here, I'm going to look at characterising such $Z$ in a few special cases.  I'm going to assume that $Z$ is a finite measure, and without loss of generality normalised to have total mass one.
+Here, I'm going to look at characterizing such $Z$ in a few special cases.  I'm going to assume that $Z$ is a finite measure, and without loss of generality normalized to have total mass one.
 
@@
 \[
-	\phi(\theta) = \mathbb{E}\left[e^{i\theta Z}\right].
+	\phi(\theta) = \mathbb{E}\left[e^{i\theta Z}\right],
 \]
+and the characteristic function of $\xi$, $\Phi(\theta)$.
+
+\section{No Selection}
+
 First, consider the case without selection, so $\mu$ is a constant.  Then, our condition on the characteristic function simplifies to
 \begin{equation}\label{CHARACTERISTIC}
-	{\textstyle \phi\left(\frac{\theta}{2}\right)^{2}} e^{-\frac{\eta^{2}\theta^{2}}{2}} = \phi(\theta).
+	{\textstyle \phi\left(\frac{\theta}{2}\right)^{2}} \Phi(\theta) %e^{-\frac{\eta^{2}\theta^{2}}{2}} 
+	= \phi(\theta),
 \end{equation}
-Now, suppose that $Z$ has a finite first moment.
+\ie $\phi(\theta)$ is a fixed point of $F$, where
+ \[
+	F(\phi)(\theta) = {\textstyle \phi\left(\frac{\theta}{2}\right)^{2}} \Phi(\theta).
+\]
+In particular, we can characterize all possible fixed points by considering the iterates of an arbitrary characteristic function, $\varphi(\theta)$, for  any distribution on $\mathbb{R}$.
+
+Iterating $F$ $n$ times, we get
+\[
+	F^{(n)}(\varphi)(\theta)
+	=  {\textstyle \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}}\prod_{k=1}^{n-1} {\textstyle \Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}},
+\]
+so that
+\[
+	\phi(\theta) = \lim_{n \to \infty} \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}\prod_{k=1}^{n-1}\Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}
+\]
+is a fixed point whenever this limit exists.
+
+Since $\varphi$ and $\Phi$ are characteristic functions, they are absolutely continuous and $\phi(0) = 1$, and thus non-zero in a neighbourhood of 0.
+In particular, $\psi(\theta) = \ln{\varphi}(\theta)$ and $\Psi(\theta) = \ln{\Phi(\theta)}$ exist in some neighbourhood of 0 as well.
+From the above, we have
+\[
+	\ln{F^{(n)}(\varphi)(\theta)} = {\textstyle 2^{n} \psi\left(\frac{\theta}{2^{n}}\right)} 
+		+ \sum_{k=1}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}.
+\]
+
+\subsection{Noise Terms}
+
+We now analyse the component arising from the noise $\xi$.
+First, we consider the case when $\Psi(\theta)$ is twice continuously differentiable, so $\xi$ has finite mean $m = \Psi'(0)$ and variance $\eta^{2 = \Psi''(0)}$.
+Taylor's theorem then tells us that for $\theta$ sufficiently small,
+\[
+	\Psi(\theta) = m\theta + \frac{\eta^{2}}{2} \theta^{2} + r(\theta)\theta^{2},
+\]
+where $r(\theta) \to 0$ as $\theta \to 0$.  Then,
+\begin{align*}
+	 \sum_{k=1}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}
+	 &=  \sum_{k=1}^{n-1} 2^{k} {\textstyle \left(m \frac{\theta}{2^{k}} - \frac{\eta^{2}}{2}  \frac{\theta^{2}}{2^{2k}}
+		+ r\left(\frac{\theta}{2^{k}}\right)\frac{\theta^{2}}{2^{2k}}\right)}\\
+	&= (n-1) m - \eta^{2} \theta^{2} (1 - 2^{-n+1}) 
+		+ \sum_{k=1}^{n-1} {\textstyle r\left(\frac{\theta}{2^{k}}\right)\frac{\theta^{2}}{2^{k}}}
+ \end{align*}
+
+Immediately, we see that for the limit to exist, we must have $m = 0$; if we assume that $r(\theta) \equiv 0$, so that the noise is $N(0,\eta^{2})$, then the limiting distribution is $N(0,2\eta^{2})$.
+We note that when  $r(\theta)$ is non-zero, we don't expect the higher order terms to vanish.
+
+Now, consider the case when $\Psi(\theta)$ corresponds to a stable law, so
+\[
+	\Psi(\theta) = im\theta + c|\theta|^{\alpha} (1-i\beta \text{sgn}(\theta)\omega_{\alpha}(\theta))
+\]
+for $\alpha \in (0,2]$, $\beta \in [-1,1]$, $c \geq 0$, and $m \in \mathbb{R}$ and
+\[
+	\omega_{\alpha}(\theta) = \begin{cases}
+		\tan{\frac{\pi\alpha}{2}} & \text{if $\alpha \neq 1$, and}\\
+		-\frac{2}{\pi}\ln{|\theta|} & \text{if $\alpha = 1$.}
+	\end{cases}
+\]
+We then have
+\begin{multline*}
+	 \sum_{k=1}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}
+	 =  \sum_{k=1}^{n-1} im\theta + 2^{(1-\alpha)k} c|\theta|^{\alpha} {\textstyle\left(1-i \beta\text{sgn}(\theta)\left(\frac{\theta}{2^{k}}\right)
+	 	 \omega_{\alpha}\left(\frac{\theta}{2^{k}}\right)\right)}\\
+	= 
+	\begin{cases}
+		 im(n-1)\theta + c|\theta|^{\alpha}\frac{1-2^{(1-\alpha)n}}{1-2^{1-\alpha}}\left(1-i\beta \text{sgn}(\theta) \tan{\frac{\pi\alpha}{2}}\right) & \text{if $\alpha \neq 1$, and}\\	
+		 \begin{multlined}
+		 im(n-1)\theta + c|\theta|^{\alpha}  \left(\frac{1-2^{(1-\alpha)n}}{1-2^{1-\alpha}}
+		 \left(1+i \beta \text{sgn}(\theta)\frac{2}{\pi}\ln{|\theta|} \right)\right.\\
+		\left. -i\beta \text{sgn}(\theta) \frac{2^{1-\alpha} + (n(2^{1-\alpha}-1)-1)2^{(1-\alpha)n})}{(1-2^{1-\alpha})^{2}}\frac{2}{\pi}\ln{2}
+		\right)    
+		\end{multlined}
+		& \text{if $\alpha = 1$.}
+	\end{cases}
+\end{multline*}
+This converges as $n \to \infty$ if and only if $m = 0$ and $\alpha > 1$, in which case the limit is 
+\[
+	 \frac{c}{1-2^{1-\alpha}}|\theta|^{\alpha}\left(1-i\beta \text{sgn}(\theta) \tan{\frac{\pi\alpha}{2}}\right), 
+\]
+which corresponds to a $\left(\alpha,\beta, \frac{c}{1-2^{1-\alpha}},0\right)$-stable law.
+
+%\subsection{Gaussian Distribution}
+
+%Now, suppose that $Z$ has a finite first moment.
 Then, since $\phi$ is a characteristic function, $\phi(0) = 1$, $\phi$ is continuously differentiable at 0, and $\phi$ is non-zero in a neighbourhood of 0.
-In particular, $\psi(\theta) = \ln{\phi}(\theta)$ and its derivative are well defined in a neighbourhood 0, and the derivative is continuous at 0.  Then,
+In particular, $\psi(\theta) = \ln{\phi}(\theta)$ and its derivative are well defined in a neighbourhood 0, and the derivative is continuous at 0.
+Then,
+%\[
+%	{\textstyle 2\psi\left(\frac{\theta}{2}\right)} -\frac{\eta^{2}\theta^{2}}{2} = \psi(\theta),
+%\]
+%and
+%\[ 
+%	\psi'\left(\frac{\theta}{2}\right) - \eta^{2} \theta = \psi(\theta)'.
+%\]
+%Iterating the latter, we have 
+%\[
+%	\psi(\theta)' = \psi'\left(\frac{\theta}{2^{n+1}}\right) - \eta^{2} \theta \sum_{i=0}^{n} 2^{-n},
+%\]
+%and taking limits as $n \to \infty$, we have
+%\[
+%	\psi'(\theta) = \psi'(0) - 2\eta^{2} \theta.
+%\]
+%Since $\phi(0) = 1$, we have $\psi'(0) = \phi'(0) = -i \mathbb{E}[Z] = m$, so 
+%\[
+%	\phi(\theta) = e^{-i m \theta - \eta^{2} \theta^{2}},
+%\]
+%which we recognize as the characteristic function of a point mass at $m$ if $\eta = 0$, and a Gaussian with mean $m$ and variance $2\eta^{2}$ otherwise.
+
+\subsection{Reproductive Terms}
+
+We now consider the terms $2^{n} \psi\left(\frac{\theta}{2^{n}}\right)$, which we call the reproductive terms.  Proceeding as below, we consider the limit
+%We note that this does not exclude the possibility of distributions without a mean; % \eg if $\eta =0$, then \eqref{CHARACTERISTIC} is satisfied by the family of Cauchy distributions with characteristic functions $\phi(\theta) = e^{i m \theta - \gamma |\theta|}$; one can check that these are the only stable laws satisfying \eqref{CHARACTERISTIC}.%  One can similarly exclude the stable laws without mean ($\alpha \leq 1$) by inspection in the case of $\eta > 0$.
+%Now, to exhaust all possibilities, observe that by iterating, we we also have that 
+%\[
+%	\psi(\theta) = 2^{n}\psi\left(\frac{\theta}{2^{n}}\right) 
+%\]
+%\ie we must have 
 \[
-	{\textstyle 2\psi\left(\frac{\theta}{2}\right)} -\frac{\eta^{2}\theta^{2}}{2} = \psi(\theta),
+		\alpha(\theta) = \lim_{n \to \infty} 2^{n}\psi\left(\frac{\theta}{2^{n}}\right).
 \]
-and
+Because $\psi(\theta)$ is the logarithm characteristic function, it must be uniformly continuous and vanish at 0, and thus $\alpha(\theta)$ as well.  We also have that for all integers $k$, 
 \[
-	\psi'\left(\frac{\theta}{2}\right) - \eta^{2} \theta = \psi(\theta)'.
+	\alpha(2^{k} \theta) = \lim_{n \to \infty} 2^{n}\psi\left(\frac{2^{k} \theta}{2^{n}}\right) =  \lim_{n \to \infty} 2^{k} 2^{n-k} \psi\left(\frac{\theta}{2^{n-k}}\right) = 2^{k} \alpha(\theta).
 \]
-Iterating the latter, we have 
+In particular, we see that given any continuous function $\alpha(\theta)$, $\theta \in [1,2]$, such that $\alpha(2) = 2\alpha(1)$, $\alpha$ is defined for all positive reals.
+We can extend the definition to the negative reals via the Hermitian property for $\phi(\theta)$: $\phi(-\theta) = \overline{\phi(\theta)}$ implies that $\alpha(-\theta) = \overline{\alpha(\theta)}$.
+
+Now, from the above, 
 \[
-	\psi(\theta)' = \psi'\left(\frac{\theta}{2^{n+1}}\right) - \eta^{2} \theta \sum_{i=0}^{n} 2^{-n},
+	\frac{\alpha(\theta)}{\theta} = \frac{\alpha(2^{k} \theta)}{2^{k}\theta} \to \alpha'(0+),
 \]
-and taking limits as $n \to \infty$, we have
+as $k \to - \infty$ if the latter exists (\ie there is a unique left-hand limit at zero).  In this case, setting $-\gamma + i m = \alpha'(0+)$, we have
 \[
-	\psi(\theta)' = \psi'(0) - 2\eta^{2} \theta.
+	\alpha(\theta) = \begin{cases} 
+		i m\theta - \gamma \theta & \text{if $\theta \geq 0$, and}\\
+		i m\theta + \gamma \theta & \text{if $\theta < 0$}.
+	\end{cases}
 \]
-Since $\phi(0) = 1$, we have $\psi'(0) = \phi'(0) = -i \mathbb{E}[Z] = m_{1}$, so 
+\ie if $\gamma > 0$, then $e^{\alpha(\theta)} = e^{im \theta -\gamma|\theta|}$ is the characteristic function of a Cauchy distribution and $\phi(\theta)$ is the characteristic function of the sum of Cauchy and Gaussian random variables, whereas if $\gamma = 0$, then $e^{\alpha(\theta)} = e^{im \theta}$ is the Fourier transform of a Dirac mass at $m$, so the limiting random variable is the Gaussian $N(m,\eta^{2})$.
+
+Finally, we remark that if $\frac{\alpha(\theta)}{\theta}$ is non-constant for $\theta > 0$, then $\alpha'(0+)$ does not exist.
+
+\subsection{Other Solutions?}
+
+We would naturally like to know if these two exhaust the possible solutions.
+To this end, we recall Bochner's theorem, which tells us that a function $\phi(\theta)$ is the characteristic function of a random variable if and only if
+\begin{enumerate}[(i)]
+\item $\phi(\theta)$ is uniformly continuous,
+\item $\phi(0) = 1$,
+\item $\phi(-\theta) = \overline{\phi(\theta)}$, and
+\item $\phi(\theta)$ is \textit{positive definite}: for all $n$ and all $\theta_{1},\ldots,\theta_{n} \in \mathbb{R}$, the matrix with entries
 \[
-	\phi(\theta) = e^{-i m_{1} \theta - \eta^{2} \theta^{2}},
+	a_{ij} = \phi(\theta_{i}-\theta_{j})
 \]
-which we recognise as the characteristic function of a point mass at $m_{1}$ if $\eta = 0$, and a Gaussian with mean $m_{1}$ and variance $2\eta^{2}$ otherwise.
+is positive definite.
+\end{enumerate}
+Given $\alpha(\theta)$ as above, the corresponding function $\phi(\theta) = e^{\alpha(\theta)-\eta^{2}\theta^{2}}$ 
+or $\phi(\theta) = e^{\alpha(\theta) - \frac{c}{1-2^{1-\alpha}}|\theta|^{\alpha}\left(1-i\beta \text{sgn}(\theta) \tan{\frac{\pi\alpha}{2}}\right)}$ satisfies the first three criteria.
+We will now seek to identify which $\alpha(\theta)$ result in positive definite $\phi(\theta)$.
+When $\alpha(\theta) = im \theta -\gamma|\theta|$ for $\gamma > 0$, this is immediate, as $e^{\alpha(\theta)}$ is then the characteristic function of a Cauchy random variable.
 
-We note that this does not exclude the possibility of distributions without a mean; \eg if $\eta =0$, then\eqref{CHARACTERISTIC} is satisfied by the family of Cauchy distributions with characteristic functions $\phi(\theta) = e^{i m_{1} \theta - \gamma |\theta|}$; one can check that these are the only stable laws satisfying \eqref{CHARACTERISTIC}.
-One can similarly exclude the stable laws without mean ($\alpha \leq 1$) by inspection in the case of $\eta > 0$.
+We can also easily exclude the case $\gamma < 0$ using a simple consequence of positive definiteness in the case when $n=2$ (name criterion): if $\phi(\theta)$ is positive definite, then $|\phi(\theta)| \leq 1$.
+For us, this requires that
+\[
+	\Re \alpha(\theta) \leq  \eta^{2} \theta^{2},\; resp. \; < \frac{c}{1-2^{1-\alpha}}|\theta|^{\alpha}
+\]
+which is violated for small values of $\theta$ when $\gamma < 0$ (recall $\alpha > 1$).
 
-Now, consider the case when $1+ \mu(x) = e^{a x^{2}}$.  In this case there is a Gaussian solution: suppose $Z$ is Gaussian with mean $m_{1}$ and variance $\sigma^{2}$.  Then,
+\noindent\emph{Question:} how can we use positive definiteness to exclude $\alpha(\theta)$ with $\Re \alpha(\theta) \leq  \eta^{2} \theta^{2}$ so that $\alpha'(0+)$ does not exist?
+
+\begin{rem}
+Polya has shown that if $\phi(\theta)$ is convex and satisfies the first three criteria of Bochner's theorem, then $\phi(\theta)$ is a characteristic function.
+In our case, if $\alpha(\theta)$ is convex, then $e^{\alpha(\theta)}$ is as well; a proof-by-picture, however, shows that the only choice of $\alpha(\theta)$ that is convex (or concave, for that matter) for $\theta > 0$ is a linear function, so Polya's criterion doesn't give us any new characteristic functions.
+\end{rem}
+
+%Further, we note that if $\alpha$ is differentiable at some $\theta_{0}> 0$, then 
+%\[
+%	 \alpha'(\theta_{0}) = \alpha'(2^{k} \theta_{0}) 
+%\]
+%for all integers $k$, so that either $\alpha'(\theta_{0}) =  \alpha'(0+)$, or $\alpha'(0+)$ is not defined.
+Similarly.
+for all $\theta_{0} < 0$, such that $\alpha'(\theta_{0})$ exists, either $\alpha'(\theta_{0}) =  \alpha'(0-)$, or $\alpha'(0-)$ is not defined.
+
+%(equivalently, let $\beta(x)$ be continuous and periodic with period $\ln{2}$, and set $\alpha(\theta) = e^{\beta(\ln{\theta})}$.
+
+%Moreover, provided we choose $\alpha(\theta) < 0$ for $\theta \in [1,2]$, we see that the corresponding  function
+
+%\subsection{Non-Gaussian Noise}
+
+%Now, suppose we assume some arbitrary distribution for the noise with characteristic function $\Phi(\theta)$. Then, our functional equation becomes
+%\[
+%	\phi(\theta) = {\textstyle \phi\left(\frac{\theta}{2}\right)^{2}} \Phi(\theta).
+%\]
+%Iterating this relation $n$ times, we get
+%\[
+%	\phi(\theta) =  {\textstyle \phi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}}\prod_{k=1}^{n-1} {\textstyle \Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}},
+%\]
+%or, taking logarithms, now taking $\Psi(\theta) = \ln{\psi(\theta)}$ as well, we have
+%
+%We have already seen that if $\Phi(\theta) = e^{-\frac{\eta^{2} \theta^{2}}{2}}$, then the latter sum converges to $-\eta^{2}\theta^{2}$, 
+%so 
+%\[
+%	\lim_{n \to \infty} \prod_{k=1}^{n-1} {\textstyle \Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}}
+%\]
+%is the characteristic function of a Gaussian of variance $2\eta^{2}$.
+
+%Now, suppose that $\Phi(\theta) = e^{-\gamma |\theta|^{\alpha}}$ is a stable law of Kolmogorov-Gnedenko type.  Then, 
+%\[
+%	\sum_{k=1}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}
+%	= -\gamma |\theta|^{\alpha} \sum_{k=1}^{n-1} 2^{k(1-\alpha)}
+%	= -\gamma |\theta|^{\alpha} \frac{1-2^{(1-\alpha)n}}{1-2^{1-\alpha}},
+%\]	
+%which has a finite limit of $-\frac{\gamma}{1-2^{1-\alpha}} |\theta|^{\alpha}$ if and only if $\alpha > 1$.
+We recognize this as a stable law as well; in particular, our previous analysis shows that we can find a fixed point that is now the sum of a Cauchy and a $\left(\frac{\gamma}{1-2^{1-\alpha}},\alpha\right)$-stable law.
+
+%\nb All of the analysis above holds if we had started with an arbitrary random $\phi(\theta)$ and looked at it's iterates under the map
+
+\section{Stabilizing Selection}
+
+\subsection{Gaussian Solution}
+
+Now, consider the case when $1+ \mu(x) = e^{a x^{2}}$.  In this case there is a Gaussian solution: suppose $Z$ is Gaussian with mean $m$ and variance $\sigma^{2}$.  Then,
 \[
-	\phi(\theta) = e^{i m_{1} \theta - \frac{1}{2} \sigma^{2} \theta^{2}}.
+	\phi(\theta) = e^{i m \theta - \frac{1}{2} \sigma^{2} \theta^{2}}.
 \]
-Moreover, provided $a < \frac{1}{2\sigma^{2}}$, we can explicitly compute the characteristic function of $\frac{(1+\mu) Z}{\langle 1+ \mu, Z \rangle}$ to conclude that
+Moreover, provided $a < \frac{1}{2\sigma^{2}}$, we can explicitly compute $\frac{\mathbb{E}\left[e^{i \theta Z}(1+\mu(Z))\right]}{\mathbb{E}[1+\mu(Z)}]$  to conclude that
 \[
-	e^{i m_{1} \theta - \frac{1}{2} \left(\frac{\sigma^{2}}{2} + \eta^{2}\right)\theta^{2}}
-	= e^{i \frac{m_{1}}{1-2 a \sigma^{2}}\theta 
+	e^{i m \theta - \frac{1}{2} \left(\frac{\sigma^{2}}{2} + \eta^{2}\right)\theta^{2}}
+	= e^{i \frac{m}{1-2 a \sigma^{2}}\theta 
 		- \frac{1}{2} \frac{\sigma^{2}}{1-2 a \sigma^{2}}\theta^{2}},
 \]
-which, equating real and complex components, has a solution provided $m_{1} = 0$ and
+which, equating real and complex components, has a solution provided $m = 0$ and
 \[
 	\sigma^{2} = \frac{\sqrt{16 a^{2} \eta^{4} + 24 a \eta^{2}+1} - 4a \eta^{2} -1}{4a} 
-	=  2\eta^{2}(1-8a\eta^{2}) + O(a^{2}).
+	=  2\eta^{2}(1-4a\eta^{2}) + O(a^{2}).
 \]
 
-For uniqueness: suppose $1+\mu$ has a Fourier transform $\widehat{(1+\mu)}(-\theta)$. Then, the characteristic function becomes 
+\subsection{Cauchy Solution}
+
+In the case of a Cauchy random variable, we again have an explicit expression for the probability distribution, which, proceeding as above, allows us to construct an explicit solution when $1+ \mu(x) = \frac{\delta^{2}}{\gamma^{2}} \frac{\gamma^{2} + x^{2}}{\delta^{2} + x^{2}}$ for $\delta > \gamma > 0$ and the noise is given by a $\text{Cauchy}(0,\delta-\gamma)$ random variable (\ie corresponding to a stable $(0,0,\delta-\gamma,0)$ law).
+
+If we now suppose that $Z$ is a $\text{Cauchy}(0,\gamma)$ random variable.  Then $Z$ has probability density function $\frac{\gamma}{\pi} \frac{1}{\gamma^{2} + x^{2}}$, whereas 
 \[
-	{\textstyle \phi\left(\frac{\theta}{2}\right)^{2}} e^{-\frac{\eta^{2}\theta^{2}}{2}} = 
-	\frac{1}{\langle 1+\mu, Z\rangle} (\widehat{(1+\mu)} * \phi)(\theta).
+	\frac{\mathbb{E}\left[e^{i \theta Z}(1+\mu(Z))\right]}{\mathbb{E}[1+\mu(Z)]}
+	= e^{-\delta |\theta|} = e^{-2\gamma \left|\frac{\theta}{2}\right|-(\delta-\gamma)|\theta|}
 \]
-Can this be used to construct other examples or exclude those with a transform?
+which we recognize as being equal to $\phi\left(\frac{\theta}{2}\right)^{2} \Phi(\theta)$.
+
+%(here, see a saturating cost to being away from the optimal phenotype $x = 0$: the mortality is maximized at  $\frac{\beta^{2}}{\alpha^{2}}$.
+
+\bigskip
+
+\noindent\emph{Question:} other stable laws?
+
+\section{Disruptive Selection}
+
+Now suppose $\mu$ has a Fourier transform $\hat{\mu}(\theta)$, \ie that $\mu(x)$ is integrable and thus vanishes at infinity.
+
+Then, the characteristic function satisfies 
+\[
+	{\textstyle \phi\left(\frac{\theta}{2}\right)^{2}} \Phi(\theta) =
+	\frac{ \phi(\theta) + (\widehat{\mu} * \phi)(\theta)} {\mathbb{E}[1+\mu(Z)]}.
+\]
+
+\noindent\emph{Question:} Can this be used to construct other examples or exclude some with a transform?
+
+\noindent\emph{Question:} If we assume that $\mu(x) = |x|^{\delta}$ (this is about the biggest selection function for which we can hope that $\langle 1+\mu, Z\rangle < \infty$), and that $\hat{\phi}(-x) = p(x)$,  then
+\[
+	\widehat{\mu(x)p(x)}  = \frac{\partial^{\delta}\phi}{\partial |\theta|^{\delta}} \quad \text{(Riesz derivative).}
+\]
+This then gives us 
+\[
+	{\textstyle \phi\left(\frac{\theta}{2}\right)^{2}} \Phi(\theta) =
+	\frac{ \phi(\theta) +\frac{\partial^{\delta}\phi}{\partial |\theta|^{\delta}}} {\mathbb{E}[1+\mu(Z)]}.
+\]
+Can we use this to find a solution?
 

### 2023-03-15 sims/parse_output.R
more sims

@@ -1,0 +1,59 @@
+#!/usr/bin/Rscript
+
+
+usage <- "
+   parse_output.R (tsv file) (number of time bins) 
+where (tsv file) should have a first line with parameters,
+and remaining lines of the form
+  (time) (trait values) (maternal trait values) (paternal trait values)
+"
+
+library(jsonlite)
+
+args <- commandArgs(TRUE)
+if (length(args) != 2) {
+    stop(usage)
+}
+
+fname <- args[1]
+num_plots <- as.numeric(args[2])
+
+infile <- file(fname, "r")
+params <- fromJSON(readLines(infile, 1))
+repro <- do.call(rbind, lapply(strsplit(readLines(infile), "\t"), function (x) {
+           x <- as.numeric(x); n <- (length(x) - 1) / 3; data.frame(t=x[1], x=x[1+(1:n)], ma=x[1+n+(1:n)], pa=x[1+2*n+(1:n)])
+}))
+close(infile)
+
+repro$seg <- repro$x - (repro$ma + repro$pa)/2
+repro$plot <- cut(repro$t, num_plots)
+
+if (params$type == "normal") {
+    qfun <- qnorm
+} else if (params$type == "cauchy") {
+    qfun <- qcauchy
+}
+
+pdf(paste0(fname, ".pdf"), width=8, height=2*num_plots, pointsize=10)
+layout(matrix(1:(3*num_plots), ncol=3, byrow=TRUE))
+xlim <- range(repro$seg)
+xh <- hist(repro$seg, breaks=200, plot=FALSE)
+for (k in levels(repro$plot)) {
+    title <- if (k == levels(repro$plot)[1]) toJSON(params) else ""
+    sub <- repro[repro$plot == k,]
+    hist(sub$seg, breaks=xh$breaks,
+         main=sprintf("times %0.2f-%0.2f", min(sub$t), max(sub$t)),
+         xlab="trait - midparent",
+         xlim=xlim
+    )
+    qqplot(qfun(ppoints(sub$seg)), sub$seg, ylab="trait value", xlab="cauchy quantiles", main=title)
+    qqline(sub$seg, distribution = function(p) qfun(p), probs = c(0.1, 0.6), col = 2)
+    x <- sort(abs(sub$seg))
+    tail_prob <- 1 - seq_along(x)/(length(x) + 1)
+    k <- max(sum(x==0) + 1, floor(length(x) * 3/4)) : length(x)
[…]

### 2023-05-31 writeup.tex
add document

\documentclass{article}
\usepackage{fullpage}
\usepackage{amsmath, amssymb}
\usepackage[hidelinks]{hyperref}
\usepackage{natbib}

\newcommand{\E}{\mathbb{E}}
\renewcommand{\P}{\mathbb{P}}

\newcommand{\comment}[1]{{\color{blue} \it #1}}

\begin{document}

% 1. *Intro*: Motivation about how many results about quant gen and methods assume Gaussian distributions.
%     Mention Levy walk macroevolution papers.
%     Point out that usual quant gen has *no* sweeps, while stable distributions would have sweeps (since some s would be bigger than 1/N).
%     Motivate with mixture model of effect sizes.
%     If stable distributions are a better model, what does that tell us? We can better understand the *noise* in evolution, i.e., predictability and saltation-ness.
%     (PETER)
% 
% 2. Motivation: empirical SNPs distributions in real data (PETER)
% 
% 3. Introducing infinitesimal model (in this case) (TODD)
% 
% 4. Proof following BEV in stable case (TODD)
% 
% 5. Theoretical results on stationary distributions (TODD)
% 
% 6. Simulations (PETER)

\title{\large{\bf
    Quantitative genetics with stable laws
}}

\author{ \begin{small}
    Todd Parsons \&
    Peter Ralph
\end{small}}

\date{\today}
\maketitle

\begin{abstract}
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

Much of the work to give a quantitative understanding to the observation that
offspring of sexual organisms resemble their parent,
has focused on \emph{quantitative genetics},
which is interestingly intermediate between a mechanistic model of genetics
and a purely phenomenological model.
The action of natural selection on genetic variants in the absence of genetic variation
is relatively well-understood (cite ewens),
but tractable mathematical models of evolution for recombining species
are elusive,
at least with some results of when the model might be a good description of reality.
For instance, a great deal of conceptual understanding and even quantitative predictions
have been made assuming that the distribution of trait values in a population is Gaussian
(cite Lande etcetera)
despite the absence of a plausible model that would make this true (cite turelli and barton).
One of the oldest and most concrete models of heritable trait evolution is the \emph{infinitesimal model}
which assumes a simple, explicit form for how an offspring's trait is inherited from their parents':
it is Normally distributed, with mean equal to the average of the parents' values.
Despite the inherent plausibility of this model,
(thanks to the Central Limit Theorem),
it was not until \citet{barton2017infinitesimal} that conditions under which this would be a good approximation
were made explicit,
through a formal convergence proof.

As suggested by its name, under the infinitesimal model the effects of any given genetic variant on the trait
is small (indeed, vanishing),
and so selection-driven trait shifts are mediated by small changes in frequencies of many alleles of small effect.
However, empirical understanding of the genetic basis of trait variation in real-world species
often finds a significant portion of heritable variation explained by variants at only a few loci.
There are many possible explanations for this observation,
including selection/publication bias,
filters for larger-effect variants due to spatially varying selection (cite),
ond/or grouping together of concordant alleles by linkage or inversions (cite).
However, it seems clear that in at least those cases where we understand the underlying molecular mechanisms
that mutations of large effect can be important in practice
(cite examples: insecticide resistance, coloration, etc),
and the consensus in the field is that traits exist on a spectrum between monogenic and polygenic,
and that many - if not most - are firmly in the intermediate, ``oligogenic'' region (cite),
where heritable trait variation is due both to segregating large- and small-effect variants.
Indeed, some sensible verbal models of the biological mechanisms by which a genetic change
comes to have an effect on an organism's trait
describe a hierarchy or range of ``proximities'' to the trait,
by which changes to more ``nearby'' genes have the possibility of causing larger changes in the trait
(cite omnigenic and others).
This suggests a mixture model for the effect size distribution:
perhaps the distribution of effect sizes within each gene is Normal,
but with a standard deviation that varies with ``proximity'' to the trait;
such mixtures can easily have large proportions of extreme values.

The infinitesimal model gains a great deal of generality by appealing to the Central Limit Theorem:
as shown by \citet{barton2017infinitesimal},
the distribution of segregation noise will be Normal, regardless of most of the details,
thanks to the universality of adding up a great many very small and independent things.
However, as is well known to probabilists,
there is a wider class of Central Limit Theorems,
which describe the distributions of sums of a great many independent and exchangeable things:
the \emph{stable laws}.
Roughly speaking, these say that if we add together $n$ independent copies of a random variable $X$
for which $\P\{ X > t \} \propto t^{-\alpha}$ for some $0 < \alpha < 2$,
then -- almost regardless of the details -- the result, scaled by $n^{1/\alpha}$,
is well-approximated by a universal form, the $\alpha$-stable distribution (and becomes exact, as $n \to \infty$).
In such ensembles, sinle large entries are important:
concretely, the largest of the $n$ copies will be of order $n^{1/\alpha}$,
so that e.g., for $\alpha=1$ a positive proportion of the sum of $n$ entries will be contributed by a single one.

This suggests exploring whether stable distributions might reasonably stand in for the Normal
in the infinitesimal model.
In this paper, we make rather preliminary explorations in that direction:
Does the universality shown by \citet{barton2017infinitesimal}
carry over when the distribution of effect sizes falls in the domain of attraction of a stable law?
What can be said about equilibrium trait distributions in this case?
And, is there any biological evidence that this theory might be relevant?

Stable distributions are by no means new to population biology.
For instance, it is a long-standing question how often the motions of organisms
are better modeled by essentially Brownian random walks or by L\'evy processes (cite),
and a growing literature studies the effects of such ``fat-tailed'' distributions
on the shapes and dynamics of traveling waves such as biological invasions (cite).
\comment{Other examples?}
More directly analogous to the problem we study here, \citet{landis2012phylogenetic}
found that an $\alpha$-stable process (with $\alpha \approx 1.5$) fit the evolution of log endocranial volume
on the primate phylogeny
better than either a Brownian model.
\comment{more}

\subsection*{Orphaned text}
The ways in which genetic variants combine to determine an organisms' trait can be complex,
and epistatic interactions between alleles and between loci can be important in practice.
However, the question is sufficiently complex in the purely additive case
that a trait can be decomposed into a sum across loci of the contributions of the alleles from each locus.
Under this additive assumption it makes sense to discuss the ``effect size distribution'',
i.e., the distribution of how much alleles affect the trait value.

An offspring's trait differs from their parents for many reasons,
often most importantly because of differences in the environment in which they live.
The focus of quantitative genetics is on the contribution of heritable material to this variation:
the offspring's genetic material

\comment{Will need this notation below:}
We focus on the \emph{infinitesimal model},
as described by \citet{barton2017infinitesimal},
which supposes that the trait value of an offspring
is equal to the average of their parents' trait values,
plus independent Gaussian noise with fixed variance,
\begin{equation} \label{eqn:offspring_model}
    X_o = \frac{X_m + X_f}{2} + \epsilon,
\end{equation}
where $\epsilon$, the ``segregation noise'',
is Gaussian with mean zero and variance $V_0$.
\comment{should call this something besides $\epsilon$}

%%%%%%%%%%%%%%%%%%%%%%
\section{Empirical observations}

It is mathematically intriguing to speculate what effect a non-Gaussian distribution
for $\epsilon$ would have on the dynamics induced by~\eqref{eqn:offspring_model}.
But, is there any evidence that non-Gaussian segregation noise is important in practice?
In this paper we focus primarily on the theory,
but first take a brief and imperfect look at what some data have to say on the topic.
To do this, we downloaded the GWAS results provided by~\citet{biobankSNPs}
for 7,221 phenotypes from 361,194 individuals from the March 2018 release of the UK Biobank.
Taken at face value, these provide for each phenotype a set of estimates of the additive effect
of the alternative allele at a set of SNPs on the phenotype relative to the reference allele.
The phenotypes we consider were all coded as 0/1,
so estimates could be thought of as additive adjustments to the risk
(so, \emph{not} on a logit scale).
These estimates were obtained as the coefficient for the SNP in a simple linear model
fit to each phenotype for each SNP,
with the first 20 principal components and all combinations of age, age squared, and inferred sex as covariates.
There are a great many possible issues with these data;
however, the results seem consistent across many SNPs and phenotypes,
and we know of no reason that possible methodological artifacts
(e.g., confounding by uncorrected variables)
would specifically affect the decay of the tail of effect sizes.

Suppose that the phenotype of each individual is additive: $X_f = \sum_i a_i$ and $X_m = \sum_i b_i$, for instance,
where each $a_i = 0$ if the individual is homozygous for the reference allele,
and that each contribution is inherited by the offspring with probability 1/2 (as if these were additive effects of rare, unlinked SNPs).
Then we would have the deviation of the offspring phenotype
from the midparent value equal to
$$
\epsilon = \sum_i (\theta_i a_i + \eta_i b_i)/2,
$$
where each $\theta_i$ and $\eta_j$ are independent Bernoulli(1/2).
So, the appropriate central limit theorem for $\epsilon$
is determined by the distribution of $\theta a_i$,
i.e.,

%%%%%%%%%%%%%%%%%%%%%%
\section{Extending the infinitesimal model}

%%%%%%%%%%%%%%%%%%%%%%
\section{Stationary distributions}

%%%%%%%%%%%%%%%%%%%%%%
\section{Simulations}

\bibliographystyle{plainnat}
\bibliography{refs}

\end{document}

### 2023-06-06 snp_effects/cauchy_distrn.Rmd
snp effects code

Let's check this works for the Cauchy.

```{r fake_data}
n <- 1e3
gwas <- data.frame(minor_AF=runif(n),
                   beta=rcauchy(n)/100)

min_num <- 20
max_prop <- 0.5
```

Now let's let $(P, Z)$ be a random draw from this table, all $Z > 0$,
and $X = Z$ with probability $P$, and $X=0$ otherwise.
Here's the complementary cumulative distribution function of $X$,
i.e., if $N$ is the number of rows in the table,
$$\begin{aligned}
   \mathbb{P}\{X > x \}  
   &= \frac{1}{N} \sum_i P_i \mathbb{1}_{Z_i > x} \\
   &= \mathbb{P}\{X > 0\} - \frac{1}{N} \sum_i P_i \mathbb{1}_{Z_i \le x} \\
   &= \frac{1}{N}\sum_i P_i - \frac{1}{N} \sum_i P_i \mathbb{1}_{Z_i \le x} .
\end{aligned}$$

```{r plotit}
estim_alpha <- function (gwas, num_breaks=1001) {
    xx <- c(seq(0, 2 * quantile(abs(gwas$beta), 0.99), length.out=num_breaks), Inf)
    disc_beta <- cut(abs(gwas$beta), xx)
    num_points <- table(disc_beta)
    dpx <- tapply(gwas$minor_AF, disc_beta, sum)
    dpx[is.na(dpx)] <- 0
    p_zero <- mean(gwas$minor_AF)
    tail_prob <- p_zero - cumsum(dpx) / nrow(gwas)
    tail_nums <- nrow(gwas) - cumsum(num_points)
    x <- xx[-1]
    xlm <- lm(log(tail_prob) ~ log(x),
              subset=(is.finite(log(tail_prob))
                      & tail_nums > min_num
                      & tail_nums < max_prop * nrow(gwas)))
    return(xlm)
}

xlm <- estim_alpha(gwas)
logx <- xlm$model[[1]]
logy <- xlm$model[[2]]
plot(logx, logy, ylab='log(prob > x)', xlab='log(x)')
abline(coef(xlm))

summary(xlm)
```

## Power

Huh, how good of an estimator is this?
```{r boot}
boots <- replicate(200, {
    gwas <- data.frame(minor_AF=runif(n),
                       beta=rcauchy(n)/100)
    xlm <- estim_alpha(gwas)
    return(coef(xlm)) })
hist(boots[2,], breaks=30)
abline(v=-1, col='red')
```
Looks ok.

## Gaussian

And, what if we give this Gaussian data?
```{r gboot}
boots <- replicate(200, {
    gwas <- data.frame(minor_AF=runif(n),
                       beta=rnorm(n)/100)
    xlm <- estim_alpha(gwas)
    return(coef(xlm)) })
hist(boots[2,], breaks=30)
abline(v=-1, col='red')
```
Looks ok.

### 2023-06-06 writeup.tex
one more section
snp effects code

@@
 \usepackage{natbib}
+\usepackage{xcolor}
 
@@
 
+\comment{TODO expand, put somewhere:}
+There are several justifications for why a ``stable infinitesimal model'' might be useful:
+(1) as a phenomenological model;
+(2) if it matches empirical observations; or
+(3) if there is a central limit theorem suggesting it would be a good approximation in a wide range of situations.
+
 \subsection*{Orphaned text}
@@
 so estimates could be thought of as additive adjustments to the risk
-(so, \emph{not} on a logit scale).
+(so, effects are \emph{not} on a logit scale).
 These estimates were obtained as the coefficient for the SNP in a simple linear model
@@
 
-Suppose that the phenotype of each individual is additive: $X_f = \sum_i a_i$ and $X_m = \sum_i b_i$, for instance,
-where each $a_i = 0$ if the individual is homozygous for the reference allele,
-and that each contribution is inherited by the offspring with probability 1/2 (as if these were additive effects of rare, unlinked SNPs).
-Then we would have the deviation of the offspring phenotype
-from the midparent value equal to
+Suppose that the phenotype of each individual is additive:
+the maternal trait value is $x = \sum_i x_{i,1} + x_{i,2}$
+and the paternal trait value is $y = \sum_i y_{i,1} + y_{i,2}$,
+where $x_{i,1}$ and $x_{i,2}$ are the effects of the alleles carried
+by the mother on each of the two chromosomes
+at the $i^\text{th}$ locus, respectively.
+Then the offspring's trait is $z = \sum_i z_{i,m} + z_{i,p}$,
+where $z_{i,m} = \theta_{i,m} x_{i,1} + (1-\theta_{i,m}) x_{i,2}$,
+and $\theta_{i,m} = 1$ with probability 1/2 and $\theta_{i,m} = 0$ otherwise;
+and likewise $z_{i,p} = \theta_{i,p} y_{i,1} + (1-\theta_{i,p}) y_{i,2}$.
+With this notation, the segregation noise $z - \frac{x+y}{2}$
+is a sum of terms of the form
 $$
-\epsilon = \sum_i (\theta_i a_i + \eta_i b_i)/2,
+    \left( \theta_{i,m} - 1/2 \right) x_{i,1}
+    + \left( 1/2 - \theta_{i,m} \right) x_{i,2} 
 $$
-where each $\theta_i$ and $\eta_j$ are independent Bernoulli(1/2).
-So, the appropriate central limit theorem for $\epsilon$
-is determined by the distribution of $\theta a_i$,
-i.e.,
+(and similar terms with $y$ in place of $x$).
+Now let $X$ be the effect of a uniformly chosen allele carried by a uniformly chosen individual
+at a uniformly chosen locus.
+The appropriate central limit theorem is determined by the tail behavior of $|(\theta - 1/2) X| = |X|/2$;
+i.e., we'd like to know
+$$
+    \lim_{t \to \infty} \frac{ \log \P\{ |X| > 2 t \} }{ \log t } = \alpha ,
+$$
+if the limit exists.
+Let $p_i$ be the minor allele frequency at the $i^\text{th}$ locus;
+since the data we have assigns effect 0 to the major allele,
+$x_i = 0$ with probability $1-p_i$.
+Therefore, if we have estimated effects for $N$ SNPs
+with minor allele frequency $p_i$ and effect size $e_i$ for the minor allele,
+for $1 \le i \le N$,
+then
+$$
+    \P\{ |X| > 2t \} \approx \frac{1}{N} \sum_{i : |e_i| > 2t} p_i .
+$$
+So, to obtain a per-trait estimate of $\alpha$,
+we plot $\frac{1}{N} \sum_{i : |e_i| > t} p_i$ against $t$
+for a sequence of values of $t$,
+and fit a linear model to the tail
+(which we take to be all values of $t$ for which the number of SNPs with absolute effect above $t$
+is between 20 and 10\% of all SNPs).
+Examples are shown in figure \comment{XXX}.
 
 %%%%%%%%%%%%%%%%%%%%%%
-\section{Extending the infinitesimal model}
+\section{Extending the infinitesimal model}  % TODD
 
 %%%%%%%%%%%%%%%%%%%%%%
-\section{Stationary distributions}
+\section{Convergence to the non-infinitesimal model}  % TODD
 
 %%%%%%%%%%%%%%%%%%%%%%
-\section{Simulations}
+\section{Stationary distributions}  % TODD
+
+%%%%%%%%%%%%%%%%%%%%%%
+\section{Simulations}  % PETER
 

### 2023-06-06 snp_effects/get_results.R

@@ -1,0 +1,26 @@
+#!/usr/bin/env Rscript
+
+## Original run with tail prob 25%
+results <- data.frame(code=setdiff(list.dirs(".", recursive=FALSE, full.names=FALSE), ".git"), stringsAsFactors=FALSE)
+coefs <- t(sapply(results$code, function (x) { f <- file.path(x, "xlm.RData"); if (file.exists(f)) { load(f); coef(xlm) } else { c(NA, NA) } } ))
+results$intercept <- coefs[,1]
+results$slope <- coefs[,2]
+results$nsnps <- sapply(results$code, function (x) { f <- file.path(x, "done"); if (file.exists(f)) { scan(f) } else { NA } } )
+write.table(results, file="snp_results_25.tsv")
+
+## Next run with tail prob 10%
+results <- data.frame(code=setdiff(list.dirs(".", recursive=FALSE, full.names=FALSE), ".git"), stringsAsFactors=FALSE)
+
+max_prop <- 0.1
+filebase <- sprintf("tail_%d_", floor(100*max_prop))
+basename <- file.path(results$code, filebase)
+donefile <- paste0(basename, "done")
+xlm_file <- paste0(basename, "xlm.RData")
+
+coefs <- t(sapply(xlm_file, function (f) { if (file.exists(f)) { load(f); coef(xlm) } else { c(NA, NA) } } ))
+results$intercept <- coefs[,1]
+results$slope <- coefs[,2]
+results$nsnps <- sapply(donefile, function (f) { if (file.exists(f)) { scan(f) } else { NA } } )
+write.table(results, file="snp_results_10.tsv")
+
+

### 2023-06-06 snp_effects/plot_results.R

@@ -1,0 +1,23 @@
+#!/usr/bin/env Rscript
+
+## First run with prop 25%
+results <- read.table("snp_results_25.tsv")
+pdf(file="results_25.pdf", width=6, height=8, pointsize=10)
+layout((1:2))
+with(subset(results, slope > -4), {
+    hist(slope, breaks=50, xlab='exponent', main="tail proportion: 25%")
+    plot(nsnps, slope, pch=20, cex=0.5, col=adjustcolor('black', 0.5))
+})
+dev.off()
+
+
+## Next run with prop 10%
+results <- read.table("snp_results_10.tsv")
+pdf(file="results_10.pdf", width=6, height=8, pointsize=10)
+layout((1:2))
+with(subset(results, slope > -4), {
+    hist(slope, breaks=50, xlab='exponent', main="tail proportion: 10%")
+    plot(nsnps, slope, pch=20, cex=0.5, col=adjustcolor('black', 0.5))
+})
+dev.off()
+

### 2023-06-06 snp_effects/snp_distrn.R

@@ -1,0 +1,120 @@
+library(data.table)
+
+# The distribution we are drawing X from is:
+#  1. pick a SNP at random
+#  2. flip a coin; with prob 1-minor_AF return 0
+#  3. otherwise return beta
+# and so with N SNPs
+#   P(X > x)  = sum(minor_AF[beta > x])/N
+# We are looking at abs(beta).
+
+pval_cutoff <- 1e-8
+min_num <- 20
+# max_prop <- 0.25 ## FIRST VERSION WAS THIS
+max_prop <- 0.1
+
+filebase <- sprintf("tail_%d_", floor(100*max_prop))
+
+estim_alpha <- function (gwas, num_breaks=1001) {
+    # For a sequence of value xi,
+    # fits the linear model
+    #  log(pxi) ~ log(xi)
+    # where
+    #  pxi = prob( a randomly chosen locus has effect above xi )
+    # to only the 'tail', defined as those xi such that there
+    # are at least min_num loci with effect above xi
+    # and at most a proportion max_prop of loci with effect above xi.
+    xx <- c(seq(0, 2 * quantile(abs(gwas$beta), 0.99), length.out=num_breaks), Inf)
+    disc_beta <- cut(abs(gwas$beta), xx)
+    num_points <- table(disc_beta)
+    dpx <- tapply(gwas$minor_AF, disc_beta, sum)
+    dpx[is.na(dpx)] <- 0
+    p_zero <- mean(gwas$minor_AF)
+    tail_prob <- p_zero - cumsum(dpx) / nrow(gwas)
+    tail_nums <- nrow(gwas) - cumsum(num_points)
+    x <- xx[-1]
+    use_these <- (is.finite(log(tail_prob))
+                      & tail_nums > min_num
+                      & tail_nums < max_prop * nrow(gwas))
+    if (sum(use_these) > min_num) {
+        xlm <- lm(log(tail_prob) ~ log(x), subset=use_these)
+    } else {
+        xlm <- NULL
+    }
+    return(xlm)
+}
+
+manifest <- subset(read.csv("gwas_manifest.csv", header=TRUE, na.string="N/A", stringsAsFactors=FALSE),
+                   Sex=="both_sexes" & !is.na(Phenotype.Code) & !grepl("_irnt", Phenotype.Code)
+                   & !grepl("^20003_", Phenotype.Code) # drug treatments
+                   & !grepl("22617_", Phenotype.Code) # job codes
[…]

### 2023-06-06 snp_effects/snp_distrn.R
.

@@ -86,17 +86,20 @@ for (k in 1:nrow(manifest)) {
     stopifnot(all(gwas$pval < pval_cutoff))
     if (ncol(gwas) == 0) {
         cat("Failed to load", code, "\n")
+        cat(nrow(gwas), "\n", file=donefile)
         next
     }
     gwas <- subset(gwas, !is.na(beta) & !is.na(minor_AF) & (beta != 0) & (minor_AF != 0))
     if (nrow(gwas) <= min_num) {
         cat("Not enough hits in", code, "\n")
+        cat(nrow(gwas), "\n", file=donefile)
         next
     }
 
     xlm <- estim_alpha(gwas)
     if (is.null(xlm)) {
-        cat("Failed to do regression for", code, "\n")
+        cat("Failed to fit linear model for", code, "\n")
+        cat(nrow(gwas), "\n", file=donefile)
         next
     }
     save(xlm, file=xlm_file)

### 2023-06-06 snp_effects/snp_distrn.R
.

@@ -60,15 +60,16 @@ translation <- read.csv("gwas_manifest_lookup.csv", header=FALSE)
 
 for (k in 1:nrow(manifest)) {
     code <- manifest$Phenotype.Code[k]
-    if (!dir.exists(code)) dir.create(code)
-    basename <- file.path(code, filebase)
+    code_dir <- file.path("phenotypes", "code")
+    if (!dir.exists(code_dir)) dir.create(code_dir)
+    basename <- file.path(code_dir, filebase)
     donefile <- paste0(basename, "done")
     xlm_file <- paste0(basename, "xlm.RData")
     plot_file <- paste0(basename, "plot.pdf")
 
     gwas_file <- file.path(code, manifest$File[k])
     if (!file.exists(gwas_file)) {
-        wgot <- system(paste("cd", code, "&&", manifest$wget.command[k]))
+        wgot <- system(paste("cd", code_dir, "&&", manifest$wget.command[k]))
         if (wgot != 0) {
             cat("Failed to get", code, "\n")
             next

### 2023-06-06 snp_effects/plot_results.R
more analysis of effects

@@ -1,5 +1,15 @@
 #!/usr/bin/env Rscript
 
+manifest <- subset(read.csv("gwas_manifest.csv", header=TRUE, na.string="N/A", stringsAsFactors=FALSE),
+                   Sex=="both_sexes" & !is.na(Phenotype.Code) & !grepl("_irnt", Phenotype.Code)
+                   & !grepl("^20003_", Phenotype.Code) # drug treatments
+                   & !grepl("22617_", Phenotype.Code) # job codes
+                   & !grepl("22601_", Phenotype.Code) # also job codes
+            )
+# remove doubles
+doubles <- names(table(manifest$Phenotype.Code))[table(manifest$Phenotype.Code) == 2]
+manifest <- manifest[-(match(doubles,manifest$Phenotype.Code)),]
+
 ## First run with prop 25%
 results <- read.table("snp_results_25.tsv")
 pdf(file="results_25.pdf", width=6, height=8, pointsize=10)
@@ -13,6 +23,10 @@ dev.off()
 
 ## Next run with prop 10%
 results <- read.table("snp_results_10.tsv")
+results$name <- manifest$Phenotype.Description[match(results$code, manifest$Phenotype.Code)]
+results <- subset(results, !grepl("Job code", results$name))
+results <- subset(results, !grepl("Type of", results$name)) # seems to be all diet-related
+
 pdf(file="results_10.pdf", width=6, height=8, pointsize=10)
 layout((1:2))
 with(subset(results, slope > -4), {
@@ -21,3 +35,20 @@ with(subset(results, slope > -4), {
 })
 dev.off()
 
+categories <- c(
+    "Non-cancer illness code",
+    "Cancer code",
+    "neoplasm",
+    "ICD10"
+)
+
+pdf("cat_results_10.pdf", width=6, height=9)
+for (cname in categories) {
+    with(subset(results, grepl(cname, name)), {
+        layout((1:3))
+        hist(slope, breaks=50, xlab='exponent', main=cname)
+        hist(slope[slope > -4], breaks=50, xlab='exponent', main="tail proportion: 10%")
+        plot(nsnps, slope, pch=20, cex=0.5, col=adjustcolor('black', 0.5))
+    })
+}
+dev.off()

### 2023-06-07 snp_effects/plot_results.R
more plotting

@@ -10,6 +10,8 @@ manifest <- subset(read.csv("gwas_manifest.csv", header=TRUE, na.string="N/A", s
 doubles <- names(table(manifest$Phenotype.Code))[table(manifest$Phenotype.Code) == 2]
 manifest <- manifest[-(match(doubles,manifest$Phenotype.Code)),]
 
+phenotypes <- read.table("phenotypes.both_sexes.tsv", stringsAsFactors=FALSE, sep="\t", quote="", header=TRUE, comment.char="")
+
 ## First run with prop 25%
 results <- read.table("snp_results_25.tsv")
 pdf(file="results_25.pdf", width=6, height=8, pointsize=10)
@@ -24,31 +26,97 @@ dev.off()
 ## Next run with prop 10%
 results <- read.table("snp_results_10.tsv")
 results$name <- manifest$Phenotype.Description[match(results$code, manifest$Phenotype.Code)]
+results$name[results$name == "NA"] <- NA
 results <- subset(results, !grepl("Job code", results$name))
 results <- subset(results, !grepl("Type of", results$name)) # seems to be all diet-related
 
-pdf(file="results_10.pdf", width=6, height=8, pointsize=10)
-layout((1:2))
-with(subset(results, slope > -4), {
-    hist(slope, breaks=50, xlab='exponent', main="tail proportion: 10%")
-    plot(nsnps, slope, pch=20, cex=0.5, col=adjustcolor('black', 0.5))
-})
-dev.off()
+pheno_names <- c("phenotype", "n_non_missing", "n_missing", "n_controls", "n_cases")
+pheno <- phenotypes[,pheno_names]
+names(pheno) <- c("code", pheno_names[-1])
+stopifnot(all(results$code %in% pheno$code))
+results <- merge(results, pheno, all.x=TRUE)
+# very common things (with > 10000 cases) have Something Different going on
+results$pruned <- !(results$n_non_missing > 300000 & results$n_cases < 10000)
 
 categories <- c(
     "Non-cancer illness code",
-    "Cancer code",
     "neoplasm",
-    "ICD10"
+    "ICD10",
+    "other"
+)
+results$category <- ifelse(
+    grepl(categories[1], results$name), categories[1],
+    ifelse(
+        grepl(categories[2], results$name), categories[2],
+        ifelse(
+            grepl(categories[3], results$name), categories[3], categories[4]
+        )
[…]

### 2023-06-07 snp_effects/plot_results.R
page of examples

@@ -24,11 +24,12 @@ dev.off()
 
 
 ## Next run with prop 10%
-results <- read.table("snp_results_10.tsv")
+results <- read.table("snp_results_10.tsv", stringsAsFactors=FALSE)
 results$name <- manifest$Phenotype.Description[match(results$code, manifest$Phenotype.Code)]
 results$name[results$name == "NA"] <- NA
 results <- subset(results, !grepl("Job code", results$name))
 results <- subset(results, !grepl("Type of", results$name)) # seems to be all diet-related
+results <- subset(results, !grepl("Own or rent", results$name))
 
 pheno_names <- c("phenotype", "n_non_missing", "n_missing", "n_controls", "n_cases")
 pheno <- phenotypes[,pheno_names]
@@ -89,17 +90,33 @@ for (cname in categories) {
 }
 dev.off()
 
-plot_code <- function (code) {
+clean_name <- function (x) {
+    clean <- c("Non-cancer illness code, self-reported: ", 
+               "Diagnoses - main ICD10: ",
+               "Cancer code, self-reported: "
+    )
+    for (u in clean) {
+        x <- gsub(u, "", x)
+    }
+    return(x)
+}
+
+plot_code <- function (code, lab) {
     ij <- match(code, results$code)
     if (is.na(ij)) stop(sprintf("Code %s not in results.", code))
     code_dir <- file.path("phenotypes", code)
     xlm_file <- file.path(code_dir, paste0("tail_10_", "xlm.RData"))
     if (!file.exists(xlm_file)) stop(sprintf("File %s does not exist.", xlm_file))
+    load(xlm_file)
     logx <- xlm$model[["log(x)"]]
     logy <- xlm$model[["log(tail_prob)"]]
-    plot(logx, logy, ylab='log(prob > x)', xlab='log(x)', main=results$name[ij], sub=code)
-    abline(coef(xlm))
-    legend("topright", lty=c(1, NA, NA),
+    plot(logx, logy, ylab='log(prob > x)', xlab='log(x)', type='l')#, sub=code)
+    cnp <- strsplit(clean_name(results$name[ij]), " ")[[1]]
+    title(paste(c(lab, cnp[1:min(5, length(cnp))]), collapse=" "), cex.main=0.9)
+    if (length(cnp) > 5)
+        title(paste(cnp[6:length(cnp)], collapse=" "), line=0.2, cex.main=0.9)
[…]

### 2023-06-07 snp_effects/snp_distrn.R

@@ -60,7 +60,7 @@ translation <- read.csv("gwas_manifest_lookup.csv", header=FALSE)
 
 for (k in 1:nrow(manifest)) {
     code <- manifest$Phenotype.Code[k]
-    code_dir <- file.path("phenotypes", "code")
+    code_dir <- file.path("phenotypes", code)
     if (!dir.exists(code_dir)) dir.create(code_dir)
     basename <- file.path(code_dir, filebase)
     donefile <- paste0(basename, "done")

### 2023-06-11 writeup.tex
histograms of results

@@
 \usepackage{xcolor}
+\usepackage{graphics}
 
@@
 for 7,221 phenotypes from 361,194 individuals from the March 2018 release of the UK Biobank.
+After filtering (see details below) we were left with 695 binary illness-related phenotypes.
 Taken at face value, these provide for each phenotype a set of estimates of the additive effect
 of the alternative allele at a set of SNPs on the phenotype relative to the reference allele.
-The phenotypes we consider were all coded as 0/1,
+The phenotypes we consider were all binary and coded as 0/1,
 so estimates could be thought of as additive adjustments to the risk
-(so, effects are \emph{not} on a logit scale).
+(i.e., effects are \emph{not} on a logit scale).
 These estimates were obtained as the coefficient for the SNP in a simple linear model
 fit to each phenotype for each SNP,
-with the first 20 principal components and all combinations of age, age squared, and inferred sex as covariates.
+with the first 20 principal components and all combinations of age, age squared, and inferred sex as covariates;
+we considered effects of those SNPs with a reported $p$-value less than $10^{-8}$.
 There are a great many possible issues with these data;
@@
 at a uniformly chosen locus.
-The appropriate central limit theorem is determined by the tail behavior of $|(\theta - 1/2) X| = |X|/2$;
-i.e., we'd like to know
+The appropriate central limit theorem is determined by the tail behavior of $|(\theta - 1/2) X| = |X|/2$.
+So, we'd like to know the value of
 $$
@@
 So, to obtain a per-trait estimate of $\alpha$,
-we plot $\frac{1}{N} \sum_{i : |e_i| > t} p_i$ against $t$
+we plot $\log \frac{1}{N} \sum_{i : |e_i| > t} p_i$ against $\log t$
 for a sequence of values of $t$,
@@
 is between 20 and 10\% of all SNPs).
-Examples are shown in figure \comment{XXX}.
+Examples are shown in figure~\ref{fig:example_snps}.
+
+\begin{figure}
+    \begin{center}
+    \includegraphics{snp_effects/examples}
+    \end{center}
+    \caption{
+        TODO: cut this down to just 2 SNPs and move this figure to the supplement; change 'x' to 't' in axis labels.
+        Tail distributions of frequency-weighted effects
+        for twelve randomly chosen phenotypes.
+        In each, for a phenotype with $N$ SNPs,
+        effect sizes $e_i$ and frequencies $p_i$,
+        the black line shows
+        $\log \frac{1}{N} \sum_{i : |e_i| > t} p_i$ (vertical axis)
+        against $\log t$ (horizontal axis);
+        the red line shows the least-squares linear fit;
+        each plot only covers the 10\% largest-effect SNPs.
+        Legends show the coefficients of the red line; the slope on `x' is $\alpha$,
+        the estimated exponent;
+        ``num snps'' is the number of SNPs for the phenotype with $p$-value less than $10^{-8}$,
+        and ``num cases'' is the reported number of the rougly 360,000 individuals
+        recorded as having the listed phenotype.
+        \label{fig:example_snps}
+    }
+\end{figure}
+
+We applied this procedure to results from models fit on both sexes,
+after removed phenotypes related to employment, drug treatments, and diet.
+This left us with 1,108 phenotypes for which which could estimate the tail exponent.
+For the main results,
+we then restricted to phenotypes having non-missing numbers of ``cases'' and ``controls''
+and a sample size of at least 300,000 (i.e., not more than 61,194 of 361,194 missing phenotypes).
+We also removed
+the 65 remaining phenotypes that are common (more than 10,000 cases),
+as these often showed very different patterns
+(e.g., effect size distributions).
+This left us with 695 phenotypes;
+plots including the 413 ``filtered'' phenotypes are provided in
+Supplementary Figure~\ref{sfig:unfiltered_hist}.
+
+Resulting values of the tail exponent $\alpha$ are shown in Figure~\ref{fig:exponent_hist}.
+Estimated tail exponents range from about 1 to 2.5,
+and 28.9\% had estimated values less than 1.5.
+A concern could be that ``noisier'' phenotypes,
+for which estimated effect sizes are less reliable,
+might lead to larger estimated tail exponents;
+for this reasons, we looked for an association between
+tail exponent and two proxies for power,
+number of SNPs (i.e., number of SNPs with $p$-value less than $10^{-8}$)
+and number of cases.
+Interestingly, the result in nonmonotonic,
+with exponents closer to $\alpha=2$ for phenotypes with around 1000 cases,
+and lower exponents for rarer \emph{and} more common phenotypes.
+(A similar pattern is seen for ``number of SNPs'',
+but this may be due to association with number of cases.)
+This may indicate a statistical artifact,
+or it may be a result of biological differences in genetic architecture
+between more and less common phenotypes.
+
+\begin{figure}
+    \begin{center}
+    \includegraphics{snp_effects/results_10}
+    \end{center}
+    \caption{
+        Estimated values of the tail exponent, $\alpha$,
+        for 775 illness-related binary phenotypes:
+        \textbf{(A)} distribution of values; and plotted against
+        \textbf{(B)} number of SNPs, and
+        \textbf{(C)} number of cases.
+        \label{fig:exponent_hist}
+    }
+\end{figure}
 
@@
 
+\appendix
+
+\begin{figure}
+    \begin{center}
+    \includegraphics{snp_effects/unfiltered_results_10}
+    \end{center}
+    \caption{
+        Estimated values of the tail exponent, $\alpha$,
+        for 1108 illness-related binary phenotypes,
+        including the 333 phenotypes removed in filtering,
+        which are shown as red points in lower plots.
+        \textbf{(A)} distribution of values; and plotted against
+        \textbf{(B)} number of SNPs,
+        \textbf{(C)} number of cases, and
+        \textbf{(C)} number of non-missing subjects.
+        \label{fig:unfiltered_hist}
+    }
+\end{figure}
+
 \end{document}

### 2023-06-11 snp_effects/plot_results.R

@@ -25,11 +25,17 @@ dev.off()
 
 ## Next run with prop 10%
 results <- read.table("snp_results_10.tsv", stringsAsFactors=FALSE)
+results <- subset(results, !is.na(slope))
 results$name <- manifest$Phenotype.Description[match(results$code, manifest$Phenotype.Code)]
 results$name[results$name == "NA"] <- NA
 results <- subset(results, !grepl("Job code", results$name))
 results <- subset(results, !grepl("Type of", results$name)) # seems to be all diet-related
 results <- subset(results, !grepl("Own or rent", results$name))
+results <- subset(results, !grepl("employment", results$name))
+results <- subset(results, !grepl("Milk type", results$name))
+results <- subset(results, !grepl("Spread type", results$name))
+results <- subset(results, !grepl("Bread type", results$name))
+results <- subset(results, !grepl("Types of transport", results$name))
 
 pheno_names <- c("phenotype", "n_non_missing", "n_missing", "n_controls", "n_cases")
 pheno <- phenotypes[,pheno_names]
@@ -37,7 +43,7 @@ names(pheno) <- c("code", pheno_names[-1])
 stopifnot(all(results$code %in% pheno$code))
 results <- merge(results, pheno, all.x=TRUE)
 # very common things (with > 10000 cases) have Something Different going on
-results$pruned <- !(results$n_non_missing > 300000 & results$n_cases < 10000)
+results$pruned <- !(!is.na(results$n_cases) & results$n_non_missing > 300000 & results$n_cases < 10000)
 
 categories <- c(
     "Non-cancer illness code",
@@ -59,24 +65,40 @@ for (cname in categories) {
     pdf(sprintf("%s_results_10.pdf", gsub(" ", "_", cname)), width=7, height=6)
     with(subset(results, category == cname), {
         layout(matrix(c(1,2,1,3,1,4), nrow=2))
-        hist(slope, breaks=30, xlab='exponent', main=cname)
-        plot(nsnps, slope, pch=20, cex=0.5, xlab='number of snps',
+        hist(slope, breaks=30, xlab='tail exponent', main=cname)
+        plot(nsnps, slope, pch=20, cex=0.5, xlab='number of snps', ylab='tail exponent',
[…]

### 2023-06-11 writeup.tex
copied over text

@@
 \usepackage{fullpage}
-\usepackage{amsmath, amssymb}
+\usepackage{amsmath,amssymb,amsthm,bm,enumerate,mathtools}
 \usepackage[hidelinks]{hyperref}
@@
 \renewcommand{\P}{\mathbb{P}}
+\newcommand{\mL}{\mathcal{L}}
+\newcommand{\ip}[2]{\left\langle #1, #2 \right\rangle}
+\newcommand{\ie}{\textit{i.e.,}\,}
+\newcommand{\eg}{\textit{e.g.,}\,}
+\newcommand{\nb}{\textit{n.b.,}\, }
+\newcommand{\defn}{:=}
 
 \newcommand{\comment}[1]{{\color{blue} \it #1}}
+
+\newtheorem{thm}{Theorem}
+\newtheorem{lem}{Lemma}
+\newtheorem{prop}{Proposition}
+\newtheorem{cor}{Corollary}
+\theoremstyle{remark} 
+	\newtheorem{rem}{Remark}
+	\newtheorem{assn}{Assumption}
+\theoremstyle{definition} 
+	\newtheorem{mydef}{Definition} 
+	\newtheorem{exmp}{Example} 
+	\newtheorem{cond}{Condition}
+	\newtheorem{conj}{Conjecture}
 
@@
 
+\comment{Copied from \texttt{eugene\_notes.tex}:}
+
+Let $X$ be a population of size $N$, recorded as the empirical distribution on phenotypes
+\begin{align*}
+    X = \frac{1}{N} \sum_{i=1}^N \delta_{x_i} .
+\end{align*}
+This evolves as a Moran process under the infinitesimal model
+with death rates depending on state $x$.
+Concretely, an individual whose phenotype is $x$
+dies at rate $1 + \mu(x)$
+and is replaced by a new individual of phenotype
+\begin{align*}
+  \frac{y + z}{2} + \xi ,
+\end{align*}
+where $y$ and $z$ are independent draws from $X$
+and $\xi$ is an independent $N(0, \eta^2)$.
+Let $\nu(X)$ denote the distribution of $\frac{y + z}{2} + \xi$.
+This has generator
+\begin{align*}
+    \mL \Phi(X)
+    &=
+    \int \int \int \int
+    \left( \Phi(X + \frac{1}{N} \delta_{(y+z)/2 + \xi} - \frac{1}{N} \delta_x) - \Phi(X) \right)
+    (1 + \mu(x)) X(dx) X(dy) X(dz) e^{-(\xi/\eta)^2/2} d\xi \\
+    &=
+    \int \int
+    \left( \Phi(X + \frac{1}{N} \delta_u - \frac{1}{N} \delta_x) - \Phi(X) \right)
+    (1 + \mu(x)) X(dx) \nu(X)(du) .
+\end{align*}
+
+For a test function $\phi$, setting $\Phi(X) = \ip{\phi}{X}$
+we get the first moment
+\begin{align*}
+    \mL \ip{\phi}{X}
+    &=
+    \frac{1}{N} \int \int \int \int
+    \left( \phi \left( \frac{y+z}{2} + \xi \right)
+         - \phi(x) \right)
+    (1 + \mu(x)) X(dx) X(dy) X(dz) e^{-(\xi/\eta)^2/2} d\xi  \\
+    &=
+    \frac{1}{N} \ip{\phi}{\nu(X)}\ip{1+\mu}{X}
+    - \frac{1}{N} \ip{\phi}{(1+\mu) X} .
+\end{align*}
+Note that $\ip{\phi}{\partial_{\nu(X)} X} = \ip{\phi}{\nu(X)}$,
+so that this is a first derivative in the direction of
+$\ip{1+\mu}{X} \nu(X) - (1+\mu) X$.
+
+More generally, if $\Phi(X) = \prod_i \ip{\phi_i}{X}$ then
+\begin{align*}
+    \mL \Phi(X)
+    &=
+    \frac{1}{N} \sum_i \left( 
+        \ip{1+\mu}{X} \ip{\phi_i}{\nu(X)}  - \ip{\phi_i}{(1+\mu)X}
+    \right)
+    \prod_{j \neq i} \ip{\phi_j}{X} 
+    + O(1/N^2) .
+\end{align*}
+This implies that after rescaling time by $N$
+and suitible other caveats,
+the limiting process is a deterministic flow in the direction of 
+$\nu(X) \ip{1+\mu}{X} - (1+\mu)X$.
+
+\paragraph{Stable measures}
+If $Z$ is a fixed point of this flow,
+then $\nu(Z)\ip{1+\mu}{Z} = (1+\mu)Z$.
+Any measure $Z$ that is a fixed point has the property that
+(the average of two independent draws plus segregation noise)
+is equal in distribution to
+(a $1+\mu$-weighted draw).
+Concretely, with $A$, $B$, and $C$ independent draws from $Z$,
+$$\begin{aligned}
+    \E\left[ f\left( \frac{A+B}{2} + \xi \right) \right]
+    &=
+    \frac{\E\left[ f(C) (1+\mu(C)) \right]}{\E[1+\mu(C)]} .
+\end{aligned}$$
+
+\comment{Copied from \texttt{MeanFlow.tex}:}
+
+Assume that $Z$ is a finite measure, and without loss of generality normalized to have total mass one.
+
+Throughout, I'll be using the characteristic function of $Z$, which I'll denote by 
+\[
+	\phi(\theta) = \mathbb{E}\left[e^{i\theta Z}\right],
+\]
+and the characteristic function of $\xi$, $\Phi(\theta)$.
+
+\section{No Selection}
+
+First, consider the case without selection, so $\mu$ is a constant.  Then, our condition on the characteristic function simplifies to
+\begin{equation}\label{CHARACTERISTIC}
+	{\textstyle \phi\left(\frac{\theta}{2}\right)^{2}} \Phi(\theta) %e^{-\frac{\eta^{2}\theta^{2}}{2}} 
+	= \phi(\theta),
+\end{equation}
+\ie $\phi(\theta)$ is a fixed point of $F$, where
+ \[
+	F(\phi)(\theta) = {\textstyle \phi\left(\frac{\theta}{2}\right)^{2}} \Phi(\theta).
+\]
+In particular, we can characterize all possible fixed points by considering the iterates of an arbitrary characteristic function, $\varphi(\theta)$, for  any distribution on $\mathbb{R}$.
+
+Iterating $F$ $n$ times, we get
+\[
+	F^{(n)}(\varphi)(\theta)
+	=  {\textstyle \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}}\prod_{k=1}^{n-1} {\textstyle \Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}},
+\]
+so that
+\[
+	\phi(\theta) = \lim_{n \to \infty} \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}\prod_{k=1}^{n-1}\Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}
+\]
+is a fixed point whenever this limit exists.
+
+Since $\varphi$ and $\Phi$ are characteristic functions, they are absolutely continuous and $\phi(0) = 1$, and thus non-zero in a neighbourhood of 0.
+In particular, $\psi(\theta) = \ln{\varphi}(\theta)$ and $\Psi(\theta) = \ln{\Phi(\theta)}$ exist in some neighbourhood of 0 as well.
+From the above, we have
+\[
+	\ln{F^{(n)}(\varphi)(\theta)} = {\textstyle 2^{n} \psi\left(\frac{\theta}{2^{n}}\right)} 
+		+ \sum_{k=1}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}.
+\]
+
+\subsection{Noise Terms}
+
+We now analyse the component arising from the noise $\xi$.
+First, we consider the case when $\Psi(\theta)$ is twice continuously differentiable, so $\xi$ has finite mean $m = \Psi'(0)$ and variance $\eta^{2 = \Psi''(0)}$.
+Taylor's theorem then tells us that for $\theta$ sufficiently small,
+\[
+	\Psi(\theta) = m\theta + \frac{\eta^{2}}{2} \theta^{2} + r(\theta)\theta^{2},
+\]
+where $r(\theta) \to 0$ as $\theta \to 0$.  Then,
+\begin{align*}
+	 \sum_{k=1}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}
+	 &=  \sum_{k=1}^{n-1} 2^{k} {\textstyle \left(m \frac{\theta}{2^{k}} - \frac{\eta^{2}}{2}  \frac{\theta^{2}}{2^{2k}}
+		+ r\left(\frac{\theta}{2^{k}}\right)\frac{\theta^{2}}{2^{2k}}\right)}\\
+	&= (n-1) m - \eta^{2} \theta^{2} (1 - 2^{-n+1}) 
+		+ \sum_{k=1}^{n-1} {\textstyle r\left(\frac{\theta}{2^{k}}\right)\frac{\theta^{2}}{2^{k}}}
+ \end{align*}
+
+Immediately, we see that for the limit to exist, we must have $m = 0$; if we assume that $r(\theta) \equiv 0$, so that the noise is $N(0,\eta^{2})$, then the limiting distribution is $N(0,2\eta^{2})$.
+We note that when  $r(\theta)$ is non-zero, we don't expect the higher order terms to vanish.
+
+Now, consider the case when $\Psi(\theta)$ corresponds to a stable law, so
+\[
+	\Psi(\theta) = im\theta + c|\theta|^{\alpha} (1-i\beta \text{sgn}(\theta)\omega_{\alpha}(\theta))
+\]
+for $\alpha \in (0,2]$, $\beta \in [-1,1]$, $c \geq 0$, and $m \in \mathbb{R}$ and
+\[
+	\omega_{\alpha}(\theta) = \begin{cases}
+		\tan{\frac{\pi\alpha}{2}} & \text{if $\alpha \neq 1$, and}\\
+		-\frac{2}{\pi}\ln{|\theta|} & \text{if $\alpha = 1$.}
+	\end{cases}
+\]
+We then have
+\begin{multline*}
+	 \sum_{k=1}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}
+	 =  \sum_{k=1}^{n-1} im\theta + 2^{(1-\alpha)k} c|\theta|^{\alpha} {\textstyle\left(1-i \beta\text{sgn}(\theta)\left(\frac{\theta}{2^{k}}\right)
+	 	 \omega_{\alpha}\left(\frac{\theta}{2^{k}}\right)\right)}\\
+	= 
+	\begin{cases}
+		 im(n-1)\theta + c|\theta|^{\alpha}\frac{1-2^{(1-\alpha)n}}{1-2^{1-\alpha}}\left(1-i\beta \text{sgn}(\theta) \tan{\frac{\pi\alpha}{2}}\right) & \text{if $\alpha \neq 1$, and}\\	
+		 \begin{multlined}
+		 im(n-1)\theta + c|\theta|^{\alpha}  \left(\frac{1-2^{(1-\alpha)n}}{1-2^{1-\alpha}}
+		 \left(1+i \beta \text{sgn}(\theta)\frac{2}{\pi}\ln{|\theta|} \right)\right.\\
+		\left. -i\beta \text{sgn}(\theta) \frac{2^{1-\alpha} + (n(2^{1-\alpha}-1)-1)2^{(1-\alpha)n})}{(1-2^{1-\alpha})^{2}}\frac{2}{\pi}\ln{2}
+		\right)    
+		\end{multlined}
+		& \text{if $\alpha = 1$.}
+	\end{cases}
+\end{multline*}
+This converges as $n \to \infty$ if and only if $m = 0$ and $\alpha > 1$, in which case the limit is 
+\[
+	 \frac{c}{1-2^{1-\alpha}}|\theta|^{\alpha}\left(1-i\beta \text{sgn}(\theta) \tan{\frac{\pi\alpha}{2}}\right), 
+\]
+which corresponds to a $\left(\alpha,\beta, \frac{c}{1-2^{1-\alpha}},0\right)$-stable law.
+
+\subsection{Reproductive Terms}
+
+We now consider the terms $2^{n} \psi\left(\frac{\theta}{2^{n}}\right)$, which we call the reproductive terms.  Proceeding as below, we consider the limit
+%We note that this does not exclude the possibility of distributions without a mean; % \eg if $\eta =0$, then \eqref{CHARACTERISTIC} is satisfied by the family of Cauchy distributions with characteristic functions $\phi(\theta) = e^{i m \theta - \gamma |\theta|}$; one can check that these are the only stable laws satisfying \eqref{CHARACTERISTIC}.%  One can similarly exclude the stable laws without mean ($\alpha \leq 1$) by inspection in the case of $\eta > 0$.
+%Now, to exhaust all possibilities, observe that by iterating, we we also have that 
+%\[
+%	\psi(\theta) = 2^{n}\psi\left(\frac{\theta}{2^{n}}\right) 
+%\]
+%\ie we must have 
+\[
+		\alpha(\theta) = \lim_{n \to \infty} 2^{n}\psi\left(\frac{\theta}{2^{n}}\right).
+\]
+Because $\psi(\theta)$ is the logarithm characteristic function, it must be uniformly continuous and vanish at 0, and thus $\alpha(\theta)$ as well.  We also have that for all integers $k$, 
+\[
+	\alpha(2^{k} \theta) = \lim_{n \to \infty} 2^{n}\psi\left(\frac{2^{k} \theta}{2^{n}}\right) =  \lim_{n \to \infty} 2^{k} 2^{n-k} \psi\left(\frac{\theta}{2^{n-k}}\right) = 2^{k} \alpha(\theta).
+\]
+In particular, we see that given any continuous function $\alpha(\theta)$, $\theta \in [1,2]$, such that $\alpha(2) = 2\alpha(1)$, $\alpha$ is defined for all positive reals.
+We can extend the definition to the negative reals via the Hermitian property for $\phi(\theta)$: $\phi(-\theta) = \overline{\phi(\theta)}$ implies that $\alpha(-\theta) = \overline{\alpha(\theta)}$.
+
+Now, from the above, 
+\[
+	\frac{\alpha(\theta)}{\theta} = \frac{\alpha(2^{k} \theta)}{2^{k}\theta} \to \alpha'(0+),
+\]
+as $k \to - \infty$ if the latter exists (\ie there is a unique left-hand limit at zero).  In this case, setting $-\gamma + i m = \alpha'(0+)$, we have
+\[
+	\alpha(\theta) = \begin{cases} 
+		i m\theta - \gamma \theta & \text{if $\theta \geq 0$, and}\\
+		i m\theta + \gamma \theta & \text{if $\theta < 0$}.
+	\end{cases}
+\]
+\ie if $\gamma > 0$, then $e^{\alpha(\theta)} = e^{im \theta -\gamma|\theta|}$ is the characteristic function of a Cauchy distribution and $\phi(\theta)$ is the characteristic function of the sum of Cauchy and Gaussian random variables, whereas if $\gamma = 0$, then $e^{\alpha(\theta)} = e^{im \theta}$ is the Fourier transform of a Dirac mass at $m$, so the limiting random variable is the Gaussian $N(m,\eta^{2})$.
+
+Finally, we remark that if $\frac{\alpha(\theta)}{\theta}$ is non-constant for $\theta > 0$, then $\alpha'(0+)$ does not exist.
+
+\subsection{Other Solutions?}
+
+We would naturally like to know if these two exhaust the possible solutions.
+To this end, we recall Bochner's theorem, which tells us that a function $\phi(\theta)$ is the characteristic function of a random variable if and only if
+\begin{enumerate}[(i)]
+\item $\phi(\theta)$ is uniformly continuous,
+\item $\phi(0) = 1$,
+\item $\phi(-\theta) = \overline{\phi(\theta)}$, and
+\item $\phi(\theta)$ is \textit{positive definite}: for all $n$ and all $\theta_{1},\ldots,\theta_{n} \in \mathbb{R}$, the matrix with entries
+\[
+	a_{ij} = \phi(\theta_{i}-\theta_{j})
+\]
+is positive definite.
+\end{enumerate}
+Given $\alpha(\theta)$ as above, the corresponding function $\phi(\theta) = e^{\alpha(\theta)-\eta^{2}\theta^{2}}$ 
+or $\phi(\theta) = e^{\alpha(\theta) - \frac{c}{1-2^{1-\alpha}}|\theta|^{\alpha}\left(1-i\beta \text{sgn}(\theta) \tan{\frac{\pi\alpha}{2}}\right)}$ satisfies the first three criteria.
+We will now seek to identify which $\alpha(\theta)$ result in positive definite $\phi(\theta)$.
+When $\alpha(\theta) = im \theta -\gamma|\theta|$ for $\gamma > 0$, this is immediate, as $e^{\alpha(\theta)}$ is then the characteristic function of a Cauchy random variable.
+
+We can also easily exclude the case $\gamma < 0$ using a simple consequence of positive definiteness in the case when $n=2$ (name criterion): if $\phi(\theta)$ is positive definite, then $|\phi(\theta)| \leq 1$.
+For us, this requires that
+\[
+	\Re \alpha(\theta) \leq  \eta^{2} \theta^{2},\; resp. \; < \frac{c}{1-2^{1-\alpha}}|\theta|^{\alpha}
+\]
+which is violated for small values of $\theta$ when $\gamma < 0$ (recall $\alpha > 1$).
+
+\noindent\emph{Question:} how can we use positive definiteness to exclude $\alpha(\theta)$ with $\Re \alpha(\theta) \leq  \eta^{2} \theta^{2}$ so that $\alpha'(0+)$ does not exist?
+
+\begin{rem}
+Polya has shown that if $\phi(\theta)$ is convex and satisfies the first three criteria of Bochner's theorem, then $\phi(\theta)$ is a characteristic function.
+In our case, if $\alpha(\theta)$ is convex, then $e^{\alpha(\theta)}$ is as well; a proof-by-picture, however, shows that the only choice of $\alpha(\theta)$ that is convex (or concave, for that matter) for $\theta > 0$ is a linear function, so Polya's criterion doesn't give us any new characteristic functions.
+\end{rem}
+
+\subsection{Stabilizing Selection}
+
+\subsubsection{Gaussian Solution}
+
+Now, consider the case when $1+ \mu(x) = e^{a x^{2}}$.  In this case there is a Gaussian solution: suppose $Z$ is Gaussian with mean $m$ and variance $\sigma^{2}$.  Then,
+\[
+	\phi(\theta) = e^{i m \theta - \frac{1}{2} \sigma^{2} \theta^{2}}.
+\]
+Moreover, provided $a < \frac{1}{2\sigma^{2}}$, we can explicitly compute $\frac{\mathbb{E}\left[e^{i \theta Z}(1+\mu(Z))\right]}{\mathbb{E}[1+\mu(Z)}]$  to conclude that
+\[
+	e^{i m \theta - \frac{1}{2} \left(\frac{\sigma^{2}}{2} + \eta^{2}\right)\theta^{2}}
+	= e^{i \frac{m}{1-2 a \sigma^{2}}\theta 
+		- \frac{1}{2} \frac{\sigma^{2}}{1-2 a \sigma^{2}}\theta^{2}},
+\]
+which, equating real and complex components, has a solution provided $m = 0$ and
+\[
+	\sigma^{2} = \frac{\sqrt{16 a^{2} \eta^{4} + 24 a \eta^{2}+1} - 4a \eta^{2} -1}{4a} 
+	=  2\eta^{2}(1-4a\eta^{2}) + O(a^{2}).
+\]
+
+\subsubsection{Cauchy Solution}
+
+In the case of a Cauchy random variable, we again have an explicit expression for the probability distribution, which, proceeding as above, allows us to construct an explicit solution when $1+ \mu(x) = \frac{\delta^{2}}{\gamma^{2}} \frac{\gamma^{2} + x^{2}}{\delta^{2} + x^{2}}$ for $\delta > \gamma > 0$ and the noise is given by a $\text{Cauchy}(0,\delta-\gamma)$ random variable (\ie corresponding to a stable $(0,0,\delta-\gamma,0)$ law).
+
+If we now suppose that $Z$ is a $\text{Cauchy}(0,\gamma)$ random variable.  Then $Z$ has probability density function $\frac{\gamma}{\pi} \frac{1}{\gamma^{2} + x^{2}}$, whereas 
+\[
+	\frac{\mathbb{E}\left[e^{i \theta Z}(1+\mu(Z))\right]}{\mathbb{E}[1+\mu(Z)]}
+	= e^{-\delta |\theta|} = e^{-2\gamma \left|\frac{\theta}{2}\right|-(\delta-\gamma)|\theta|}
+\]
+which we recognize as being equal to $\phi\left(\frac{\theta}{2}\right)^{2} \Phi(\theta)$.
+
+%(here, see a saturating cost to being away from the optimal phenotype $x = 0$: the mortality is maximized at  $\frac{\beta^{2}}{\alpha^{2}}$.
+
+\bigskip
+
+\noindent\emph{Question:} other stable laws?
+
 %%%%%%%%%%%%%%%%%%%%%%

### 2023-06-11 snp_effects/plot_results.R

@@ -65,12 +65,12 @@ for (cname in categories) {
     pdf(sprintf("%s_results_10.pdf", gsub(" ", "_", cname)), width=7, height=6)
     with(subset(results, category == cname), {
         layout(matrix(c(1,2,1,3,1,4), nrow=2))
-        hist(slope, breaks=30, xlab='tail exponent', main=cname)
-        plot(nsnps, slope, pch=20, cex=0.5, xlab='number of snps', ylab='tail exponent',
+        hist(-slope, breaks=30, xlab='tail exponent', main=cname)
+        plot(nsnps, -slope, pch=20, cex=0.5, xlab='number of snps', ylab='tail exponent',
              col=adjustcolor(ifelse(pruned, "red", "black"), 0.5))
-        plot(n_cases, slope, pch=20, cex=0.5, xlab='number of cases', ylab='tail exponent',
+        plot(n_cases, -slope, pch=20, cex=0.5, xlab='number of cases', ylab='tail exponent',
              col=adjustcolor(ifelse(pruned, "red", "black"), 0.5))
-        plot(n_non_missing, slope, pch=20, cex=0.5, xlab='number of non-missing subjects', ylab='tail exponent',
+        plot(n_non_missing, -slope, pch=20, cex=0.5, xlab='number of non-missing subjects', ylab='tail exponent',
              col=adjustcolor(ifelse(pruned, "red", "black"), 0.5))
     })
     dev.off()
@@ -79,12 +79,12 @@ for (cname in categories) {
 pdf("unfiltered_results_10.pdf", width=7, height=6, pointsize=10)
 with(results, {
     layout(matrix(c(1,2,1,3,1,4), nrow=2))
-    hist(slope, breaks=30, xlab='tail exponent', main='')
-    plot(nsnps, slope, pch=20, cex=0.5, xlab='number of snps', ylab='tail exponent',
+    hist(-slope, breaks=30, xlab='tail exponent', main='')
+    plot(nsnps, -slope, pch=20, cex=0.5, xlab='number of snps', ylab='tail exponent',
          col=adjustcolor(ifelse(pruned, "red", "black"), 0.5))
-    plot(n_cases, slope, pch=20, cex=0.5, xlab='number of cases', ylab='tail exponent',
+    plot(n_cases, -slope, pch=20, cex=0.5, xlab='number of cases', ylab='tail exponent',
[…]

### 2023-06-11 writeup.tex
start at simulation whatnot

@@
 \subsection{Stabilizing Selection}
+    \label{sec:stabilizing_selection}
 
 \subsubsection{Gaussian Solution}
+    \label{sec:stabilizing_gaussian}
 
@@
 \]
-Moreover, provided $a < \frac{1}{2\sigma^{2}}$, we can explicitly compute $\frac{\mathbb{E}\left[e^{i \theta Z}(1+\mu(Z))\right]}{\mathbb{E}[1+\mu(Z)}]$  to conclude that
+Moreover, provided $a < \frac{1}{2\sigma^{2}}$, we can explicitly compute $\frac{\mathbb{E}\left[e^{i \theta Z}(1+\mu(Z))\right]}{\mathbb{E}[1+\mu(Z)]}$  to conclude that
+if $\xi$ is Gaussian with mean zero and variance $\eta^2$,
+then equation~\eqref{CHARACTERISTIC} is
+\comment{something not right here? need a factor of 2 on the left?}
 \[
@@
 \subsubsection{Cauchy Solution}
+    \label{sec:stabilizing_cauchy}
 
-In the case of a Cauchy random variable, we again have an explicit expression for the probability distribution, which, proceeding as above, allows us to construct an explicit solution when $1+ \mu(x) = \frac{\delta^{2}}{\gamma^{2}} \frac{\gamma^{2} + x^{2}}{\delta^{2} + x^{2}}$ for $\delta > \gamma > 0$ and the noise is given by a $\text{Cauchy}(0,\delta-\gamma)$ random variable (\ie corresponding to a stable $(0,0,\delta-\gamma,0)$ law).
+In the case of a Cauchy random variable, we again have an explicit expression for the probability distribution, which, proceeding as above, allows us to construct an explicit solution when $1+ \mu(x) = \frac{\delta^{2}}{\gamma^{2}} \frac{\gamma^{2} + x^{2}}{\delta^{2} + x^{2}}$ for $\delta > \gamma > 0$ and the noise $\xi$ is given by a $\text{Cauchy}(0,\delta-\gamma)$ random variable (\ie corresponding to a stable $(0,0,\delta-\gamma,0)$ law).
 
@@
 \]
-which we recognize as being equal to $\phi\left(\frac{\theta}{2}\right)^{2} \Phi(\theta)$.
+which we recognize as being equal to $\phi\left(\frac{\theta}{2}\right)^{2} \Phi(\theta)$,
+and so solves equation~\eqref{CHARACTERISTIC}.
 
@@
 \section{Simulations}  % PETER
+
+In Section~\ref{sec:stabilizing_selection}
+we have found two explicit conditions in which
+if the form of segregation noise and stabilizing selection
+satisfy certain conditions,
+then the limiting phenotype distribution
+will be either Gaussian (Subsection~\ref{sec:stabilizing_gaussian}
+or Cauchy (Subsection~\ref{sec:stabilizing_cauchy}.
+Furthermore, we expect based on
+\citet{barton2017infinitesimal} and arguments above
+that the distribution of segreation noise
+is determined by the tail exponent of the distribution of effect sizes
+(under appropriate assumptions; e.g., that most mutations are of small effect).
+However, it is not clear how selection would affect these arguments,
+especially in the stable case where some number of mutations are expected to
+\emph{not} have small effect (and therefore such mutations will not be nearly neutral).
+
+Therefore,
+to study how this works in practice, we turn to simulation.
+We used SLiM v3 \comment{(cite)}
+to simulate a discrete-time approximation to the continuous-time Moran model described above
+as follows.
+Individuals are diploid and sexual;
+the genome is of lenth $10^8$bp with a uniform recombination rate of $10^{-8}$ crossovers per bp;
+mutations occur at a rate of $m$ per bp.
+Each new mutation is independently assigned an ``effect''
+drawn from the ``effect size distribution'' (and will be either Gaussian or Cauchy).
+The phenotype of each individual is determined additively
+by summing the effects of all mutations they carry across both chromosomes.
+Let $f(z) = 1 + \mu(z)$
+be the ``fitness'' of an individual with phenotype $z$,
+and let $dt$ be a small constant (we took $dt=0.01$).
+Then, at each time step
+each individual dies with probability $1 - \exp(-dt f(z))$;
+if there are $k$ such individuals chosen to die in a given time step,
+then we choose $k$ individuals uniformly and without replacement
+to produce new offspring;
+for each such reproduction event the parent chooses a mate uniformly at random from the population.
+This maintains the population at a fixed size, $N$.
+(Note that individuals may self and individuals chosen to die may also reproduce,
+but in a large population these details should be unimportant.)
+
+We compared two situations:
+(1) the effect size distribution was Gaussian with mean zero and standard deviation $\sigma_g$,
+and $f(z) = e^{\alpha z^2}$; and
+(2) the effect size distribution was Cauchy, centered at zero and with width $\delta - \gamma$
+and $f(z) = (\delta/\gamma) (\gamma + z^2) / (\delta + z^2)$.
+\comment{TODO: just refer to equations above?}
+We ran simulations of each with $N=XXX$ for XXX time steps of ``burn-in''
+(judged to be long enough as the number of segregating mutations had plateaued),
+following which we recorded the phenotypes of each new offspring and their two parents
+for XXX time steps (a total of around YYY offspring).
+
+Figure XXX shows the resulting distributions of segregation noise:
+i.e., the distribution of child phenotype minus average of parents' phenotypes,
+across all offspring following burn-in.
 

### 2023-06-12 sims/make_figs.R
more plots

@@ -1,0 +1,96 @@
+#!/usr/bin/Rscript
+
+usage <- "
+   parse_output.R (basename)
+where basename.repro.tsv and basename.pop.tsv and basename.fix.tsv should be files,
+and the 'repro' file should have a first line with parameters,
+and remaining lines of the form
+  (time) (trait values) (maternal trait values) (paternal trait values)
+the 'pop' file should have 'trait' and 'age' columns,
+and the 'fix' file should have 'time' and 'num_fixations' columns.
+"
+
+library(jsonlite)
+
+args <- commandArgs(TRUE)
+if (length(args) != 1) {
+    stop(usage)
+}
+
+basename <- args[1]
+
+infile <- file(paste0(basename, ".repro.tsv"), "r")
+params <- fromJSON(readLines(infile, 1))
+repro <- do.call(rbind, lapply(strsplit(readLines(infile), "\t"), function (x) {
+           x <- as.numeric(x); n <- (length(x) - 1) / 3; data.frame(t=x[1], x=x[1+(1:n)], ma=x[1+n+(1:n)], pa=x[1+2*n+(1:n)])
+}))
+close(infile)
+
+pop <- read.table(paste0(basename, ".pop.tsv"), sep="\t", header=TRUE)
+pop$age <- pop$age * params$DT
+fix <- read.table(paste0(basename, ".fix.tsv"), sep="\t", header=TRUE)
+
+repro$midp <- (repro$ma + repro$pa)/2
+repro$seg <- repro$x - repro$midp
+
+# pdf(sprintf("%s_independence.pdf", basename), width=6.5, height=4, pointsize=10)
+png(sprintf("%s_independence.png", basename), width=6.5, height=4, pointsize=10, units='in', res=144)
+{
+    layout(matrix(c(1,1,1,2,2,2,3,3,4,4,5,5), ncol=2), widths=c(1.5,3))
+    par(mar=c(4,4,1,1), mgp=c(2.5, 1, 0))
+    # subset for plotting in a way that won't affect the plot
+    # ut <- sample.int(nrow(repro), 3e4)
+    # ut <- sort(unique(c(ut, which(
+    #            pmax(rank(abs(repro$midp)), rank(abs(repro$seg))) > nrow(repro) * .99
+    # ))))
+    ut <- sample.int(nrow(repro),
+                prob=pmax(rank(abs(repro$midp)), rank(abs(repro$seg)))^3,
+                size=3e4
[…]

### 2023-06-21 sims/do_figs.sh
sim params

@@ -1,0 +1,12 @@
+#!/bin/bash
+
+for type in normal cauchy
+do
+    for xopt in 0.0 1.0 2.0 10.0
+    do
+        for seed in 123 456 789
+        do
+            sbatch run_sim.srun "-s ${seed} -d BURNIN=40000 -d NUMGENS=10000 -d N=10000 -d XOPT=${xopt} ${type}.slim"
+        done
+    done
+done

### 2023-06-21 writeup.tex
minor edits

@@
 \renewcommand{\P}{\mathbb{P}}
+\newcommand{\R}{\mathbb{R}}
 \newcommand{\mL}{\mathcal{L}}
@@
 
+\comment{Proposal: change $1 + \mu$ to $f$ (or, something else, but a single letter).}
+
 \comment{Copied from \texttt{eugene\_notes.tex}:}
@@
 \end{align*}
+So, $X$ is a point measure of total mass 1 on $\R$.
 This evolves as a Moran process under the infinitesimal model
@@
 where $y$ and $z$ are independent draws from $X$
-and $\xi$ is an independent $N(0, \eta^2)$.
-Let $\nu(X)$ denote the distribution of $\frac{y + z}{2} + \xi$.
+and $\xi$ is an independent draw from the segregation noise distribution.
+Let $p_\xi$ denote the distribution of $\xi$,
+and $\nu(X)$ denote the distribution of $\frac{y + z}{2} + \xi$.
 This has generator
@@
     \int \int \int \int
-    \left( \Phi(X + \frac{1}{N} \delta_{(y+z)/2 + \xi} - \frac{1}{N} \delta_x) - \Phi(X) \right)
-    (1 + \mu(x)) X(dx) X(dy) X(dz) e^{-(\xi/\eta)^2/2} d\xi \\
+    \left( \Phi\left( X + \frac{1}{N} \delta_{(y+z)/2 + \xi} - \frac{1}{N} \delta_x \right) - \Phi(X) \right)
+    (1 + \mu(x)) X(dx) X(dy) X(dz) p_\xi(d\xi) \\
     &=
     \int \int
-    \left( \Phi(X + \frac{1}{N} \delta_u - \frac{1}{N} \delta_x) - \Phi(X) \right)
+    \left( \Phi\left(X + \frac{1}{N} \delta_u - \frac{1}{N} \delta_x\right) - \Phi(X) \right)
     (1 + \mu(x)) X(dx) \nu(X)(du) .
@@
 
-For a test function $\phi$, setting $\Phi(X) = \ip{\phi}{X}$
+For a test function $\phi$, setting $\Phi(X) = \ip{\phi}{X} := \int \phi(x) X(dx)$
 we get the first moment
@@
          - \phi(x) \right)
-    (1 + \mu(x)) X(dx) X(dy) X(dz) e^{-(\xi/\eta)^2/2} d\xi  \\
+    (1 + \mu(x)) X(dx) X(dy) X(dz) p_\xi(d\xi) \\
     &=
@@
 \end{align*}
-Note that $\ip{\phi}{\partial_{\nu(X)} X} = \ip{\phi}{\nu(X)}$,
-so that this is a first derivative in the direction of
-$\ip{1+\mu}{X} \nu(X) - (1+\mu) X$.
+Note that for $Y$ another finite measure on $\R$,
+the Fr\'echet derivative \comment{(is that the right name?)} of $X$ in the direction of $Y$
+satisfies $\ip{\phi}{\partial_Y X} = \lim_{\epsilon \to 0} \epsilon^{-1} (\ip{\phi}{X + \epsilon Y} - \ip{\phi}{X}) = \ip{\phi}{Y}$,
+so that this is a first derivative in the direction of the measure
+$$ \ip{1+\mu}{X} \nu(X) - (1+\mu) X . $$
 
@@
 then $\nu(Z)\ip{1+\mu}{Z} = (1+\mu)Z$.
-Any measure $Z$ that is a fixed point has the property that
+In other words, any measure $Z$ that is a fixed point has the property that
 (the average of two independent draws plus segregation noise)
@@
 Concretely, with $A$, $B$, and $C$ independent draws from $Z$,
-$$\begin{aligned}
+\begin{align} \label{eqn:ABC_condition}
     \E\left[ f\left( \frac{A+B}{2} + \xi \right) \right]
@@
     \frac{\E\left[ f(C) (1+\mu(C)) \right]}{\E[1+\mu(C)]} .
-\end{aligned}$$
+\end{align}
 
@@
 
-We now analyse the component arising from the noise $\xi$.
-First, we consider the case when $\Psi(\theta)$ is twice continuously differentiable, so $\xi$ has finite mean $m = \Psi'(0)$ and variance $\eta^{2 = \Psi''(0)}$.
-Taylor's theorem then tells us that for $\theta$ sufficiently small,
+We now analyse the component arising from the noise $\xi$.  First, we consider the case when $\Psi(\theta)$ is twice continuously differentiable, so $\xi$ has finite mean $m = \Psi'(0)$
+and variance $\eta^{2} = \Psi''(0)$.   Taylor's theorem then tells us that for $\theta$ sufficiently small, 
 \[
@@
 
-Now, consider the case when $1+ \mu(x) = e^{a x^{2}}$.  In this case there is a Gaussian solution: suppose $Z$ is Gaussian with mean $m$ and variance $\sigma^{2}$.  Then,
+Now, consider the case when $1+ \mu(x) = e^{a x^{2}}$. 
+In this case there is a Gaussian solution: suppose $Z$ is Gaussian with mean $m$ and variance $\sigma^{2}$.  Then,
 \[
@@
 if $\xi$ is Gaussian with mean zero and variance $\eta^2$,
-then equation~\eqref{CHARACTERISTIC} is
-\comment{something not right here? need a factor of 2 on the left?}
+then equation~\eqref{eqn:ABC_condition} with $f(x)=e^{i\theta x}$ is
 \[
@@
 then the limiting phenotype distribution
-will be either Gaussian (Subsection~\ref{sec:stabilizing_gaussian}
-or Cauchy (Subsection~\ref{sec:stabilizing_cauchy}.
+will be either Gaussian (Subsection~\ref{sec:stabilizing_gaussian})
+or Cauchy (Subsection~\ref{sec:stabilizing_cauchy}).
 Furthermore, we expect based on
@@
 to study how this works in practice, we turn to simulation.
-We used SLiM v3 \comment{(cite)}
+We used SLiM v4 \citep{haller2022slim4}
 to simulate a discrete-time approximation to the continuous-time Moran model described above
 as follows.
-Individuals are diploid and sexual;
-the genome is of lenth $10^8$bp with a uniform recombination rate of $10^{-8}$ crossovers per bp;
+Individuals are diploid, hermaphroditic, and sexual;
+the genome is of length $10^8$bp with a uniform recombination rate of $10^{-8}$ crossovers per bp;
 mutations occur at a rate of $m$ per bp.

### 2023-06-21 sims/do_figs.sh
.

@@ -2,11 +2,11 @@
 
 for type in normal cauchy
 do
-    for xopt in 0.0 1.0 2.0 10.0
+    for xopt in 0 1 2 10
     do
         for seed in 123 456 789
         do
-            sbatch run_sim.srun "-s ${seed} -d BURNIN=40000 -d NUMGENS=10000 -d N=10000 -d XOPT=${xopt} ${type}.slim"
+            sbatch run_sim.srun "-s ${seed}${xopt} -d BURNIN=40000 -d NUMGENS=10000 -d N=10000 -d XOPT=${xopt}.0 ${type}.slim"
         done
     done
 done

### 2023-06-30 sims/make_neutral_figs.R
.

@@ -14,16 +14,16 @@ and the 'fix' file should have 'time' and 'num_fixations' columns.
 library(jsonlite)
 library(pracma)
 
-if (!exists("basename")) {
+if (!exists("base")) {
     args <- commandArgs(TRUE)
     if (length(args) != 1) {
         stop(usage)
     }
 
-    basename <- args[1]
+    base <- args[1]
 }
 
-infile <- file(paste0(basename, ".repro.tsv"), "r")
+infile <- file(paste0(base, ".repro.tsv"), "r")
 params <- fromJSON(readLines(infile, 1))
 repro <- read.table(infile, sep="\t", header=TRUE)
 close(infile)
@@ -39,11 +39,11 @@ repro_pa <- repro[,grepl("pa_", names(repro))] |> revCumRowSums()
 midp <- (repro_ma + repro_pa)/2
 seg <- repro_self - midp
 
-pop <- read.table(paste0(basename, ".pop.tsv"), sep="\t", header=TRUE)
+pop <- read.table(paste0(base, ".pop.tsv"), sep="\t", header=TRUE)
 pop$age <- pop$age * params$DT
-fix <- read.table(paste0(basename, ".fix.tsv"), sep="\t", header=TRUE)
+fix <- read.table(paste0(base, ".fix.tsv"), sep="\t", header=TRUE)
 
-png(sprintf("%s_stratified.png", basename), width=6.5, height=length(params$EPSILON)*1.5, pointsize=10, units='in', res=144)
+png(sprintf("%s_stratified.png", base), width=6.5, height=length(params$EPSILON)*1.5, pointsize=10, units='in', res=144)
 {
     layout(matrix(seq_along(params$EPSILON), ncol=2, byrow=TRUE))
     par(mar=c(4,4,1,1), mgp=c(2.5, 1, 0))
@@ -58,7 +58,7 @@ png(sprintf("%s_stratified.png", basename), width=6.5, height=length(params$EPSI
 }
 dev.off()
 
-plot_independence <- function (midp, seg) {
+plot_conditional_hists <- function (midp, seg) {
     layout(1:3)
     par(mar=c(4,4,1,1), mgp=c(2.5, 1, 0))
     # lims <- quantile(abs(seg), 0.98)
@@ -77,10 +77,37 @@ plot_independence <- function (midp, seg) {
 }
 
[…]

### 2023-07-03 writeup.tex
independence bits

@@
 \newcommand{\defn}{:=}
+\newcommand{\var}{\mathop{\mbox{Var}}}
+\newcommand{\cov}{\mathop{\mbox{Cov}}}
 
@@
 
+% Intro:
+%  - the infinitesimal model is useful because it is a "trait-only" model: it allows for predictions and modeling without knowing the underlying genetics.
+%  - Its key properties are: (a) independence of seg noise, and (b) Gaussian-ness.
+%  - It is basically a CLT, and in particular, (b) derives from a the CLT deriving from similar effect sizes.
+%  - Motivating Q: there are other CLTs; can we generalize (b) to non-Gaussian effects, while keeping independence of seg variation?
+% 
+% Data motivation:
+%  First, to see if this is worth asking we look at some data
+%  ... and find that yeah, maybe effect sizes don't fall into the Gaussian domain of attraction.
+% 
+% First answer to the motivating question:  "No"
+% Non-Gaussian effects (sometimes) imply nonindependence of segregation noise:
+%   In at least some cases, natural non-gaussian models produce nonindependence:
+%   1. all loci are uniformly small except there's one big one
+%   2. loci are Cauchy
+%   3. (is there a simple statement we can make more generally?)
+% 
+% Second answer, weaker but also "no":
+%  - Observation: Let A and B be the contributions to parent phenotype, with A inherited by the offspring and B not.
+If A and B are independent, and seg noise is independent of parent phenotype, then A and B are Gaussian.
+%  - When would A and B be asymptotically independent?
+One case is the rare-mutation limit that obtains a "thinned Poisson process" limit for A and B (I think: if all allele frequencies are at frequency O(1/M)?)
+%  - (I have other thoughts to be sorted out here)
+%  - Open Q (maybe?): is there any non-Gaussian model where segregation noise is independent of parental traits?
+% 
+% Simulations: let's see how this stuff works in practice
+%  - neutral simulations
+%  - also, stabilizing selection with theoretical motivation for form of stabilizing selection
+%  - plot: (A) trait mean over time and trait distribution at the end
+%  - plot: (B) (distrn of seg noise), (seg noise vs midparent), and (Radon-Nicodym deriv of conditional seg noise for a few values of midparent)
+% 
+% If we remove 'big loci' is what's left the infinitesimal model?
+%  - people do this in practice (animal breeding papers)
+%  - theory:  form of limiting conditional distribution
+%  - theory: something about independence  (?? if we get there ??)
+%  - plots (A) & (B) above from simulation, but for only small effects
+
+%%%% LAST VERSION
 % 1. *Intro*: Motivation about how many results about quant gen and methods assume Gaussian distributions.
@@
 \title{\large{\bf
-    Quantitative genetics with stable laws
+    Abnormal quantitative genetics?
 }}
@@
 which assumes a simple, explicit form for how an offspring's trait is inherited from their parents':
-it is Normally distributed, with mean equal to the average of the parents' values.
-Despite the inherent plausibility of this model,
+it is Normally distributed, with mean equal to the average of the parents' values
+and a variance independent of their values.
+Despite the plausibility of this model,
 (thanks to the Central Limit Theorem),
@@
 
-The infinitesimal model gains a great deal of generality by appealing to the Central Limit Theorem:
-as shown by \citet{barton2017infinitesimal},
-the distribution of segregation noise will be Normal, regardless of most of the details,
-thanks to the universality of adding up a great many very small and independent things.
+Concretely, the infinitesimal model postulates that
+if the parents' genetic contributions to the trait (i.e., their \emph{breeding values})
+are $z^{[1]}$ and $z^{[2]}$,
+and the offspring's genetic contribution to the trait is $z$,
+and $R = z - (z^{[1]} + z^{[2]})/2$ is the deviation from the midparent
+(which we call the \emph{segregation noise}), then
+\begin{enumerate}
+    \item $R$ is independent of $z^{[1]}$ and $z^{[2]}$, and
+    \item $R$ has a Gaussian distribution.
+\end{enumerate}
+(This is not a full definition,
+since the model makes broader statements involving conditioning on the full pedigree;
+see \citet{barton2017infinitesimal}.)
+We set out here to see whether (2) can be generalized,
+and find that (1) is apparently inexitrably connected.
+
+The Gaussian distribution appears thanks to ``the'' Central Limit Theorem,
+i.e., the universality
+of adding up a great many very small and independent things.
 However, as is well known to probabilists,
@@
 
-This suggests exploring whether stable distributions might reasonably stand in for the Normal
+This suggests exploring whether stable distributions might reasonably stand in for the Gaussian
 in the infinitesimal model.
-In this paper, we make rather preliminary explorations in that direction:
+In this paper, we make preliminary explorations in that direction:
 Does the universality shown by \citet{barton2017infinitesimal}
@@
 \end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%
+\section{Non-Gaussian effects}
+
+We generally follow the notation of \citet{barton2017infinitesimal},
+although in a simpler situation,
+as we consider only the haploid offspring of a single set of haploid parents
+rather than full pedigrees,
+and ignore non-genetic contributions to the trait;
+so, below we say ``trait'' to mean ``genetic contribution to the trait''.
+
+Suppose that the focal individual's trait is
+$$
+    Z_M = \sum_{\ell=1}^M \eta_\ell / M^{1/\alpha},
+$$
+where $\eta_\ell$ is the contribution of the allele carried at the $\ell^\text{th}$ locus.
+The effects of the alleles carried by the two parents
+are $\eta_\ell^{[1]}$ and $\eta_\ell^{[2]}$, respectively,
+and their traits are
+$Z_M^{[1]} = \sum_\ell \eta_\ell^{[1]} / M^{1/\alpha}$ and
+$Z_M^{[2]} = \sum_\ell \eta_\ell^{[2]} / M^{1/\alpha}$.
+We assume the loci are unlinked and inheritance is fair, so
+letting $X_\ell$ be a family of Bernoulli(1/2) random variables that are independent of everything else,
+$$
+    \eta_\ell = X_\ell \eta_\ell^{[1]} + (1 - X_\ell) \eta_\ell^{[1]} .
+$$
+
+We know from \citet{barton2017infinitesimal} that if the $\eta^{[i]}_\ell$ are
+independent with mean zero and $\sum_{\ell=1}^M \var[\eta_\ell]/M \to \sigma^2$,
+then with $\alpha=2$,
+the traits $(Z_M, Z_M^{[1]}, Z_M^{[2]})$ converge in distribution to a multivariate Gaussian distribution
+$(Z, Z^{[1]}, Z^{[2]})$ for which $Z - (Z^{[1]} + Z^{[2]})/2$ is independent of $Z^{[1]}$ and $Z^{[2]}$.
+
+Now let us choose the distribution of allelic effects so that
+the trait distributions converge to a nontrivial limit after rescaling by a different exponent $\alpha$.
+Can the property of independence be preserved?
+The following calculations suggest the answer is ``no``.
+
+\begin{lem} \label{lem:nongaussian_nonindep}
+    \comment{SOMETHING LIKE THE FOLLOWING?}
+    Suppose that the allelic effects are such that
+    $$
+        \lim_{M \to \infty} \log \E[ \prod_{\ell=1}^M e^{it\eta^{[i]}} ]
+        = c |t|^\alpha .
+    $$
+    Then $(Z_M, Z_M^{[1]}, Z_M^{[2]})$ converges in distribution
+    to $(Z, Z^{[1]}, Z^{[2]})$ for which each of the three traits
+    marginally has a standard $\alpha$-stable distribution,
+    but the conditional distribution of $R = Z - (Z^{[1]} - Z^{[2]}))/2$
+    given $Z^{[1]}=z_1$ and/or $Z^{[2]}=z_2$ depends nontrivially on $z_1$ and $z_2$.
+\end{lem}
+
+\begin{proof}[of Lemma~\ref{lem:nongaussian_nonindep}]
+    \comment{calculations with characteristic functions?}
+\end{proof}
+
+This result should perhaps not be surprising:
+there is a deep connection between independence and the Gaussian distribution.
+For instance, a proof of the basic properties of the (Gaussian) infinitesimal model
+might be sketched as follows:
+divide the parental loci into groups according to the results of inheritance,
+defining
+\begin{align*}
+    U_M^{[i]} = \sum_{\ell=1}^M X_\ell \eta_\ell^{[i]} 
+    \qquad \text{ and } \qquad
+    V_M^{[i]} = \sum_{\ell=1}^M (1-X_\ell) \eta_\ell^{[i]} ,
+\end{align*}
+so that 
+$Z_M^{[1]} = U_M^{[1]} + V_M^{[1]}$ and
+$Z_M^{[2]} = U_M^{[2]} + V_M^{[2]}$, so
+\begin{align*}
+    Z_M = U_M^{[1]} + V_M^{[2]} \qquad \text{ and } \qquad
+    R_M = \frac{U_M^{[1]} - V_M^{[1]} - U_M^{[2]} + V_M^{[2]}}{ 2 } .
+\end{align*}
+Suppose we then can prove (as is indeed true in the Gaussian case) that
+(a) $(U_M^{[1]}, V_M^{[1]}, U_M^{[2]}, V_M^{[2]})$ converge in distribution;
+(b) the limiting random variables are independent; and
+defining $R$, $Z^{[1]}$ and $Z^{[2]}$ as is natural in the limit,
+(c) $R$ is independent of $Z^{[1]}$ and $Z^{[2]}$.
+The Kac-Bernstein theorem states that if two independent random variables
+have the property that if their sum and difference are independent of each other,
+then the random variables have a Gaussian distribution
+(\citet{kac1939characterization,bernstein1941property},
+but see \citet{kagan1973characterization} for many much more general results).
+It is not hard to see that (c) implies that $U^{[1]}$ and $V^{[1]}$
+are such a pair of random variables, and therefore Gaussian.
+
+Is there a converse? In other words,
+does independence of segregation noise imply a Gaussian distribution of trait values?
+It is not clear to us in how much generality such a statement might be made,
+but we can at least prove the following:
+
+\begin{lemma}\label{lem:parental_contribs}
+    Under the conditions of Lemma~\ref{lem:nongaussian_nonindep},
+    $U^{[1]}$ and $V^{[1]}$ are not independent.
+    \comment{OR SOMETHING?}
+\end{lemma}
+
+\begin{proof}[of Lemma~\ref{lem:parental_contribs}]
+    \comment{more calcs with characteristic functions?}
+\end{proof}
 

### 2023-07-03 sims/make_neutral_figs.R

@@ -103,6 +103,7 @@ plot_conditional_ratios <- function (midp, seg, do_legend=TRUE, title=NA, ...) {
 }
 
 pdf(sprintf("%s_cond.pdf", base), width=9, height=6, pointsize=10)
+layout(matrix(seq_along(params$EPSILON), nrow=2))
 par(mar=c(4,4,1,1), mgp=c(2.5, 1, 0))
 for (j in 1:ncol(seg)) {
     plot_conditional_ratios(midp[,j], seg[,j], title=sprintf("effects <= %.2f", c(params$EPSILON, Inf)[j+1]))
@@ -169,13 +170,13 @@ dev.off()
 
 # compare simulated distributions to observed
 gvals <- seq(0.5, 4, length.out=11)
-simvals <- do.call(cbind, lapply(gvals, function (g) sort(rtcauchy(1e5, scale=g, gamma=pi/(2*g)))))
+simvals <- do.call(cbind, lapply(gvals, function (g) sort(rtcauchy(min(1e5,nrow(seg)), scale=g, gamma=pi/(2*g)))))
 
 png(file=sprintf("%s_sim_observed.png", base), width=9, height=6, pointsize=10, units='in', res=144)
 par(mar=c(4,4,1,1), mgp=c(2.5, 1, 0))
 layout(matrix(seq_along(params$EPSILON), nrow=2))
 for (j in seq_along(params$EPSILON)) {
-    matplot(sort(sample(seg[,j], min(nrow(simvals), nrow(seg))))/sqrt(seg_var[j]), simvals, pch=20, xlab='observed', ylab='simulated')
+    matplot(sort(sample(seg[,j], nrow(simvals)))/sqrt(seg_var[j]), simvals, pch=20, xlab='observed', ylab='simulated')
     abline(0, 1)
 }
 dev.off()

### 2023-07-03 writeup.tex
.

@@
 %%%%%%%%%%%%%%%%%%%%%%
-\section{Non-Gaussian effects}
+\section{Independence and the Gaussian assumption}
 
@@
 
-\begin{proof}[of Lemma~\ref{lem:nongaussian_nonindep}]
+\begin{proof}[Proof of Lemma~\ref{lem:nongaussian_nonindep}]
     \comment{calculations with characteristic functions?}
@@
 
-This result should perhaps not be surprising:
-there is a deep connection between independence and the Gaussian distribution.
-For instance, a proof of the basic properties of the (Gaussian) infinitesimal model
-might be sketched as follows:
-divide the parental loci into groups according to the results of inheritance,
+Is there a converse? In other words,
+does independence of segregation noise imply a Gaussian distribution of trait values?
+It is not clear to us in how much generality such a statement might be made,
+but we can at least prove the following.
+Divide the parental loci into groups according to the results of inheritance,
 defining
@@
 \end{align*}
-Suppose we then can prove (as is indeed true in the Gaussian case) that
-(a) $(U_M^{[1]}, V_M^{[1]}, U_M^{[2]}, V_M^{[2]})$ converge in distribution;
-(b) the limiting random variables are independent; and
-defining $R$, $Z^{[1]}$ and $Z^{[2]}$ as is natural in the limit,
-(c) $R$ is independent of $Z^{[1]}$ and $Z^{[2]}$.
+It turns out that independence of the portions inherited and not inherited by the offspring
+implies a Gaussian distribution:
+
+\begin{lem}\label{lem:parental_contribs}
+    Suppose that
+    $(U_M^{[1]}, V_M^{[1]}, U_M^{[2]}, V_M^{[2]})$ converge in distribution as $M \to \infty$
+    to independent random variables $(U^{[1]}, V^{[1]}, U^{[2]}, V^{[2]})$
+    (and so $Z_M$, $R_M$, $Z_M^{[1]}$ and $Z_M^{[2]}$ converge to $Z$, $R$, $Z^{[1]}$ and $Z^{[2]}$
+    defined in the natural way).
+    Then if $R$ is independent of $Z^{[1]}$ and $Z^{[2]}$,
+    $(U^{[1]}, V^{[1]}, U^{[2]}, V^{[2]})$ are jointly Gaussian.
+\end{lem}
+
+\begin{proof}[Proof of Lemma~\ref{lem:parental_contribs}]
     The Kac-Bernstein theorem states that if two independent random variables
@@
     but see \citet{kagan1973characterization} for many much more general results).
-It is not hard to see that (c) implies that $U^{[1]}$ and $V^{[1]}$
+    The assumptions imply that $U^{[1]}$ and $V^{[1]}$
     are such a pair of random variables, and therefore Gaussian.
+\end{proof}
 
-Is there a converse? In other words,
-does independence of segregation noise imply a Gaussian distribution of trait values?
-It is not clear to us in how much generality such a statement might be made,
-but we can at least prove the following:
+It is clear in the Gaussian case
+that $U^{[1]}$ and $V^{[1]}$ are asymptotically independent.
+More generally, independence depends on how the model is set up.
+If the allelic effects $\eta_\ell^{[i]}$ are i.i.d.,
+then even for finite $M$, the parental contributions
+$(U_M^{[1]}, V_M^{[1]}, U_M^{[2]}, V_M^{[2]})$ are independent.
+If the distribution of $\eta_\ell^{[i]}$ is chosen to obtain a non-Gaussian limit,
+then Lemma~\ref{lem:parental_contribs} implies that $R$ cannot be independent of the parental traits.
 
-\begin{lemma}\label{lem:parental_contribs}
-    Under the conditions of Lemma~\ref{lem:nongaussian_nonindep},
-    $U^{[1]}$ and $V^{[1]}$ are not independent.
-    \comment{OR SOMETHING?}
-\end{lemma}
-
-\begin{proof}[of Lemma~\ref{lem:parental_contribs}]
-    \comment{more calcs with characteristic functions?}
-\end{proof}
+However, the presence of a large effect polymorphic allele
+makes $(U^{[1]}, V^{[1]}, U^{[2]}, V^{[2]})$ dependent:
+consider the extreme case where $\eta^{[i]}_\ell$ are i.i.d.{} with variance 1 for $\ell \ge 2$,
+while $\eta^{[i]}_1 = \pm 10$ with probability 1/2 each,
+and $Z^{[1]} = \eta_1^{[1]} + \sum_{\ell=2}^M \eta_\ell^{[1]} / \sqrt{M}$.
+Then in the limit $U^{[i]}$ and $V^{[i]}$ are Gaussian with mean either 0 or $\pm 10$ and variance 1/2,
+and are highly dependent: $|U^{[i]}| \approx 10$ or $|V^{[i]}| \approx 10$ but not both.
+(Segregation noise, meanwhile, depends strongly on parental trait values:
+if $Z^{[1]} \approx Z^{[2]}$ then most likely
+$\eta_1^{[1]} = \eta_1^{[2]}$ and so $R$ is Normal with variance 1/2;
+but in the other case then $R$ has equal probabilities of being close to $+5$ or $-5$.)
 

### 2023-07-03 writeup.tex

@@
 \end{multline*}
-This converges as $n \to \infty$ if and only if $m = 0$ and $\alpha > 1$, in which case the limit is 
+This converges as $n \to \infty$ if and only if $m = 0$ and $\alpha \ge 1$, in which case the limit is 
 \[
@@
 the genome is of length $10^8$bp with a uniform recombination rate of $10^{-8}$ crossovers per bp;
-mutations occur at a rate of $m$ per bp.
+mutations occur at a rate of $10^{-9}$ per bp.
 Each new mutation is independently assigned an ``effect''
@@
 
+\subsection{Notes}
+
+Let $x_\ell$ and $p_\ell$ be sequences of numbers.
+When is, for given $s_1$, $s_2$, $t_1$, and $t_2$,
+\begin{align*}
+    &
+    \prod_{\ell=1}^M
+    \left\{
+    \frac{1}{2}
+        \left( 1 + p_\ell(e^{i s_1 x_\ell/M^{1/\alpha}} - 1) \right)
+        \left( 1 + p_\ell(e^{i s_2 x_\ell/M^{1/\alpha}} - 1) \right)
+        +
+    \frac{1}{2}
+        \left( 1 + p_\ell(e^{i t_1 x_\ell/M^{1/\alpha}} - 1) \right)
+        \left( 1 + p_\ell(e^{i t_2 x_\ell/M^{1/\alpha}} - 1) \right)
+    \right\}
+\\ & \qquad \approx
+    \prod_{\ell=1}^M
+        \left( 1 + p_\ell (e^{i s_1 x_\ell/M^{1/\alpha}} - 1)/2 \right)
+        \left( 1 + p_\ell (e^{i s_2 x_\ell/M^{1/\alpha}} - 1)/2 \right)
+        \left( 1 + p_\ell (e^{i t_1 x_\ell/M^{1/\alpha}} - 1)/2 \right)
+        \left( 1 + p_\ell (e^{i t_2 x_\ell/M^{1/\alpha}} - 1)/2 \right)
+\end{align*}
+
 \end{document}

### 2023-07-03 writeup.tex
step things

@@
 
+\begin{figure}
+    \begin{center}
+        \includegraphics{sims/neutral_trait_traces}
+    \end{center}
+    \caption{
+        Median trait values over time \textbf{(A and C)}
+        and distributions of final trait values \textbf{(B and D)}
+        in neutral simulations
+        with Normal \textbf{(A and B)} and Cauchy \textbf{(C and D)}
+        effect size distributions.
+        Simulations were begun with no genetic variation,
+        had populations of size $N=1000$,
+        a genome of length $L=10^8$ and mutation rate of $\mu=10^{-9}$;
+        the effect size distributions were
+        Normal with mean 0 and standard deviation $1/\sqrt{4NL\mu}$
+        and Cauchy with center 0 and scale $1/(4NL\mu)$, respectively.
+        \label{fig:trait_distrns}
+    }
+\end{figure}
+
+\begin{figure}
+    \begin{center}
+        \includegraphics{sims/selected_trait_traces}
+    \end{center}
+    \caption{
+        Median trait values over time \textbf{(A and C)}
+        and distributions of final trait values \textbf{(B and D)}
+        in neutral simulations
+        with Normal \textbf{(A and B)} and Cauchy \textbf{(C and D)}
+        effect size distributions.
+        Simulations were begun with no genetic variation,
+        had populations of size $N=1000$,
+        a genome of length $L=10^8$ and mutation rate of $\mu=10^{-9}$;
+        the effect size distributions were
+        Normal with mean 0 and standard deviation $1/\sqrt{4NL\mu}$
+        and Cauchy with center 0 and scale $1/(4NL\mu)$, respectively.
+        \label{fig:trait_distrns}
+    }
+\end{figure}
+
 \bibliographystyle{plainnat}

### 2023-07-03 sims/helpers.R

@@ -1,0 +1,6 @@
+
+revCumRowSums <- function (x) {
+    for (j in ncol(x):2) x[,j] <- rowSums(x[,1:j])
+    return(x)
+}
+

### 2023-07-03 sims/make_neutral_figs.R

@@ -13,6 +13,7 @@ and the 'fix' file should have 'time' and 'num_fixations' columns.
 
 library(jsonlite)
 library(pracma)
+source("helpers.R")
 
 if (!exists("base")) {
     args <- commandArgs(TRUE)
@@ -27,11 +28,6 @@ infile <- file(paste0(base, ".repro.tsv"), "r")
 params <- fromJSON(readLines(infile, 1))
 repro <- read.table(infile, sep="\t", header=TRUE)
 close(infile)
-
-revCumRowSums <- function (x) {
-    for (j in ncol(x):2) x[,j] <- rowSums(x[,1:j])
-    return(x)
-}
 
 repro_self <- repro[,grepl("self_", names(repro))] |> revCumRowSums()
 repro_ma <- repro[,grepl("ma_", names(repro))] |> revCumRowSums()

### 2023-07-03 sims/make_paper_figs.R

@@ -1,0 +1,78 @@
+#!/usr/bin/Rscript
+library(jsonlite)
+source("helpers.R")
+
+load_files <- function (base) {
+    infile <- file(paste0(base, ".repro.tsv"), "r")
+    params <- fromJSON(readLines(infile, 1))
+    repro <- read.table(infile, sep="\t", header=TRUE)
+    close(infile)
+    if (! "DT" %in% names(params)) params$DT <- 1.0
+    pop <- read.table(paste0(base, ".pop.tsv"), sep="\t", header=TRUE)
+    pop$age <- pop$age * params$DT
+    fix <- read.table(paste0(base, ".fix.tsv"), sep="\t", header=TRUE)
+    repro_self <- repro[,grepl("self_", names(repro))] |> revCumRowSums()
+    repro_ma <- repro[,grepl("ma_", names(repro))] |> revCumRowSums()
+    repro_pa <- repro[,grepl("pa_", names(repro))] |> revCumRowSums()
+    midp <- (repro_ma + repro_pa)/2
+    seg <- repro_self - midp
+    return(list(
+                basename=base,
+                params=params,
+                repro_self=repro_self,
+                midp=midp,
+                seg=seg,
+                pop=pop,
+                fix=fix
+    ))
+}
+
+# 1. Plots of median trait value over time and final trait distribution
+#  for neutral Normal and Cauchy examples
+fname <- "neutral_trait_traces.pdf"
+examples <- lapply(list(
+                 "neutral Normal" = "sim_neutral_normal_831",
+                 "neutral Cauchy" = "sim_neutral_cauchy_831"
+    ), load_files)
+labels <- list(c("(A)", "(B)"), c("(C)", "(D)"))
+names(labels) <- names(examples)
+
+pdf(file=fname, width=6.5, height=3, pointsize=10)
+layout(matrix(1:4, nrow=2), heights=c(1.5,1))
+for (x in names(examples)) {
+    par(mar=c(4,4,2,0)+.1, mgp=c(2.5,1,0))
+        plot(med_trait ~ time, data=examples[[x]][['fix']], type='l',
+             main=x, xlab='time', ylab='median trait value')
+    mtext(labels[[x]][1], 3, adj=-0.0, line=0.2)
+    par(mar=c(4,4,0,0)+.1, mgp=c(2.5,1,0))
+        hist(examples[[x]][['pop']]$trait, breaks=100,
[…]

### 2023-07-03 sims/make_paper_figs.sh

@@ -1,0 +1,5 @@
+slim -s 831 -d BURNIN=20000 -d NUMGENS=1000 -d N=1000 neutral_cauchy.slim
+slim -s 831 -d BURNIN=20000 -d NUMGENS=1000 -d N=1000 neutral_normal.slim
+
+slim -s 707 -d BURNIN=20000 -d NUMGENS=1000 -d N=1000 -d STEP=20 cauchy.slim
+slim -s 707 -d BURNIN=20000 -d NUMGENS=1000 -d N=1000 -d STEP=20 normal.slim

### 2023-07-03 sims/parse_output.R

@@ -38,6 +38,13 @@ layout(1:2)
     hist(diff(fix$num_fixations), xlab='number of fixations/tick', breaks=40)
 dev.off()
 
+pdf(sprintf("%s.traits.pdf", basename), width=3, height=4, pointsize=10)
+layout((1:2))
+par(mar=c(4,4,0,0)+.1, mgp=c(2.5,1,0))
+    plot(fix$time, fix$med_trait, type='l', xlab='time', ylab='median trait value')
+    hist(pop$trait, xlab='trait values at final time', breaks=100, main="")
+dev.off()
+
 pdf(sprintf("%s.ages.pdf", basename), width=6, height=8, pointsize=10)
 layout(1:3)
     hist(pop$age, main=toJSON(params), xlab="age", breaks=30)

### 2023-07-03 sims/helpers.R
plotting fixups

@@ -4,3 +4,16 @@ revCumRowSums <- function (x) {
     return(x)
 }
 
+
+plot_tail <- function (x, minp=0.05, ...) {
+    x <- abs(x)
+    pp <- exp(seq(log(minp), log(.0001), length.out=201))
+    qq <- quantile(x, 1 - pp)
+    xlm <- lm(log(pp) ~ log(qq))
+    plot(qq, pp, log='xy', type='l', ylab="tail probability", ...)
+    # abline(coef(xlm), col='red')
+    lines(qq, exp(coef(xlm)[1] + coef(xlm)[2] * log(qq)), col='red')
+    legend("topright", lty=1, col='red', bty='n',
+           legend=sprintf("y = %.2f %.2f x", coef(xlm)[1], coef(xlm)[2])
+    )
+}

### 2023-07-03 sims/make_figs.R

@@ -11,6 +11,7 @@ and the 'fix' file should have 'time' and 'num_fixations' columns.
 "
 
 library(jsonlite)
+source("helpers.R")
 
 args <- commandArgs(TRUE)
 if (length(args) != 1) {
@@ -21,10 +22,14 @@ basename <- args[1]
 
 infile <- file(paste0(basename, ".repro.tsv"), "r")
 params <- fromJSON(readLines(infile, 1))
-repro <- do.call(rbind, lapply(strsplit(readLines(infile), "\t"), function (x) {
-           x <- as.numeric(x); n <- (length(x) - 1) / 3; data.frame(t=x[1], x=x[1+(1:n)], ma=x[1+n+(1:n)], pa=x[1+2*n+(1:n)])
-}))
+repro <- read.table(infile, sep="\t", header=TRUE)
 close(infile)
+
+repro_self <- repro[,grepl("self_", names(repro))] |> revCumRowSums()
+repro_ma <- repro[,grepl("ma_", names(repro))] |> revCumRowSums()
+repro_pa <- repro[,grepl("pa_", names(repro))] |> revCumRowSums()
+midp <- (repro_ma + repro_pa)/2
+seg <- repro_self - midp
 
 if (! "DT" %in% names(params)) params$DT <- 1.0
 
@@ -32,8 +37,11 @@ pop <- read.table(paste0(basename, ".pop.tsv"), sep="\t", header=TRUE)
 pop$age <- pop$age * params$DT
 fix <- read.table(paste0(basename, ".fix.tsv"), sep="\t", header=TRUE)
 
-repro$midp <- (repro$ma + repro$pa)/2
-repro$seg <- repro$x - repro$midp
+# repro <- do.call(rbind, lapply(strsplit(readLines(infile), "\t"), function (x) {
+#            x <- as.numeric(x); n <- (length(x) - 1) / 3; data.frame(t=x[1], x=x[1+(1:n)], ma=x[1+n+(1:n)], pa=x[1+2*n+(1:n)])
+# }))
+# repro$midp <- (repro$ma + repro$pa)/2
+# repro$seg <- repro$x - repro$midp
 
 # pdf(sprintf("%s_independence.pdf", basename), width=6.5, height=4, pointsize=10)
 png(sprintf("%s_independence.png", basename), width=6.5, height=4, pointsize=10, units='in', res=144)
@@ -51,11 +59,12 @@ png(sprintf("%s_independence.png", basename), width=6.5, height=4, pointsize=10,
[…]

### 2023-07-03 sims/parse_output.R

@@ -23,10 +23,14 @@ num_plots <- as.numeric(args[2])
 
 infile <- file(paste0(basename, ".repro.tsv"), "r")
 params <- fromJSON(readLines(infile, 1))
-repro <- do.call(rbind, lapply(strsplit(readLines(infile), "\t"), function (x) {
-           x <- as.numeric(x); n <- (length(x) - 1) / 3; data.frame(t=x[1], x=x[1+(1:n)], ma=x[1+n+(1:n)], pa=x[1+2*n+(1:n)])
-}))
+repro <- read.table(infile, sep="\t", header=TRUE)
 close(infile)
+
+repro_self <- repro[,grepl("self_", names(repro))] |> revCumRowSums()
+repro_ma <- repro[,grepl("ma_", names(repro))] |> revCumRowSums()
+repro_pa <- repro[,grepl("pa_", names(repro))] |> revCumRowSums()
+midp <- (repro_ma + repro_pa)/2
+seg <- repro_self - midp
 
 pop <- read.table(paste0(basename, ".pop.tsv"), sep="\t", header=TRUE)
 pop$age <- pop$age * params$DT
@@ -52,8 +56,7 @@ layout(1:3)
     hist(log(abs(pop$trait)), xlab='log|trait|', main='log|trait|', breaks=30)
 dev.off()
 
-repro$seg <- repro$x - (repro$ma + repro$pa)/2
-repro$plot <- cut(repro$t, num_plots)
+repro_plot <- cut(repro$time, num_plots)
 
 if (params$type == "normal") {
     qfun <- qnorm
@@ -66,13 +69,14 @@ if (params$type == "normal") {
 # pdf(sprintf("%s.seg.pdf", basename), width=8, height=2*num_plots, pointsize=10)
 png(sprintf("%s.seg.png", basename), width=8, height=2*num_plots, pointsize=10, units='in', res=144)
 layout(matrix(1:(3*num_plots), ncol=3, byrow=TRUE))
-xlim <- range(repro$seg)
-xh <- hist(repro$seg, breaks=200, plot=FALSE)
-for (k in levels(repro$plot)) {
-    title <- if (k == levels(repro$plot)[1]) toJSON(params) else ""
-    sub <- repro[repro$plot == k,]
-    hist(sub$seg, breaks=xh$breaks,
-         main=sprintf("times %0.2f-%0.2f", min(sub$t), max(sub$t)),
+j <- ncol(seg)
+xlim <- range(seg[,j])
[…]

### 2023-07-03 sims/helpers.R
paper plots

@@ -17,3 +17,43 @@ plot_tail <- function (x, minp=0.05, ...) {
            legend=sprintf("y = %.2f %.2f x", coef(xlm)[1], coef(xlm)[2])
     )
 }
+
+plot_conditional_hists <- function (midp, seg, ...) {
+    layout(1:3)
+    par(mar=c(4,4,1,1), mgp=c(2.5, 1, 0))
+    # lims <- quantile(abs(seg), 0.98)
+    lims <- max(abs(seg))
+    for (q in c(.1, .5, .9)) {
+        ut <- (abs(rank(midp)/length(midp) - q) < 0.025)
+        outl <- (seg > -lims) & (seg < lims)
+        hist(seg[ut & outl], breaks=100, main=sprintf("midparent %0.0f%%", 100*q),
+             # xlab="segregation noise",
+             xlab='',
+             xlim=c(-1,1) * lims,
+            ...
+        )
+        legend("bottomleft", title=sprintf("tail: %.2f%% < %.0f", 100 * mean(seg[ut] < -lims), -lims), legend='', bty='n')
+        legend("bottomright", title=sprintf("tail: %.2f%% > %.0f", 100 * mean(seg[ut] > lims), lims), legend='', bty='n')
+    }
+    mtext("segregation noise", 1, line=2.5, cex=0.75)
+}
+
+plot_conditional_ratios <- function (midp, seg, do_legend=TRUE, title=NA, ylim=c(0,3), ...) {
+    pbreaks <- pmax(0.01, pmin(0.99, seq(0, 1, length.out=51)))
+    qbreaks <- c(-Inf, quantile(seg, pbreaks), Inf)
+    total_xh <- hist(seg, breaks=qbreaks, plot=FALSE)
+    uk <- seq(2, length(qbreaks)-1)
+    par(mar=c(4,4,1,1), mgp=c(2.5, 1, 0))
+    plot(0, 0, type='n', xlim=range(qbreaks, finite=TRUE), ylim=ylim,
+         xlab="", ylab="enrichment", ...)
+    cols = c(2, 1, 3)
+    qvals <- c(.1, .5, .9)
+    for (k in seq_along(qvals)) {
+        q <- qvals[k]
+        ut <- (abs(rank(midp)/length(midp) - q) < 0.025)
+        xh <- hist(seg[ut], breaks=qbreaks, plot=FALSE)
[…]

### 2023-07-03 sims/make_neutral_figs.R

@@ -54,48 +54,10 @@ png(sprintf("%s_stratified.png", base), width=6.5, height=length(params$EPSILON)
 }
 dev.off()
 
-plot_conditional_hists <- function (midp, seg) {
-    layout(1:3)
-    par(mar=c(4,4,1,1), mgp=c(2.5, 1, 0))
-    # lims <- quantile(abs(seg), 0.98)
-    lims <- max(abs(seg))
-    for (q in c(.1, .5, .9)) {
-        ut <- (abs(rank(midp)/length(midp) - q) < 0.025)
-        outl <- (seg > -lims) & (seg < lims)
-        hist(seg[ut & outl], breaks=100, main=sprintf("midparent %0.0f%%", 100*q),
-             # xlab="segregation noise",
-             xlab='',
-             xlim=c(-1,1) * lims)
-        legend("bottomleft", title=sprintf("tail: %.2f%% < %.0f", 100 * mean(seg[ut] < -lims), -lims), legend='', bty='n')
-        legend("bottomright", title=sprintf("tail: %.2f%% > %.0f", 100 * mean(seg[ut] > lims), lims), legend='', bty='n')
-    }
-    mtext("segregation noise", 1, line=2.5, cex=0.75)
-}
-
 for (j in seq_along(params$EPSILON)) {
     png(sprintf("%s_ind_%d.png", base, j), width=3.5, height=5, pointsize=10, units='in', res=144)
     plot_conditional_hists(midp[,j], seg[,j])
     dev.off()
-}
-
-plot_conditional_ratios <- function (midp, seg, do_legend=TRUE, title=NA, ...) {
-    pbreaks <- pmax(0.01, pmin(0.99, seq(0, 1, length.out=51)))
-    qbreaks <- c(-Inf, quantile(seg, pbreaks), Inf)
-    total_xh <- hist(seg, breaks=qbreaks, plot=FALSE)
-    uk <- seq(2, length(qbreaks)-1)
-    par(mar=c(4,4,1,1), mgp=c(2.5, 1, 0))
-    plot(0, 0, type='n', xlim=range(qbreaks, finite=TRUE), ylim=c(0, 2), 
-         xlab="", ylab="enrichment", ...)
-    cols = c(2, 1, 3)
-    qvals <- c(.1, .5, .9)
-    for (k in seq_along(qvals)) {
-        q <- qvals[k]
-        ut <- (abs(rank(midp)/length(midp) - q) < 0.025)
[…]

### 2023-07-03 sims/make_paper_figs.R

@@ -75,4 +75,87 @@ for (x in names(examples)) {
 }
 dev.off()
 
+# 3. Segregation noise:
+# (distrn of seg noise), (seg noise vs midparent), and (Radon-Nicodym deriv of conditional seg noise for a few values of midparent)
+
+plot_seg_noise <- function (fname, examples, labels, seg_col, main_append="") {
+    png(file=fname, width=6.5, height=6, pointsize=10, units='in', res=288)
+    layout(matrix(1:6, nrow=3), heights=c(1.5,1, 1))
+    for (x in names(examples)) {
+        j <- if (!missing(seg_col)) seg_col else ncol(examples[[x]][['seg']])
+        main <- paste0(x, 
+                       if (missing(seg_col)) ""
+                       else
+                       sprintf(", effects <= %.2f", c(examples[[x]][['params']][['EPSILON']],Inf)[j+1])
+        )
+        this_seg <- examples[[x]][['seg']][,j]
+        this_midp <- examples[[x]][['midp']][,j]
+        par(mar=c(4,4,2,0)+.1, mgp=c(2.5,1,0))
+            plot(this_midp, this_seg, pch=20, col=adjustcolor("black", 0.5), asp=1,
+                 xlab="midparent trait", ylab="segregation noise",
+                 main=main)
+        mtext(labels[[x]][1], 3, adj=-0.0, line=0.2)
+        par(mar=c(4,4,0,0)+.1, mgp=c(2.5,1,0))
+            hist(this_seg, breaks=40, main='',
+                 xlab='segreation noise')
+        mtext(labels[[x]][2], 3, adj=-0.0, line=0.2)
+            plot_conditional_ratios(this_midp, this_seg, do_legend=(x == names(examples)[1]))
+        mtext(labels[[x]][3], 3, adj=-0.0, line=0.2)
+    }
+    dev.off()
+}
+
+# Neutral
+
+fname <- "neutral_seg_noise.png"
+examples <- lapply(list(
+                 "neutral Normal" = "sim_neutral_normal_831",
+                 "neutral Cauchy" = "sim_neutral_cauchy_831"
+    ), load_files)
+labels <- list(c("(A)", "(B)", "(C)"), c("(D)", "(E)", "(F)"))
+names(labels) <- names(examples)
+
+plot_seg_noise(fname, examples, labels)
+
+# Selected
+
+fname <- "selected_seg_noise.png"
+examples <- lapply(list(
[…]

### 2023-07-04 writeup.tex
paper plots
.

@@
 
-Now let us choose the distribution of allelic effects so that
-the trait distributions converge to a nontrivial limit after rescaling by a different exponent $\alpha$.
-Can the property of independence be preserved?
-The following calculations suggest the answer is ``no``.
-
-\begin{lem} \label{lem:nongaussian_nonindep}
-    \comment{SOMETHING LIKE THE FOLLOWING?}
-    Suppose that the allelic effects are such that
-    $$
-        \lim_{M \to \infty} \log \E[ \prod_{\ell=1}^M e^{it\eta^{[i]}} ]
-        = c |t|^\alpha .
-    $$
-    Then $(Z_M, Z_M^{[1]}, Z_M^{[2]})$ converges in distribution
-    to $(Z, Z^{[1]}, Z^{[2]})$ for which each of the three traits
-    marginally has a standard $\alpha$-stable distribution,
-    but the conditional distribution of $R = Z - (Z^{[1]} - Z^{[2]}))/2$
-    given $Z^{[1]}=z_1$ and/or $Z^{[2]}=z_2$ depends nontrivially on $z_1$ and $z_2$.
-\end{lem}
-
-\begin{proof}[Proof of Lemma~\ref{lem:nongaussian_nonindep}]
-    \comment{calculations with characteristic functions?}
-\end{proof}
-
 Is there a converse? In other words,
@@
 It turns out that independence of the portions inherited and not inherited by the offspring
+combined with independence of segregation noise and midparent value
 implies a Gaussian distribution:
@@
 \begin{lem}\label{lem:parental_contribs}
-    Suppose that
-    $(U_M^{[1]}, V_M^{[1]}, U_M^{[2]}, V_M^{[2]})$ converge in distribution as $M \to \infty$
-    to independent random variables $(U^{[1]}, V^{[1]}, U^{[2]}, V^{[2]})$
-    (and so $Z_M$, $R_M$, $Z_M^{[1]}$ and $Z_M^{[2]}$ converge to $Z$, $R$, $Z^{[1]}$ and $Z^{[2]}$
-    defined in the natural way).
-    Then if $R$ is independent of $Z^{[1]}$ and $Z^{[2]}$,
-    $(U^{[1]}, V^{[1]}, U^{[2]}, V^{[2]})$ are jointly Gaussian.
+    Suppose that $\{\eta_\ell^{[i]}\}$ are independent for all $\ell$ and $i$,
+    that the distribution of $\eta_\ell^{[i]}$ only depends on $\ell$,
+    and that $\{X_\ell\}$ are independent of $\{\eta_\ell^{[i]}\}$.
+    If the segregation noise is independent of the midparent value
+    (i.e., $R_M$ is independent of $(Z^{[1]} + Z^{[2]})/2$)
+    then $R_M$ is Gaussian.
 \end{lem}
@@
 \begin{proof}[Proof of Lemma~\ref{lem:parental_contribs}]
-    The Kac-Bernstein theorem states that if two independent random variables
+    Note that if inheritance (the $X_\ell$) is independent of the per-locus effects ($\eta_\ell^{[i]}$),
+    the latter are all independent of each other,
+    and all effects at a given locus are drawn from the same distribution,
+    then the inherited allele, 
+    $\eta_\ell^{[1]} X_\ell + \eta_\ell^{[2]} (1-X_\ell)$
+    is independent of (and has the same distribution as)
+    the non-inherited allele,
+    $\eta_\ell^{[1]} (1-X_\ell) + \eta_\ell^{[2]} X_\ell$
+    -- they are simply two independent draws from the distribution of effects at locus $\ell$.
+    Therefore, the sum of inherited alleles, $U_M^{[1]} + V_M^{[2]}$
+    are independent of the non-inherited alleles, $U_M^{[2]} + V_M^{[1]}$.
+
+    Now, the Kac-Bernstein theorem states that if two independent random variables
     have the property that if their sum and difference are independent of each other,
-    then the random variables have a Gaussian distribution
+    then the random variables have a joint Gaussian distribution
     (\citet{kac1939characterization,bernstein1941property},
     but see \citet{kagan1973characterization} for many much more general results).
-    The assumptions imply that $U^{[1]}$ and $V^{[1]}$
-    are such a pair of random variables, and therefore Gaussian.
+    Since $R_M$ is (half) the difference and $(Z^{[1]} + Z^{[2]})/2$ is (half) the sum
+    of $U_M^{[1]} + V_M^{[2]}$ and $U_M^{[2]} + V_M^{[1]}$,
+    both are therefore jointly Gaussian.
 \end{proof}
 
+Since models can be easily set up for which $Z$ is not Gaussian,
+the content of this is that for such models, independence of segregation noise
+is not likely a good assumption.
+Note that the proof does not depend on independence of the $X_\ell$:
+indeed, alleles may be inherited together,
+although in practice this would quickly create dependencies between nearby allelic effects.
+Also note that it is \emph{not} usually true that the components --
+for instance, $U_M^{[1]}$ and $U_M^{[2]}$ --
+are independent of each other.
+
+\comment{remove the following?}
 It is clear in the Gaussian case
@@
 
-\subsection{Notes}
-
-Let $x_\ell$ and $p_\ell$ be sequences of numbers.
-When is, for given $s_1$, $s_2$, $t_1$, and $t_2$,
-\begin{align*}
-    &
-    \prod_{\ell=1}^M
-    \left\{
-    \frac{1}{2}
-        \left( 1 + p_\ell(e^{i s_1 x_\ell/M^{1/\alpha}} - 1) \right)
-        \left( 1 + p_\ell(e^{i s_2 x_\ell/M^{1/\alpha}} - 1) \right)
-        +
-    \frac{1}{2}
-        \left( 1 + p_\ell(e^{i t_1 x_\ell/M^{1/\alpha}} - 1) \right)
-        \left( 1 + p_\ell(e^{i t_2 x_\ell/M^{1/\alpha}} - 1) \right)
-    \right\}
-\\ & \qquad \approx
-    \prod_{\ell=1}^M
-        \left( 1 + p_\ell (e^{i s_1 x_\ell/M^{1/\alpha}} - 1)/2 \right)
-        \left( 1 + p_\ell (e^{i s_2 x_\ell/M^{1/\alpha}} - 1)/2 \right)
-        \left( 1 + p_\ell (e^{i t_1 x_\ell/M^{1/\alpha}} - 1)/2 \right)
-        \left( 1 + p_\ell (e^{i t_2 x_\ell/M^{1/\alpha}} - 1)/2 \right)
-\end{align*}
-
 \end{document}

### 2023-07-04 sims/make_paper_figs.R

@@ -27,57 +27,6 @@ load_files <- function (base) {
     ))
 }
 
-# 1. Plots of median trait value over time and final trait distribution
-#  for neutral Normal and Cauchy examples
-fname <- "neutral_trait_traces.pdf"
-examples <- lapply(list(
-                 "neutral Normal" = "sim_neutral_normal_831",
-                 "neutral Cauchy" = "sim_neutral_cauchy_831"
-    ), load_files)
-labels <- list(c("(A)", "(B)"), c("(C)", "(D)"))
-names(labels) <- names(examples)
-
-pdf(file=fname, width=6.5, height=3, pointsize=10)
-layout(matrix(1:4, nrow=2), heights=c(1.5,1))
-for (x in names(examples)) {
-    par(mar=c(4,4,2,0)+.1, mgp=c(2.5,1,0))
-        plot(med_trait ~ time, data=examples[[x]][['fix']], type='l',
-             main=x, xlab='time', ylab='median trait value')
-    mtext(labels[[x]][1], 3, adj=-0.0, line=0.2)
-    par(mar=c(4,4,0,0)+.1, mgp=c(2.5,1,0))
-        hist(examples[[x]][['pop']]$trait, breaks=100,
-             xlab='trait value', main='')
-    mtext(labels[[x]][2], 3, adj=-0.0, line=0.2)
-}
-dev.off()
-
-# 2. Plots of median trait value over time and final trait distribution
-#  for selected Normal and Cauchy examples
-fname <- "selected_trait_traces.pdf"
-examples <- lapply(list(
-                 "selected Normal" = "sim_normal_707",
-                 "selected Cauchy" = "sim_cauchy_707"
-    ), load_files)
-labels <- list(c("(A)", "(B)"), c("(C)", "(D)"))
-names(labels) <- names(examples)
-
-pdf(file=fname, width=6.5, height=3, pointsize=10)
-layout(matrix(1:4, nrow=2), heights=c(1.5,1))
-for (x in names(examples)) {
-    par(mar=c(4,4,2,0)+.1, mgp=c(2.5,1,0))
-        plot(med_trait ~ time, data=examples[[x]][['fix']], type='l',
-             main=x, xlab='time', ylab='median trait value')
-    mtext(labels[[x]][1], 3, adj=-0.0, line=0.2)
-    par(mar=c(4,4,0,0)+.1, mgp=c(2.5,1,0))
-        hist(examples[[x]][['pop']]$trait, breaks=100,
[…]

### 2023-07-04 sims/do_figs.sh
.

@@ -2,11 +2,12 @@
 
 for type in normal cauchy
 do
-    for xopt in 0 1 2 10
+    for seed in 8135 8791
     do
-        for seed in 1024 4057 7090
+        sbatch run_sim.srun "-s ${seed}66${xopt} -d BURNIN=40000 -d NUMGENS=10000 -d N=10000 neutral_${type}.slim"
+        for xopt in 0 1 2 10
         do
-            sbatch run_sim.srun "-s ${seed}88${xopt} -d BURNIN=40000 -d NUMGENS=1000 -d N=10000 -d STEP=10 -d XOPT=${xopt}.0 ${type}.slim"
+            sbatch run_sim.srun "-s ${seed}55${xopt} -d BURNIN=40000 -d NUMGENS=10000 -d N=10000 -d STEP=10 -d XOPT=${xopt}.0 ${type}.slim"
         done
     done
 done

### 2023-07-04 sims/make_paper_figs.sh

@@ -1,5 +1,9 @@
-slim -s 831 -d BURNIN=20000 -d NUMGENS=1000 -d N=1000 neutral_cauchy.slim
-slim -s 831 -d BURNIN=20000 -d NUMGENS=1000 -d N=1000 neutral_normal.slim
+#!/bin/bash
+SLIM="sbatch run_sim.srun"
+# SLIM=slim
 
-slim -s 707 -d BURNIN=20000 -d NUMGENS=1000 -d N=1000 -d STEP=20 cauchy.slim
-slim -s 707 -d BURNIN=20000 -d NUMGENS=1000 -d N=1000 -d STEP=20 normal.slim
+$SLIM -s 831 -d BURNIN=20000 -d NUMGENS=1000 -d N=1000 -d 'EPSILON=c(0,0.1,10)' neutral_cauchy.slim
+$SLIM -s 831 -d BURNIN=20000 -d NUMGENS=1000 -d N=1000 -d 'EPSILON=c(0,0.1,10)' neutral_normal.slim
+
+$SLIM -s 707 -d BURNIN=20000 -d NUMGENS=1000 -d N=1000 -d STEP=20 -d 'EPSILON=c(0,0.1,10)' cauchy.slim
+$SLIM -s 707 -d BURNIN=20000 -d NUMGENS=1000 -d N=1000 -d STEP=20 -d 'EPSILON=c(0,0.1,10)' normal.slim

### 2023-07-04 sims/helpers.R
.

@@ -38,8 +38,8 @@ plot_conditional_hists <- function (midp, seg, ...) {
     mtext("segregation noise", 1, line=2.5, cex=0.75)
 }
 
-plot_conditional_ratios <- function (midp, seg, do_legend=TRUE, title=NA, ylim=c(0,3), ...) {
-    pbreaks <- pmax(0.01, pmin(0.99, seq(0, 1, length.out=51)))
+plot_conditional_ratios <- function (midp, seg, w=0.025, n=51, do_legend=TRUE, title=NA, ylim=c(0,3), ...) {
+    pbreaks <- pmax(0.01, pmin(0.99, seq(0, 1, length.out=n)))
     qbreaks <- c(-Inf, quantile(seg, pbreaks), Inf)
     total_xh <- hist(seg, breaks=qbreaks, plot=FALSE)
     uk <- seq(2, length(qbreaks)-1)
@@ -50,7 +50,7 @@ plot_conditional_ratios <- function (midp, seg, do_legend=TRUE, title=NA, ylim=c
     qvals <- c(.1, .5, .9)
     for (k in seq_along(qvals)) {
         q <- qvals[k]
-        ut <- (abs(rank(midp)/length(midp) - q) < 0.025)
+        ut <- (abs(rank(midp)/length(midp) - q) < w)
         xh <- hist(seg[ut], breaks=qbreaks, plot=FALSE)
         lines(total_xh$mids[uk], xh$density[uk] / total_xh$density[uk], col=cols[k])
     }

### 2023-07-04 sims/make_paper_figs.R

@@ -27,8 +27,7 @@ load_files <- function (base) {
     ))
 }
 
-plot_seg_noise <- function (fname, examples, labels, seg_col, main_append="") {
-    png(file=fname, width=6.5, height=6, pointsize=10, units='in', res=288)
+plot_seg_noise <- function (examples, labels, seg_col, main_append="", ...) {
     layout(matrix(1:6, nrow=3), heights=c(1.5,1, 1))
     for (x in names(examples)) {
         j <- if (!missing(seg_col)) seg_col else ncol(examples[[x]][['seg']])
@@ -48,10 +47,9 @@ plot_seg_noise <- function (fname, examples, labels, seg_col, main_append="") {
             hist(this_seg, breaks=40, main='',
                  xlab='segreation noise')
         mtext(labels[[x]][2], 3, adj=-0.0, line=0.2)
-            plot_conditional_ratios(this_midp, this_seg, do_legend=(x == names(examples)[1]))
+            plot_conditional_ratios(this_midp, this_seg, do_legend=(x == names(examples)[1]), ...)
         mtext(labels[[x]][3], 3, adj=-0.0, line=0.2)
     }
-    dev.off()
 }
 
 
@@ -88,7 +86,9 @@ fname <- "neutral_seg_noise.png"
 labels <- list(c("(A)", "(B)", "(C)"), c("(D)", "(E)", "(F)"))
 names(labels) <- names(examples)
 
-plot_seg_noise(fname, examples, labels)
+png(file=fname, width=6.5, height=6, pointsize=10, units='in', res=288)
+plot_seg_noise(examples, labels)
+dev.off()
 
 # 4. Segregation noise for small effects
 # (distrn of seg noise), (seg noise vs midparent), and (Radon-Nicodym deriv of conditional seg noise for a few values of midparent)
@@ -98,7 +98,9 @@ fname <- "neutral_seg_noise_small.png"
 labels <- list(c("(A)", "(B)", "(C)"), c("(D)", "(E)", "(F)"))
 names(labels) <- names(examples)
 
-plot_seg_noise(fname, examples, labels, seg_col=1)
+png(file=fname, width=6.5, height=6, pointsize=10, units='in', res=288)
+plot_seg_noise(examples, labels, seg_col=1)
+dev.off()
 
 
 ##########
@@ -136,7 +138,9 @@ fname <- "selected_seg_noise.png"
[…]

### 2023-07-04 writeup.tex
.
low mem plotting

@@
 
-%%%% LAST VERSION
-% 1. *Intro*: Motivation about how many results about quant gen and methods assume Gaussian distributions.
-%     Mention Levy walk macroevolution papers.
-%     Point out that usual quant gen has *no* sweeps, while stable distributions would have sweeps (since some s would be bigger than 1/N).
-%     Motivate with mixture model of effect sizes.
-%     If stable distributions are a better model, what does that tell us? We can better understand the *noise* in evolution, i.e., predictability and saltation-ness.
-%     (PETER)
-% 
-% 2. Motivation: empirical SNPs distributions in real data (PETER)
-% 
-% 3. Introducing infinitesimal model (in this case) (TODD)
-% 
-% 4. Proof following BEV in stable case (TODD)
-% 
-% 5. Theoretical results on stationary distributions (TODD)
-% 
-% 6. Simulations (PETER)
-
 \title{\large{\bf
-    Abnormal quantitative genetics?
+    Normality and large effect loci in quantitative genetics
 }}
@@
 
-\comment{remove the following?}
+\comment{remove or shorten the following?}
 It is clear in the Gaussian case
@@
 %%%%%%%%%%%%%%%%%%%%%%
-\section{Convergence to the non-infinitesimal model}  % TODD
-
-%%%%%%%%%%%%%%%%%%%%%%
-\section{Stationary distributions}  % TODD
-
-%%%%%%%%%%%%%%%%%%%%%%
-\section{Simulations}  % PETER
+\section{Simulations}
 
@@
 
-We compared two situations:
-(1) the effect size distribution was Gaussian with mean zero and standard deviation $\sigma_g$,
-and $f(z) = e^{\alpha z^2}$; and
-(2) the effect size distribution was Cauchy, centered at zero and with width $\delta - \gamma$
-and $f(z) = (\delta/\gamma) (\gamma + z^2) / (\delta + z^2)$.
-\comment{TODO: just refer to equations above?}
-We ran simulations of each with $N=XXX$ for XXX time steps of ``burn-in''
-(judged to be long enough as the number of segregating mutations had plateaued),
-following which we recorded the phenotypes of each new offspring and their two parents
-for XXX time steps (a total of around YYY offspring).
-
-Figure XXX shows the resulting distributions of segregation noise:
-i.e., the distribution of child phenotype minus average of parents' phenotypes,
-across all offspring following burn-in.
-
 \begin{figure}
@@
         in neutral simulations
-        with Normal \textbf{(A and B)} and Cauchy \textbf{(C and D)}
+        with Gaussian \textbf{(A and B)} and Cauchy \textbf{(C and D)}
         effect size distributions.
@@
     \begin{center}
+        \includegraphics{sims/neutral_seg_noise}
+    \end{center}
+    Segregation noise with Gaussian \textbf{(A-C)} and Cauchy \textbf{(D-F)} effect size distributions,
+    from the same simulations shown in Figure~\ref{fig:trait_distrns},
+    for all individuals born during the last 1,000 time units of the simulation.
+    \textbf{(A,D)} Segregation noise (offspring trait minus midparent) plotted against midparent.
+    \textbf{(B,E)} Histograms of segregation noise.
+    \textbf{(B,E)} Enrichment in quantiles of segregation noise conditional on midparent value.
+    To compute the latter,
+    let $q_0, q_2, q_4, \ldots, q_{98}, q_{100}$ divide the distribution of segregation noise
+    into 50 bins with roughly equal numbers in each (i.e., the quantiles),
+    let $p_k(a,b)$ denote the proportion of the offspring of parents with midparent value
+    is in $[a, b)$ whose segregation noise is in $[q_{k-1}, q_k)$,
+    and let $p_k$ be the proportion of all offspring with segreation noise in $[q_{k-1}, q_k)$
+    (so, $p_k \approx 0.02$).
+    Then, each line shows $p_k(a,b)/p_k$
+    plotted against the midpoint of the corresponding quantile bin $(q_{k-1},q_k)$,
+    for $(a,b)$ chosen to span the 5\% of midparents centered on:
+    (red) the 10\% quantile,
+    (black) the median, and
+    (green) the 90\% quantile of midparent value.
+    \caption{
+        \label{fig:seg_noise}
+    }
+\end{figure}
+
+We first consider the neutral case (i.e., with $\mu \equiv 0$).
+As shown in Figure~\ref{fig:trait_distrns},
+a Cauchy effect size distribution, unsurprisingly,
+produces a distribution of trait values with more extreme values
+and a rougher path of population median trait value over time
+than in the Gaussian case.
+Indeed, the sum of all effects along the path from the start of the simulation
+to an arbitrary individual will approximate either Brownian motion (in the Gaussian case)
+or a Cauchy process (in the Cauchy case).
+Figure~\ref{fig:seg_noise} shows how segregation noise
+depends on midparent value.
+In the Gaussian case, segregation noise is Gaussian and independent of midparent value,
+as shown by the flat lines at 1 in Figure~\ref{fig:seg_noise}C. 
+On the other hand, in the Cauchy case we can see from either Figure~\ref{fig:seg_noise}D or F
+that segregation noise depends strongly on midparent trait,
+as we would expect in the presence of several large-effect alleles.
+
+\begin{figure}
+    \begin{center}
         \includegraphics{sims/selected_trait_traces}
@@
     \caption{
-        Median trait values over time \textbf{(A and C)}
-        and distributions of final trait values \textbf{(B and D)}
-        in neutral simulations
-        with Normal \textbf{(A and B)} and Cauchy \textbf{(C and D)}
-        effect size distributions.
-        Simulations were begun with no genetic variation,
-        had populations of size $N=1000$,
-        a genome of length $L=10^8$ and mutation rate of $\mu=10^{-9}$;
-        the effect size distributions were
-        Normal with mean 0 and standard deviation $1/\sqrt{4NL\mu}$
-        and Cauchy with center 0 and scale $1/(4NL\mu)$, respectively.
-        \label{fig:trait_distrns}
+        As in Figure~\ref{fig:trait_distrns},
+        except under a model of stabilizing selection (see text for details).
+        \label{fig:sel_trait_distrns}
     }
 \end{figure}
+
+\begin{figure}
+    \begin{center}
+        \includegraphics{sims/selected_seg_noise}
+    \end{center}
+    \caption{
+        As in Figure~\ref{fig:seg_noise},
+        except under a model of stabilizing selection (see text for details).
+        \label{fig:sel_seg_noise}
+    }
+\end{figure}
+
+We next compared two situations of stabilizing selection, motivated by the results above
+\comment{TODO: refer to where?}:
+(1) the effect size distribution was Gaussian with mean zero and standard deviation $\sigma_g$,
+and $f(z) = e^{\alpha z^2}$; and
+(2) the effect size distribution was Cauchy, centered at zero and with width $\delta - \gamma$
+and $f(z) = (\delta/\gamma) (\gamma + z^2) / (\delta + z^2)$.
+Figure XXX shows the resulting distributions of segregation noise:
+i.e., the distribution of child phenotype minus average of parents' phenotypes,
+across all offspring following burn-in.
 

### 2023-07-04 sims/make_paper_figs.R

@@ -19,7 +19,6 @@ load_files <- function (base) {
     return(list(
                 basename=base,
                 params=params,
-                repro_self=repro_self,
                 midp=midp,
                 seg=seg,
                 pop=pop,
@@ -27,17 +26,71 @@ load_files <- function (base) {
     ))
 }
 
+load_params <- function (base) {
+    infile <- file(paste0(base, ".repro.tsv"), "r")
+    params <- fromJSON(readLines(infile, 1))
+    close(infile)
+    if (! "DT" %in% names(params)) params$DT <- 1.0
+    return(params)
+}
+
+load_fix <- function (base) {
+    fix <- read.table(paste0(base, ".fix.tsv"), sep="\t", header=TRUE)
+    return(fix)
+}
+
+load_pop <- function (base) {
+    params <- load_params(base)
+    pop <- read.table(paste0(base, ".pop.tsv"), sep="\t", header=TRUE)
+    pop$age <- pop$age * params$DT
+    return(pop)
+}
+
+load_seg_midp <- function (base) {
+    infile <- file(paste0(base, ".repro.tsv"), "r")
+    params <- fromJSON(readLines(infile, 1))
+    repro <- read.table(infile, sep="\t", header=TRUE)
+    close(infile)
+    if (! "DT" %in% names(params)) params$DT <- 1.0
+    repro_self <- repro[,grepl("self_", names(repro))] |> revCumRowSums()
+    repro_ma <- repro[,grepl("ma_", names(repro))] |> revCumRowSums()
+    repro_pa <- repro[,grepl("pa_", names(repro))] |> revCumRowSums()
+    midp <- (repro_ma + repro_pa)/2
+    seg <- repro_self - midp
+    return(list(midp=midp, seg=seg))
+}
+
+plot_trait_traces <- function (examples, labels) {
+    layout(matrix(1:4, nrow=2), heights=c(1.5,1))
+    for (x in names(examples)) {
+        fix <- load_fix(examples[[x]])
+        pop <- load_pop(examples[[x]])
+        par(mar=c(4,4,2,0)+.1, mgp=c(2.5,1,0))
+            plot(med_trait ~ time, data=fix, type='l',
+                 main=x, xlab='time', ylab='median trait value')
+        mtext(labels[[x]][1], 3, adj=-0.0, line=0.2)
+        par(mar=c(4,4,0,0)+.1, mgp=c(2.5,1,0))
[…]

### 2023-07-04 sims/do_figs.sh
.

@@ -4,7 +4,7 @@ for type in normal cauchy
 do
     for seed in 8135 8791
     do
-        sbatch run_sim.srun "-s ${seed}66${xopt} -d BURNIN=40000 -d NUMGENS=10000 -d N=10000 neutral_${type}.slim"
+        sbatch run_sim.srun "-s ${seed} -d BURNIN=40000 -d NUMGENS=10000 -d N=10000 neutral_${type}.slim"
         for xopt in 0 1 2 10
         do
             sbatch run_sim.srun "-s ${seed}55${xopt} -d BURNIN=40000 -d NUMGENS=10000 -d N=10000 -d STEP=10 -d XOPT=${xopt}.0 ${type}.slim"

### 2023-07-04 sims/make_paper_figs.R

@@ -98,7 +98,8 @@ plot_seg_noise <- function (examples, labels, seg_col, main_append="", ...) {
         mtext(labels[[x]][1], 3, adj=-0.0, line=0.2)
         par(mar=c(4,4,0,0)+.1, mgp=c(2.5,1,0))
             hist(this_seg, breaks=40, main='',
-                 xlab='segreation noise')
+                 xlab='segregation noise')
+
         mtext(labels[[x]][2], 3, adj=-0.0, line=0.2)
             plot_conditional_ratios(this_midp, this_seg, do_legend=(x == names(examples)[1]), ...)
         mtext(labels[[x]][3], 3, adj=-0.0, line=0.2)
@@ -109,8 +110,10 @@ plot_seg_noise <- function (examples, labels, seg_col, main_append="", ...) {
 ###########
 # Neutral examples
 examples <- list(
-                 "neutral Normal" = "sim_neutral_normal_831",
-                 "neutral Cauchy" = "sim_neutral_cauchy_831"
+                 # "neutral Normal" = "sim_neutral_normal_831",
+                 # "neutral Cauchy" = "sim_neutral_cauchy_831"
+                 "neutral Normal" = "sim_neutral_normal_813566",
+                 "neutral Cauchy" = "sim_neutral_cauchy_81356610"
 )
 
 # 1. Plots of median trait value over time and final trait distribution

### 2023-07-04 sims/make_paper_figs.sh

@@ -2,8 +2,8 @@
 SLIM="sbatch run_sim.srun"
 # SLIM=slim
 
-$SLIM -s 831 -d BURNIN=20000 -d NUMGENS=1000 -d N=1000 -d 'EPSILON=c(0,0.1,10)' neutral_cauchy.slim
-$SLIM -s 831 -d BURNIN=20000 -d NUMGENS=1000 -d N=1000 -d 'EPSILON=c(0,0.1,10)' neutral_normal.slim
+$SLIM -s 831 -d BURNIN=20000 -d NUMGENS=1000 -d N=1000 -d 'EPSILON=c(0,0.2,1.0,10)' neutral_cauchy.slim
+$SLIM -s 831 -d BURNIN=20000 -d NUMGENS=1000 -d N=1000 -d 'EPSILON=c(0,0.2,1.0,10)' neutral_normal.slim
 
-$SLIM -s 707 -d BURNIN=20000 -d NUMGENS=1000 -d N=1000 -d STEP=20 -d 'EPSILON=c(0,0.1,10)' cauchy.slim
-$SLIM -s 707 -d BURNIN=20000 -d NUMGENS=1000 -d N=1000 -d STEP=20 -d 'EPSILON=c(0,0.1,10)' normal.slim
+$SLIM -s 707 -d BURNIN=20000 -d NUMGENS=1000 -d N=1000 -d STEP=20 -d 'EPSILON=c(0,0.2,1.0,10)' cauchy.slim
+$SLIM -s 707 -d BURNIN=20000 -d NUMGENS=1000 -d N=1000 -d STEP=20 -d 'EPSILON=c(0,0.2,1.0,10)' normal.slim

### 2023-07-04 writeup.tex
not too many rows plz

@@
 and $f(z) = (\delta/\gamma) (\gamma + z^2) / (\delta + z^2)$.
-Figure XXX shows the resulting distributions of segregation noise:
-i.e., the distribution of child phenotype minus average of parents' phenotypes,
-across all offspring following burn-in.
+Figures~\ref{fig:sel_trait_distrns} and~\ref{fig:sel_seg_noise}
+show that the picture is largely unchanged by stabilizing selection
+except that, unsurprisingly, the median trait is no longer constrained to wander
+(but still shows substantially larger fluctuations under the Cauchy model).
+
+\begin{figure}
+    \begin{center}
+        \includegraphics{sims/selected_seg_noise_small}
+    \end{center}
+    \caption{
+        As in Figure~\ref{fig:seg_noise},
+        except computed using only mutations with absolute value effect size less than 0.1.
+        \label{fig:seg_noise_small}
+    }
+\end{figure}
+
+However, as suggested above,
+the two models look much more similar when considering only alleles of small effect.
+Figure~\ref{fig:seg_noise_small} shows that ``midparent value''
+and ``segregation noise'' computed using only mutations with effect size less than 0.2
+(excluding less than 1\% of mutations)
+are quite close to independent.
+The result does not depend strongly on the cutoff chosen,
+and are similar even if less 0.1\% of mutations are excluded.
 

### 2023-07-04 sims/make_paper_figs.R

@@ -2,30 +2,6 @@
 library(jsonlite)
 source("helpers.R")
 
-load_files <- function (base) {
-    infile <- file(paste0(base, ".repro.tsv"), "r")
-    params <- fromJSON(readLines(infile, 1))
-    repro <- read.table(infile, sep="\t", header=TRUE)
-    close(infile)
-    if (! "DT" %in% names(params)) params$DT <- 1.0
-    pop <- read.table(paste0(base, ".pop.tsv"), sep="\t", header=TRUE)
-    pop$age <- pop$age * params$DT
-    fix <- read.table(paste0(base, ".fix.tsv"), sep="\t", header=TRUE)
-    repro_self <- repro[,grepl("self_", names(repro))] |> revCumRowSums()
-    repro_ma <- repro[,grepl("ma_", names(repro))] |> revCumRowSums()
-    repro_pa <- repro[,grepl("pa_", names(repro))] |> revCumRowSums()
-    midp <- (repro_ma + repro_pa)/2
-    seg <- repro_self - midp
-    return(list(
-                basename=base,
-                params=params,
-                midp=midp,
-                seg=seg,
-                pop=pop,
-                fix=fix
-    ))
-}
-
 load_params <- function (base) {
     infile <- file(paste0(base, ".repro.tsv"), "r")
     params <- fromJSON(readLines(infile, 1))
@@ -46,10 +22,10 @@ load_pop <- function (base) {
     return(pop)
 }
 
-load_seg_midp <- function (base) {
+load_seg_midp <- function (base, max_n=as.integer(1e6)) {
     infile <- file(paste0(base, ".repro.tsv"), "r")
     params <- fromJSON(readLines(infile, 1))
-    repro <- read.table(infile, sep="\t", header=TRUE)
+    repro <- read.table(infile, sep="\t", header=TRUE, nrows=max_n)
     close(infile)
     if (! "DT" %in% names(params)) params$DT <- 1.0
     repro_self <- repro[,grepl("self_", names(repro))] |> revCumRowSums()
@@ -99,7 +75,6 @@ plot_seg_noise <- function (examples, labels, seg_col, main_append="", ...) {
         par(mar=c(4,4,0,0)+.1, mgp=c(2.5,1,0))
             hist(this_seg, breaks=40, main='',
                  xlab='segregation noise')
-
         mtext(labels[[x]][2], 3, adj=-0.0, line=0.2)
[…]

### 2023-07-04 writeup.tex
edits through intro

@@
 \title{\large{\bf
-    Normality and large effect loci in quantitative genetics
+    Large effects and the infinitesimal model
 }}
@@
 \begin{abstract}
+    The infinitesimal model of quantitative genetics
+    relies on the Central Limit Theorem to stipulate that
+    under additive models of quantitative traits
+    determined by many loci having similar effect size,
+    the difference between an offspring's genetic trait component
+    and the average of their two parents' genetic trait components
+    is Normally distributed and independent of the parents' values.
+    Here, we investigate how the assumption of similar effect sizes
+    affects the model:
+    in particular, if the tail of the effect size distribution is polynomial
+    with exponent $\alpha < 2$, then sums of effects
+    should be well-approximated by a ``stable distribution'',
+    and find tail exponents between 1 and 2 in 
+    effect sizes estimated by genome-wide association studies
+    of many human disease-related traits.
+    We show that the independence of offspring trait deviations from parental averages
+    in many cases implies a Gaussian distribution,
+    suggesting that non-Gaussian models of trait evolution must expicitly
+    track the underlying genetics, at least for loci of large effect.
+    We also characterize possible limiting trait distributions
+    of the infinitesimal model with infinitely divisible noise distributions,
+    and compare our results to simulations.
 \end{abstract}
@@
 
-Much of the work to give a quantitative understanding to the observation that
-offspring of sexual organisms resemble their parent,
-has focused on \emph{quantitative genetics},
-which is interestingly intermediate between a mechanistic model of genetics
-and a purely phenomenological model.
-The action of natural selection on genetic variants in the absence of genetic variation
-is relatively well-understood (cite ewens),
-but tractable mathematical models of evolution for recombining species
-are elusive,
-at least with some results of when the model might be a good description of reality.
-For instance, a great deal of conceptual understanding and even quantitative predictions
-have been made assuming that the distribution of trait values in a population is Gaussian
-(cite Lande etcetera)
-despite the absence of a plausible model that would make this true (cite turelli and barton).
-One of the oldest and most concrete models of heritable trait evolution is the \emph{infinitesimal model}
-which assumes a simple, explicit form for how an offspring's trait is inherited from their parents':
-it is Normally distributed, with mean equal to the average of the parents' values
-and a variance independent of their values.
-Despite the plausibility of this model,
-(thanks to the Central Limit Theorem),
-it was not until \citet{barton2017infinitesimal} that conditions under which this would be a good approximation
-were made explicit,
-through a formal convergence proof.
+Much of the work to 
+understand the evolution of quantitative traits in sexual species
+has focused on models in quantitative genetics
+that describe evolution of trait values without explicitly representing underlying genetic variation,
+but in a manner deriving from our understanding of how genetic variation works in practice.
+Such models are therefore
+interestingly intermediate between mechanistic models of genetics and purely phenomenological.
+% The action of natural selection on genetic variants in the absence of genetic variation is relatively well-understood (cite ewens),
+% but tractable mathematical models of evolution for recombining species are elusive,
+% at least with some results of when the model might be a good description of reality.
+% For instance, a great deal of conceptual understanding and even quantitative predictions
+% have been made assuming that the distribution of trait values in a population is Gaussian (cite Lande etcetera)
+% despite the absence of a plausible model that would make this true (cite turelli and barton).
+One concrete model of heritable trait evolution is the \emph{infinitesimal model},
+which assumes a simple, explicit form for how the genetic component of offsprings' traits are determined by their parents:
+they are Normally distributed, with the mean for each offspring equal to the average of the parents' values,
+and a covariance independent of the parental trait values (but that depends on how they are related).
+\citet{barton2017infinitesimal} give a history of the model
+and a formal proof of conditions under which this would be a good approximation to reality,
+which they extended to include dominance in \citet{barton2022infinitesimal}.
 
-As suggested by its name, under the infinitesimal model the effects of any given genetic variant on the trait
+\comment{TODO: citations below}
+
+As suggested by its name,
+under the infinitesimal model the effects of any given genetic variant on the trait
 is small (indeed, vanishing),
-and so selection-driven trait shifts are mediated by small changes in frequencies of many alleles of small effect.
+and selection-driven trait shifts are mediated by small changes in frequencies of many alleles of small effect.
 However, empirical understanding of the genetic basis of trait variation in real-world species
@@
 There are many possible explanations for this observation,
-including selection/publication bias,
+including selection or publication bias,
 filters for larger-effect variants due to spatially varying selection (cite),
-ond/or grouping together of concordant alleles by linkage or inversions (cite).
-However, it seems clear that in at least those cases where we understand the underlying molecular mechanisms
-that mutations of large effect can be important in practice
-(cite examples: insecticide resistance, coloration, etc),
-and the consensus in the field is that traits exist on a spectrum between monogenic and polygenic,
+and/or grouping together of concordant alleles by linkage or inversions (cite).
+However, it seems clear that in at least those cases where we understand the underlying molecular mechanisms,
+mutations of large effect can be important in practice
+(cite examples: insecticide resistance, coloration, etcetera).
+The consensus in the field seems to be that traits exist on a spectrum between monogenic and polygenic,
 and that many - if not most - are firmly in the intermediate, ``oligogenic'' region (cite),
@@
 
-Concretely, the infinitesimal model postulates that
-if the parents' genetic contributions to the trait (i.e., their \emph{breeding values})
-are $z^{[1]}$ and $z^{[2]}$,
-and the offspring's genetic contribution to the trait is $z$,
-and $R = z - (z^{[1]} + z^{[2]})/2$ is the deviation from the midparent
-(which we call the \emph{segregation noise}), then
-\begin{enumerate}
-    \item $R$ is independent of $z^{[1]}$ and $z^{[2]}$, and
-    \item $R$ has a Gaussian distribution.
-\end{enumerate}
-(This is not a full definition,
-since the model makes broader statements involving conditioning on the full pedigree;
-see \citet{barton2017infinitesimal}.)
-We set out here to see whether (2) can be generalized,
-and find that (1) is apparently inexitrably connected.
-
 The Gaussian distribution appears thanks to ``the'' Central Limit Theorem,
@@
 of adding up a great many very small and independent things.
-However, as is well known to probabilists,
-there is a wider class of Central Limit Theorems,
+However, there is a wider class of Central Limit Theorems,
 which describe the distributions of sums of a great many independent and exchangeable things:
@@
 Roughly speaking, these say that if we add together $n$ independent copies of a random variable $X$
-for which $\P\{ X > t \} \propto t^{-\alpha}$ for some $0 < \alpha < 2$,
+for which $\P\{ X > t \}$ is proportional to $t^{-\alpha}$ for some $0 < \alpha < 2$,
 then -- almost regardless of the details -- the result, scaled by $n^{1/\alpha}$,
-is well-approximated by a universal form, the $\alpha$-stable distribution (and becomes exact, as $n \to \infty$).
-In such ensembles, sinle large entries are important:
+is well-approximated by a universal form, an $\alpha$-stable distribution
+(and the approximation becomes exact, as $n \to \infty$).
+In such ensembles, single large entries can be important:
 concretely, the largest of the $n$ copies will be of order $n^{1/\alpha}$,
@@
 In this paper, we make preliminary explorations in that direction:
-Does the universality shown by \citet{barton2017infinitesimal}
+Do the results of \citet{barton2017infinitesimal}
 carry over when the distribution of effect sizes falls in the domain of attraction of a stable law?
-What can be said about equilibrium trait distributions in this case?
-And, is there any biological evidence that this theory might be relevant?
+The immediate answer to this will be ``no'' --
+in fact, we will see that independence of offsprings' deviations and midparent values \emph{implies} a Gaussian distribution in some sense.
+This independence is a key reason why the infinitesimal model is tractable,
+and so a conclusion is that additive models with non-Gaussian effects may not be well-approximated
+by ``trait-only'' models that do not explicitly represent underlying genetic variation.
+Having thus shown that ``infinitesimal models with non-Gaussian noise'' is on weak footing as a model of reality,
+we then go on to analyze exactly these models.
+(As motivation,
+we have not shown that such models are \emph{not} good approximations to reality,
+and it is interesting and perhaps informative to understand how the noise distribution affects results.)
 
@@
 are better modeled by essentially Brownian random walks or by L\'evy processes (cite),
-and a growing literature studies the effects of such ``fat-tailed'' distributions
+and a growing literature studies the effects of such power law distributions
 on the shapes and dynamics of traveling waves such as biological invasions (cite).
@@
 better than either a Brownian model.
-\comment{more}
+It is also important to note that
+the infinitesimal model implies a Gaussian distribution for the difference between an offspring's trait and their parents' average,
+\emph{not} for the entire population.
+The question of when the trait distribution in an evolving population
+might be modeled as a Gaussian has been discussed extensively, e.g., by \comment{cite}.
 
-\comment{TODO expand, put somewhere:}
-There are several justifications for why a ``stable infinitesimal model'' might be useful:
-(1) as a phenomenological model;
-(2) if it matches empirical observations; or
-(3) if there is a central limit theorem suggesting it would be a good approximation in a wide range of situations.
+The ways in which genetic variants affect an organisms' trait can be complex,
+and interactions between alleles, between loci, and with the environment are often important in practice.
+In fact, many traits claimed to be genetically determined
+in fact only appear heritable through non-genetic means,
+e.g., systematic inequalities in resource distribution partitioned by class or race. \comment{e.g., cite}
+In this paper, we follow the literature in working with the relatively simple purely additive case
+of a trait that can be decomposed into a sum across loci of the contributions of the alleles from each locus,
+even omitting an independent effect of ``the environment''.
+However, we emphasize that such models should \emph{not} always be the default for understanding biological variation,
+especially for human traits.
 
-\subsection*{Orphaned text}
-The ways in which genetic variants combine to determine an organisms' trait can be complex,
-and epistatic interactions between alleles and between loci can be important in practice.
-However, the question is sufficiently complex in the purely additive case
-that a trait can be decomposed into a sum across loci of the contributions of the alleles from each locus.
-Under this additive assumption it makes sense to discuss the ``effect size distribution'',
-i.e., the distribution of how much alleles affect the trait value.
+%%%%%%%%%%%%%%%%%%%%%%
+\subsection*{Notation}
 
-An offspring's trait differs from their parents for many reasons,
-often most importantly because of differences in the environment in which they live.
-The focus of quantitative genetics is on the contribution of heritable material to this variation:
-the offspring's genetic material
+We generally follow the notation of \citet{barton2017infinitesimal},
+although in a simpler situation,
+as we do not consider full pedigrees
+but rather only the haploid offspring of a single set of haploid parents
+(as produced, for instance, by a round of meiosis in a diploid stage),
+and ignore non-genetic contributions to the trait;
+so, below we say ``trait'' to mean ``genetic contribution to the trait''.
 
-\comment{Will need this notation below:}
-We focus on the \emph{infinitesimal model},
-as described by \citet{barton2017infinitesimal},
-which supposes that the trait value of an offspring
-is equal to the average of their parents' trait values,
-plus independent Gaussian noise with fixed variance,
-\begin{equation} \label{eqn:offspring_model}
-    X_o = \frac{X_m + X_f}{2} + \epsilon,
-\end{equation}
-where $\epsilon$, the ``segregation noise'',
-is Gaussian with mean zero and variance $V_0$.
-\comment{should call this something besides $\epsilon$}
+Suppose that the focal individual's trait is
+$$
+    Z_M = \sum_{\ell=1}^M \eta_\ell / M^{1/\alpha},
+$$
+where $\eta_\ell$ is the contribution of the allele carried at the $\ell^\text{th}$ locus.
+The effects of the alleles carried by the two parents
+are $\eta_\ell^{[1]}$ and $\eta_\ell^{[2]}$, respectively,
+and their traits are
+$Z_M^{[1]} = \sum_\ell \eta_\ell^{[1]} / M^{1/\alpha}$ and
+$Z_M^{[2]} = \sum_\ell \eta_\ell^{[2]} / M^{1/\alpha}$.
+We assume the loci are unlinked and inheritance is fair, so
+letting $X_\ell$ be a family of Bernoulli(1/2) random variables that are independent of everything else,
+$$
+    \eta_\ell = X_\ell \eta_\ell^{[1]} + (1 - X_\ell) \eta_\ell^{[2]} .
+$$
+The \emph{transmission noise}, $R_M$, is the offspring's trait minus the midparent value, $\bar Z_M$:
+\begin{align*}
+    \bar Z_M &= \frac{Z_M^{[1]} + Z_M^{[2]}}{2} \\
+    R_M &= Z_M - \bar Z_M \\
+        &= \sum_{\ell=1}^M
+            \frac{
+                X_\ell (\eta_\ell^{[1]} - \eta_\ell^{[2]})
+                + (1 - X_\ell) (\eta_\ell^{[2]} - \eta_\ell^{[1]})
+            }{ 2 } .
+\end{align*}
 
@@
 plots including the 413 ``filtered'' phenotypes are provided in
-Supplementary Figure~\ref{sfig:unfiltered_hist}.
+Supplementary Figure~\ref{fig:unfiltered_hist}.
 
@@
     If the segregation noise is independent of the midparent value
-    (i.e., $R_M$ is independent of $(Z^{[1]} + Z^{[2]})/2$)
+    (i.e., $R_M$ is independent of $\bar Z_M$)
     then $R_M$ is Gaussian.
@@
     \end{center}
+    \caption{
         Segregation noise with Gaussian \textbf{(A-C)} and Cauchy \textbf{(D-F)} effect size distributions,
@@
         (green) the 90\% quantile of midparent value.
-    \caption{
         \label{fig:seg_noise}
@@
     \begin{center}
-        \includegraphics{sims/selected_seg_noise_small}
+        \includegraphics{sims/neutral_seg_noise_small}
     \end{center}
@@
         except computed using only mutations with absolute value effect size less than 0.1.
+        The same quantities for the simulations of Figure~\ref{fig:sel_seg_noise}
+        with stabilizing selection are shown in Figure~\ref{fig:sel_seg_noise_small}.
         \label{fig:seg_noise_small}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%
+\section{Discussion}
+
+Not assuming all effects are the same order of magnitude seems very bioloically realistic,
+and so it'd be nice to figure out how evolution works and how to do quant gen in such cases.
+Such models can always be simulated,
+but the way forwards with theoretical understanding seems murky.
+It would be interesting to see how changing the distribution of segregation noise
+in the infinitesimal model
+affects properties like the rate of adaptive evolution,
+levels of genetic load,
+and the rate of fixation.
+However,
+we have seen that this assumption imples a sort of ``blending inheritance''
+for the large effect alleles,
+and so such work much be accompanied by simulations
+to see if the predictions are borne out under concrete models of genetic inheritance.
+Another open question is whether there is some set of assumptions
+(perhaps involving linkage, pleiotropy, and/or environmental interactions with genetic effects)
+which leads to these non-Gaussian infinitesimal models.
+Finally, it may well be that there is a clever way to set up a trait-only model of evolution
+that is both insensitive to the underlying genetics
+and does not rely on the Gaussian central limit theorem.
+
+Above we've mostly used the Cauchy (e.g., in simulations)
+not because it seems most realistic
+(indeed data suggests $\alpha=3/2$ might be better)
+but for convenience and to make effects more obvious.
+
 \bibliographystyle{plainnat}
@@
 \appendix
+\setcounter{table}{0}
+\renewcommand{\thetable}{S\arabic{table}}
+\setcounter{figure}{0}
+\renewcommand{\thefigure}{S\arabic{figure}}
 
@@
 
+\begin{figure}
+    \begin{center}
+        \includegraphics{sims/selected_seg_noise_small}
+    \end{center}
+    \caption{
+        As in Figure~\ref{fig:seg_noise},
+        except computed using only mutations with absolute value effect size less than 0.1.
+        \label{fig:sel_seg_noise_small}
+    }
+\end{figure}
+
 \end{document}

### 2023-07-05 writeup.tex
peter edits

@@
 % First answer to the motivating question:  "No"
-% Non-Gaussian effects (sometimes) imply nonindependence of segregation noise:
+% Non-Gaussian effects (sometimes) imply nonindependence of transmission noise:
 %   In at least some cases, natural non-gaussian models produce nonindependence:
@@
 %  - (I have other thoughts to be sorted out here)
-%  - Open Q (maybe?): is there any non-Gaussian model where segregation noise is independent of parental traits?
+%  - Open Q (maybe?): is there any non-Gaussian model where transmission noise is independent of parental traits?
 % 
@@
     in many cases implies a Gaussian distribution,
-    suggesting that non-Gaussian models of trait evolution must expicitly
+    suggesting that non-Gaussian models of trait evolution must explicitly
     track the underlying genetics, at least for loci of large effect.
@@
 $$
-The \emph{transmission noise}, $R_M$, is the offspring's trait minus the midparent value, $\bar Z_M$:
+What we will call the \emph{transmission noise}, $R_M$,
+is the offspring's trait minus the midparent value, $\bar Z_M$:
 \begin{align*}
@@
 
-It is mathematically intriguing to speculate what effect a non-Gaussian distribution
-for $\epsilon$ would have on the dynamics induced by~\eqref{eqn:offspring_model}.
-But, is there any evidence that non-Gaussian segregation noise is important in practice?
+It is mathematically intriguing to speculate what effect a power law effect size distribution
+for would have on the induced evolutionary dynamics.
+But, is there any evidence that non-Gaussian transmission noise is important in practice?
 In this paper we focus primarily on the theory,
@@
 To do this, we downloaded the GWAS results provided by~\citet{biobankSNPs}
-for 7,221 phenotypes from 361,194 individuals from the March 2018 release of the UK Biobank.
+for 7,221 phenotypes from 361,194 human individuals from the March 2018 release of the UK Biobank.
 After filtering (see details below) we were left with 695 binary illness-related phenotypes.
 Taken at face value, these provide for each phenotype a set of estimates of the additive effect
-of the alternative allele at a set of SNPs on the phenotype relative to the reference allele.
+of the alternative allele at a set of single nucleotide polymorphisms (SNPs)
+on the phenotype relative to the reference allele.
 The phenotypes we consider were all binary and coded as 0/1,
 so estimates could be thought of as additive adjustments to the risk
-(i.e., effects are \emph{not} on a logit scale).
+(so, effects are \emph{not} on a logit scale).
 These estimates were obtained as the coefficient for the SNP in a simple linear model
@@
 
-Suppose that the phenotype of each individual is additive:
-the maternal trait value is $x = \sum_i x_{i,1} + x_{i,2}$
-and the paternal trait value is $y = \sum_i y_{i,1} + y_{i,2}$,
-where $x_{i,1}$ and $x_{i,2}$ are the effects of the alleles carried
-by the mother on each of the two chromosomes
-at the $i^\text{th}$ locus, respectively.
-Then the offspring's trait is $z = \sum_i z_{i,m} + z_{i,p}$,
-where $z_{i,m} = \theta_{i,m} x_{i,1} + (1-\theta_{i,m}) x_{i,2}$,
-and $\theta_{i,m} = 1$ with probability 1/2 and $\theta_{i,m} = 0$ otherwise;
-and likewise $z_{i,p} = \theta_{i,p} y_{i,1} + (1-\theta_{i,p}) y_{i,2}$.
-With this notation, the segregation noise $z - \frac{x+y}{2}$
-is a sum of terms of the form
+Since the transmission noise is a sum of terms of the form
 $$
-    \left( \theta_{i,m} - 1/2 \right) x_{i,1}
-    + \left( 1/2 - \theta_{i,m} \right) x_{i,2} 
+    (X_\ell - 1/2) \eta_\ell^{[1]} + (1/2 - X_\ell) \eta_\ell^{[2]},
 $$
-(and similar terms with $y$ in place of $x$).
-Now let $X$ be the effect of a uniformly chosen allele carried by a uniformly chosen individual
-at a uniformly chosen locus.
-The appropriate central limit theorem is determined by the tail behavior of $|(\theta - 1/2) X| = |X|/2$.
-So, we'd like to know the value of
+where $\eta_\ell^{[i]}$ is the effect of the allele in parent $i$ at locus $\ell$,
+the appropriate central limit theorem for the sum is determined by
+the tail behavior of $|(X - 1/2) \eta| = |\eta|/2$,
+where $X$ is Bernoulli(1/2) and $\eta$ is the effect
+of a uniformly chosen allele at a uniformly chosen locus.
+So, we'd like to know
+if $\P\{|X|>2t\}$ is proportional to $t^{-\alpha}$,
+i.e., the value of
 $$
@@
 if the limit exists.
-Let $p_i$ be the minor allele frequency at the $i^\text{th}$ locus;
+Let $p_\ell$ be the minor allele frequency at the $\ell^\text{th}$ locus;
 since the data we have assigns effect 0 to the major allele,
-$x_i = 0$ with probability $1-p_i$.
-Therefore, if we have estimated effects for $N$ SNPs
-with minor allele frequency $p_i$ and effect size $e_i$ for the minor allele,
-for $1 \le i \le N$,
+$\eta_\ell = 0$ with probability $1-p_\ell$.
+Therefore, if we have estimated effects for $M$ SNPs
+with minor allele frequency $p_\ell$ and effect size $e_\ell$ for the minor allele,
+for $1 \le \ell \le M$,
 then
 $$
-    \P\{ |X| > 2t \} \approx \frac{1}{N} \sum_{i : |e_i| > 2t} p_i .
+    \P\{ |\eta| > 2t \} \approx \frac{1}{M} \sum_{\ell : |e_\ell| > 2t} p_\ell .
 $$
 So, to obtain a per-trait estimate of $\alpha$,
-we plot $\log \frac{1}{N} \sum_{i : |e_i| > t} p_i$ against $\log t$
+we plot $\log \frac{1}{M} \sum_{\ell : |e_\ell| > t} p_\ell$ against $\log t$
 for a sequence of values of $t$,
@@
 is between 20 and 10\% of all SNPs).
-Examples are shown in figure~\ref{fig:example_snps}.
+Examples are shown in Figure~\ref{fig:example_snps}.
 
@@
     \caption{
-        TODO: cut this down to just 2 SNPs and move this figure to the supplement; change 'x' to 't' in axis labels.
+        % TODO: cut this down to just 2 SNPs and move this figure to the supplement; change 'x' to 't' in axis labels.
         Tail distributions of frequency-weighted effects
@@
         ``num snps'' is the number of SNPs for the phenotype with $p$-value less than $10^{-8}$,
-        and ``num cases'' is the reported number of the rougly 360,000 individuals
+        and ``num cases'' is the reported number of the roughly 360,000 individuals
         recorded as having the listed phenotype.
@@
 We applied this procedure to results from models fit on both sexes,
-after removed phenotypes related to employment, drug treatments, and diet.
-This left us with 1,108 phenotypes for which which could estimate the tail exponent.
-For the main results,
-we then restricted to phenotypes having non-missing numbers of ``cases'' and ``controls''
-and a sample size of at least 300,000 (i.e., not more than 61,194 of 361,194 missing phenotypes).
-We also removed
-the 65 remaining phenotypes that are common (more than 10,000 cases),
+after removing phenotypes related to non-medical traits
+(e.g., employment, drug treatments, and dietary choices).
+This left us with 1,108 phenotypes for which we had enough data (e.g., enough significant SNPs)
+to estimate the tail exponent.
+We then restricted to phenotypes having non-missing numbers of ``cases'' and ``controls''
+and a sample size of at least 300,000 (i.e., not more than 61,194 of 361,194 missing phenotypes),
+and removed the 65 remaining phenotypes that are common (more than 10,000 cases),
 as these often showed very different patterns
@@
 This left us with 695 phenotypes;
-plots including the 413 ``filtered'' phenotypes are provided in
+plots including the 413 phenotypes removed by these last steps are provided in
 Supplementary Figure~\ref{fig:unfiltered_hist}.
@@
 and 28.9\% had estimated values less than 1.5.
-A concern could be that ``noisier'' phenotypes,
+A concern is that ``noisier'' phenotypes,
 for which estimated effect sizes are less reliable,
 might lead to larger estimated tail exponents;
-for this reasons, we looked for an association between
+for this reason, we looked for an association between
 tail exponent and two proxies for power,
@@
 and number of cases.
-Interestingly, the result in nonmonotonic,
+Interestingly, the relationship is nonmonotonic,
 with exponents closer to $\alpha=2$ for phenotypes with around 1000 cases,
-and lower exponents for rarer \emph{and} more common phenotypes.
+and lower exponents for both rarer and more common phenotypes.
 (A similar pattern is seen for ``number of SNPs'',
-but this may be due to association with number of cases.)
+but this may be due to association with power and hence number of cases.)
 This may indicate a statistical artifact,
@@
 
-We generally follow the notation of \citet{barton2017infinitesimal},
-although in a simpler situation,
-as we consider only the haploid offspring of a single set of haploid parents
-rather than full pedigrees,
-and ignore non-genetic contributions to the trait;
-so, below we say ``trait'' to mean ``genetic contribution to the trait''.
-
-Suppose that the focal individual's trait is
-$$
-    Z_M = \sum_{\ell=1}^M \eta_\ell / M^{1/\alpha},
-$$
-where $\eta_\ell$ is the contribution of the allele carried at the $\ell^\text{th}$ locus.
-The effects of the alleles carried by the two parents
-are $\eta_\ell^{[1]}$ and $\eta_\ell^{[2]}$, respectively,
-and their traits are
-$Z_M^{[1]} = \sum_\ell \eta_\ell^{[1]} / M^{1/\alpha}$ and
-$Z_M^{[2]} = \sum_\ell \eta_\ell^{[2]} / M^{1/\alpha}$.
-We assume the loci are unlinked and inheritance is fair, so
-letting $X_\ell$ be a family of Bernoulli(1/2) random variables that are independent of everything else,
-$$
-    \eta_\ell = X_\ell \eta_\ell^{[1]} + (1 - X_\ell) \eta_\ell^{[1]} .
-$$
-
 We know from \citet{barton2017infinitesimal} that if the $\eta^{[i]}_\ell$ are
@@
 
-Is there a converse? In other words,
-does independence of segregation noise imply a Gaussian distribution of trait values?
-It is not clear to us in how much generality such a statement might be made,
-but we can at least prove the following.
-Divide the parental loci into groups according to the results of inheritance,
+Perhaps surprisingly, there is a converse --
+in at least some situations,
+independence of transmission noise implies a Gaussian distribution.
+To see this,
+divide the parental loci into groups according to the results of inheritance,
 defining
 \begin{align*}
-    U_M^{[i]} = \sum_{\ell=1}^M X_\ell \eta_\ell^{[i]} 
-    \qquad \text{ and } \qquad
-    V_M^{[i]} = \sum_{\ell=1}^M (1-X_\ell) \eta_\ell^{[i]} ,
+    U_M^{[1]} &= \sum_{\ell=1}^M X_\ell \eta_\ell^{[1]}  ,
+    \qquad \qquad
+    V_M^{[1]} = \sum_{\ell=1}^M (1-X_\ell) \eta_\ell^{[1]} ,  \\
+    U_M^{[2]} &= \sum_{\ell=1}^M (1-X_\ell) \eta_\ell^{[2]} ,
+    \qquad \qquad
+    V_M^{[2]} = \sum_{\ell=1}^M X_\ell \eta_\ell^{[2]}  ,
 \end{align*}
@@
 $Z_M^{[1]} = U_M^{[1]} + V_M^{[1]}$ and
-$Z_M^{[2]} = U_M^{[2]} + V_M^{[2]}$, so
+$Z_M^{[2]} = U_M^{[2]} + V_M^{[2]}$, while
+$Z_M = U_M^{[1]} + U_M^{[2]}$ is the transmitted portion of the parental genomes,
+and we can define $V_M = V_M^{[1]} + V_M^{[2]}$ to be the untransmitted portion.
+Furthermore,
 \begin{align*}
-    Z_M = U_M^{[1]} + V_M^{[2]} \qquad \text{ and } \qquad
-    R_M = \frac{U_M^{[1]} - V_M^{[1]} - U_M^{[2]} + V_M^{[2]}}{ 2 } .
+    \bar Z_M = \frac{Z_M + V_M}{2}
+    \qquad \text{ and } \qquad
+    R_M = \frac{Z_M - V_M}{ 2 } .
 \end{align*}
-It turns out that independence of the portions inherited and not inherited by the offspring
-combined with independence of segregation noise and midparent value
-implies a Gaussian distribution:
-
-\begin{lem}\label{lem:parental_contribs}
-    Suppose that $\{\eta_\ell^{[i]}\}$ are independent for all $\ell$ and $i$,
-    that the distribution of $\eta_\ell^{[i]}$ only depends on $\ell$,
-    and that $\{X_\ell\}$ are independent of $\{\eta_\ell^{[i]}\}$.
-    If the segregation noise is independent of the midparent value
-    (i.e., $R_M$ is independent of $\bar Z_M$)
-    then $R_M$ is Gaussian.
-\end{lem}
-
-\begin{proof}[Proof of Lemma~\ref{lem:parental_contribs}]
-    Note that if inheritance (the $X_\ell$) is independent of the per-locus effects ($\eta_\ell^{[i]}$),
+
+Now, note that if inheritance (the $X_\ell$) is independent of the per-locus effects ($\eta_\ell^{[i]}$),
 the latter are all independent of each other,
@@
 -- they are simply two independent draws from the distribution of effects at locus $\ell$.
-    Therefore, the sum of inherited alleles, $U_M^{[1]} + V_M^{[2]}$
-    are independent of the non-inherited alleles, $U_M^{[2]} + V_M^{[1]}$.
-
-    Now, the Kac-Bernstein theorem states that if two independent random variables
+Therefore, the inherited alleles that sum to $Z_M$
+are independent of the non-inherited alleles that sum to $V_M$.
+
+The Kac-Bernstein theorem states that if two independent random variables
 have the property that if their sum and difference are independent of each other,
@@
 (\citet{kac1939characterization,bernstein1941property},
-    but see \citet{kagan1973characterization} for many much more general results).
-    Since $R_M$ is (half) the difference and $(Z^{[1]} + Z^{[2]})/2$ is (half) the sum
-    of $U_M^{[1]} + V_M^{[2]}$ and $U_M^{[2]} + V_M^{[1]}$,
-    both are therefore jointly Gaussian.
-\end{proof}
-
-Since models can be easily set up for which $Z$ is not Gaussian,
-the content of this is that for such models, independence of segregation noise
+and see \citet{kagan1973characterization} for many much more general results).
+Since $R_M$ is (half) the difference and $\bar Z_M$ is (half) the sum
+of $Z_M$ and $V_M$, and $Z_M$ and $V_M$ are independent,
+if we assume that also $R_M$ and $\bar Z_M$ are independent,
+then both are therefore jointly Gaussian.
+
+We have shown the following:
+
+\begin{prop}\label{prop:parental_contribs}
+    Suppose that $\{\eta_\ell^{[i]}\}$ are independent for all $\ell$ and $i$,
+    that the distribution of $\eta_\ell^{[i]}$ only depends on $\ell$,
+    and that $\{X_\ell\}$ are independent of $\{\eta_\ell^{[i]}\}$.
+    If the transmission noise $R_M$ is independent of the midparent value $\bar Z_M$,
+    then $R_M$ and $\bar Z_M$ are jointly Gaussian.
+\end{prop}
+
+% A similar statement holds for the offspring of two parents in a pedigree,
+% using the Skitovich-Darmois theorem \citet{kagan1973characterization},
+% except that the portion of the genome shared in common by the parents is unconstrained.
+
+Since models can be easily set up for which trait distributions are not Gaussian,
+the implication of this is that for such models, independence of transmission noise
 is not likely a good assumption.
@@
 
-\comment{remove or shorten the following?}
-It is clear in the Gaussian case
-that $U^{[1]}$ and $V^{[1]}$ are asymptotically independent.
-More generally, independence depends on how the model is set up.
-If the allelic effects $\eta_\ell^{[i]}$ are i.i.d.,
-then even for finite $M$, the parental contributions
-$(U_M^{[1]}, V_M^{[1]}, U_M^{[2]}, V_M^{[2]})$ are independent.
-If the distribution of $\eta_\ell^{[i]}$ is chosen to obtain a non-Gaussian limit,
-then Lemma~\ref{lem:parental_contribs} implies that $R$ cannot be independent of the parental traits.
-
-However, the presence of a large effect polymorphic allele
-makes $(U^{[1]}, V^{[1]}, U^{[2]}, V^{[2]})$ dependent:
+The presence of a large effect polymorphic allele
+makes $R_M$ and $\bar Z_M$ dependent:
 consider the extreme case where $\eta^{[i]}_\ell$ are i.i.d.{} with variance 1 for $\ell \ge 2$,
 while $\eta^{[i]}_1 = \pm 10$ with probability 1/2 each,
-and $Z^{[1]} = \eta_1^{[1]} + \sum_{\ell=2}^M \eta_\ell^{[1]} / \sqrt{M}$.
-Then in the limit $U^{[i]}$ and $V^{[i]}$ are Gaussian with mean either 0 or $\pm 10$ and variance 1/2,
-and are highly dependent: $|U^{[i]}| \approx 10$ or $|V^{[i]}| \approx 10$ but not both.
-(Segregation noise, meanwhile, depends strongly on parental trait values:
-if $Z^{[1]} \approx Z^{[2]}$ then most likely
+and $Z^{[i]} = \eta_1^{[i]} + \sum_{\ell=2}^M \eta_\ell^{[i]} / \sqrt{M}$.
+Then in the limit $R_M$ depends strongly on parental trait values:
+if $Z^{[1]} \approx Z^{[2]}$ and so $\bar Z \approx \pm 10$ then most likely
 $\eta_1^{[1]} = \eta_1^{[2]}$ and so $R$ is Normal with variance 1/2;
-but in the other case then $R$ has equal probabilities of being close to $+5$ or $-5$.)
+but if $\eta_1^{[1]} \neq \eta_1^{[2]}$ (and so $\bar Z \approx 0$)
+then $R$ is close to $+5$ or $-5$ with equal probability.
 
@@
 where $y$ and $z$ are independent draws from $X$
-and $\xi$ is an independent draw from the segregation noise distribution.
+and $\xi$ is an independent draw from the transmission noise distribution.
 Let $p_\xi$ denote the distribution of $\xi$,
@@
 This implies that after rescaling time by $N$
-and suitible other caveats,
+and suitable other caveats,
 the limiting process is a deterministic flow in the direction of 
@@
 In other words, any measure $Z$ that is a fixed point has the property that
-(the average of two independent draws plus segregation noise)
+(the average of two independent draws plus transmission noise)
 is equal in distribution to
@@
 
-In Section~\ref{sec:stabilizing_selection}
-we have found two explicit conditions in which
-if the form of segregation noise and stabilizing selection
-satisfy certain conditions,
-then the limiting phenotype distribution
-will be either Gaussian (Subsection~\ref{sec:stabilizing_gaussian})
-or Cauchy (Subsection~\ref{sec:stabilizing_cauchy}).
-Furthermore, we expect based on
-\citet{barton2017infinitesimal} and arguments above
-that the distribution of segreation noise
-is determined by the tail exponent of the distribution of effect sizes
-(under appropriate assumptions; e.g., that most mutations are of small effect).
-However, it is not clear how selection would affect these arguments,
-especially in the stable case where some number of mutations are expected to
-\emph{not} have small effect (and therefore such mutations will not be nearly neutral).
-
-Therefore,
-to study how this works in practice, we turn to simulation.
+Next we turn to simulation,
+in which we can easily include realistic recombination (instead of unlinked loci) and selection.
 We used SLiM v4 \citep{haller2022slim4}
@@
 by summing the effects of all mutations they carry across both chromosomes.
-Let $f(z) = 1 + \mu(z)$
-be the ``fitness'' of an individual with phenotype $z$,
+Let $f(z)$ be the death rate of an individual with phenotype $z$,
 and let $dt$ be a small constant (we took $dt=0.01$).
+In all plots, one time unit is approximately one generation.
 Then, at each time step
@@
         had populations of size $N=1000$,
-        a genome of length $L=10^8$ and mutation rate of $\mu=10^{-9}$;
+        a genome of length $L=10^8$ bp and mutation rate of $\mu=10^{-9}$ per bp per generation;
         the effect size distributions were
@@
         \textbf{(A,D)} Segregation noise (offspring trait minus midparent) plotted against midparent.
-        \textbf{(B,E)} Histograms of segregation noise.
-        \textbf{(B,E)} Enrichment in quantiles of segregation noise conditional on midparent value.
+        \textbf{(B,E)} Histograms of transmission noise.
+        \textbf{(B,E)} Enrichment in quantiles of transmission noise conditional on midparent value.
         To compute the latter,
-        let $q_0, q_2, q_4, \ldots, q_{98}, q_{100}$ divide the distribution of segregation noise
+        let $q_0, q_2, q_4, \ldots, q_{98}, q_{100}$ divide the distribution of transmission noise
         into 50 bins with roughly equal numbers in each (i.e., the quantiles),
         let $p_k(a,b)$ denote the proportion of the offspring of parents with midparent value
-        is in $[a, b)$ whose segregation noise is in $[q_{k-1}, q_k)$,
-        and let $p_k$ be the proportion of all offspring with segreation noise in $[q_{k-1}, q_k)$
+        is in $[a, b)$ whose transmission noise is in $[q_{k-1}, q_k)$,
+        and let $p_k$ be the proportion of all offspring with segregation noise in $[q_{k-1}, q_k)$
         (so, $p_k \approx 0.02$).
@@
 
-We first consider the neutral case (i.e., with $\mu \equiv 0$).
+We first consider the neutral case (i.e., with $f \equiv 1$).
 As shown in Figure~\ref{fig:trait_distrns},
-a Cauchy effect size distribution, unsurprisingly,
-produces a distribution of trait values with more extreme values
+a simulation with a Cauchy effect size distribution, unsurprisingly,
+has a distribution of trait values with more extreme values
 and a rougher path of population median trait value over time
-than in the Gaussian case.
+than does a simulation with a Gaussian effect size distribution.
 Indeed, the sum of all effects along the path from the start of the simulation
@@
 or a Cauchy process (in the Cauchy case).
-Figure~\ref{fig:seg_noise} shows how segregation noise
+Figure~\ref{fig:seg_noise} shows how transmission noise
 depends on midparent value.
-In the Gaussian case, segregation noise is Gaussian and independent of midparent value,
-as shown by the flat lines at 1 in Figure~\ref{fig:seg_noise}C. 
+In the Gaussian case, transmission noise is Gaussian and independent of midparent value,
+as shown by the horizontal lines at 1.0 in Figure~\ref{fig:seg_noise}C. 
 On the other hand, in the Cauchy case we can see from either Figure~\ref{fig:seg_noise}D or F
-that segregation noise depends strongly on midparent trait,
+that transmission noise depends strongly on midparent trait,
 as we would expect in the presence of several large-effect alleles.
@@
 \comment{TODO: refer to where?}:
-(1) the effect size distribution was Gaussian with mean zero and standard deviation $\sigma_g$,
+(1) the effect size distribution was Gaussian with mean zero and standard deviation 1.0
 and $f(z) = e^{\alpha z^2}$; and
 (2) the effect size distribution was Cauchy, centered at zero and with width $\delta - \gamma$
-and $f(z) = (\delta/\gamma) (\gamma + z^2) / (\delta + z^2)$.
+and $f(z) = (\delta/\gamma) (\gamma + z^2) / (\delta + z^2)$ with $\delta=0.5$ and $\gamma=0.2$.
 Figures~\ref{fig:sel_trait_distrns} and~\ref{fig:sel_seg_noise}
 show that the picture is largely unchanged by stabilizing selection
-except that, unsurprisingly, the median trait is no longer constrained to wander
+except that, unsurprisingly, the median trait no longer wanders unconstrained
 (but still shows substantially larger fluctuations under the Cauchy model).
@@
 
-However, as suggested above,
+However,
 the two models look much more similar when considering only alleles of small effect.
 Figure~\ref{fig:seg_noise_small} shows that ``midparent value''
-and ``segregation noise'' computed using only mutations with effect size less than 0.2
+and ``transmission noise'' computed using only mutations with effect size less than 0.2
 (excluding less than 1\% of mutations)
@@
 
-Not assuming all effects are the same order of magnitude seems very bioloically realistic,
-and so it'd be nice to figure out how evolution works and how to do quant gen in such cases.
-Such models can always be simulated,
-but the way forwards with theoretical understanding seems murky.
-It would be interesting to see how changing the distribution of segregation noise
-in the infinitesimal model
-affects properties like the rate of adaptive evolution,
+A substantial challenge for the field is to develop tractable models of trait evolution
+that include contributions of both ``large'' and ``small'' effect loci.
+In practice, breeders already embellish the ``animal model'' (essentially, the infinitesimal model)
+to predict traits
+by combining both effects of known large effect loci
+and calculations using a relatedness matrix.  \comment{cite}
+We can study any model with simulation,
+but the way forwards to theoretical understanding seems murky.
+
+Further analysis of the infinitesimal model with other noise distributions,
+as we have begun here,
+could be interesting,
+e.g., to see how changing the distribution of transmission noise
+affects the rate of adaptive evolution,
 levels of genetic load,
-and the rate of fixation.
+the rate of fixation,
+and the strength of linked selection.
 However,
-we have seen that this assumption imples a sort of ``blending inheritance''
+we have seen that the assumption of the infinitesimal model
+implies a sort of ``blending inheritance''
 for the large effect alleles,
-and so such work much be accompanied by simulations
+and so such work must be accompanied by simulations
 to see if the predictions are borne out under concrete models of genetic inheritance.
@@
 
-Above we've mostly used the Cauchy (e.g., in simulations)
+Finally: above we've most often used the Cauchy distribution
 not because it seems most realistic
-(indeed data suggests $\alpha=3/2$ might be better)
+(indeed data suggests the L\'evy stable distribution with $\alpha=3/2$ might be better)
 but for convenience and to make effects more obvious.
+
+\subsection*{Acknowledgements}
+Thanks go to Nate Pope for useful discussion
+and to Gregor Gorjanc for references to the literature.
 

### 2023-07-05 sims/do_figs.sh
code for figs

@@ -1,4 +1,6 @@
 #!/bin/bash
+# OLD VERSION: takes a loooong time to run the selected sims
+# see make_paper_figs.sh
 
 for type in normal cauchy
 do

### 2023-07-05 sims/do_figs_local.sh

@@ -1,18 +1,0 @@
-#!/bin/bash
-
-if [ $# -lt 1 ]
-then
-    echo "Usage: $0 seed"
-fi
-
-seed=$1
-
-for xopt in 0 1 2 10
-do
-    for type in normal cauchy
-    do
-        echo "starting at $(date)" >> slurm.log
-        bash run_sim.srun "-s ${seed}99${xopt} -d BURNIN=10000 -d NUMGENS=1000 -d N=1000 -d XOPT=${xopt}.0 ${type}.slim"
-        echo "ending: $(date)" >> slurm.log
-    done
-done

### 2023-07-05 sims/make_paper_figs.R

@@ -87,8 +87,8 @@ plot_seg_noise <- function (examples, labels, seg_col, main_append="", ...) {
 examples <- list(
                  # "neutral Normal" = "sim_neutral_normal_831",
                  # "neutral Cauchy" = "sim_neutral_cauchy_831"
-                 "neutral Normal" = "sim_neutral_normal_813566",
-                 "neutral Cauchy" = "sim_neutral_cauchy_81356610"
+                 "neutral Normal" = "sim_neutral_normal_8135",
+                 "neutral Cauchy" = "sim_neutral_cauchy_8135"
 )
 
 # 1. Plots of median trait value over time and final trait distribution

### 2023-07-05 sims/make_paper_figs.sh

@@ -2,8 +2,11 @@
 SLIM="sbatch run_sim.srun"
 # SLIM=slim
 
-$SLIM -s 831 -d BURNIN=20000 -d NUMGENS=1000 -d N=1000 -d 'EPSILON=c(0,0.2,1.0,10)' neutral_cauchy.slim
-$SLIM -s 831 -d BURNIN=20000 -d NUMGENS=1000 -d N=1000 -d 'EPSILON=c(0,0.2,1.0,10)' neutral_normal.slim
+$SLIM -s 831 -d BURNIN=20000 -d NUMGENS=1000 -d N=1000 -d 'EPSILON=c(0,0.2,1.0,4.0,10)' neutral_cauchy.slim
+$SLIM -s 831 -d BURNIN=20000 -d NUMGENS=1000 -d N=1000 -d 'EPSILON=c(0,0.2,1.0,4.0,10)' neutral_normal.slim
 
-$SLIM -s 707 -d BURNIN=20000 -d NUMGENS=1000 -d N=1000 -d STEP=20 -d 'EPSILON=c(0,0.2,1.0,10)' cauchy.slim
-$SLIM -s 707 -d BURNIN=20000 -d NUMGENS=1000 -d N=1000 -d STEP=20 -d 'EPSILON=c(0,0.2,1.0,10)' normal.slim
+$SLIM -s 707 -d BURNIN=20000 -d NUMGENS=1000 -d N=1000 -d STEP=20 -d 'EPSILON=c(0,0.2,1.0,4.0,10)' cauchy.slim
+$SLIM -s 707 -d BURNIN=20000 -d NUMGENS=1000 -d N=1000 -d STEP=20 -d 'EPSILON=c(0,0.2,1.0,4.0,10)' normal.slim
+
+$SLIM -s 8135 -d BURNIN=40000 -d NUMGENS=1000 -d N=10000 -d 'EPSILON=c(0,0.2,1.0,4.0,10)' neutral_normal.slim
+$SLIM -s 8135 -d BURNIN=40000 -d NUMGENS=1000 -d N=10000 -d 'EPSILON=c(0,0.2,1.0,4.0,10)' neutral_cauchy.slim

### 2023-07-05 writeup.tex
stable sim

@@
 that include contributions of both ``large'' and ``small'' effect loci.
-In practice, breeders already embellish the ``animal model'' (essentially, the infinitesimal model)
+In practice, breeders have long embellished the ``animal model'' (essentially, the infinitesimal model)
 to predict traits
 by combining both effects of known large effect loci
-and calculations using a relatedness matrix.  \comment{cite}
+and calculations using a relatedness matrix
+\citep[e.g.,][]{fernando1989marker,teissier2018weighted,bernardo2014genomewide,rice2019evaluation}.
 We can study any model with simulation,
@@
 Thanks go to Nate Pope for useful discussion
-and to Gregor Gorjanc for references to the literature.
+and to Gregor Gorjanc for insight into modern uses
+of the animal model and genomic selection.
 

### 2023-07-05 writeup.tex
citations

@@
 
-\comment{TODO: citations below}
-
 As suggested by its name,
@@
 including selection or publication bias,
-filters for larger-effect variants due to spatially varying selection (cite),
-and/or grouping together of concordant alleles by linkage or inversions (cite).
-However, it seems clear that in at least those cases where we understand the underlying molecular mechanisms,
-mutations of large effect can be important in practice
-(cite examples: insecticide resistance, coloration, etcetera).
+filters for larger-effect variants due to spatially varying selection
+\citep{barton1987establishment,yeaman2011establishment,yeaman2011genetic},
+and/or grouping together of concordant alleles by linkage or inversions
+\citep{kirkpatrick2006chromosome}.
+However,
+it seems clear that in at least those cases where we understand the underlying molecular mechanisms,
+mutations of large effect can be important in practice.
+% \citep[e.g.,][]{cyclodiene-resistance,hoekstra2006genetics}.
 The consensus in the field seems to be that traits exist on a spectrum between monogenic and polygenic,
-and that many - if not most - are firmly in the intermediate, ``oligogenic'' region (cite),
+and that many - if not most - lie in the intermediate region,
 where heritable trait variation is due both to segregating large- and small-effect variants.
@@
 by which changes to more ``nearby'' genes have the possibility of causing larger changes in the trait
-(cite omnigenic and others).
+\citep{kopp2012transcriptional,boyle2017expanded}.
 This suggests a mixture model for the effect size distribution:
@@
 The Gaussian distribution appears thanks to ``the'' Central Limit Theorem,
-i.e., the universality
-of adding up a great many very small and independent things.
+i.e., the universality of adding up a great many very small and independent things.
 However, there is a wider class of Central Limit Theorems,
 which describe the distributions of sums of a great many independent and exchangeable things:
-the \emph{stable laws}.
+the \emph{stable laws} \citep{gnedenko1968limit}
 Roughly speaking, these say that if we add together $n$ independent copies of a random variable $X$
@@
 For instance, it is a long-standing question how often the motions of organisms
-are better modeled by essentially Brownian random walks or by L\'evy processes (cite),
+are better modeled by essentially Brownian random walks or by L\'evy processes \citep{benhamou2007animals}
 and a growing literature studies the effects of such power law distributions
-on the shapes and dynamics of traveling waves such as biological invasions (cite).
-\comment{Other examples?}
+on spatial genetic and dynamics
+\citep[e.g.,][]{paulose2019spatial,smith2023isolation}.
 More directly analogous to the problem we study here, \citet{landis2012phylogenetic}
@@
 The question of when the trait distribution in an evolving population
-might be modeled as a Gaussian has been discussed extensively, e.g., by \comment{cite}.
+might be modeled as a Gaussian has been discussed extensively
+\citep[e.g.,][]{turelli1994genetic,dbarre2015evolution}.
 
@@
 and interactions between alleles, between loci, and with the environment are often important in practice.
-In fact, many traits claimed to be genetically determined
-in fact only appear heritable through non-genetic means,
-e.g., systematic inequalities in resource distribution partitioned by class or race. \comment{e.g., cite}
+% In fact, many traits claimed to be genetically determined
+% in fact only appear heritable through non-genetic means,
+% e.g., systematic inequalities in resource distribution partitioned by class or race
+% \citep[e.g.,][]{davenport1916hereditary}.
 In this paper, we follow the literature in working with the relatively simple purely additive case

### 2023-07-05 sims/gen_stable.R

@@ -1,0 +1,26 @@
+# From Devroye:
+# 1. If U ~ Unif(0,pi) and E ~ Exp(1) then
+#    ( ( sin(alpha*U) / sin(U) )^{1/(1-alpha)} * sin((1-alpha)*U) / (E * sin(alpha*U) ) )^{(1-alpha)/alpha}
+#  = ( sin(alpha*U) / sin(U) )^(1/alpha) * ( sin((1-alpha)*U) / (E * sin(alpha*U) ) )^((1-alpha)/alpha)
+# is Stable(alpha, 1)
+# 2. If S ~ Stable(alpha/2, 1) and N ~ N(0,1) then
+#   N * sqrt(2*S) ~ Stable(alpha, 0)
+
+rstable <- function(n, alpha) {
+    U <- runif(n, 0, pi)
+    E <- rexp(n)
+    N <- rnorm(n)
+    a <- alpha/2
+    S <- ( sin(a*U) / sin(U) )^(1/a) * ( sin((1-a)*U) / (E * sin(a*U) ) )^((1-a)/a)
+    return(N * sqrt(2*S))
+}
+
+rstable2 <- function (n, alpha) {
+    a <- alpha/2
+    U <- runif(n, 0, pi)
+    return(
+        rnorm(n) * sqrt(2 * 
+            ( sin(a*U) / sin(U) )^(1/a) * ( sin((1-a)*U) / (rexp(n) * sin(a*U) ) )^((1-a)/a)
+        )
+    )
+}

### 2023-07-06 writeup.tex
.

@@
 $$
+\comment{TODO: Gregor suggests ``mendelian sampling term''}
 What we will call the \emph{transmission noise}, $R_M$,
@@
 %%%%%%%%%%%%%%%%%%%%%%
-\section{Extending the infinitesimal model}  % TODD
-
-\comment{Proposal: change $1 + \mu$ to $f$ (or, something else, but a single letter).}
+\section{Abnormal quantitative genetics?}
+
+\comment{
+    Proposal: describe model, pointing out it's the infinitesimal model with gaussian noise;
+    say something about the Cauchy case (maintains diversity with blending inheritance?);
+    and jump to equation (1);
+    then present section 5.4 which we use to parameterize simulation.
+}
+
+\comment{Proposal: change $1 + \mu$ to $\mu$ (or, something else, but a single letter).}
 

### 2023-07-07 sims/make_paper_figs.R
stable(3/2) figs

@@ -37,7 +37,7 @@ load_seg_midp <- function (base, max_n=as.integer(1e6)) {
 }
 
 plot_trait_traces <- function (examples, labels) {
-    layout(matrix(1:4, nrow=2), heights=c(1.5,1))
+    layout(matrix(1:(2*length(examples)), nrow=2), heights=c(1.5,1))
     for (x in names(examples)) {
         fix <- load_fix(examples[[x]])
         pop <- load_pop(examples[[x]])
@@ -55,7 +55,7 @@ plot_trait_traces <- function (examples, labels) {
 }
 
 plot_seg_noise <- function (examples, labels, seg_col, main_append="", ...) {
-    layout(matrix(1:6, nrow=3), heights=c(1.5,1, 1))
+    layout(matrix(1:(3*length(examples)), nrow=3), heights=c(1.5,1, 1))
     for (x in names(examples)) {
         params <- load_params(examples[[x]])
         sm <- load_seg_midp(examples[[x]])
@@ -162,7 +162,33 @@ fname <- "selected_seg_noise_small.png"
 labels <- list(c("(A)", "(B)", "(C)"), c("(D)", "(E)", "(F)"))
 names(labels) <- names(examples)
 
-png(file=fname, width=6.5, height=6, pointsize=10, units='in', res=288)
+png(file=fname, width=8.5, height=6, pointsize=10, units='in', res=288)
 plot_seg_noise(examples, labels, seg_col=1)
+dev.off()
+
+
+# 5. with stable(3/2)
+examples <- list(
+                 "neutral Normal" = "sim_neutral_normal_8135",
+                 "neutral Stable(3/2)" = "sim_neutral_stable_8135",
+                 "neutral Cauchy" = "sim_neutral_cauchy_8135"
+)
+
+fname <- "neutral_trait_traces2.pdf"
+labels <- list(c("(A)", "(B)"), c("(C)", "(D)"), c("(E)", "(F)"))
+names(labels) <- names(examples)
+
+pdf(file=fname, width=6.5, height=3, pointsize=10)
+plot_trait_traces(examples, labels)
+dev.off()
+
+# 6. Segregation noise:
+
+fname <- "neutral_seg_noise2.png"
+labels <- list(c("(A)", "(B)", "(C)"), c("(D)", "(E)", "(F)"), c("(G)", "(H)", "(I)"))
+names(labels) <- names(examples)
+
[…]

### 2023-07-07 writeup.tex
.
Clean up.

@@
 \newcommand{\comment}[1]{{\color{blue} \it #1}}
+
+\DeclareMathOperator*{\Log}{Log}
+\DeclareMathOperator*{\sgn}{sgn}
 
@@
 and interactions between alleles, between loci, and with the environment are often important in practice.
-% In fact, many traits claimed to be genetically determined
-% in fact only appear heritable through non-genetic means,
-% e.g., systematic inequalities in resource distribution partitioned by class or race
+In fact, many traits claimed to be genetically determined
+in fact only appear heritable through non-genetic means,
+e.g., systematic inequalities in resource distribution partitioned by class or race.
 % \citep[e.g.,][]{davenport1916hereditary}.
@@
 
-\comment{Proposal: change $1 + \mu$ to $\mu$ (or, something else, but a single letter).}
+As observed in \cite{turelli2017commentary}, there are at least three interpretations of what is called ``the infinitesimal model''.
+The maximalist of these, what \cite{turelli2017commentary} calls the Gaussian-population approximation, assumes a large number of small-effect loci,  parental trait-independent Gaussian Mendelian sampling terms  (as rigorously justified in \cite{barton2017infinitesimal,barton2022infinitesimal}), and, moreover, that  population trait values are in deterministic mutation-selection balance, with a Gaussian empirical distribution (\ie the \emph{fraction} of the population with trait value in $[z,z+dz)$ is approximately $\frac{1}{2\pi\sigma^{2}} e^{-\frac{(z-m)^{2}}{2\sigma^{2}}}$ for suitable choices of the empirical mean and variance $m$ and $\sigma^{2}$.
+Leaving aside -- as the Gaussian-population approximation frequently does -- the question of whether or not some combination of effect sizes, dominance, and epistasis can justify non-Gaussian, trait-independent,  Mendelian sampling terms, we might reasonably ask when (or if) deterministic mutation selection-balance is possible in an non-Gaussian model.
 
-\comment{Copied from \texttt{eugene\_notes.tex}:}
+Concretely, consider a  Moran process where an individual whose trait value is $Z$ dies at rate $\mu(Z)$ and is replaced by a new individual of phenotype
+\begin{align*}
+  \frac{Z^{1} +Z^{2}}{2} + R ,
+\end{align*}
+where $Z^{1}$ and $Z^{2}$ are independent draws from the empirical population distribution and $R$ is an independent draw from the transmission noise distribution (independent of $Z^{1}$ and $Z^{2}$).
+Mutation selection balance demands that \comment{I'm struggling a bit to find a clear, short, and simple argument to jump directly to this point} that $\frac{Z^{1} +Z^{2}}{2} + R$ be equal in distribution to the selection-biased choice of $Z$, \ie for all bounded and continuous functions $f$,
+\begin{equation} \label{eqn:MSB_condition}
+    \E\left[ f\left(\frac{Z^{1} +Z^{2}}{2} + R \right) \right] = \frac{\E\left[ f(Z) \mu(Z) \right]}{\E[\mu(Z)]},
+\end{equation}
+where the expectation is with respect to the joint law of the population empirical distribution (the law of $Z$, $Z^{1}$ and $Z^{2}$), and $R$.
 
-Let $X$ be a population of size $N$, recorded as the empirical distribution on phenotypes
+Since the set of functions $\mathcal{S} = \{e^{i \theta x} : \theta \in \R\}$ is separating (see \eg \S 3.4 in \cite{Ethier+Kurtz1986}), it suffices to show that \eqref{eqn:MSB_condition} holds for $f \in \mathcal{S}$, \ie
+\begin{align} \label{eqn:MSB_CF}
+    \E\left[ e^{i\theta\left(\frac{Z^{1} +Z^{2}}{2} + R \right)}\right] = \frac{\E\left[ e^{i \theta Z}\mu(Z) \right]}{\E[\mu(Z)]}.
+\end{align}
+Now, we assumed that $Z^{1}$, $Z^{2}$, and $R$ are independent, so we can write \eqref{eqn:MSB_CF} as 
 \begin{align*}
-    X = \frac{1}{N} \sum_{i=1}^N \delta_{x_i} .
+	\frac{\E\left[ f(Z) \mu(Z) \right]}{\E[\mu(Z)]} &= \E\left[ e^{i\theta\left(\frac{Z^{1} +Z^{2}}{2} + R \right)}\right]\\
+	&= \E\left[ e^{i\frac{theta}{2} Z^{1}}e^{i\frac{theta}{2} Z^{2}}e^{\i\theta R}\right]\\
+	&= \E\left[e^{i\frac{\theta}{2} Z^{1}}\right]\E\left[e^{i\frac{\theta}{2} Z^{2}}\right]\E\left[e^{\i\theta R}\right]\\
+	&= \phi\left(\frac{\theta}{2}\right)	^{2} \Phi(\theta),
 \end{align*}
-So, $X$ is a point measure of total mass 1 on $\R$.
-This evolves as a Moran process under the infinitesimal model
-with death rates depending on state $x$.
-Concretely, an individual whose phenotype is $x$
-dies at rate $1 + \mu(x)$
-and is replaced by a new individual of phenotype
-\begin{align*}
-  \frac{y + z}{2} + \xi ,
-\end{align*}
-where $y$ and $z$ are independent draws from $X$
-and $\xi$ is an independent draw from the transmission noise distribution.
-Let $p_\xi$ denote the distribution of $\xi$,
-and $\nu(X)$ denote the distribution of $\frac{y + z}{2} + \xi$.
-This has generator
-\begin{align*}
-    \mL \Phi(X)
-    &=
-    \int \int \int \int
-    \left( \Phi\left( X + \frac{1}{N} \delta_{(y+z)/2 + \xi} - \frac{1}{N} \delta_x \right) - \Phi(X) \right)
-    (1 + \mu(x)) X(dx) X(dy) X(dz) p_\xi(d\xi) \\
-    &=
-    \int \int
-    \left( \Phi\left(X + \frac{1}{N} \delta_u - \frac{1}{N} \delta_x\right) - \Phi(X) \right)
-    (1 + \mu(x)) X(dx) \nu(X)(du) .
-\end{align*}
+where
+\begin{equation}\label{eq:phi}
+	\phi(\theta) = \mathbb{E}\left[e^{i\theta Z^{1}}\right] = \mathbb{E}\left[e^{i\theta Z^{2}}\right]
+\end{equation}
+and
+\begin{equation}\label{eq:Phi}
+	\phi(\theta) = \mathbb{E}\left[e^{i\theta R}\right]
+\end{equation}
+are the characteristic functions of $Z^{j}$ and $R$, respectively.  In what follows, we will explore some non-Gaussian solutions $\phi(\theta)$ to \eqref{eqn:MSB_CF}.
 
-For a test function $\phi$, setting $\Phi(X) = \ip{\phi}{X} := \int \phi(x) X(dx)$
-we get the first moment
-\begin{align*}
-    \mL \ip{\phi}{X}
-    &=
-    \frac{1}{N} \int \int \int \int
-    \left( \phi \left( \frac{y+z}{2} + \xi \right)
-         - \phi(x) \right)
-    (1 + \mu(x)) X(dx) X(dy) X(dz) p_\xi(d\xi) \\
-    &=
-    \frac{1}{N} \ip{\phi}{\nu(X)}\ip{1+\mu}{X}
-    - \frac{1}{N} \ip{\phi}{(1+\mu) X} .
-\end{align*}
-Note that for $Y$ another finite measure on $\R$,
-the Fr\'echet derivative \comment{(is that the right name?)} of $X$ in the direction of $Y$
-satisfies $\ip{\phi}{\partial_Y X} = \lim_{\epsilon \to 0} \epsilon^{-1} (\ip{\phi}{X + \epsilon Y} - \ip{\phi}{X}) = \ip{\phi}{Y}$,
-so that this is a first derivative in the direction of the measure
-$$ \ip{1+\mu}{X} \nu(X) - (1+\mu) X . $$
-
-More generally, if $\Phi(X) = \prod_i \ip{\phi_i}{X}$ then
-\begin{align*}
-    \mL \Phi(X)
-    &=
-    \frac{1}{N} \sum_i \left( 
-        \ip{1+\mu}{X} \ip{\phi_i}{\nu(X)}  - \ip{\phi_i}{(1+\mu)X}
-    \right)
-    \prod_{j \neq i} \ip{\phi_j}{X} 
-    + O(1/N^2) .
-\end{align*}
-This implies that after rescaling time by $N$
-and suitable other caveats,
-the limiting process is a deterministic flow in the direction of 
-$\nu(X) \ip{1+\mu}{X} - (1+\mu)X$.
-
-\paragraph{Stable measures}
-If $Z$ is a fixed point of this flow,
-then $\nu(Z)\ip{1+\mu}{Z} = (1+\mu)Z$.
-In other words, any measure $Z$ that is a fixed point has the property that
-(the average of two independent draws plus transmission noise)
-is equal in distribution to
-(a $1+\mu$-weighted draw).
-Concretely, with $A$, $B$, and $C$ independent draws from $Z$,
-\begin{align} \label{eqn:ABC_condition}
-    \E\left[ f\left( \frac{A+B}{2} + \xi \right) \right]
-    &=
-    \frac{\E\left[ f(C) (1+\mu(C)) \right]}{\E[1+\mu(C)]} .
-\end{align}
-
-\comment{Copied from \texttt{MeanFlow.tex}:}
-
-Assume that $Z$ is a finite measure, and without loss of generality normalized to have total mass one.
-
-Throughout, I'll be using the characteristic function of $Z$, which I'll denote by 
-\[
-	\phi(\theta) = \mathbb{E}\left[e^{i\theta Z}\right],
-\]
-and the characteristic function of $\xi$, $\Phi(\theta)$.
-
-\section{No Selection}
+\subsection{The Neutral Case}
 
 First, consider the case without selection, so $\mu$ is a constant.  Then, our condition on the characteristic function simplifies to
-\begin{equation}\label{CHARACTERISTIC}
+\begin{equation}\label{eq:CF_neutral}
 	{\textstyle \phi\left(\frac{\theta}{2}\right)^{2}} \Phi(\theta) %e^{-\frac{\eta^{2}\theta^{2}}{2}} 
@@
  \[
-	F(\phi)(\theta) = {\textstyle \phi\left(\frac{\theta}{2}\right)^{2}} \Phi(\theta).
+	F(\phi)(\theta) = {\textstyle \phi\left(\frac{\theta}{2}\right)^{2}} \Phi(\theta),
 \]
-In particular, we can characterize all possible fixed points by considering the iterates of an arbitrary characteristic function, $\varphi(\theta)$, for  any distribution on $\mathbb{R}$.
-
+whereas any fixed point must be the limit of an arbitrary characteristic function, $\varphi(\theta)$, under repeated applications of $F$:
+\[
+	\phi(\theta) = \lim_{n \to \infty} F^{(n)}(\varphi)(\theta).
+\]
 Iterating $F$ $n$ times, we get
 \[
-	F^{(n)}(\varphi)(\theta)
-	=  {\textstyle \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}}\prod_{k=1}^{n-1} {\textstyle \Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}},
+	F^{(n)}(\varphi)(\theta) =  {\textstyle \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}}\prod_{k=1}^{n-1} {\textstyle \Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}},
 \]
@@
 Since $\varphi$ and $\Phi$ are characteristic functions, they are absolutely continuous and $\phi(0) = 1$, and thus non-zero in a neighbourhood of 0.
-In particular, $\psi(\theta) = \ln{\varphi}(\theta)$ and $\Psi(\theta) = \ln{\Phi(\theta)}$ exist in some neighbourhood of 0 as well.
+In particular, $\psi(\theta) = \Log{\varphi}(\theta)$ and $\Psi(\theta) = \Log{\Phi(\theta)}$ exist in some neighbourhood of 0 as well\footnote{We use $\Log$ to denote the principal value of the complex logarithm, for which the imaginary part always lies in the interval $(-\pi,\pi]$}.
 From the above, we have
 \[
-	\ln{F^{(n)}(\varphi)(\theta)} = {\textstyle 2^{n} \psi\left(\frac{\theta}{2^{n}}\right)} 
+	\Log{F^{(n)}(\varphi)(\theta)} = {\textstyle 2^{n} \psi\left(\frac{\theta}{2^{n}}\right)} 
 		+ \sum_{k=1}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}.
 \]
+We will briefly explore what limiting values are possible -- and which limits are valid characteristic functions.
+We refer to the first term on the left hand side as the \emph{reproduction term} and the remaining sum as the \emph{noise term}.
+For the sake of simplicity, we will limit ourselves to the caae when both have well-posed limits independent of the other.
 
-\subsection{Noise Terms}
+\subsubsection{Reproductive Terms}
 
-We now analyse the component arising from the noise $\xi$.  First, we consider the case when $\Psi(\theta)$ is twice continuously differentiable, so $\xi$ has finite mean $m = \Psi'(0)$
-and variance $\eta^{2} = \Psi''(0)$.   Taylor's theorem then tells us that for $\theta$ sufficiently small, 
-\[
-	\Psi(\theta) = m\theta + \frac{\eta^{2}}{2} \theta^{2} + r(\theta)\theta^{2},
-\]
-where $r(\theta) \to 0$ as $\theta \to 0$.  Then,
-\begin{align*}
-	 \sum_{k=1}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}
-	 &=  \sum_{k=1}^{n-1} 2^{k} {\textstyle \left(m \frac{\theta}{2^{k}} - \frac{\eta^{2}}{2}  \frac{\theta^{2}}{2^{2k}}
-		+ r\left(\frac{\theta}{2^{k}}\right)\frac{\theta^{2}}{2^{2k}}\right)}\\
-	&= (n-1) m - \eta^{2} \theta^{2} (1 - 2^{-n+1}) 
-		+ \sum_{k=1}^{n-1} {\textstyle r\left(\frac{\theta}{2^{k}}\right)\frac{\theta^{2}}{2^{k}}}
- \end{align*}
-
-Immediately, we see that for the limit to exist, we must have $m = 0$; if we assume that $r(\theta) \equiv 0$, so that the noise is $N(0,\eta^{2})$, then the limiting distribution is $N(0,2\eta^{2})$.
-We note that when  $r(\theta)$ is non-zero, we don't expect the higher order terms to vanish.
-
-Now, consider the case when $\Psi(\theta)$ corresponds to a stable law, so
-\[
-	\Psi(\theta) = im\theta + c|\theta|^{\alpha} (1-i\beta \text{sgn}(\theta)\omega_{\alpha}(\theta))
-\]
-for $\alpha \in (0,2]$, $\beta \in [-1,1]$, $c \geq 0$, and $m \in \mathbb{R}$ and
-\[
-	\omega_{\alpha}(\theta) = \begin{cases}
-		\tan{\frac{\pi\alpha}{2}} & \text{if $\alpha \neq 1$, and}\\
-		-\frac{2}{\pi}\ln{|\theta|} & \text{if $\alpha = 1$.}
-	\end{cases}
-\]
-We then have
-\begin{multline*}
-	 \sum_{k=1}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}
-	 =  \sum_{k=1}^{n-1} im\theta + 2^{(1-\alpha)k} c|\theta|^{\alpha} {\textstyle\left(1-i \beta\text{sgn}(\theta)\left(\frac{\theta}{2^{k}}\right)
-	 	 \omega_{\alpha}\left(\frac{\theta}{2^{k}}\right)\right)}\\
-	= 
-	\begin{cases}
-		 im(n-1)\theta + c|\theta|^{\alpha}\frac{1-2^{(1-\alpha)n}}{1-2^{1-\alpha}}\left(1-i\beta \text{sgn}(\theta) \tan{\frac{\pi\alpha}{2}}\right) & \text{if $\alpha \neq 1$, and}\\	
-		 \begin{multlined}
-		 im(n-1)\theta + c|\theta|^{\alpha}  \left(\frac{1-2^{(1-\alpha)n}}{1-2^{1-\alpha}}
-		 \left(1+i \beta \text{sgn}(\theta)\frac{2}{\pi}\ln{|\theta|} \right)\right.\\
-		\left. -i\beta \text{sgn}(\theta) \frac{2^{1-\alpha} + (n(2^{1-\alpha}-1)-1)2^{(1-\alpha)n})}{(1-2^{1-\alpha})^{2}}\frac{2}{\pi}\ln{2}
-		\right)    
-		\end{multlined}
-		& \text{if $\alpha = 1$.}
-	\end{cases}
-\end{multline*}
-This converges as $n \to \infty$ if and only if $m = 0$ and $\alpha \ge 1$, in which case the limit is 
-\[
-	 \frac{c}{1-2^{1-\alpha}}|\theta|^{\alpha}\left(1-i\beta \text{sgn}(\theta) \tan{\frac{\pi\alpha}{2}}\right), 
-\]
-which corresponds to a $\left(\alpha,\beta, \frac{c}{1-2^{1-\alpha}},0\right)$-stable law.
-
-\subsection{Reproductive Terms}
-
-We now consider the terms $2^{n} \psi\left(\frac{\theta}{2^{n}}\right)$, which we call the reproductive terms.  Proceeding as below, we consider the limit
-%We note that this does not exclude the possibility of distributions without a mean; % \eg if $\eta =0$, then \eqref{CHARACTERISTIC} is satisfied by the family of Cauchy distributions with characteristic functions $\phi(\theta) = e^{i m \theta - \gamma |\theta|}$; one can check that these are the only stable laws satisfying \eqref{CHARACTERISTIC}.%  One can similarly exclude the stable laws without mean ($\alpha \leq 1$) by inspection in the case of $\eta > 0$.
-%Now, to exhaust all possibilities, observe that by iterating, we we also have that 
-%\[
-%	\psi(\theta) = 2^{n}\psi\left(\frac{\theta}{2^{n}}\right) 
-%\]
-%\ie we must have 
+First consider the limit
 \[
@@
 \]
-Because $\psi(\theta)$ is the logarithm characteristic function, it must be uniformly continuous and vanish at 0, and thus $\alpha(\theta)$ as well.  We also have that for all integers $k$, 
+Because $\psi(\theta)$ is the logarithm characteristic function, it must be continuous in some neighbourhood of, and vanish at 0, and thus $\alpha(\theta)$ as well.
+Furthermore, for all integers $k$,
 \[
@@
 \]
-as $k \to - \infty$ if the latter exists (\ie there is a unique left-hand limit at zero).  In this case, setting $-\gamma + i m = \alpha'(0+)$, we have
+as $k \to - \infty$ if the latter exists (\ie there is a unique left-hand limit at zero; this is the case if \eg we require $\alpha(\theta)$, and thus $\phi(\theta)$, to be convex functions).
+In this case, setting $-\gamma + i m = \alpha'(0+)$, we have
 \[
@@
 \]
-\ie if $\gamma > 0$, then $e^{\alpha(\theta)} = e^{im \theta -\gamma|\theta|}$ is the characteristic function of a Cauchy distribution and $\phi(\theta)$ is the characteristic function of the sum of Cauchy and Gaussian random variables, whereas if $\gamma = 0$, then $e^{\alpha(\theta)} = e^{im \theta}$ is the Fourier transform of a Dirac mass at $m$, so the limiting random variable is the Gaussian $N(m,\eta^{2})$.
+Now, if $\phi(\theta)$ is the characteristic function of a random variable $Z$, Jensen's inequality tells us that for all $\theta$,
+\[
+	|\phi(\theta)| = \left|\E\left[e^{i\theta Z}\right]\right| \leq \E\left[\left|e^{i\theta Z}\right|\right] = 1
+\]
+so $\gamma \geq 0$.
+If $\gamma > 0$, then we recognize $e^{\alpha(\theta)} = e^{im \theta -\gamma|\theta|}$ as the characteristic function of a Cauchy distribution with median 0 and mean absolute deviation $\gamma$, whereas if $\gamma = 0$, then $e^{\alpha(\theta)} = e^{im \theta}$ is the Fourier transform of a Dirac mass at $m$, \ie the characteristic function for a monomorphic population with all individuals having trait value equal to $m$.
 
-Finally, we remark that if $\frac{\alpha(\theta)}{\theta}$ is non-constant for $\theta > 0$, then $\alpha'(0+)$ does not exist.
+\paragraph{Lamarkian inheritance redux.} 
+Note that the characteristic functions constructed above are solutions to \eqref{eq:CF_neutral} when transmission is noiseless, and the offspring have exactly the parental mean trait \ie Lamarkian blending inheritance.
+The solution $\phi(\theta) = e^{i\theta m}$ is exactly the reversion to the mean that Darwin's detractors used to argue against the theory of evolution \citep{Provine1971}.
+The solution $\phi(\theta) = e^{im \theta -\gamma|\theta|}$ offers an amusing counter-factual: if a Lamarkian population could somehow achieved a Cauchy-distributed trait distribution, it would maintain that diversity indefinitely.
 
-\subsection{Other Solutions?}
-
+\paragraph{Other solutions?}
 We would naturally like to know if these two exhaust the possible solutions.
-To this end, we recall Bochner's theorem, which tells us that a function $\phi(\theta)$ is the characteristic function of a random variable if and only if
+Above, we exhausted all solutions for which $\alpha'(0+)$ exists.
+However, if $\frac{\alpha(\theta)}{\theta}$ is non-constant for $\theta > 0$, then $\alpha'(0+)$ does not exist.
+This does not, however, in and of itself exclude $e^{\alpha(\theta)}$ from being a characteristic function: Bochner's theorem tells us that a function $\phi(\theta)$ is the characteristic function of a random variable if and only if
 \begin{enumerate}[(i)]
-\item $\phi(\theta)$ is uniformly continuous,
+\item $\phi(\theta)$ is continuous,
 \item $\phi(0) = 1$,
 \item $\phi(-\theta) = \overline{\phi(\theta)}$, and
-\item $\phi(\theta)$ is \textit{positive definite}: for all $n$ and all $\theta_{1},\ldots,\theta_{n} \in \mathbb{R}$, the matrix with entries
+\item $\phi(\theta)$ is \textit{positive definite}: for all $n$ and all $\theta_{1},\ldots,\theta_{n} \in \mathbb{R}$, the matrix with entries $a_{ij} = \phi(\theta_{i}-\theta_{j})$
+\end{enumerate}
+One can easily choose $\alpha(\theta)$ so that $e^{\alpha(\theta)}$ satisfies the first three criteria, \eg $\alpha(\theta) = -|\theta|\left(\sin\left(\pi\frac{\log{|\theta|}}{\log{2}}\right)+\frac{3}{2}\right)$, but excluding such functions via Bochner's positive definite criterion is non-trivial.
+
+\subsubsection{Noise Terms}
+
+We now turn our attention to the noise terms.
+A necessary condition for $\sum_{k=1}^{\infty} 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)$ to converge is that $2^{k} \Psi\left(\frac{\theta}{2^{k}}\right) \to 0$; since $\Psi(0) = 0$, this implies that if it exists, $\Psi'(0+) = 0$.
+In particular, as we would expect, if $R$ has a mean, it must be 0.
+If, on the other hand, $\Psi(\theta)$ is twice differentiable, so $R$ has mean 0 and finite variance, say $\Psi''(0) = \eta^{2}$.
+Taylor's theorem then tells us that for $\theta$ sufficiently small,
 \[
-	a_{ij} = \phi(\theta_{i}-\theta_{j})
+	\Psi(\theta) \frac{\eta^{2}}{2} \theta^{2} + r(\theta)\theta^{2},
 \]
-is positive definite.
-\end{enumerate}
-Given $\alpha(\theta)$ as above, the corresponding function $\phi(\theta) = e^{\alpha(\theta)-\eta^{2}\theta^{2}}$ 
-or $\phi(\theta) = e^{\alpha(\theta) - \frac{c}{1-2^{1-\alpha}}|\theta|^{\alpha}\left(1-i\beta \text{sgn}(\theta) \tan{\frac{\pi\alpha}{2}}\right)}$ satisfies the first three criteria.
-We will now seek to identify which $\alpha(\theta)$ result in positive definite $\phi(\theta)$.
-When $\alpha(\theta) = im \theta -\gamma|\theta|$ for $\gamma > 0$, this is immediate, as $e^{\alpha(\theta)}$ is then the characteristic function of a Cauchy random variable.
+where $r(\theta) \to 0$ as $\theta \to 0$.  Thus,
+\begin{equation*}
+	\frac{\textstyle 2^{k+1} \Psi\left(\frac{\theta}{2^{k+1}}\right)}{\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}
+	= \frac{1}{2} \frac{\eta^{2} \theta^{2} + {\textstyle  r\left(\frac{\theta}{2^{k+1}}\right)}\theta^{2}}{\eta^{2} \theta^{2} + {\textstyle  r\left(\frac{\theta}{2^{k}}\right)}\theta^{2}}
+	\to \frac{1}{2},
+\end{equation*}
+as $k \to \infty$ and the sum converges by the ratio test.  In this case,
+\begin{align*}
+	 \sum_{k=1}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}
+	 &=  \sum_{k=1}^{n-1} 2^{k} {\textstyle \left(- \frac{\eta^{2}}{2}  \frac{\theta^{2}}{2^{2k}}
+		+ r\left(\frac{\theta}{2^{k}}\right)\frac{\theta^{2}}{2^{2k}}\right)}\\
+	&=- \eta^{2} \theta^{2} (1 - 2^{-n+1})  + \sum_{k=1}^{n-1} {\textstyle r\left(\frac{\theta}{2^{k}}\right)\frac{\theta^{2}}{2^{k}}}
+\end{align*}
+If we assume that $r(\theta) \equiv 0$, so that the noise is $N(0,\eta^{2})$, then the limiting distribution is $N(0,2\eta^{2})$.
+Note that when  $r(\theta)$ is non-zero, we don't expect the higher order terms to vanish.
 
-We can also easily exclude the case $\gamma < 0$ using a simple consequence of positive definiteness in the case when $n=2$ (name criterion): if $\phi(\theta)$ is positive definite, then $|\phi(\theta)| \leq 1$.
-For us, this requires that
+It is unclear what more can be said in general.
+If however, we specialize to the case when $R$ is a stable law, we can explicitly compute the limiting noise term.
+We recall that if $\Psi(\theta)$ corresponds to a stable law(see \eg \cite{breiman1992probability}), then
 \[
-	\Re \alpha(\theta) \leq  \eta^{2} \theta^{2},\; resp. \; < \frac{c}{1-2^{1-\alpha}}|\theta|^{\alpha}
+	\Psi(\theta) = im\theta + c|\theta|^{\alpha} (1-i\beta \sgn(\theta)\omega_{\alpha}(\theta))
 \]
-which is violated for small values of $\theta$ when $\gamma < 0$ (recall $\alpha > 1$).
+for $\alpha \in (0,2]$, $\beta \in [-1,1]$, $c \geq 0$, and $m \in \mathbb{R}$ and
+\[
+	\omega_{\alpha}(\theta) = \begin{cases}
+		\tan{\frac{\pi\alpha}{2}} & \text{if $\alpha \neq 1$, and}\\
+		-\frac{2}{\pi}\log{|\theta|} & \text{if $\alpha = 1$.}
+	\end{cases}
+\]
+By our general remarks above, we can assume $m = 0$ and exclude the case $\alpha = 1$, $\beta = 0$, for which  $\Psi'(0+) = c$. We then have
+\begin{align*}
+	 \sum_{k=1}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}
+	 &=  \sum_{k=1}^{n-1} im\theta + 2^{(1-\alpha)k} c|\theta|^{\alpha} {\textstyle\left(1-i \beta\sgn(\theta)\left(\frac{\theta}{2^{k}}\right)
+	 	 \omega_{\alpha}\left(\frac{\theta}{2^{k}}\right)\right)}\\
+	&= \begin{cases}
+		 im(n-1)\theta + c|\theta|^{\alpha}\frac{1-2^{(1-\alpha)n}}{1-2^{1-\alpha}}\left(1-i\beta \sgn(\theta) \tan{\frac{\pi\alpha}{2}}\right) & \text{if $\alpha \neq 1$, and}\\	
+		 c|\theta|^{\alpha}  \left(1+\frac{2i\beta}{\pi}\sgn(\theta)\left(\log{|\theta|} - \frac{n(n-1)}{2} \log{2}\right)\right) & \text{if $\alpha = 1$,}
+	\end{cases}
+\end{align*}
+which converges as $n \to \infty$ if and only if $\alpha > 1$, in which case the limit is 
+\[
+	 \frac{c}{1-2^{1-\alpha}}|\theta|^{\alpha}\left(1-i\beta \sgn(\theta) \tan{\frac{\pi\alpha}{2}}\right), 
+\]
+which corresponds to a $\left(\alpha,\beta, \frac{c}{1-2^{1-\alpha}},0\right)$-stable law.
 
-\noindent\emph{Question:} how can we use positive definiteness to exclude $\alpha(\theta)$ with $\Re \alpha(\theta) \leq  \eta^{2} \theta^{2}$ so that $\alpha'(0+)$ does not exist?
-
-\begin{rem}
-Polya has shown that if $\phi(\theta)$ is convex and satisfies the first three criteria of Bochner's theorem, then $\phi(\theta)$ is a characteristic function.
-In our case, if $\alpha(\theta)$ is convex, then $e^{\alpha(\theta)}$ is as well; a proof-by-picture, however, shows that the only choice of $\alpha(\theta)$ that is convex (or concave, for that matter) for $\theta > 0$ is a linear function, so Polya's criterion doesn't give us any new characteristic functions.
-\end{rem}
+More generally, one will get convergence if the noise is a mixture of components with finite variance and stable laws as above, whereas the population distribution will be a mixture of a (possibly degenerate) Cauchy distribution and such noise terms -- we remark that in the presence of noise, a pure Cauchy distributed solution is \emph{not} possible in the neutral case.
 
@@
 
+Unfortunately, once selection is included, \eqref{eqn:MSB_CF} is less amenable to the sort of simple arguments above.
+We will content ourselves with presenting two examples in the case of stabilizing selection, one already familiar to enthusiasts of quantitative genetics, the other a Cauchy solution obtained by trial-and-error, which illustrates how selection can stabilize an otherwise unviable solution.
+
 \subsubsection{Gaussian Solution}
@@
 
-Now, consider the case when $1+ \mu(x) = e^{a x^{2}}$. 
-In this case there is a Gaussian solution: suppose $Z$ is Gaussian with mean $m$ and variance $\sigma^{2}$.  Then,
+First, consider the case when $1+ \mu(x) = e^{a x^{2}}$.   In this case there is a Gaussian solution: suppose $Z$ is Gaussian with mean $m$ and variance $\sigma^{2}$.  Then,
 \[
@@
 \]
-Moreover, provided $a < \frac{1}{2\sigma^{2}}$, we can explicitly compute $\frac{\mathbb{E}\left[e^{i \theta Z}(1+\mu(Z))\right]}{\mathbb{E}[1+\mu(Z)]}$  to conclude that
-if $\xi$ is Gaussian with mean zero and variance $\eta^2$,
-then equation~\eqref{eqn:ABC_condition} with $f(x)=e^{i\theta x}$ is
+Moreover, provided $a < \frac{1}{2\sigma^{2}}$, we can explicitly compute $\frac{\mathbb{E}\left[e^{i \theta Z}(1+\mu(Z))\right]}{\mathbb{E}[1+\mu(Z)]}$  using the Gaussian probability density function to conclude that
+if $R$ is Gaussian with mean zero and variance $\eta^2$, then equation~\eqref{eqn:MSB_CF} becomes
 \[
@@
 
+Note that we do not require very weak selection to guarantee a Gaussian steady state, only that $a < \frac{1}{2\sigma^{2}}$.
+However, this does not imply that the transient distribution is also Gaussian, as is sometimes assumed \comment{cite any number of Lande papers?}.
+The latter does indeed require very weak selection.
+
 \subsubsection{Cauchy Solution}
@@
 
-In the case of a Cauchy random variable, we again have an explicit expression for the probability distribution, which, proceeding as above, allows us to construct an explicit solution when $1+ \mu(x) = \frac{\delta^{2}}{\gamma^{2}} \frac{\gamma^{2} + x^{2}}{\delta^{2} + x^{2}}$ for $\delta > \gamma > 0$ and the noise $\xi$ is given by a $\text{Cauchy}(0,\delta-\gamma)$ random variable (\ie corresponding to a stable $(0,0,\delta-\gamma,0)$ law).
+Almost uniquely among the stable laws, the Cauchy distribution has a probability density function expressible in elementary functions (the other exceptions being the L\'evy and Gaussian distributions): for a Cauchy random variable with median (location) $x_{0}$ and mean absolute deviation (scale) $\gamma$, the density is $f(x;x_{0},\gamma) = \frac{1}{\pi\gamma\left[1+\left(\frac{x-x_{0}}{\gamma}\right)^{2}\right]}$.
+Armed with this density function, we can explicitly compute the right hand size of \eqref{eqn:MSB_CF}, assuming that
+\begin{enumerate}[(i)]
+\item the population is in mutation-selection balance corresponding to a Cauchy density with median 0 and scale $\gamma > 0$,
+\item selection acts as $\mu(x) = \frac{\delta^{2}}{\gamma^{2}} \frac{\gamma^{2} + x^{2}}{\delta^{2} + x^{2}}$ for some $\delta > \gamma > 0$, and
+\item the transmission noise $R$ is given by a $\text{Cauchy}(0,\delta-\gamma)$ random variable (\ie corresponding to a stable $(0,0,\delta-\gamma,0)$ law).
+\end{enumerate}
 
-If we now suppose that $Z$ is a $\text{Cauchy}(0,\gamma)$ random variable.  Then $Z$ has probability density function $\frac{\gamma}{\pi} \frac{1}{\gamma^{2} + x^{2}}$, whereas 
-\[
+Under these assumptions,  
+\begin{align*}
 	\frac{\mathbb{E}\left[e^{i \theta Z}(1+\mu(Z))\right]}{\mathbb{E}[1+\mu(Z)]} 
-	= e^{-\delta |\theta|} = e^{-2\gamma \left|\frac{\theta}{2}\right|-(\delta-\gamma)|\theta|}
-\]
-which we recognize as being equal to $\phi\left(\frac{\theta}{2}\right)^{2} \Phi(\theta)$,
-and so solves equation~\eqref{CHARACTERISTIC}.
+	&= e^{-\delta |\theta|}\\
+	&= e^{-2\gamma \left|\frac{\theta}{2}\right|-(\delta-\gamma)|\theta|},
+\end{align*} 
+which we recognize as being equal to $\phi\left(\frac{\theta}{2}\right)^{2} \Phi(\theta)$, where $\phi(\theta) = e^{i\gamma|\theta|}$ and $\Phi(\theta) = e^{i(\delta - \gamma)|\theta|}$ are the characteristic functions of the population density and the transmission noise, respectively, giving us a.non-Gaussian solution to \eqref{eqn:MSB_CF} with both noise and selection.
+
+As remarked above, we found this solution by inspection and trial-and-error; it is interesting to ask if any similar examples can be constructed for other stable laws.
+The lack of an elementary probability density function unfortunately rules out the na\"ive approach used here.
 
 %(here, see a saturating cost to being away from the optimal phenotype $x = 0$: the mortality is maximized at  $\frac{\beta^{2}}{\alpha^{2}}$.
-
-\bigskip
-
-\noindent\emph{Question:} other stable laws?
 

### 2023-07-07 writeup.tex
peter edits

@@
 $$
-\comment{TODO: Gregor suggests ``mendelian sampling term''}
+\comment{TODO: Gregor suggests ``Mendelian sampling term''}
 What we will call the \emph{transmission noise}, $R_M$,
@@
 
-\comment{
-    Proposal: describe model, pointing out it's the infinitesimal model with gaussian noise;
-    say something about the Cauchy case (maintains diversity with blending inheritance?);
-    and jump to equation (1);
-    then present section 5.4 which we use to parameterize simulation.
-}
+As observed in \cite{turelli2017commentary}, there are at least three interpretations of what is called ``the infinitesimal model''.
+The maximalist of these, what \cite{turelli2017commentary} calls the Gaussian-population approximation, assumes a large number of small-effect loci,  parental trait-independent Gaussian Mendelian sampling terms  (as rigorously justified in \cite{barton2017infinitesimal,barton2022infinitesimal}), and, moreover, that  population trait values are in deterministic mutation-selection balance, with a Gaussian empirical distribution
+(\ie the \emph{fraction} of the population with trait value in $[z,z+dz)$
+is approximately $\frac{1}{2\pi\sigma^{2}} e^{-(z-m)^{2}/(2\sigma^{2})} dz$
+for suitable choices of the empirical mean and variance $m$ and $\sigma^{2}$).
+Leaving aside
+% -- as the Gaussian-population approximation frequently does --
+the question of whether or not some combination of effect sizes, dominance, and epistasis can justify non-Gaussian, trait-independent,  Mendelian sampling terms, we might reasonably ask when (or if)
+deterministic mutation-selection balance is possible in a non-Gaussian model
+that retains independence of the Mendelian sampling terms.
+(Indeed, above we have put substantial restrictions on what models of inheritance
+might lead to such models.)
+We might also ask whether such a model
+might still reasonably approximate an additive model with non-Gaussian effects,
+and for the purpose of investigating this with simulations,
+some calculations are helpful.
 
-As observed in \cite{turelli2017commentary}, there are at least three interpretations of what is called ``the infinitesimal model''.
-The maximalist of these, what \cite{turelli2017commentary} calls the Gaussian-population approximation, assumes a large number of small-effect loci,  parental trait-independent Gaussian Mendelian sampling terms  (as rigorously justified in \cite{barton2017infinitesimal,barton2022infinitesimal}), and, moreover, that  population trait values are in deterministic mutation-selection balance, with a Gaussian empirical distribution (\ie the \emph{fraction} of the population with trait value in $[z,z+dz)$ is approximately $\frac{1}{2\pi\sigma^{2}} e^{-\frac{(z-m)^{2}}{2\sigma^{2}}}$ for suitable choices of the empirical mean and variance $m$ and $\sigma^{2}$.
-Leaving aside -- as the Gaussian-population approximation frequently does -- the question of whether or not some combination of effect sizes, dominance, and epistasis can justify non-Gaussian, trait-independent,  Mendelian sampling terms, we might reasonably ask when (or if) deterministic mutation selection-balance is possible in an non-Gaussian model.
-
-Concretely, consider a  Moran process where an individual whose trait value is $Z$ dies at rate $\mu(Z)$ and is replaced by a new individual of phenotype
+Concretely, consider a  Moran process where an individual whose trait value is $Z$ dies at rate $\mu(Z)$
+and is replaced by a the offspring of two randomly chosen individuals,
+i.e., a new individual of phenotype
 \begin{align*}
@@
 \end{align*}
-where $Z^{1}$ and $Z^{2}$ are independent draws from the empirical population distribution and $R$ is an independent draw from the transmission noise distribution (independent of $Z^{1}$ and $Z^{2}$).
-Mutation selection balance demands that \comment{I'm struggling a bit to find a clear, short, and simple argument to jump directly to this point} that $\frac{Z^{1} +Z^{2}}{2} + R$ be equal in distribution to the selection-biased choice of $Z$, \ie for all bounded and continuous functions $f$,
-\begin{equation} \label{eqn:MSB_condition}
-    \E\left[ f\left(\frac{Z^{1} +Z^{2}}{2} + R \right) \right] = \frac{\E\left[ f(Z) \mu(Z) \right]}{\E[\mu(Z)]},
-\end{equation}
-where the expectation is with respect to the joint law of the population empirical distribution (the law of $Z$, $Z^{1}$ and $Z^{2}$), and $R$.
-
-Since the set of functions $\mathcal{S} = \{e^{i \theta x} : \theta \in \R\}$ is separating (see \eg \S 3.4 in \cite{Ethier+Kurtz1986}), it suffices to show that \eqref{eqn:MSB_condition} holds for $f \in \mathcal{S}$, \ie
+where $Z^{1}$ and $Z^{2}$ are independent draws from the empirical population distribution
+and $R$ is an independent draw from the Mendelian sampling distribution
+(independent of $Z^{1}$ and $Z^{2}$). 
+Mutation-selection balance demands that that
+the distribution of individuals removed by death
+and the distribution of individuals added by birth are equal,
+\ie that $\frac{Z^{1} +Z^{2}}{2} + R$ is equal in distribution
+to a $\mu$-weighted choice of $Z$.
+% In other words, for all bounded and continuous functions $f$,
+% \begin{equation} \label{eqn:MSB_condition}
+%     \E\left[ f\left(\frac{Z^{1} +Z^{2}}{2} + R \right) \right] = \frac{\E\left[ f(Z) \mu(Z) \right]}{\E[\mu(Z)]},
+% \end{equation}
+% where the expectation is with respect to the joint law of the population empirical distribution (the law of $Z$, $Z^{1}$ and $Z^{2}$), and $R$.
+% 
+% Since the set of functions $\mathcal{S} = \{e^{i \theta x} : \theta \in \R\}$ is separating (see \eg \S 3.4 in \cite{Ethier+Kurtz1986}), it suffices to show that \eqref{eqn:MSB_condition} holds for $f \in \mathcal{S}$, \ie
+Taking the characteristic function of both these distributions,
+deterministic mutation-selection balance is therefore equivalent to
 \begin{align} \label{eqn:MSB_CF}
@@
 \begin{align*}
-	\frac{\E\left[ f(Z) \mu(Z) \right]}{\E[\mu(Z)]} &= \E\left[ e^{i\theta\left(\frac{Z^{1} +Z^{2}}{2} + R \right)}\right]\\
-	&= \E\left[ e^{i\frac{theta}{2} Z^{1}}e^{i\frac{theta}{2} Z^{2}}e^{\i\theta R}\right]\\
-	&= \E\left[e^{i\frac{\theta}{2} Z^{1}}\right]\E\left[e^{i\frac{\theta}{2} Z^{2}}\right]\E\left[e^{\i\theta R}\right]\\
-	&= \phi\left(\frac{\theta}{2}\right)	^{2} \Phi(\theta),
+    \frac{\E\left[ e^{i\theta Z} \mu(Z) \right]}{\E[\mu(Z)]} &= \E\left[ e^{i\theta\left(\frac{Z^{1} +Z^{2}}{2} + R \right)}\right]\\
+	% &= \E\left[ e^{i\frac{\theta}{2} Z^{1}}e^{i\frac{\theta}{2} Z^{2}}e^{\i\theta R}\right]\\
+	&= \E\left[e^{i\frac{\theta}{2} Z^{1}}\right]\E\left[e^{i\frac{\theta}{2} Z^{2}}\right]\E\left[e^{i\theta R}\right]\\
+	&= \phi\left(\frac{\theta}{2}\right)^{2} \Phi(\theta),
 \end{align*}
 where
-\begin{equation}\label{eq:phi}
-	\phi(\theta) = \mathbb{E}\left[e^{i\theta Z^{1}}\right] = \mathbb{E}\left[e^{i\theta Z^{2}}\right]
-\end{equation}
+$ \phi(\theta) = \mathbb{E}\left[e^{i\theta Z^{1}}\right] = \mathbb{E}\left[e^{i\theta Z^{2}}\right] $
 and
-\begin{equation}\label{eq:Phi}
-	\phi(\theta) = \mathbb{E}\left[e^{i\theta R}\right]
-\end{equation}
+$ \Phi(\theta) = \mathbb{E}\left[e^{i\theta R}\right] $
 are the characteristic functions of $Z^{j}$ and $R$, respectively.  In what follows, we will explore some non-Gaussian solutions $\phi(\theta)$ to \eqref{eqn:MSB_CF}.
@@
 
-First, consider the case without selection, so $\mu$ is a constant.  Then, our condition on the characteristic function simplifies to
+First, consider the case without selection, so $\mu$ is a constant
+which we may take equal to 1.  Then, our condition on the characteristic function simplifies to
 \begin{equation}\label{eq:CF_neutral}
 	{\textstyle \phi\left(\frac{\theta}{2}\right)^{2}} \Phi(\theta) %e^{-\frac{\eta^{2}\theta^{2}}{2}} 
-	= \phi(\theta),
+	= \phi(\theta) ,
 \end{equation}
@@
 \]
-whereas any fixed point must be the limit of an arbitrary characteristic function, $\varphi(\theta)$, under repeated applications of $F$:
-\[
-	\phi(\theta) = \lim_{n \to \infty} F^{(n)}(\varphi)(\theta).
-\]
-Iterating $F$ $n$ times, we get
-\[
-	F^{(n)}(\varphi)(\theta) =  {\textstyle \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}}\prod_{k=1}^{n-1} {\textstyle \Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}},
-\]
-so that
-\[
-	\phi(\theta) = \lim_{n \to \infty} \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}\prod_{k=1}^{n-1}\Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}
-\]
+whereas any fixed point must be the limit of an arbitrary characteristic function, $\varphi(\theta)$, under repeated applications of $F$,
+\ie
+\begin{align*}
+	\phi(\theta) 
+    &= \lim_{n \to \infty} F^{(n)}(\varphi)(\theta) \\
+    &= \lim_{n \to \infty} \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}\prod_{k=1}^{n-1}\Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}} .
+\end{align*}
+% Iterating $F$ $n$ times, we get
+% \[
+% 	F^{(n)}(\varphi)(\theta) =  {\textstyle \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}}\prod_{k=1}^{n-1} {\textstyle \Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}},
+% \]
+% so that
+% \[
+% 	\phi(\theta) = \lim_{n \to \infty} \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}\prod_{k=1}^{n-1}\Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}
+% \]
 is a fixed point whenever this limit exists.
@@
 Since $\varphi$ and $\Phi$ are characteristic functions, they are absolutely continuous and $\phi(0) = 1$, and thus non-zero in a neighbourhood of 0.
-In particular, $\psi(\theta) = \Log{\varphi}(\theta)$ and $\Psi(\theta) = \Log{\Phi(\theta)}$ exist in some neighbourhood of 0 as well\footnote{We use $\Log$ to denote the principal value of the complex logarithm, for which the imaginary part always lies in the interval $(-\pi,\pi]$}.
+In particular, $\psi(\theta) = \Log{\varphi}(\theta)$ and $\Psi(\theta) = \Log{\Phi(\theta)}$ exist in some neighbourhood of 0 as well\footnote{We use $\Log$ to denote the principal value of the complex logarithm, for which the imaginary part always lies in the interval $(-\pi,\pi]$.}.
 From the above, we have
@@
 \]
-We will briefly explore what limiting values are possible -- and which limits are valid characteristic functions.
-We refer to the first term on the left hand side as the \emph{reproduction term} and the remaining sum as the \emph{noise term}.
-For the sake of simplicity, we will limit ourselves to the caae when both have well-posed limits independent of the other.
+We briefly explore what possible limiting values are valid characteristic functions.
+We refer to the first term on the left hand side as the \emph{reproduction term} and the remaining sum as the \emph{noise term}, and only consider to the case when both have well-posed limits independent of the other.
 
-\subsubsection{Reproductive Terms}
-
+\paragraph{Reproductive Terms:}
 First consider the limit
@@
 \]
-Because $\psi(\theta)$ is the logarithm characteristic function, it must be continuous in some neighbourhood of, and vanish at 0, and thus $\alpha(\theta)$ as well.
-Furthermore, for all integers $k$,
+Because $\psi(\theta)$ is the logarithm of a characteristic function, it is continuous in some neighbourhood of, and vanish at 0.
+If $\phi$ is not differentiable at zero (and so the corresponding distribution
+does not have a well-defined mean),
+the limit does not exist.
+On the other hand if $\psi$ is differentiable at zero,
+then $\alpha$ is continuous there,
+and for all integers $k$, 
 \[
@@
 \]
-as $k \to - \infty$ if the latter exists (\ie there is a unique left-hand limit at zero; this is the case if \eg we require $\alpha(\theta)$, and thus $\phi(\theta)$, to be convex functions).
+as $k \to - \infty$ if the latter exists (\ie there is a unique right-hand limit at zero; this is the case if \eg we require $\alpha(\theta)$, and thus $\phi(\theta)$, to be convex functions).
 In this case, setting $-\gamma + i m = \alpha'(0+)$, we have
@@
 \paragraph{Lamarkian inheritance redux.} 
-Note that the characteristic functions constructed above are solutions to \eqref{eq:CF_neutral} when transmission is noiseless, and the offspring have exactly the parental mean trait \ie Lamarkian blending inheritance.
+Note that the characteristic functions constructed above are solutions to \eqref{eq:CF_neutral} when transmission is noiseless, and the offspring have exactly the parental mean trait -- \ie Lamarkian blending inheritance.
 The solution $\phi(\theta) = e^{i\theta m}$ is exactly the reversion to the mean that Darwin's detractors used to argue against the theory of evolution \citep{Provine1971}.
@@
 
-\paragraph{Other solutions?}
-We would naturally like to know if these two exhaust the possible solutions.
+% \paragraph{Other solutions?}
+% We would naturally like to know if these two exhaust the possible solutions.
 Above, we exhausted all solutions for which $\alpha'(0+)$ exists.
@@
 This does not, however, in and of itself exclude $e^{\alpha(\theta)}$ from being a characteristic function: Bochner's theorem tells us that a function $\phi(\theta)$ is the characteristic function of a random variable if and only if
-\begin{enumerate}[(i)]
-\item $\phi(\theta)$ is continuous,
-\item $\phi(0) = 1$,
-\item $\phi(-\theta) = \overline{\phi(\theta)}$, and
-\item $\phi(\theta)$ is \textit{positive definite}: for all $n$ and all $\theta_{1},\ldots,\theta_{n} \in \mathbb{R}$, the matrix with entries $a_{ij} = \phi(\theta_{i}-\theta_{j})$
-\end{enumerate}
-One can easily choose $\alpha(\theta)$ so that $e^{\alpha(\theta)}$ satisfies the first three criteria, \eg $\alpha(\theta) = -|\theta|\left(\sin\left(\pi\frac{\log{|\theta|}}{\log{2}}\right)+\frac{3}{2}\right)$, but excluding such functions via Bochner's positive definite criterion is non-trivial.
+% \begin{enumerate}[(i)]
+% \item $\phi(\theta)$ is continuous,
+% \item $\phi(0) = 1$,
+% \item $\phi(-\theta) = \overline{\phi(\theta)}$, and
+% \item $\phi(\theta)$ is \textit{positive definite}: for all $n$ and all $\theta_{1},\ldots,\theta_{n} \in \mathbb{R}$, the matrix with entries $a_{ij} = \phi(\theta_{i}-\theta_{j})$
+% \end{enumerate}
+% One can easily choose $\alpha(\theta)$ so that $e^{\alpha(\theta)}$ satisfies the first three criteria, \eg $\alpha(\theta) = -|\theta|\left(\sin\left(\pi\frac{\log{|\theta|}}{\log{2}}\right)+\frac{3}{2}\right)$, but excluding such functions via Bochner's positive definite criterion is non-trivial.
 
-\subsubsection{Noise Terms}
-
+\paragraph{Noise Terms}
 We now turn our attention to the noise terms.

### 2023-07-07 writeup.tex
peter edited throuh

@@
 If however, we specialize to the case when $R$ is a stable law, we can explicitly compute the limiting noise term.
-We recall that if $\Psi(\theta)$ corresponds to a stable law(see \eg \cite{breiman1992probability}), then
+We recall that if $\Psi(\theta)$ corresponds to a stable law (see \eg \cite{breiman1992probability}), then
 \[
@@
 	 \sum_{k=1}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}
-	 &=  \sum_{k=1}^{n-1} im\theta + 2^{(1-\alpha)k} c|\theta|^{\alpha} {\textstyle\left(1-i \beta\sgn(\theta)\left(\frac{\theta}{2^{k}}\right)
+	 &=  \sum_{k=1}^{n-1} 2^{(1-\alpha)k} c|\theta|^{\alpha} {\textstyle\left(1-i \beta\sgn(\theta)\left(\frac{\theta}{2^{k}}\right)
 	 	 \omega_{\alpha}\left(\frac{\theta}{2^{k}}\right)\right)}\\
 	&= \begin{cases}
-		 im(n-1)\theta + c|\theta|^{\alpha}\frac{1-2^{(1-\alpha)n}}{1-2^{1-\alpha}}\left(1-i\beta \sgn(\theta) \tan{\frac{\pi\alpha}{2}}\right) & \text{if $\alpha \neq 1$, and}\\	
-		 c|\theta|^{\alpha}  \left(1+\frac{2i\beta}{\pi}\sgn(\theta)\left(\log{|\theta|} - \frac{n(n-1)}{2} \log{2}\right)\right) & \text{if $\alpha = 1$,}
+		 c|\theta|^{\alpha}\frac{1-2^{(1-\alpha)n}}{1-2^{1-\alpha}}\left(1-i\beta \sgn(\theta) \tan{\frac{\pi\alpha}{2}}\right) & \text{if $\alpha \neq 1$, and}\\	
+         (n-1) c|\theta|^{\alpha}  \left(1+\frac{2i\beta}{\pi}\sgn(\theta)\left(\log{|\theta|} - \frac{n}{2} \log{2}\right)\right) & \text{if $\alpha = 1$,}
 	\end{cases}
@@
 
-More generally, one will get convergence if the noise is a mixture of components with finite variance and stable laws as above, whereas the population distribution will be a mixture of a (possibly degenerate) Cauchy distribution and such noise terms -- we remark that in the presence of noise, a pure Cauchy distributed solution is \emph{not} possible in the neutral case.
+More generally, one will get convergence if the noise is a mixture of components with finite variance and stable laws as above, whereas the population distribution will be a mixture of a (possibly degenerate) Cauchy distribution and such noise terms.
+Note that in the presence of noise, a pure Cauchy-distributed solution is \emph{not} possible in the neutral case.
 
@@
 Unfortunately, once selection is included, \eqref{eqn:MSB_CF} is less amenable to the sort of simple arguments above.
-We will content ourselves with presenting two examples in the case of stabilizing selection, one already familiar to enthusiasts of quantitative genetics, the other a Cauchy solution obtained by trial-and-error, which illustrates how selection can stabilize an otherwise unviable solution.
+We will content ourselves with presenting two examples of stabilizing selection, one already familiar to quantitative genetics, the other a Cauchy solution obtained by trial-and-error, which illustrates how selection can stabilize an otherwise unviable solution.
 
-\subsubsection{Gaussian Solution}
-    \label{sec:stabilizing_gaussian}
-
-First, consider the case when $1+ \mu(x) = e^{a x^{2}}$.   In this case there is a Gaussian solution: suppose $Z$ is Gaussian with mean $m$ and variance $\sigma^{2}$.  Then,
-\[
-	\phi(\theta) = e^{i m \theta - \frac{1}{2} \sigma^{2} \theta^{2}}.
-\]
-Moreover, provided $a < \frac{1}{2\sigma^{2}}$, we can explicitly compute $\frac{\mathbb{E}\left[e^{i \theta Z}(1+\mu(Z))\right]}{\mathbb{E}[1+\mu(Z)]}$  using the Gaussian probability density function to conclude that
+\paragraph{Gaussian Solution:}
+First, consider the case when $\mu(x) = e^{a x^{2}}$.   In this case there is a Gaussian solution: suppose $Z$ is Gaussian with mean $m$ and variance $\sigma^{2}$.  Then,
+$ \phi(\theta) = \exp(i m \theta - \frac{1}{2} \sigma^{2} \theta^{2}).  $
+Moreover, provided $a < \frac{1}{2\sigma^{2}}$, we can explicitly compute $\frac{\mathbb{E}\left[e^{i \theta Z}\mu(Z)\right]}{\mathbb{E}[\mu(Z)]}$  using the Gaussian probability density function to conclude that
 if $R$ is Gaussian with mean zero and variance $\eta^2$, then equation~\eqref{eqn:MSB_CF} becomes
@@
 Note that we do not require very weak selection to guarantee a Gaussian steady state, only that $a < \frac{1}{2\sigma^{2}}$.
-However, this does not imply that the transient distribution is also Gaussian, as is sometimes assumed \comment{cite any number of Lande papers?}.
-The latter does indeed require very weak selection.
+However, this does not imply that the transient distribution is also Gaussian, as is sometimes assumed \citep[e.g.,][]{lande1976natural}.
+%  The latter does indeed require very weak selection.
 
-\subsubsection{Cauchy Solution}
-    \label{sec:stabilizing_cauchy}
-
-Almost uniquely among the stable laws, the Cauchy distribution has a probability density function expressible in elementary functions (the other exceptions being the L\'evy and Gaussian distributions): for a Cauchy random variable with median (location) $x_{0}$ and mean absolute deviation (scale) $\gamma$, the density is $f(x;x_{0},\gamma) = \frac{1}{\pi\gamma\left[1+\left(\frac{x-x_{0}}{\gamma}\right)^{2}\right]}$.
-Armed with this density function, we can explicitly compute the right hand size of \eqref{eqn:MSB_CF}, assuming that
+\paragraph{Cauchy Solution:}
+Almost uniquely among the stable laws, the Cauchy distribution has a probability density function expressible in elementary functions:
+% (the other exceptions being the L\'evy and Gaussian distributions):
+for a Cauchy random variable with median (location) $x_{0}$ and mean absolute deviation (scale) $\gamma$, the density is $f(x;x_{0},\gamma) = \frac{1}{\pi\gamma\left[1+\left(\frac{x-x_{0}}{\gamma}\right)^{2}\right]}$.
+Armed with this density function, we can explicitly compute the right hand side of \eqref{eqn:MSB_CF}, assuming that
 \begin{enumerate}[(i)]
@@
 \item selection acts as $\mu(x) = \frac{\delta^{2}}{\gamma^{2}} \frac{\gamma^{2} + x^{2}}{\delta^{2} + x^{2}}$ for some $\delta > \gamma > 0$, and
-\item the transmission noise $R$ is given by a $\text{Cauchy}(0,\delta-\gamma)$ random variable (\ie corresponding to a stable $(0,0,\delta-\gamma,0)$ law).
+\item the Mendelian sampling term $R$ is given by a $\text{Cauchy}(0,\delta-\gamma)$ random variable (\ie corresponding to a stable $(1,0,\delta-\gamma,0)$ law).
 \end{enumerate}
 
-Under these assumptions,  
+Under these assumptions, \eqref{eqn:MSB_CF} is
 \begin{align*}
-	\frac{\mathbb{E}\left[e^{i \theta Z}(1+\mu(Z))\right]}{\mathbb{E}[1+\mu(Z)]} 
+	\frac{\mathbb{E}\left[e^{i \theta Z}\mu(Z)\right]}{\mathbb{E}[\mu(Z)]} 
 	&= e^{-\delta |\theta|}\\
@@
 \end{align*} 
-which we recognize as being equal to $\phi\left(\frac{\theta}{2}\right)^{2} \Phi(\theta)$, where $\phi(\theta) = e^{i\gamma|\theta|}$ and $\Phi(\theta) = e^{i(\delta - \gamma)|\theta|}$ are the characteristic functions of the population density and the transmission noise, respectively, giving us a.non-Gaussian solution to \eqref{eqn:MSB_CF} with both noise and selection.
+which we recognize as being equal to $\phi\left(\frac{\theta}{2}\right)^{2} \Phi(\theta)$, where $\phi(\theta) = e^{i\gamma|\theta|}$ and $\Phi(\theta) = e^{i(\delta - \gamma)|\theta|}$ are the characteristic functions of the population density and the Mendelian sampling term, respectively,
+giving a.non-Gaussian solution to \eqref{eqn:MSB_CF} with both noise and selection.
 
@@
 Next we turn to simulation,
-in which we can easily include realistic recombination (instead of unlinked loci) and selection.
+in which we can easily include realistic recombination (instead of unlinked loci) and selection,
+and in which we can explore the applicability of the models of the previous section.
 We used SLiM v4 \citep{haller2022slim4}
@@
 by summing the effects of all mutations they carry across both chromosomes.
-Let $f(z)$ be the death rate of an individual with phenotype $z$,
+Let $\mu(z)$ be the death rate of an individual with phenotype $z$,
 and let $dt$ be a small constant (we took $dt=0.01$).
@@
 Then, at each time step
-each individual dies with probability $1 - \exp(-dt f(z))$;
+each individual dies with probability $1 - \exp(-dt \mu(z))$;
 if there are $k$ such individuals chosen to die in a given time step,
@@
 
-We first consider the neutral case (i.e., with $f \equiv 1$).
+We first consider the neutral case (i.e., with $\mu \equiv 1$).
 As shown in Figure~\ref{fig:trait_distrns},
@@
 
-We next compared two situations of stabilizing selection, motivated by the results above
-\comment{TODO: refer to where?}:
+We next compared two situations of stabilizing selection, motivated by the results
+in Section~\ref{sec:stabilizing_selection}:
 (1) the effect size distribution was Gaussian with mean zero and standard deviation 1.0
-and $f(z) = e^{\alpha z^2}$; and
+and $\mu(z) = e^{\alpha z^2}$; and
 (2) the effect size distribution was Cauchy, centered at zero and with width $\delta - \gamma$
-and $f(z) = (\delta/\gamma) (\gamma + z^2) / (\delta + z^2)$ with $\delta=0.5$ and $\gamma=0.2$.
+and $\mu(z) = (\delta/\gamma) (\gamma + z^2) / (\delta + z^2)$ with $\delta=0.5$ and $\gamma=0.2$.
 Figures~\ref{fig:sel_trait_distrns} and~\ref{fig:sel_seg_noise}

### 2023-07-07 writeup.tex
Added Cauchy section

--- 2023-07-07 writeup.tex
@@
 \usepackage{fullpage}
-\usepackage{amsmath,amssymb,amsthm,bm,enumerate,mathtools}
+\usepackage{amsmath,amssymb,amsthm,bm,bbm,enumerate,mathtools}
 \usepackage[hidelinks]{hyperref}
@@
 
+\newcommand{\iid}{\overset{i.i.d}{\sim}}
+\newcommand{\1}{\mathbbm{1}}
+\newcommand{\Oh}{{\mathcal O}}
+
 \newcommand{\comment}[1]{{\color{blue} \it #1}}
@@
 \DeclareMathOperator*{\Log}{Log}
+\DeclareMathOperator*{\Cauchy}{Cauchy}
+\DeclareMathOperator*{\Bernoulli}{Bernoulli}
 \DeclareMathOperator*{\sgn}{sgn}
@@
 
-Almost uniquely among the stable laws, the Cauchy distribution has a probability density function expressible in elementary functions (the other exceptions being the L\'evy and Gaussian distributions): for a Cauchy random variable with median (location) $x_{0}$ and mean absolute deviation (scale) $\gamma$, the density is $f(x;x_{0},\gamma) = \frac{1}{\pi\gamma\left[1+\left(\frac{x-x_{0}}{\gamma}\right)^{2}\right]}$.
+Almost uniquely among the stable laws, the Cauchy distribution has a probability density function expressible in elementary functions (the other exceptions being the L\'evy and Gaussian distributions): for a Cauchy random variable with median (location) $x_{0}$ and mean absolute deviation (scale) $\gamma$, the density is
+\begin{equation}\label{eq:Cpdf}
+	f(x;x_{0},\gamma) = \frac{1}{\pi\gamma\left[1+\left(\frac{x-x_{0}}{\gamma}\right)^{2}\right]}.  
+\end{equation}	
 Armed with this density function, we can explicitly compute the right hand size of \eqref{eqn:MSB_CF}, assuming that 
@@
 %(here, see a saturating cost to being away from the optimal phenotype $x = 0$: the mortality is maximized at  $\frac{\beta^{2}}{\alpha^{2}}$.
+
+\section{Cauchy-decendants}
+
+Having seen that a mutation-selection balance is possible with a Cauchy empirical distribution and Cauchy transmission noise, it is natural to ask why we couldn't have an infinitesimal model with Cauchy-descendants, \textit{sensu} \cite{turelli2017commentary}.
+To that end, it's instructive to see where the argument of \cite{barton2017infinitesimal} breaks down if instead of assuming bounded allele effects of order $\Oh\left(\frac{1}{\sqrt{M}}\right)$, we assume that the allele at locus $i$ in individual $j$ contributes $\frac{1}{M} \eta^{[j]}_{\ell}$, where the $\eta^{[j]}_{\ell}$ are $\Cauchy(0,\gamma_{\ell})$ random variables.
+In this case, the median effect size at locus $\ell$, $\frac{\gamma_{\ell}}{M}$, is asymptotically smaller than the allele effects assumed in \cite{barton2017infinitesimal}, but large effect alleles are more likely.
+
+The crucial step in the derivation of the infinitesimal model in \cite{barton2017infinitesimal} is to show that, following the arguments in \cite{Fisher1918}, knowing an individual's trait value gives vanishingly little information about the contribution of any given locus as $M \to \infty$, which in turn allows one to show that the within family segregation variance is independent of the parental traits.
+Here, we will consider how conditioning an individual's trait value affects the law of a single allelic contribution in the Cauchy setting described above.
+A similar argument can be made for any $\alpha$-stable distribution, but the lack of probability density functions for generic $\alpha \in (0,2)$ adds complexity to the argument without adding any insight.
+While is meaningless to speak of variances when the transmission noise is $\alpha$-stable, if an individual's trait value is informative of the allelic effect at any given locus, then its trait value is informative of the trait value of any offspring that inherits that allele, and we lose the independence between parental traits and transmission noise that is central to the infinitesimal model.
+
+Consider the simplest possible case of an ancestral population, of $N$ individuals, where the $j$\textsuperscript{th} individual has trait value 
+\[
+	Z^{[j]}_{M} = \frac{1}{M} \sum_{\ell = 1}^{M}  \eta^{[j]}_{\ell},
+\]
+where the $\eta^{[j]}_{\ell}$ are independent across individuals and across loci.  By the additive property of Cauchy random variables, $Z^{[j]} \iid \Cauchy(0,\bar{\gamma})$, where 
+\[
+	\bar{\gamma} = \frac{1}{M} \sum_{\ell = 1}^{M} \gamma_{\ell}
+\]
+is the mean of the scale parameters.
+
+To see what knowing the trait value $Z^{[j]}_{M}$ tells us about the contribution of any given locus, $\eta^{[j]}_{\ell}$, we follow the arguments in  \cite{Fisher1918,barton2017infinitesimal} using Bayes' Theorem.
+Proceeding informally, with the probabilities indicating probability density functions, we have
+\begin{align*}
+	\P\{\eta^{j[l]}_{\ell} = y \vert Z^{j[l]} = z\} &= \frac{\P\{\eta^{j[l]}_{\ell} = y; Z^{j[l]} = z\}}{\P\{Z^{j[l]} = z\}}\\
+		&= \frac{\P\{Z^{j[l]} = z\vert\eta^{j[l]}_{\ell} = y\}}{\P\{Z^{j[l]} = z\}}\P\{\eta^{j[l]}_{\ell} = y\}\\
+		&=  \frac{\P\{Z^{j[l]} -\frac{1}{M} \eta^{j[l]}_{\ell} = z - \frac{1}{M} y\}}{\P\{Z^{j[l]} = z\}}\P\{\eta^{j[l]}_{\ell} = y\}\\
+\end{align*}
+Now, $Z^{j[l]} -\frac{1}{M} \eta^{j[l]}_{\ell} = \frac{1}{M} \sum_{{k = 1} \atop {k \neq i}}^{M} \eta^{[j]}_{k}$ is Cauchy-distributed with scale $\bar{\gamma} - \frac{\gamma_{\ell}}{M}$.
+Using the probability density function for a Cauchy distributed random variable \eqref{eq:Cpdf}, a straightforward computation gives
+\begin{align*}
+	\P\{\eta^{j[l]}_{\ell} = y \vert Z^{j[l]} = z\} &= \frac{f\left(z - \frac{y}{M} ;0,\bar{\gamma}-\frac{\gamma_{\ell}}{M}\right)}{f(z;0,\bar{\gamma})} \P\{\eta^{j[l]}_{\ell} = y\}\\
+	&= \left(1+\frac{\gamma_{\ell}(\bar{\gamma}^{2}-z^{2})+2zy}{\gamma_{\ell}(\bar{\gamma}^{2}+z^{2})} \frac{1}{M} + \Oh\left(\frac{1}{M^{2}}\right)\right) \P\{\eta^{j[l]}_{\ell} = y\}.
+\end{align*} 
+Now, as $z \to 0$, $\frac{\gamma_{\ell}(\bar{\gamma}^{2}-z^{2})+2zy}{\gamma_{\ell}(\bar{\gamma}^{2}+z^{2})} \to \frac{\gamma_{\ell}}{\bar{\gamma}}$ and as $z \to \pm \infty$, $\frac{\gamma_{\ell}(\bar{\gamma}^{2}-z^{2})+2zy}{\gamma_{\ell}(\bar{\gamma}^{2}+z^{2})} \to -\frac{\gamma_{\ell}}{\bar{\gamma}}$,  whereas $|\frac{\bar{\gamma}^{2}-z^{2}+2zy}{\bar{\gamma}^{2}+z^{2}}|$ attains a maximum of
+\[
+	\frac{(\gamma_{\ell}^{2}+y^{2})\left(\sqrt{\gamma_{\ell}^{2}+y^{2}}-\gamma_{\ell}\right)}{\bar{\gamma}(\gamma_{\ell}^{2}+y^{2} - \gamma_{\ell}\sqrt{\gamma_{\ell}^{2}+y^{2}})},
+\]
+which is asymptotically equal to $\frac{|y|}{\bar{\gamma}}$ for large values of $|y|$, at
+\[
+	z = -\frac{\bar{\gamma}}{y}(\gamma_{\ell} \pm \sqrt{\gamma_{\ell}^{2}+y^{2}}),  
+\]
+which is itself asymptotically equal to $\mp \sgn(y) \bar{\gamma}$ for large values of $|y|$.
+
+Thus, for large values of $|y|$,
+\[
+	|\P\{\eta^{j[l]}_{\ell} = y \vert Z^{j[l]} = z\} - \P\{\eta^{j[l]}_{\ell} = y\}| \lesssim \left(\frac{|y|}{\bar{\gamma} M} + \Oh\left(\frac{1}{M}\right)\right)  \P\{\eta^{j[l]}_{\ell} = y\}.
+\]
+We remark that, unlike the case of bounded allelic effects, we obtain a bound independent of the trait value $Z^{[j]}$.
+Moreover, in the case of bounded allelic effects, the error introduced by approximating the effect-size distribution conditional on the trait value by the unconditioned distribution is proportional to $|Z^{[j]}|$, so that extreme trait values are most informative of the size of allelic effects, whereas here, the error is maximized when the trait value is approximately equal to the mean absolute deviation of its distribution, \ie the median value of $|Z^{[j]}|$.
+
+The argument above shows for a typical locus, the effect of conditioning on the trait gives negligible information about the contribution from that site.   Recalling that
+\[
+	\P\left\{|\eta^{j[l]}_{\ell}| > y\right\} =  1-\frac{2}{\pi} \arctan\left(\frac{y}{\gamma_{\ell}}\right) \sim \frac{2\gamma_{\ell}}{\pi y},
+\]
+we see that for any $\epsilon > 0$, and large values of $M$,
+\[
+	\P\left\{\frac{\eta^{j[l]}_{\ell}}{\bar{\gamma} M} > \epsilon\right\} \sim \frac{2\gamma_{\ell}}{\pi \bar{\gamma} M \epsilon}.
+\]
+so the probability the error is of order $\Oh(\epsilon)$ at any given locus decays rapidly as $M \to \infty$.
+
+If, however, we consider not any one locus, but rather all loci, we get a different picture.  The variables $\eta^{j[l]}_{\ell}$ are independent, and thus so are the Bernoulli random variables 
+\begin{equation}
+	\1^{M}_{\ell} = \begin{cases}
+		1 & \text{if $\frac{\eta^{j[l]}_{\ell}}{\bar{\gamma} M \epsilon}$, and}\\
+		0 & \text{otherwise.}
+	\end{cases}
+\end{equation}
+Now, since
+\[
+	\lim_{M \to \infty} \sum_{\ell = 1}^{M} \frac{2\gamma_{\ell}}{\pi \bar{\gamma} M \epsilon} = \frac{2}{\pi \epsilon},
+\]
+the number of loci at which the magnitude of the effect size exceeds $M \epsilon$, which is 
+\[
+	\sum_{\ell = 1}^{M} \1^{M}_{\ell},
+\]
+converges weakly as $M \to \infty$ to a Poisson random variable with rate $\frac{2}{\pi \epsilon}$ (see \eg Theorem 2.6.1 in \cite{Durrett2005}).
+In particular, for large values of $M$ and fixed $\epsilon > 0$,
+\begin{align*}
+	\P\left\{\max_{i=1}^{M} \frac{\eta^{j[l]}_{\ell}}{\bar{\gamma} M} > \epsilon\right\}  \sim 1-e^{-\frac{2}{\pi \bar{\gamma} \epsilon}},
+\end{align*}
+which approaches 1 as $\epsilon \to 0$, while the expected number of $M$ alleles such that $\frac{\eta^{j[l]}_{\ell}}{\bar{\gamma} M} > \epsilon$ is asymptotically $\frac{2}{\pi\epsilon}$.
+In particular, there is at least one locus at which knowing the parental trait gives one information about the effect size (and there are on the order of $\Oh\left(\frac{1}{\epsilon}\right)$ that give ``an $\epsilon$ of information''), and these large effect loci in turn give rise to correlations between parental and offspring traits.
 

### 2023-07-07 writeup.tex

@@
 Unfortunately, once selection is included, \eqref{eqn:MSB_CF} is less amenable to the sort of simple arguments above.
-We will content ourselves with presenting two examples in the case of stabilizing selection, one already familiar to enthusiasts of quantitative genetics, the other a Cauchy solution obtained by trial-and-error, which illustrates how selection can stabilize an otherwise unviable solution.
-
-\subsubsection{Gaussian Solution}
-    \label{sec:stabilizing_gaussian}
-
-First, consider the case when $1+ \mu(x) = e^{a x^{2}}$.   In this case there is a Gaussian solution: suppose $Z$ is Gaussian with mean $m$ and variance $\sigma^{2}$.  Then,
-\[
-	\phi(\theta) = e^{i m \theta - \frac{1}{2} \sigma^{2} \theta^{2}}.
-\]
-Moreover, provided $a < \frac{1}{2\sigma^{2}}$, we can explicitly compute $\frac{\mathbb{E}\left[e^{i \theta Z}(1+\mu(Z))\right]}{\mathbb{E}[1+\mu(Z)]}$  using the Gaussian probability density function to conclude that
+We will content ourselves with presenting two examples of stabilizing selection, one already familiar to quantitative genetics, the other a Cauchy solution obtained by trial-and-error, which illustrates how selection can stabilize an otherwise unviable solution.
+
+\paragraph{Gaussian Solution:}
+First, consider the case when $\mu(x) = e^{a x^{2}}$.   In this case there is a Gaussian solution: suppose $Z$ is Gaussian with mean $m$ and variance $\sigma^{2}$.  Then,
+$ \phi(\theta) = \exp(i m \theta - \frac{1}{2} \sigma^{2} \theta^{2}).  $
+Moreover, provided $a < \frac{1}{2\sigma^{2}}$, we can explicitly compute $\frac{\mathbb{E}\left[e^{i \theta Z}\mu(Z)\right]}{\mathbb{E}[\mu(Z)]}$  using the Gaussian probability density function to conclude that
 if $R$ is Gaussian with mean zero and variance $\eta^2$, then equation~\eqref{eqn:MSB_CF} becomes
@@
 Note that we do not require very weak selection to guarantee a Gaussian steady state, only that $a < \frac{1}{2\sigma^{2}}$.
-However, this does not imply that the transient distribution is also Gaussian, as is sometimes assumed \comment{cite any number of Lande papers?}.
-The latter does indeed require very weak selection.
+However, this does not imply that the transient distribution is also Gaussian, as is sometimes assumed \citep[e.g.,][]{lande1976natural}.
+%  The latter does indeed require very weak selection.
 
@@
 
-Almost uniquely among the stable laws, the Cauchy distribution has a probability density function expressible in elementary functions (the other exceptions being the L\'evy and Gaussian distributions): for a Cauchy random variable with median (location) $x_{0}$ and mean absolute deviation (scale) $\gamma$, the density is
+Almost uniquely among the stable laws, the Cauchy distribution has a probability density function expressible in elementary functions: 
+% (the other exceptions being the L\'evy and Gaussian distributions): 
+for a Cauchy random variable with median (location) $x_{0}$ and mean absolute deviation (scale) $\gamma$, the density is 
 \begin{equation}\label{eq:Cpdf}
@@
 \item selection acts as $\mu(x) = \frac{\delta^{2}}{\gamma^{2}} \frac{\gamma^{2} + x^{2}}{\delta^{2} + x^{2}}$ for some $\delta > \gamma > 0$, and
-\item the transmission noise $R$ is given by a $\text{Cauchy}(0,\delta-\gamma)$ random variable (\ie corresponding to a stable $(0,0,\delta-\gamma,0)$ law).
+\item the Mendelian sampling term $R$ is given by a $\text{Cauchy}(0,\delta-\gamma)$ random variable (\ie corresponding to a stable $(1,0,\delta-\gamma,0)$ law).
 \end{enumerate}
 
-Under these assumptions,  
-\begin{align*}
-	\frac{\mathbb{E}\left[e^{i \theta Z}(1+\mu(Z))\right]}{\mathbb{E}[1+\mu(Z)]} 
+Under these assumptions, \eqref{eqn:MSB_CF} is
+\begin{align*}
+	\frac{\mathbb{E}\left[e^{i \theta Z}\mu(Z)\right]}{\mathbb{E}[\mu(Z)]} 
 	&= e^{-\delta |\theta|}\\

### 2023-07-07 writeup.tex
edits to cauchy section

--- 2023-07-07 writeup.tex
@@
 $$
-\comment{TODO: Gregor suggests ``Mendelian sampling term''}
 What we will call the \emph{transmission noise}, $R_M$,
@@
 
-\paragraph{Cauchy Solution:}
+\subsubsection{Cauchy Solution}
+    \label{sec:stabilizing_cauchy}
+
 Almost uniquely among the stable laws, the Cauchy distribution has a probability density function expressible in elementary functions: 
 % (the other exceptions being the L\'evy and Gaussian distributions): 
-for a Cauchy random variable with median (location) $x_{0}$ and mean absolute deviation (scale) $\gamma$, the density is $f(x;x_{0},\gamma) = \frac{1}{\pi\gamma\left[1+\left(\frac{x-x_{0}}{\gamma}\right)^{2}\right]}$.
-Armed with this density function, we can explicitly compute the right hand side of \eqref{eqn:MSB_CF}, assuming that
+for a Cauchy random variable with median (location) $x_{0}$ and mean absolute deviation (scale) $\gamma$, the density is 
+\begin{equation}\label{eq:Cpdf}
+	f(x;x_{0},\gamma) = \frac{1}{\pi\gamma\left[1+\left(\frac{x-x_{0}}{\gamma}\right)^{2}\right]}.  
+\end{equation}	
+Armed with this density function, we can explicitly compute the right hand size of \eqref{eqn:MSB_CF}, assuming that 
 \begin{enumerate}[(i)]
@@
 
+\section{The difficulty with an abnormal quantitative genetics}
+
+Above we have seen that
+(a) in principle,
+a mutation-selection balance is possible with a Cauchy empirical distribution and Cauchy Mendelian sampling term,
+but (b)
+such a model cannot derive from a simple model with additive loci.
+To that end, it's instructive to see where the argument of \cite{barton2017infinitesimal} breaks down
+if instead of assuming bounded allele effects of order $\Oh\left(1/\sqrt{M}\right)$,
+we assume that the allele at locus $i$ in individual $j$ contributes
+$\eta^{[j]}_{\ell}/M$, where the $\eta^{[j]}_{\ell}$ are independent $\Cauchy(0,\gamma_{\ell})$ random variables. 
+In this case, the typical effect size at locus $\ell$ is of order $\frac{\gamma_{\ell}}{M}$,
+which is asymptotically smaller than the allele effects assumed in \cite{barton2017infinitesimal},
+but large effect alleles are more likely.
+
+The crucial step in the derivation of the infinitesimal model in \cite{barton2017infinitesimal} is to show that, following the arguments in \cite{Fisher1918}, knowing an individual's trait value gives vanishingly little information about the contribution of any given locus as $M \to \infty$, which in turn allows one to show that the within-family Mendelian segregation variance is independent of the parental traits.
+Here, we will consider how conditioning an individual's trait value affects the law of a single allelic contribution in the Cauchy setting described above.
+A similar argument can be made for any $\alpha$-stable distribution, but the lack of explicit probability density functions for generic $\alpha \in (0,2)$ adds complexity to the argument without adding any insight.
+
+The key observation is this:
+while is meaningless to speak of variances when the transmission noise is $\alpha$-stable, if an individual's trait value is informative of the allelic effect at any given locus,
+then its trait value is informative of the \emph{distribution} of an offsprings' trait value,
+and hence there can be no independence between parental traits and the Mendelian sampling term
+that is central to the tractability of the infinitesimal model.
+
+Consider the simplest possible case of an ancestral population, of $N$ individuals, where the $j$\textsuperscript{th} individual has trait value 
+\[
+	Z^{[j]}_{M} = \frac{1}{M} \sum_{\ell = 1}^{M}  \eta^{[j]}_{\ell},
+\]
+where the $\eta^{[j]}_{\ell}$ are independent across individuals and across loci.  By the additive property of Cauchy random variables, $Z^{[j]}_M \iid \Cauchy(0,\bar{\gamma})$, where 
+\[
+	\bar{\gamma} = \frac{1}{M} \sum_{\ell = 1}^{M} \gamma_{\ell}
+\]
+is the mean of the scale parameters.
+
+To see what knowing the parental trait value $Z^{[j]}_{M}$ tells us about the contribution of any given locus, $\eta^{[j]}_{\ell}$, we follow the arguments in  \citet{Fisher1918} and \citet{barton2017infinitesimal} using Bayes' Theorem.
+Proceeding informally, with the probabilities indicating probability density functions, we have
+\begin{align*}
+    \P\{\eta^{[j]}_{\ell} = y \vert Z^{[j]} = z\} &= \frac{\P\{\eta^{[j]}_{\ell} = y; Z^{[j]} = z\}}{\P\{Z^{[j]} = z\}}\\
+        &= \frac{\P\{Z^{[j]} = z\vert\eta^{[j]}_{\ell} = y\}}{\P\{Z^{[j]} = z\}}\P\{\eta^{[j]}_{\ell} = y\}\\
+        &=  \frac{\P\{Z^{[j]} -\frac{1}{M} \eta^{[j]}_{\ell} = z - \frac{1}{M} y\}}{\P\{Z^{[j]} = z\}}\P\{\eta^{[j]}_{\ell} = y\}\\
+\end{align*}
+Now, $Z^{[j]} -\frac{1}{M} \eta^{[j]}_{\ell} = \frac{1}{M} \sum_{k = 1 : k \neq i}^{M} \eta^{[j]}_{k}$ is Cauchy-distributed with scale $\bar{\gamma} - \frac{\gamma_{\ell}}{M}$.
+Using the probability density function for a Cauchy distributed random variable \eqref{eq:Cpdf}, a straightforward computation gives
+\begin{align*}
+    \P\{\eta^{[j]}_{\ell} = y \vert Z^{[j]} = z\}
+    &=
+    \frac{f\left(z - \frac{y}{M} ;0,\bar{\gamma}-\frac{\gamma_{\ell}}{M}\right)}{f(z;0,\bar{\gamma})} \P\{\eta^{[j]}_{\ell} = y\}\\
+    &=
+    \left(1
+        +
+        \frac{\gamma_{\ell}(\bar{\gamma}^{2}-z^{2})+2zy}{\gamma_{\ell}(\bar{\gamma}^{2}+z^{2})}
+        \frac{1}{M}
+        + \Oh\left(\frac{1}{M^{2}}\right)\right)
+        \P\{\eta^{[j]}_{\ell} = y\}.
+\end{align*} 
+Noting that $|(\bar{\gamma}^2 - z^2)/(\bar{\gamma}^2 + z^2)| \le 1$
+and $|z/(\bar{\gamma}^2 + z^2)| \le 1/(2 \bar{\gamma})$
+(since it is maximized at $z = \bar{\gamma}$),
+so
+\begin{align*}
+        \left|
+        \frac{\gamma_{\ell}(\bar{\gamma}^{2}-z^{2})+2zy}{\gamma_{\ell}(\bar{\gamma}^{2}+z^{2})}
+        \right|
+        \le \frac{|y|}{\gamma_\ell \bar{\gamma}} ,
+\end{align*}
+% % BEGIN PREVIOUS ARGUMENT
+% Now, as $z \to 0$, $\frac{\gamma_{\ell}(\bar{\gamma}^{2}-z^{2})+2zy}{\gamma_{\ell}(\bar{\gamma}^{2}+z^{2})} \to \frac{\gamma_{\ell}}{\bar{\gamma}}$ and as $z \to \pm \infty$, $\frac{\gamma_{\ell}(\bar{\gamma}^{2}-z^{2})+2zy}{\gamma_{\ell}(\bar{\gamma}^{2}+z^{2})} \to -\frac{\gamma_{\ell}}{\bar{\gamma}}$,  whereas $|\frac{\bar{\gamma}^{2}-z^{2}+2zy}{\bar{\gamma}^{2}+z^{2}}|$ attains a maximum of
+% \[
+% 	\frac{(\gamma_{\ell}^{2}+y^{2})\left(\sqrt{\gamma_{\ell}^{2}+y^{2}}-\gamma_{\ell}\right)}{\bar{\gamma}(\gamma_{\ell}^{2}+y^{2} - \gamma_{\ell}\sqrt{\gamma_{\ell}^{2}+y^{2}})},
+% \]
+% which is asymptotically equal to $\frac{|y|}{\bar{\gamma}}$ for large values of $|y|$, at
+% \[
+% 	z = -\frac{\bar{\gamma}}{y}(\gamma_{\ell} \pm \sqrt{\gamma_{\ell}^{2}+y^{2}}),  
+% \]
+% which is itself asymptotically equal to $\mp \sgn(y) \bar{\gamma}$ for large values of $|y|$.
+% 	
+we have that
+\[
+    \left|
+    \P\{\eta^{[j]}_{\ell} = y \vert Z^{[j]} = z\} - \P\{\eta^{[j]}_{\ell} = y\}
+    \right|
+    \le \left(\frac{|y|}{\gamma_\ell \bar{\gamma} M} + \Oh\left(\frac{1}{M^2}\right)\right)  \P\{\eta^{[j]}_{\ell} = y\}.
+\]
+% Unlike the case of bounded allelic effects, the bound is independent of the trait value $Z^{[j]}$. 
+In the case of bounded allelic effects, the error introduced by approximating the effect-size distribution conditional on the trait value by the unconditioned distribution is proportional to $|Z^{[j]}|$, so that extreme trait values are most informative of the size of allelic effects,
+whereas here,
+the bound is independent of the trait value,
+and the error is maximized when the trait value is approximately equal to
+the mean absolute deviation of its distribution, \ie the median value of $|Z^{[j]}|$.
+
+The argument above shows for a typical locus, the effect of conditioning on the trait gives negligible information about the contribution from that site.   Recalling that
+\[
+    \P\left\{|\eta^{[j]}_{\ell}| > y\right\} =  1-\frac{2}{\pi} \arctan\left(\frac{y}{\gamma_{\ell}}\right) \sim \frac{2\gamma_{\ell}}{\pi y},
+\]
+we see that for any $\epsilon > 0$, and large values of $M$,
+\[
+    \P\left\{\frac{\eta^{[j]}_{\ell}}{\gamma_\ell \bar{\gamma} M} > \epsilon\right\} \sim \frac{2\gamma_{\ell}}{\pi \gamma_\ell \bar{\gamma} M \epsilon}.
+\]
+so the probability the error is larger than $\epsilon$ at any given locus decays rapidly as $M \to \infty$.
+
+If, however, we consider not any one locus, but rather all loci, we get a different picture. 
+The variables $\eta^{[j]}_{\ell}$ are independent, and thus so are the Bernoulli random variables 
+\begin{equation}
+	\1^{M}_{\ell} = \begin{cases}
+        1 & \text{if $\frac{\eta^{[j]}_{\ell}}{\bar{\gamma} M} > \epsilon$, and}\\
+		0 & \text{otherwise.}
+	\end{cases}
+\end{equation}
+Now, since
+\[
+	\lim_{M \to \infty} \sum_{\ell = 1}^{M} \frac{2\gamma_{\ell}}{\pi \bar{\gamma} M \epsilon} = \frac{2}{\pi \epsilon},
+\]
+the number of loci at which the magnitude of the effect size exceeds $M \epsilon$, which is 
+$\sum_{\ell = 1}^{M} \1^{M}_{\ell}$,
+converges weakly as $M \to \infty$ to a Poisson random variable with mean $2/(\pi \epsilon)$ (see \eg Theorem 2.6.1 in \cite{Durrett2005}).
+In particular, for large values of $M$ and fixed $\epsilon > 0$,
+\begin{align*}
+    \P\left\{\max_{i=1}^{M} \frac{\eta^{[j]}_{\ell}}{\bar{\gamma} M} > \epsilon\right\}  \sim 1-e^{-\frac{2}{\pi \bar{\gamma} \epsilon}},
+\end{align*}
+which approaches 1 as $\epsilon \to 0$, while the expected number of $M$ alleles such that $\frac{\eta^{[j]}_{\ell}}{\bar{\gamma} M} > \epsilon$ is asymptotically $\frac{2}{\pi\epsilon}$.
+
+In particular, there is at least one locus at which knowing the parental trait gives one information about the effect size (and there are on the order of $\Oh\left(\frac{1}{\epsilon}\right)$ that give ``an $\epsilon$ of information''), and these large effect loci in turn give rise to correlations between
+parental traits and the deviation of the offsprings' traits from the midparent
+(i.e., the Mendelian sampling terms).
+However, the first calculations above
+suggest that for ``most'' loci,
+this is not true,
+and so perhaps a reasonable picture of Mendelian sampling terms
+in an additive model with Cauchy-distributed effects
+might be ``some large Mendelian alleles plus Gaussian noise''.
+We now turn to simulation,
+to investigate.
+
 %%%%%%%%%%%%%%%%%%%%%%
@@
 
-Next we turn to simulation,
-in which we can easily include realistic recombination (instead of unlinked loci) and selection,
-and in which we can explore the applicability of the models of the previous section.
+In simulation we can easily include realistic recombination (instead of unlinked loci) and selection.
 We used SLiM v4 \citep{haller2022slim4}

### 2023-07-07 sims/make_paper_figs.R
figure size fix

@@ -162,7 +162,7 @@ fname <- "selected_seg_noise_small.png"
 labels <- list(c("(A)", "(B)", "(C)"), c("(D)", "(E)", "(F)"))
 names(labels) <- names(examples)
 
-png(file=fname, width=8.5, height=6, pointsize=10, units='in', res=288)
+png(file=fname, width=6.5, height=6, pointsize=10, units='in', res=288)
 plot_seg_noise(examples, labels, seg_col=1)
 dev.off()
 

### 2023-07-07 writeup.tex
.

@@
 and a growing literature studies the effects of such power law distributions
-on spatial genetic and dynamics
+on spatial genetic and dynamics \comment{I'm not quite sure what you mean by that}
 \citep[e.g.,][]{paulose2019spatial,smith2023isolation}.
@@
 \]
-Because $\psi(\theta)$ is the logarithm of a characteristic function, it is continuous in some neighbourhood of, and vanish at 0.
+Because $\psi(\theta)$ is the logarithm of a characteristic function, it is continuous in some neighbourhood of, and vanishes at 0.
 If $\phi$ is not differentiable at zero (and so the corresponding distribution
@@
         As in Figure~\ref{fig:seg_noise},
-        except computed using only mutations with absolute value effect size less than 0.1.
+        except computed using only mutations with absolute value effect size less than 0.2.
         The same quantities for the simulations of Figure~\ref{fig:sel_seg_noise}
@@
         As in Figure~\ref{fig:seg_noise},
-        except computed using only mutations with absolute value effect size less than 0.1.
+        except computed using only mutations with absolute value effect size less than 0.2.
         \label{fig:sel_seg_noise_small}

### 2023-07-07 writeup.tex
FIxed typo
minor edits up to section 4

@@
 which describe the distributions of sums of a great many independent and exchangeable things:
-the \emph{stable laws} \citep{gnedenko1968limit}
+the \emph{stable laws} \citep{gnedenko1968limit}.
 Roughly speaking, these say that if we add together $n$ independent copies of a random variable $X$
@@
 and a growing literature studies the effects of such power law distributions
-on spatial genetic and dynamics \comment{I'm not quite sure what you mean by that}
+on spatial genetic patterns and evolutionary dynamics
 \citep[e.g.,][]{paulose2019spatial,smith2023isolation}.
 More directly analogous to the problem we study here, \citet{landis2012phylogenetic}
-found that an $\alpha$-stable process (with $\alpha \approx 1.5$) fit the evolution of log endocranial volume
+found that an $\alpha$-stable process (with $\alpha \approx 3/2$) fit the evolution of log endocranial volume
 on the primate phylogeny
-better than either a Brownian model.
+better than a Brownian model.
 It is also important to note that
@@
 even omitting an independent effect of ``the environment''.
-However, we emphasize that such models should \emph{not} always be the default for understanding biological variation,
+However, we emphasize that such models should not always be the default for understanding biological variation,
 especially for human traits.
@@
 $$
-What we will call the \emph{transmission noise}, $R_M$,
+What we will call the \emph{Mendelian sampling term}
+(or occasionally \emph{Mendelian noise}), $R_M$,
 is the offspring's trait minus the midparent value, $\bar Z_M$:
@@
 \end{align*}
+Note in particular that
+the contribution of locus $\ell$ to $R_M$ is $\pm |\eta_\ell^{[1]} - \eta_\ell^{[2]}|$.
 
@@
 for would have on the induced evolutionary dynamics.
-But, is there any evidence that non-Gaussian transmission noise is important in practice?
+But, is there any evidence that non-Gaussian effects are important in practice?
 In this paper we focus primarily on the theory,
@@
 
-Since the transmission noise is a sum of terms of the form
+Since the mendelian sampling term is a sum of terms of the form
 $$
@@
 in at least some situations,
-independence of transmission noise implies a Gaussian distribution.
+independence of the Mendelian sampling term implies a Gaussian distribution.
 To see this,
@@
 The Kac-Bernstein theorem states that if two independent random variables
-have the property that if their sum and difference are independent of each other,
+have the property that their sum and difference are independent of each other,
 then the random variables have a joint Gaussian distribution
@@
     and that $\{X_\ell\}$ are independent of $\{\eta_\ell^{[i]}\}$.
-    If the transmission noise $R_M$ is independent of the midparent value $\bar Z_M$,
+    If the Mendelian sampling term $R_M$ is independent of the midparent value $\bar Z_M$,
     then $R_M$ and $\bar Z_M$ are jointly Gaussian.
@@
 Since models can be easily set up for which trait distributions are not Gaussian,
-the implication of this is that for such models, independence of transmission noise
+the implication of this is that for such models, independence of the Mendelian sampling term
 is not likely a good assumption.
@@
 
-The presence of a large effect polymorphic allele
-makes $R_M$ and $\bar Z_M$ dependent:
+As we discuss more later,
+the presence of a large effect polymorphic allele
+makes $R$ and $\bar Z$ dependent:
 consider the extreme case where $\eta^{[i]}_\ell$ are i.i.d.{} with variance 1 for $\ell \ge 2$,
@@
 and $Z^{[i]} = \eta_1^{[i]} + \sum_{\ell=2}^M \eta_\ell^{[i]} / \sqrt{M}$.
-Then in the limit $R_M$ depends strongly on parental trait values:
+Then in the limit $R$ depends strongly on parental trait values:
 if $Z^{[1]} \approx Z^{[2]}$ and so $\bar Z \approx \pm 10$ then most likely
@@
 and the distribution of individuals added by birth are equal,
-\ie that $\frac{Z^{1} +Z^{2}}{2} + R$ is equal in distribution
+\ie that $(Z^{1} +Z^{2})/2 + R$ is equal in distribution
 to a $\mu$-weighted choice of $Z$.
@@
 The key observation is this:
-while is meaningless to speak of variances when the transmission noise is $\alpha$-stable, if an individual's trait value is informative of the allelic effect at any given locus,
+while is meaningless to speak of variances when the Mendelian sampling term is $\alpha$-stable, if an individual's trait value is informative of the allelic effect at any given locus,
 then its trait value is informative of the \emph{distribution} of an offsprings' trait value,
@@
         \textbf{(A,D)} Segregation noise (offspring trait minus midparent) plotted against midparent.
-        \textbf{(B,E)} Histograms of transmission noise.
-        \textbf{(B,E)} Enrichment in quantiles of transmission noise conditional on midparent value.
+        \textbf{(B,E)} Histograms of the Mendelian sampling term.
+        \textbf{(B,E)} Enrichment in quantiles of the Mendelian sampling term conditional on midparent value.
         To compute the latter,
-        let $q_0, q_2, q_4, \ldots, q_{98}, q_{100}$ divide the distribution of transmission noise
+        let $q_0, q_2, q_4, \ldots, q_{98}, q_{100}$ divide the distribution of the Mendelian sampling term
         into 50 bins with roughly equal numbers in each (i.e., the quantiles),
         let $p_k(a,b)$ denote the proportion of the offspring of parents with midparent value
-        is in $[a, b)$ whose transmission noise is in $[q_{k-1}, q_k)$,
+        is in $[a, b)$ whose Mendelian sampling term is in $[q_{k-1}, q_k)$,
         and let $p_k$ be the proportion of all offspring with segregation noise in $[q_{k-1}, q_k)$
@@
 or a Cauchy process (in the Cauchy case).
-Figure~\ref{fig:seg_noise} shows how transmission noise
+Figure~\ref{fig:seg_noise} shows how the Mendelian sampling term
 depends on midparent value.
-In the Gaussian case, transmission noise is Gaussian and independent of midparent value,
+In the Gaussian case, the Mendelian noise is Gaussian and independent of midparent value,
 as shown by the horizontal lines at 1.0 in Figure~\ref{fig:seg_noise}C. 
 On the other hand, in the Cauchy case we can see from either Figure~\ref{fig:seg_noise}D or F
-that transmission noise depends strongly on midparent trait,
+that Mendelian noise depends strongly on midparent trait,
 as we would expect in the presence of several large-effect alleles.
@@
 the two models look much more similar when considering only alleles of small effect.
-Figure~\ref{fig:seg_noise_small} shows that ``midparent value''
-and ``transmission noise'' computed using only mutations with effect size less than 0.2
+Figure~\ref{fig:seg_noise_small} shows that midparent values
+and Mendelian sampling terms computed using only mutations with effect size less than 0.2
 (excluding less than 1\% of mutations)
@@
 could be interesting,
-e.g., to see how changing the distribution of transmission noise
+e.g., to see how changing the distribution of the Mendelian sampling term
 affects the rate of adaptive evolution,

### 2023-07-08 writeup.tex
Various edits.

--- 2023-07-07 writeup.tex
@@
 \DeclareMathOperator*{\sgn}{sgn}
+\DeclareMathOperator*{\re}{Re}
 
@@
 \author{ \begin{small}
-    Todd Parsons \&
+    Todd L. Parsons (LPSM, Sorbonne Universit\'e, CNRS UMR 8001, Paris, France, 75005) \&
     Peter Ralph
@@
 
-Since $\varphi$ and $\Phi$ are characteristic functions, they are absolutely continuous and $\phi(0) = 1$, and thus non-zero in a neighbourhood of 0.
+Since $\varphi$ and $\Phi$ are characteristic functions, they are absolutely continuous and $\phi(0) = \Phi(0) = 1$, and thus both are non-zero in a neighbourhood of 0.
 In particular, $\psi(\theta) = \Log{\varphi}(\theta)$ and $\Psi(\theta) = \Log{\Phi(\theta)}$ exist in some neighbourhood of 0 as well\footnote{We use $\Log$ to denote the principal value of the complex logarithm, for which the imaginary part always lies in the interval $(-\pi,\pi]$.}.
@@
 
+\paragraph{Noise Terms}
+First, consider the noise terms.
+A necessary condition for $\sum_{k=1}^{\infty} 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)$ to converge is that $2^{k} \Psi\left(\frac{\theta}{2^{k}}\right) \to 0$; since $\Psi(0) = 0$, this implies that if it exists, $\Psi'(0+) = 0$.
+In particular, as we would expect, if $R$ has a mean, it must be 0.
+If, on the other hand, $\Psi(\theta)$ is twice differentiable, so $R$ has mean 0 and finite variance, say $\Psi''(0) = \eta^{2}$.
+Taylor's theorem then tells us that for $\theta$ sufficiently small,
+\[
+	\Psi(\theta) \frac{\eta^{2}}{2} \theta^{2} + r(\theta)\theta^{2},
+\]
+where $r(\theta) \to 0$ as $\theta \to 0$.  Thus,
+\begin{equation*}
+	\frac{\textstyle 2^{k+1} \Psi\left(\frac{\theta}{2^{k+1}}\right)}{\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}
+	= \frac{1}{2} \frac{\eta^{2} \theta^{2} + {\textstyle  r\left(\frac{\theta}{2^{k+1}}\right)}\theta^{2}}{\eta^{2} \theta^{2} + {\textstyle  r\left(\frac{\theta}{2^{k}}\right)}\theta^{2}}
+	\to \frac{1}{2},
+\end{equation*}
+as $k \to \infty$ and the sum converges by the ratio test.  In this case,
+\begin{align*}
+	 \sum_{k=1}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}
+	 &=  \sum_{k=1}^{n-1} 2^{k} {\textstyle \left(- \frac{\eta^{2}}{2}  \frac{\theta^{2}}{2^{2k}}
+		+ r\left(\frac{\theta}{2^{k}}\right)\frac{\theta^{2}}{2^{2k}}\right)}\\
+	&=- \eta^{2} \theta^{2} (1 - 2^{-n+1})  + \sum_{k=1}^{n-1} {\textstyle r\left(\frac{\theta}{2^{k}}\right)\frac{\theta^{2}}{2^{k}}}
+\end{align*}
+If we assume that $r(\theta) \equiv 0$, so that the noise is $N(0,\eta^{2})$, then the limiting distribution is $N(0,2\eta^{2})$.
+Note that when  $r(\theta)$ is non-zero, we don't expect the higher order terms to vanish.
+
+It is unclear what more can be said in general.
+If however, we specialize to the case when $R$ is a stable law, we can explicitly compute the limiting noise term.
+We recall that if $\Psi(\theta)$ corresponds to a stable law (see \eg \cite{breiman1992probability}), then
+\[
+	\Psi(\theta) = im\theta + c|\theta|^{\alpha} (1-i\beta \sgn(\theta)\omega_{\alpha}(\theta))
+\]
+for $\alpha \in (0,2]$, $\beta \in [-1,1]$, $c \geq 0$, and $m \in \mathbb{R}$ and
+\[
+	\omega_{\alpha}(\theta) = \begin{cases}
+		\tan{\frac{\pi\alpha}{2}} & \text{if $\alpha \neq 1$, and}\\
+		-\frac{2}{\pi}\log{|\theta|} & \text{if $\alpha = 1$.}
+	\end{cases}
+\]
+By our general remarks above, we can assume $m = 0$ and exclude the case $\alpha = 1$, $\beta = 0$, for which  $\Psi'(0+) = c$. We then have
+\begin{align*}
+	 \sum_{k=1}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}
+	 &=  \sum_{k=1}^{n-1} 2^{(1-\alpha)k} c|\theta|^{\alpha} {\textstyle\left(1-i \beta\sgn(\theta)\left(\frac{\theta}{2^{k}}\right)
+	 	 \omega_{\alpha}\left(\frac{\theta}{2^{k}}\right)\right)}\\
+	&= \begin{cases}
+		 c|\theta|^{\alpha}\frac{1-2^{(1-\alpha)n}}{1-2^{1-\alpha}}\left(1-i\beta \sgn(\theta) \tan{\frac{\pi\alpha}{2}}\right) & \text{if $\alpha \neq 1$, and}\\	
+         (n-1) c|\theta|^{\alpha}  \left(1+\frac{2i\beta}{\pi}\sgn(\theta)\left(\log{|\theta|} - \frac{n}{2} \log{2}\right)\right) & \text{if $\alpha = 1$,}
+	\end{cases}
+\end{align*}
+which converges as $n \to \infty$ if and only if $\alpha > 1$, in which case the limit is 
+\begin{equation}\label{eq:stablenoises}
+	 \frac{c}{1-2^{1-\alpha}}|\theta|^{\alpha}\left(1-i\beta \sgn(\theta) \tan{\frac{\pi\alpha}{2}}\right), 
+\end{equation}
+which corresponds to a $\left(\alpha,\beta, \frac{c}{1-2^{1-\alpha}},0\right)$-stable law.
+
 \paragraph{Reproductive Terms:}
-First consider the limit
-\[
-		\alpha(\theta) = \lim_{n \to \infty} 2^{n}\psi\left(\frac{\theta}{2^{n}}\right).
-\]
-Because $\psi(\theta)$ is the logarithm of a characteristic function, it is continuous in some neighbourhood of, and vanishes at 0.
-If $\phi$ is not differentiable at zero (and so the corresponding distribution
-does not have a well-defined mean),
-the limit does not exist.
-On the other hand if $\psi$ is differentiable at zero,
-then $\alpha$ is continuous there,
-and for all integers $k$, 
+We now turn our attention to the limit
+\[
+		\alpha(\theta) = \lim_{n \to \infty} 2^{n}\psi\left(\frac{\theta}{2^{n}}\right)
+\]
+If it exists, then for all integers $k$, 
 \[
@@
 \]
-In particular, we see that given any continuous function $\alpha(\theta)$, $\theta \in [1,2]$, such that $\alpha(2) = 2\alpha(1)$, $\alpha$ is defined for all positive reals.
-We can extend the definition to the negative reals via the Hermitian property for $\phi(\theta)$: $\phi(-\theta) = \overline{\phi(\theta)}$ implies that $\alpha(-\theta) = \overline{\alpha(\theta)}$.
-
-Now, from the above, 
-\[
-	\frac{\alpha(\theta)}{\theta} = \frac{\alpha(2^{k} \theta)}{2^{k}\theta} \to \alpha'(0+),
-\]
-as $k \to - \infty$ if the latter exists (\ie there is a unique right-hand limit at zero; this is the case if \eg we require $\alpha(\theta)$, and thus $\phi(\theta)$, to be convex functions).
+and, since $\alpha(0) = 0$,
+\[
+	\frac{\alpha(\theta)}{\theta} = \frac{\alpha(2^{k} \theta)}{2^{k}\theta} 
+	\to \begin{cases} 
+		\alpha'(0+) & \text{if $\theta > 0$, and}\\
+		\alpha'(0-) & \text{if $\theta < 0$,}
+	\end{cases}
+\]
+as $k \to - \infty$ if the limits exist (\ie there are unique left-and right-hand limit at zero; this is the case if \eg we require $\alpha(\theta)$, and thus $\phi(\theta)$, to be convex functions).
+We henceforth assume this.
+
+Since $\varphi(\theta)$ is a characteristic function, it has the Hermitian property: $\phi(-\theta) = \overline{\phi(\theta)}$, and thus $\psi(-\theta) = \overline{\psi(\theta)}$ and $\alpha(-\theta) = \overline{\alpha(\theta)}$.
+In particular, if one (and by extension both) limits exist, we must have $\alpha'(0-) = \overline{\alpha'(0+)}$.
 In this case, setting $-\gamma + i m = \alpha'(0+)$, we have
@@
 \]
-Now, if $\phi(\theta)$ is the characteristic function of a random variable $Z$, Jensen's inequality tells us that for all $\theta$,
-\[
-	|\phi(\theta)| = \left|\E\left[e^{i\theta Z}\right]\right| \leq \E\left[\left|e^{i\theta Z}\right|\right] = 1
-\]
-so $\gamma \geq 0$.
 If $\gamma > 0$, then we recognize $e^{\alpha(\theta)} = e^{im \theta -\gamma|\theta|}$ as the characteristic function of a Cauchy distribution with median 0 and mean absolute deviation $\gamma$, whereas if $\gamma = 0$, then $e^{\alpha(\theta)} = e^{im \theta}$ is the Fourier transform of a Dirac mass at $m$, \ie the characteristic function for a monomorphic population with all individuals having trait value equal to $m$.
+In general, it does not seem possible to exclude $\gamma < 0$.
+However, recalling that if $\phi(\theta)$ is the characteristic function of a random variable $Z$, Jensen's inequality tells us that for all $\theta$,
+\begin{equation}\label{eq:CF-J}
+	|\phi(\theta)| = \left|\E\left[e^{i\theta Z}\right]\right| \leq \E\left[\left|e^{i\theta Z}\right|\right] = 1,
+\end{equation}
+we can conclude $\gamma \geq 0$ when the limiting noise term is a mixture of Gaussians and stable laws of the form \eqref{eq:stablenoises}: we then have 
+\[
+	|\phi(\theta)| = e^{\re(\alpha(\theta) + o(|\theta|))} = e^{\gamma |\theta| + o(|\theta|))},
+\] 
+and thus is less than or equal to one in a small neighbourhood of $0$ if and only if $\gamma < 0$.
 
@@
 % One can easily choose $\alpha(\theta)$ so that $e^{\alpha(\theta)}$ satisfies the first three criteria, \eg $\alpha(\theta) = -|\theta|\left(\sin\left(\pi\frac{\log{|\theta|}}{\log{2}}\right)+\frac{3}{2}\right)$, but excluding such functions via Bochner's positive definite criterion is non-trivial.
-
-\paragraph{Noise Terms}
-We now turn our attention to the noise terms.
-A necessary condition for $\sum_{k=1}^{\infty} 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)$ to converge is that $2^{k} \Psi\left(\frac{\theta}{2^{k}}\right) \to 0$; since $\Psi(0) = 0$, this implies that if it exists, $\Psi'(0+) = 0$.
-In particular, as we would expect, if $R$ has a mean, it must be 0.
-If, on the other hand, $\Psi(\theta)$ is twice differentiable, so $R$ has mean 0 and finite variance, say $\Psi''(0) = \eta^{2}$.
-Taylor's theorem then tells us that for $\theta$ sufficiently small,
-\[
-	\Psi(\theta) \frac{\eta^{2}}{2} \theta^{2} + r(\theta)\theta^{2},
-\]
-where $r(\theta) \to 0$ as $\theta \to 0$.  Thus,
-\begin{equation*}
-	\frac{\textstyle 2^{k+1} \Psi\left(\frac{\theta}{2^{k+1}}\right)}{\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}
-	= \frac{1}{2} \frac{\eta^{2} \theta^{2} + {\textstyle  r\left(\frac{\theta}{2^{k+1}}\right)}\theta^{2}}{\eta^{2} \theta^{2} + {\textstyle  r\left(\frac{\theta}{2^{k}}\right)}\theta^{2}}
-	\to \frac{1}{2},
-\end{equation*}
-as $k \to \infty$ and the sum converges by the ratio test.  In this case,
-\begin{align*}
-	 \sum_{k=1}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}
-	 &=  \sum_{k=1}^{n-1} 2^{k} {\textstyle \left(- \frac{\eta^{2}}{2}  \frac{\theta^{2}}{2^{2k}}
-		+ r\left(\frac{\theta}{2^{k}}\right)\frac{\theta^{2}}{2^{2k}}\right)}\\
-	&=- \eta^{2} \theta^{2} (1 - 2^{-n+1})  + \sum_{k=1}^{n-1} {\textstyle r\left(\frac{\theta}{2^{k}}\right)\frac{\theta^{2}}{2^{k}}}
-\end{align*}
-If we assume that $r(\theta) \equiv 0$, so that the noise is $N(0,\eta^{2})$, then the limiting distribution is $N(0,2\eta^{2})$.
-Note that when  $r(\theta)$ is non-zero, we don't expect the higher order terms to vanish.
-
-It is unclear what more can be said in general.
-If however, we specialize to the case when $R$ is a stable law, we can explicitly compute the limiting noise term.
-We recall that if $\Psi(\theta)$ corresponds to a stable law (see \eg \cite{breiman1992probability}), then
-\[
-	\Psi(\theta) = im\theta + c|\theta|^{\alpha} (1-i\beta \sgn(\theta)\omega_{\alpha}(\theta))
-\]
-for $\alpha \in (0,2]$, $\beta \in [-1,1]$, $c \geq 0$, and $m \in \mathbb{R}$ and
-\[
-	\omega_{\alpha}(\theta) = \begin{cases}
-		\tan{\frac{\pi\alpha}{2}} & \text{if $\alpha \neq 1$, and}\\
-		-\frac{2}{\pi}\log{|\theta|} & \text{if $\alpha = 1$.}
-	\end{cases}
-\]
-By our general remarks above, we can assume $m = 0$ and exclude the case $\alpha = 1$, $\beta = 0$, for which  $\Psi'(0+) = c$. We then have
-\begin{align*}
-	 \sum_{k=1}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}
-	 &=  \sum_{k=1}^{n-1} 2^{(1-\alpha)k} c|\theta|^{\alpha} {\textstyle\left(1-i \beta\sgn(\theta)\left(\frac{\theta}{2^{k}}\right)
-	 	 \omega_{\alpha}\left(\frac{\theta}{2^{k}}\right)\right)}\\
-	&= \begin{cases}
-		 c|\theta|^{\alpha}\frac{1-2^{(1-\alpha)n}}{1-2^{1-\alpha}}\left(1-i\beta \sgn(\theta) \tan{\frac{\pi\alpha}{2}}\right) & \text{if $\alpha \neq 1$, and}\\	
-         (n-1) c|\theta|^{\alpha}  \left(1+\frac{2i\beta}{\pi}\sgn(\theta)\left(\log{|\theta|} - \frac{n}{2} \log{2}\right)\right) & \text{if $\alpha = 1$,}
-	\end{cases}
-\end{align*}
-which converges as $n \to \infty$ if and only if $\alpha > 1$, in which case the limit is 
-\[
-	 \frac{c}{1-2^{1-\alpha}}|\theta|^{\alpha}\left(1-i\beta \sgn(\theta) \tan{\frac{\pi\alpha}{2}}\right), 
-\]
-which corresponds to a $\left(\alpha,\beta, \frac{c}{1-2^{1-\alpha}},0\right)$-stable law.
 
@@
         As in Figure~\ref{fig:seg_noise},
-        except computed using only mutations with absolute value effect size less than 0.2.
+        except computed using only mutations with absolute value effect size less than 0.1.
         The same quantities for the simulations of Figure~\ref{fig:sel_seg_noise}
@@
 of the animal model and genomic selection.
+This project was started on a visit by TLP to the University of Oregon that was partially supported by the CNRS PEPS grant ``Jeunes Chercheuses et Jeunes Chercheurs''.
 
@@
         As in Figure~\ref{fig:seg_noise},
-        except computed using only mutations with absolute value effect size less than 0.2.
+        except computed using only mutations with absolute value effect size less than 0.1.
         \label{fig:sel_seg_noise_small}

### 2023-07-08 writeup.tex
.

@@
 and for the purpose of investigating this with simulations,
-some calculations are helpful.
+some calculations will be helpful.
 
 Concretely, consider a  Moran process where an individual whose trait value is $Z$ dies at rate $\mu(Z)$
-and is replaced by a the offspring of two randomly chosen individuals,
-i.e., a new individual of phenotype
+and is replaced by the offspring of two randomly chosen individuals,
+i.e., a new individual with trait
 \begin{align*}
@@
 and $R$ is an independent draw from the Mendelian sampling distribution
-(independent of $Z^{1}$ and $Z^{2}$). 
+(independent of $Z^{1}$ and $Z^{2}$, with distribution to be specified). 
 Mutation-selection balance demands that that
@@
 \begin{align} \label{eqn:MSB_CF}
-    \E\left[ e^{i\theta\left(\frac{Z^{1} +Z^{2}}{2} + R \right)}\right] = \frac{\E\left[ e^{i \theta Z}\mu(Z) \right]}{\E[\mu(Z)]}.
+    \frac{\E\left[ e^{i \theta Z}\mu(Z) \right]}{\E[\mu(Z)]}
+    =
+    \E\left[ e^{i\theta\left(\frac{Z^{1} +Z^{2}}{2} + R \right)}\right] .
 \end{align}
@@
 drawn from the ``effect size distribution'' (and will be either Gaussian or Cauchy).
-The phenotype of each individual is determined additively
+The trait of each individual is determined additively
 by summing the effects of all mutations they carry across both chromosomes.
-Let $\mu(z)$ be the death rate of an individual with phenotype $z$,
+Let $\mu(z)$ be the death rate of an individual with trait $z$,
 and let $dt$ be a small constant (we took $dt=0.01$).

### 2023-07-08 sims/make_paper_figs.R
.

@@ -81,7 +81,6 @@ plot_seg_noise <- function (examples, labels, seg_col, main_append="", ...) {
     }
 }
 
-
 ###########
 # Neutral examples
 examples <- list(
@@ -166,11 +165,10 @@ png(file=fname, width=6.5, height=6, pointsize=10, units='in', res=288)
 plot_seg_noise(examples, labels, seg_col=1)
 dev.off()
 
-
 # 5. with stable(3/2)
 examples <- list(
                  "neutral Normal" = "sim_neutral_normal_8135",
-                 "neutral Stable(3/2)" = "sim_neutral_stable_8135",
+                 "neutral Stable(3/2)" = "sim_neutral_stable_8136",
                  "neutral Cauchy" = "sim_neutral_cauchy_8135"
 )
 

### 2023-07-08 sims/make_paper_figs.sh

@@ -10,3 +10,5 @@ $SLIM -s 707 -d BURNIN=20000 -d NUMGENS=1000 -d N=1000 -d STEP=20 -d 'EPSILON=c(
 
 $SLIM -s 8135 -d BURNIN=40000 -d NUMGENS=1000 -d N=10000 -d 'EPSILON=c(0,0.2,1.0,4.0,10)' neutral_normal.slim
 $SLIM -s 8135 -d BURNIN=40000 -d NUMGENS=1000 -d N=10000 -d 'EPSILON=c(0,0.2,1.0,4.0,10)' neutral_cauchy.slim
+
+$SLIM -s 8136 -d BURNIN=40000 -d NUMGENS=1000 -d N=10000 -d 'EPSILON=c(0,0.2,1.0,4.0,10)' neutral_stable.slim

### 2023-07-08 writeup.tex
.

@@
 
-Since $\varphi$ and $\Phi$ are characteristic functions, they are absolutely continuous and $\phi(0) = \Phi(0) = 1$, and thus both are non-zero in a neighbourhood of 0.
+Since $\varphi$ and $\Phi$ are characteristic functions, they are continuous and $\phi(0) = \Phi(0) = 1$, and thus both are non-zero in a neighbourhood of 0.
 In particular, $\psi(\theta) = \Log{\varphi}(\theta)$ and $\Psi(\theta) = \Log{\Phi(\theta)}$ exist in some neighbourhood of 0 as well\footnote{We use $\Log$ to denote the principal value of the complex logarithm, for which the imaginary part always lies in the interval $(-\pi,\pi]$.}.
@@
 In particular, as we would expect, if $R$ has a mean, it must be 0.
-If, on the other hand, $\Psi(\theta)$ is twice differentiable, so $R$ has mean 0 and finite variance, say $\Psi''(0) = \eta^{2}$.
-Taylor's theorem then tells us that for $\theta$ sufficiently small,
-\[
-	\Psi(\theta) \frac{\eta^{2}}{2} \theta^{2} + r(\theta)\theta^{2},
+If, on the other hand, $\Psi(\theta)$ is twice differentiable, so $R$ has mean 0 and finite variance, say $\Psi''(0) = \eta^{2}$, then Taylor's theorem tells us that for $\theta$ sufficiently small,
+\[
+	\Psi(\theta) = \frac{\eta^{2}}{2} \theta^{2} + r(\theta)\theta^{2},
 \]
@@
 If we assume that $r(\theta) \equiv 0$, so that the noise is $N(0,\eta^{2})$, then the limiting distribution is $N(0,2\eta^{2})$.
-Note that when  $r(\theta)$ is non-zero, we don't expect the higher order terms to vanish.
+When  $r(\theta)$ is non-zero, we don't expect the higher order terms to vanish.
 
@@
 \]
-as $k \to - \infty$ if the limits exist (\ie there are unique left-and right-hand limit at zero; this is the case if \eg we require $\alpha(\theta)$, and thus $\phi(\theta)$, to be convex functions).
+as $k \to - \infty$ if the limits exist (\ie there are unique left-and right-hand limits at zero; this is the case if \eg we require $\alpha(\theta)$, and thus $\phi(\theta)$, to be convex functions).
 We henceforth assume this.
@@
 \]
-If $\gamma > 0$, then we recognize $e^{\alpha(\theta)} = e^{im \theta -\gamma|\theta|}$ as the characteristic function of a Cauchy distribution with median 0 and mean absolute deviation $\gamma$, whereas if $\gamma = 0$, then $e^{\alpha(\theta)} = e^{im \theta}$ is the Fourier transform of a Dirac mass at $m$, \ie the characteristic function for a monomorphic population with all individuals having trait value equal to $m$.
-In general, it does not seem possible to exclude $\gamma < 0$.
+If $\gamma > 0$, then we recognize $e^{\alpha(\theta)} = e^{im \theta -\gamma|\theta|}$ as the characteristic function of a Cauchy distribution with median 0 and median absolute deviation $\gamma$, whereas if $\gamma = 0$, then $e^{\alpha(\theta)} = e^{im \theta}$ is the Fourier transform of a Dirac mass at $m$, \ie the characteristic function for a monomorphic population with all individuals having trait value equal to $m$.
+(And, $\gamma < 0$ does not give a valid characteristic function.)
+% % I think we've shown this characteristic function has to be exp(im\theta - \gamma|\theta|);
+% % if $\gamma < 0$ then it is not a characteristic function; am I missing something?
+% In general, it does not seem possible to exclude $\gamma < 0$.
 However, recalling that if $\phi(\theta)$ is the characteristic function of a random variable $Z$, Jensen's inequality tells us that for all $\theta$,
-\begin{equation}\label{eq:CF-J}
-	|\phi(\theta)| = \left|\E\left[e^{i\theta Z}\right]\right| \leq \E\left[\left|e^{i\theta Z}\right|\right] = 1,
-\end{equation}
-we can conclude $\gamma \geq 0$ when the limiting noise term is a mixture of Gaussians and stable laws of the form \eqref{eq:stablenoises}: we then have 
-\[
-	|\phi(\theta)| = e^{\re(\alpha(\theta) + o(|\theta|))} = e^{\gamma |\theta| + o(|\theta|))},
-\] 
-and thus is less than or equal to one in a small neighbourhood of $0$ if and only if $\gamma < 0$.
-
-\paragraph{Lamarkian inheritance redux.} 
-Note that the characteristic functions constructed above are solutions to \eqref{eq:CF_neutral} when transmission is noiseless, and the offspring have exactly the parental mean trait -- \ie Lamarkian blending inheritance.
-The solution $\phi(\theta) = e^{i\theta m}$ is exactly the reversion to the mean that Darwin's detractors used to argue against the theory of evolution \citep{Provine1971}.
-The solution $\phi(\theta) = e^{im \theta -\gamma|\theta|}$ offers an amusing counter-factual: if a Lamarkian population could somehow achieved a Cauchy-distributed trait distribution, it would maintain that diversity indefinitely.
+% \begin{equation}\label{eq:CF-J}
+% 	|\phi(\theta)| = \left|\E\left[e^{i\theta Z}\right]\right| \leq \E\left[\left|e^{i\theta Z}\right|\right] = 1,
+% \end{equation}
+% we can conclude $\gamma \geq 0$ when the limiting noise term is a mixture of Gaussians and stable laws of the form \eqref{eq:stablenoises}: we then have 
+% \[
+% 	|\phi(\theta)| = e^{\re(\alpha(\theta) + o(|\theta|))} = e^{\gamma |\theta| + o(|\theta|))},
+% \] 
+% and thus is less than or equal to one in a small neighbourhood of $0$ if and only if $\gamma < 0$.
+
+\paragraph{Blending inheritance redux.} 
+Note that the characteristic functions constructed above are solutions to \eqref{eq:CF_neutral} when transmission is noiseless, and the offspring have exactly the parental mean trait -- \ie blending inheritance.
+The solution $\phi(\theta) = e^{i\theta m}$ is exactly the reversion to the mean that Darwin's detractors used to argue against the theory of evolution \citep{jenkin1867,Provine1971}.
+The solution $\phi(\theta) = e^{im \theta -\gamma|\theta|}$ offers an amusing counter-factual: if a Lamarkian population could somehow achieve a Cauchy-distributed trait distribution, it would maintain that diversity indefinitely.
 
@@
 \paragraph{Gaussian Solution:}
-First, consider the case when $\mu(x) = e^{a x^{2}}$.   In this case there is a Gaussian solution: suppose $Z$ is Gaussian with mean $m$ and variance $\sigma^{2}$.  Then,
+First, consider the case when $\mu(x) = e^{a x^{2}}$.   In this case there is a Gaussian solution: suppose $Z$ is Gaussian with mean $m$ and variance $\sigma^{2}$, so
 $ \phi(\theta) = \exp(i m \theta - \frac{1}{2} \sigma^{2} \theta^{2}).  $
-Moreover, provided $a < \frac{1}{2\sigma^{2}}$, we can explicitly compute $\frac{\mathbb{E}\left[e^{i \theta Z}\mu(Z)\right]}{\mathbb{E}[\mu(Z)]}$  using the Gaussian probability density function to conclude that
+Moreover, provided $a < \frac{1}{2\sigma^{2}}$, we can explicitly compute $\frac{\mathbb{E}\left[e^{i \theta Z}\mu(Z)\right]}{\mathbb{E}[\mu(Z)]}$: %  using the Gaussian probability density function to conclude that
 if $R$ is Gaussian with mean zero and variance $\eta^2$, then equation~\eqref{eqn:MSB_CF} becomes
@@
 
-Note that we do not require very weak selection to guarantee a Gaussian steady state, only that $a < \frac{1}{2\sigma^{2}}$.
+Note that we do not require very weak selection to guarantee a Gaussian steady state, only that $a < \frac{1}{4\eta^{2}}$.
 However, this does not imply that the transient distribution is also Gaussian, as is sometimes assumed \citep[e.g.,][]{lande1976natural}.
@@
 (a) in principle,
-a mutation-selection balance is possible with a Cauchy empirical distribution and Cauchy Mendelian sampling term,
+a mutation-selection balance is possible with a Cauchy empirical distribution and Cauchy Mendelian noise,
 but (b)
-such a model cannot derive from a simple model with additive loci.
+such a model cannot derive from a simple model of additive loci.
 To that end, it's instructive to see where the argument of \cite{barton2017infinitesimal} breaks down
-if instead of assuming bounded allele effects of order $\Oh\left(1/\sqrt{M}\right)$,
-we assume that the allele at locus $i$ in individual $j$ contributes
+if instead of assuming bounded allele effects of order $\Oh(1/\sqrt{M})$,
+we assume that the allele at locus $\ell$ in individual $j$ contributes
 $\eta^{[j]}_{\ell}/M$, where the $\eta^{[j]}_{\ell}$ are independent $\Cauchy(0,\gamma_{\ell})$ random variables. 
@@
 
-The key observation is this:
-while is meaningless to speak of variances when the Mendelian sampling term is $\alpha$-stable, if an individual's trait value is informative of the allelic effect at any given locus,
-then its trait value is informative of the \emph{distribution} of an offsprings' trait value,
-and hence there can be no independence between parental traits and the Mendelian sampling term
-that is central to the tractability of the infinitesimal model.
+% The key observation is this:
+% while is meaningless to speak of variances when the Mendelian sampling term is $\alpha$-stable, if an individual's trait value is informative of the allelic effect at any given locus,
+% then its trait value is informative of the \emph{distribution} of an offsprings' trait value,
+% and hence there can be no independence between parental traits and the Mendelian sampling term
+% that is central to the tractability of the infinitesimal model.
 
@@
 \end{align*}
-% % BEGIN PREVIOUS ARGUMENT
-% Now, as $z \to 0$, $\frac{\gamma_{\ell}(\bar{\gamma}^{2}-z^{2})+2zy}{\gamma_{\ell}(\bar{\gamma}^{2}+z^{2})} \to \frac{\gamma_{\ell}}{\bar{\gamma}}$ and as $z \to \pm \infty$, $\frac{\gamma_{\ell}(\bar{\gamma}^{2}-z^{2})+2zy}{\gamma_{\ell}(\bar{\gamma}^{2}+z^{2})} \to -\frac{\gamma_{\ell}}{\bar{\gamma}}$,  whereas $|\frac{\bar{\gamma}^{2}-z^{2}+2zy}{\bar{\gamma}^{2}+z^{2}}|$ attains a maximum of
+% Now, as $z \to 0$,
+% $\frac{\gamma_{\ell}(\bar{\gamma}^{2}-z^{2})+2zy}{\gamma_{\ell}(\bar{\gamma}^{2}+z^{2})} \to \frac{\gamma_{\ell}}{\bar{\gamma}}$
+% and as $z \to \pm \infty$,
+% $\frac{\gamma_{\ell}(\bar{\gamma}^{2}-z^{2})+2zy}{\gamma_{\ell}(\bar{\gamma}^{2}+z^{2})} \to -\frac{\gamma_{\ell}}{\bar{\gamma}}$,
+% \comment{I'm not sure where $\gamma_\ell$ went here}
+% whereas $|\frac{\bar{\gamma}^{2}-z^{2}+2zy}{\bar{\gamma}^{2}+z^{2}}|$ 
 % \[
@@
 % which is itself asymptotically equal to $\mp \sgn(y) \bar{\gamma}$ for large values of $|y|$.
-% 	
 we have that
@@
     \right|
-    \le \left(\frac{|y|}{\gamma_\ell \bar{\gamma} M} + \Oh\left(\frac{1}{M^2}\right)\right)  \P\{\eta^{[j]}_{\ell} = y\}.
+    \le \left(\left(1 + \frac{|y|}{\gamma_\ell \bar{\gamma}}\right) \frac{1}{M} + \Oh\left(\frac{1}{M^2}\right)\right)  \P\{\eta^{[j]}_{\ell} = y\}.
 \]
@@
     \caption{
-        Segregation noise with Gaussian \textbf{(A-C)} and Cauchy \textbf{(D-F)} effect size distributions,
+        The distribution of Mendelian sampling terms
+        with Gaussian \textbf{(A-C)} and Cauchy \textbf{(D-F)} effect size distributions,
         from the same simulations shown in Figure~\ref{fig:trait_distrns},
         for all individuals born during the last 1,000 time units of the simulation.
-        \textbf{(A,D)} Segregation noise (offspring trait minus midparent) plotted against midparent.
+        \textbf{(A,D)} Mendelian noise (offspring trait minus midparent) plotted against midparent.
         \textbf{(B,E)} Histograms of the Mendelian sampling term.
@@
         is in $[a, b)$ whose Mendelian sampling term is in $[q_{k-1}, q_k)$,
-        and let $p_k$ be the proportion of all offspring with segregation noise in $[q_{k-1}, q_k)$
+        and let $p_k$ be the proportion of all offspring with Mendelian noise in $[q_{k-1}, q_k)$
         (so, $p_k \approx 0.02$).

### 2023-07-08 writeup.tex
peter edits

@@
 and $\mu(z) = e^{\alpha z^2}$; and
-(2) the effect size distribution was Cauchy, centered at zero and with width $\delta - \gamma$
+(2) the effect size distribution was Cauchy, centered at zero and with median absolute deviation $\delta - \gamma$
 and $\mu(z) = (\delta/\gamma) (\gamma + z^2) / (\delta + z^2)$ with $\delta=0.5$ and $\gamma=0.2$.
@@
 by combining both effects of known large effect loci
-and calculations using a relatedness matrix
+and cumulative effects of many unknown loci as mediated by a relatedness matrix
 \citep[e.g.,][]{fernando1989marker,teissier2018weighted,bernardo2014genomewide,rice2019evaluation}.
@@
 However,
-we have seen that the assumption of the infinitesimal model
+we have seen that the assumption of independence in the infinitesimal model
 implies a sort of ``blending inheritance''
@@
 to see if the predictions are borne out under concrete models of genetic inheritance.
-Another open question is whether there is some set of assumptions
-(perhaps involving linkage, pleiotropy, and/or environmental interactions with genetic effects)
+Another open question is whether there is some set of realistic assumptions
+(perhaps involving epigenetic effects, pleiotropy, and/or environmental interactions with genetic effects)
 which leads to these non-Gaussian infinitesimal models.
@@
 
-Finally: above we've most often used the Cauchy distribution
+Finally, we note that above we've most often used the Cauchy distribution
 not because it seems most realistic
 (indeed data suggests the L\'evy stable distribution with $\alpha=3/2$ might be better)
-but for convenience and to make effects more obvious.
+but for mathematical convenience.
 

### 2023-11-15 review-responses.tex
reviews put in latex

%%%%%%
%%
%%  Don't reorder the reviewer points; that'll mess up the automatic referencing!
%%
%%%%%

\begin{minipage}[b]{2.5in}
  Resubmission Cover Letter \\
  {\it Theoretical Population Biology}
\end{minipage}
\hfill
\begin{minipage}[b]{2.5in}
    Todd Parsons \\
    \emph{and} Peter Ralph \\
  \today
\end{minipage}

\vskip 2em

\noindent
{\bf To the Editor(s) -- }

\vskip 1em

We are writing to submit a revision of our manuscript, ETCETERA

\noindent \hspace{4em}
\begin{minipage}{3in}
\noindent
{\bf Sincerely,}

\vskip 2em

{\bf 
Todd Parsons and
Peter Ralph
}\\
\end{minipage}

\vskip 4em

\pagebreak

%%%%%%%%%%%%%%
\section*{AE comments:}

\begin{quote}
The two reviewers find the results interesting and worth being eventually published in the special issue. However, they highlight a number of places in the text where the properties claimed or the precise questions addressed need serious clarification. As the readership of the special issue will be quite broad, it is important that the main messages are accessible to colleagues who do not necessarily know Fisher's infinitesimal model precisely. Therefore, although the mathematical results in themselves do not require deep modifications, I recommend that a major revision of the paper should be provided, which would take into account the different comments and suggestions of the reviewers.
\end{quote}

XXX

%%%%%%%%%%%%%%
\reviewersection{1}

\begin{quote}
The paper explores inheritance of an additive trait influenced by a large number of loci with effect sizes drawn from a power law distribution, also fitting effect size distributions as estimated by GWAS (for several human disease traits) to estimate plausible power law exponents. The study is motivated by Barton et al's 2017 paper which shows how in the limit of a very large number of loci with effect sizes drawn from a distribution with finite variance, inheritance is described by the the infinitesimal model: in particular, that the distribution of trait values of the offspring of two parents is multivariate normal with a variance-covariance matrix that is independent of parental trait values. This paper explores whether some analogous statements can be made if the underlying effect size distribution does not have finite variance- whether, loosely speaking, some ``generalisation'' of the infinitesimal model might still apply. Below I summarise some of my concerns about the paper.
\end{quote}

Thanks for the careful reading...

\begin{point}{}
    Paragraph starting ``This suggests exploring whether stable distributions..'' \revref:
The questions as stated: e.g., whether stable distributions can ``stand in'' for a Gaussian distribution in the infinitesimal model or whether the results of Barton et al ``carry over'' when the sum of effect sizes follows a stable law are quite vague and allow for multiple interpretations. Ideally, in order to address whether Barton et al's results (do not) carry over, one would like to at least show that with a power law distribution of effect sizes, the ``Mendelian noise term'', i.e., the difference between the offspring trait value and the midparent value, is (not) independent of the midparent value in the limit of a very large number of loci. And indeed this is what the authors address
somewhat non-rigorously (and for the case of a Cauchy distribution of effect sizes) in section 5 and using simulations : see e.g., figures 4C and 4f, figures 6C and 6F. Unsurprisingly, these simulations show that with e.g., Cauchy distribution of effect sizes, the non-independence between the segregation/noise term and the midparent trait value does not hold. However, a lot of the mathematical analysis in the paper (sections 3 and 4) explores when the distribution of trait values in a population is (not) Gaussian, which is a completely different question.
\end{point}

\reply{
}

\begin{point}{\revref}
In the same paragraph: I don't think the claim that ``independence of offsprings' deviations and midparent values implies a Gaussian distribution'' (which is also stated in the abstract) is true? Surely, ``independence of offsprings' deviations and midparent values'' is just the basic statement of the infinitesimal model (say, in Barton et al 2017) and the infinitesimal model does not imply a Gaussian distribution of trait values? See also section 3.
\end{point}

\reply{
}

\begin{point}{}
    Section 2 and figure \ref{fig:example_snps}: It is not obvious to me why one would study the ``frequency-weighted'' effect sizes. There is an assumption here I think that allele frequencies are uncorrelated with effect sizes: this will not be true if for example traits are under some kind of stabilising selection: see e.g., Simons et al (A Population Genetic Interpretation of GWAS Findings for Human Quantitative Traits) for a detailed exploration of this. Moreover, if SNPs are not truly causal but only tag causal alleles, then weighting by allele frequencies can be even more problematic?

As an alternative, would it make sense to look at what fraction of loci have effect size difference between alternative alleles to be $>2t$ without weighing by allele frequency?
More broadly, how do we interpret the frequency-weighted vs. unweighted distributions?
\end{point}

\reply{
}

\begin{point}{}
I was also wondering about the ``goodness of fit'' of observed distributions to power laws. I am not necessarily recommending a detailed quantitative analysis, but nevertheless it may be useful to comment on whether one has power to say reject alternative distributions (e.g., a mixture of exponentials) that are often considered. If not, then perhaps the exact exponents inferred are unimportant, and what the data points to more broadly, is a broad distribution of effect sizes.
\end{point}

\reply{
}

\begin{point}{}
    Section 3, Sentence starting ``We know from Barton et al (2017)...'' \revref: As far as I understand, the statement of Barton et al is that the joint distribution of trait values of offspring is multivaraite normal, conditioned on the two parents. I don't think the joint distribution of two randomly picked parents and their offspring is mutivariate normal, as this would require some kind of normality assumption about trait values in the population as a whole. It'd be helpful to distinguish more carefully between conditioned and unconditioned distributions, especially in this section.
\end{point}

\reply{
}

\begin{point}{}
    In general, I find this section rather confused: the infinitesimal model basically implies that the ``Mendelian segregation term'' $R_{M}$ (which is the difference between an individual and the midparent value $\bar{Z}_{M}$) is normally distributed with mean zero and variance that depends on the relatedness between parents but is independent of the midparent value $\bar{Z}_{M}$. However, in Proposition \ref{prop:parental_contribs}, the authors claim that $\bar{Z}_{M}$ and $R_{M}$ are jointly Gaussian (which is equivalent to saying that the trait value of an individual and the mean of its parents' trait values is jointly Gaussian), which is in general not true.
\end{point}

\reply{
}

\begin{point}{}
    I think it is also not correct to claim that ``Since models can be easily set up for which trait distributions are not Gaussian, the implication of this is that for such models, independence of the Mendelian sampling term is not likely a good assumption.'' \revref. As far as I understand, a non-Gaussian distribution of trait values can arise quite easily under the infinitesimal model (where the Mendelian segregation term is independent of the midparent value) e.g, due to strong or non-Gaussian selection or migration or some combination of the two (see, e.g., Figure 1 in Barton et al).
\end{point}

\reply{
}

\begin{point}{}
    Section 4, first paragraph \revref: I don't think the third assumption (that trait values in the population follow a Gaussian distribution) is a component of the ``infinitesimal model''. I think Turelli (2017) is also quite clear about this stating that a Gaussian distribution of trait values should only emerge for Gaussian or very weak selection on the population.
\end{point}

\reply{
}

\begin{point}{}
More generally, I found it a bit hard to follow the overall logical flow of section 4.1, especially once the authors launch into an exploration of the ``reproduction'' and ``noise'' terms (pages 9 and 10).
    It would be useful to clarify what the biological/intuitive meaning of the noise and reproduction terms at the outset. It would also be useful to state (at the beginning) what the main goal of these explorations is: (is the goal to identify when both terms ``have well-posed limits independent of the other'' as stated at \llname{limit_remark}), and also to summarise (at the end) to what extent one can do this in all generality.
\end{point}

\reply{
}

\begin{point}{}
    Last inequality on Page 13: Is this the distribution of the largest of the $M$ alleles carried by an individual, conditioned on the trait value $Z$ of the individual (and this is independent of $Z$?), or is it just the largest of $M$ iid draws from a Cauchy distribution? If the former, then it'd be good to say this explicitly. If the latter, then this does not quite answer the question posed at the beginning of the section about how much information knowing the trait value of an individual gives about underlying allelic effects.
\end{point}

\reply{
}

\begin{point}{}
Also, maybe worth specifying what the corresponding distribution for the largest of $M$ alleles looks like when the effect size distribution has finite variance, in order to highlight the contrast between ``well-behaved'' and heavy-tailed effect size distributions.
\end{point}

\reply{
}

\subsubsection{Minor comments:}

\begin{point}{}
    Introduction: Also worth citing Fisher, Bulmer etc. (the original references) when introducing the infinitesimal model?
\end{point}

\reply{
}

\begin{point}{\revref}
    ``Perhaps the distribution of effect sizes within each gene is Normal'': I am confused by this. Do you mean that the distribution of effect sizes across all genes (with roughly the same ``proximity'' to the trait) is normal. If yes, then the above phrasing is a bit misleading. If what is meant that is really that the effect sizes of different alleles within a gene are normally distributed, then I am not sure if this is entirely plausible: see various papers by Turelli (e.g., ``Heritable Genetic Variation via Mutation-Selection Balance: Lerch's Zeta Meets the Abdominal Bristle.''). Presumably, a normal distribution of effect sizes for a sufficiently large genomic region is a good approximation but it is unclear what is large enough: smaller or larger than a typical gene...?
\end{point}

\reply{
}

\begin{point}{\revref}
    Sentence starting ``In fact, many traits claimed...'': This phrasing here is rather cryptic and I do not understand what is being said. Maybe rephrase? Also a reference would help...
\end{point}

\reply{
}

\begin{point}{}
    Figure \ref{fig:exponent_hist}: Is there also a correlation between the number of cases and the number of SNPs? In other words, to what extent are (b) and (c) independent?
\end{point}

\reply{
}

\begin{point}{\revref}
    Sentence starting ``To see what knowing the parental trait value...'': In what sense is this the ``parental'' trait value??
\end{point}

\reply{
}

\begin{point}{}
    Last inequality on Page 13: should the subscript be $j$ instead of $i$?
\end{point}

\reply{
}

\begin{point}{}
    Simulations: Why not choose the same effect size distributions in the neutral case and the case with stabilizing selection? Alternatively, if different effect size distributions are used, it may be good to plot the distribution of (appropriately) scaled trait values. Right now the figures give the impression that more variation is maintained under stabilizing selection..
\end{point}

\reply{
}

\begin{point}{}
    Figure \ref{fig:seg_noise} caption: I find the phrasing ``(a,b) chosen to span the 5\% of midparents centered on the 10\% quantile of midparent value'' a bit confusing. Does this mean a and b are respectively the 7.5\% and 12.5\% quantile of the distribution of midparent trait values?
\end{point}

\reply{
}

\begin{point}{}
    Discussion: In the context of theory combining large and small-effect loci, the 2008 paper by Chevin and Hospital: ``Selective Sweep at a Quantitative Trait Locus in the Presence of Background Genetic Variation'' and the 1983 work by Lande: ``The response to selection on major and minor mutations affecting a metrical trait'' may also be somewhat relevant.
\end{point}

\reply{
}

%%%%%%%%%%%%%%
\reviewersection{2}

\begin{quote}
The authors investigate the possibility of extending results of Barton et al (2017) on the
justification of the infinitesimal model to distributions of effect sizes with a large tail (i.e.,
polynomial decay with exponent $\alpha < 2$). This is biologically interesting and important
because there are clear indications that variation in some traits is caused by a mixture
of polygenic (infinitesimal) contributions and one or few loci of large effect. It also a
mathematically difficult topic because it leads into the realm of generalized Central Limit
Theorems, namely stable laws.

After a brief, partly rough (see detailed comments below) introduction the ms starts with
a statistical analysis of the effect size distributions of a large number of traits (all human
disease related) for which GWAS data are available. The evaluation shows that almost
all estimated tail exponents are between 1 and 2.5 with a quite flat distribution between
about 1.3 and 2 (some potential statistical artifacts are noted). However, the authors do
not discuss the influence of selection, which could be essential because the traits are disease
related.
\end{quote}

We now discuss selection XXX WHERE

\begin{quote}
In Section 3, the authors show a kind of negative result concerning the possibility of $\alpha$-
stable distributions with $\alpha < 2$ (Proposition 1): If the deviations of the offspring trait
from the midparent value (here called the Mendelian sampling term, $R$) are independent
of the midparent value $\bar Z$ (part of the core result in Barton et al (2017)), then the joint
distribution of $R$ and $\bar Z$ is Gaussian. Thus, this independence is not compatible with ‘$\alpha$-
stable laws’. In fact, as they show (by example) and discuss more generally in Sect. 4, a
single polymorphic large-effect locus causes dependence of $R$ and $\bar Z$.

In Section 4, a version of the infinitesimal model is investigated that proposes a much
stronger result than a Gaussian Mendelian sampling term independent of the parental
trait distribution, namely a Gaussian distribution of the trait (within the population)
under mutation-selection balance. Although such a model has occasionally been assumed
in the literature, it has never been justified rigorously and, presumably, is not justifiable
(on the basis of realistic assumptions – as clearly noted by Turelli’s brief review and
previous extensive work by Turelli, Barton, and others). Nevertheless, the analysis of the
authors is interesting from a theoretical point of view because they investigate whether
such deterministic mutation-selection balance is possible in a non-Gaussian model that
retains independence of the Mendelian sampling terms, and if such a model might yield a
reasonable approximation to an additive model with non-Gaussian effects.
\end{quote}

We don't actually assume a Gaussian distribution....

\begin{quote}
The authors use a simplistic Moran model to investigate these questions. However, mu-
tation doesn’t seem to be modeled explicitly. Is it mutation if a parent is replaced by an
offspring of the type described in the display equation on p 7? Maybe yes, but not neces-
sarily because of segregation and recombination contribute substantially to $R$. Doesn’t the
present assumption imply that the mutation rate is one and that mutation is modeled as a
kind of diffusion? In the latter case, a Gaussian distribution can result. With explicit mu-
tation models, where mutations occur per locus and a finite number of loci determines the
trait additively, a Gaussian distribution never results (and in no known limit; see above).
How does population size enter? Is this a deterministic Moran model? Anyway, the results
in Sect 4 are of interest because they strongly constrain the possibilities to justify such
‘Gaussian-population approximation’ by any infinitesimal-like model (essentially to mix-
tures of Cauchy distributions. The noted relation to blending inheritance is intriguing!).
My main concern is that especially this section needs are clearer description of the model,
and clear distinction between this ‘maximalist’ model (which has little theoretical sup-
port) and the model investigated by Barton et al (2017). Also a much clearer summary
and discussion of the implications of the results shown here are needed. It should not be
assumed that readers have read Turelli (2017) and even less that they have studied Barton
et al (2017) in considerable detail!
\end{quote}

We now explain stuff more...

\begin{quote}
In Section 5, it is pointed out why the Barton et al. (2017) fails in the case of ‘Cauchy
noise’ (and presumably for any other limit law with $\alpha < 2$). In the absence of giving the
estimate in App D in Barton et al, the explanation at the bottom of p 12 is difficult to
follow. Please provide their estimate for easier comparison (do not assume that readers
studied Barton et al; assume that they scanned (parts) of it).
\end{quote}

\begin{quote}
Finally, Section 6 presents an insightful comparison of the distributions of trait values
resulting from either Gaussian or Cauchy mutation-effect distributions. This seems to be
an explicit genetic model, but again, what is the mutation rate? Clearly, the mutation
rate determines the variance (or more generally variability) of the population distribution.
Please clarify the model. It also needs to be pointed out clearly that such results seem to
be at variance with all (?) explicit multilocus models that have been studied so far

In summary, this ms contains several interesting aspects and will deserve publication after
improving the presentation and clarifying a number of issues.
\end{quote}

\begin{point}{}
Intro, p 2: Not too many readers will be familiar with $\alpha$-stable distributions. It may be
useful to clearly point out the difference to the Gaussian case and what happens if $\alpha \ge 2$
(obviously, this is of relevance).
\end{point}

\reply{
}

\begin{point}{}
p 3, top: State that M is the number of loci.
\end{point}

\reply{
}

\begin{point}{\revref}
Should this read ‘between 20\% and 10\%’? Why this choice?
\end{point}

\reply{
}

\begin{point}{\revref}
First line in Sect 3: add ‘as $M \to \infty$’. In the next line, this is first time that $\alpha = 2$ is
mentioned. Clarify earlier that this case is very different from $\alpha < 2$ (maybe already below
the last display eq on p3).
\end{point}

\reply{
}

\begin{point}{\revref}
What means ‘nearby alleles’ ? Do you mean physically linked
loci? Another 3 lines below: alleles are not polymorphic; loci can be polymorphic.
\end{point}

\reply{
}

\begin{point}{}
p 7, Sect 4: As noted above, please describe this model in more detail to distinguish it
from that in Barton et al (2017), which makes much weaker assumptions. Note that it is
not justifiable except under quite extreme assumptions.
\end{point}

\reply{
}

\begin{point}{}
Third display equation in Sect 4.1 \revref: The index k in the product should start at k = 0.
\end{point}

\reply{
}

\begin{point}{}
p 8, last line: Delete ‘to’
\end{point}

\reply{
}

\begin{point}{}
p 10: It might be useful to mention earlier that the Cauchy distribution is an archetypical
example for an $\alpha$-stable distribution. The enthusiasm among biologists may be limited
though, because all the empirical trait distributions they find have mean and variance. I
don’t know if there are examples in which the variance grows (significantly) as the sample
size increases. In the last para of Sect 4.1, please specify ‘one will get convergence’. I
suppose this refers again to $M \to \infty$.
\end{point}

\reply{
}

\begin{point}{}
p 12, last display equation: Please relate this explicitly to the corresponding result in
Barton et al. (e.g., by stating their estimate).
\end{point}

\reply{
}

\begin{point}{}
p 16: This discussion is too vague. At least, adding a summary of the main findings
would be appropriate (always noting for which version of the infinitesimal model).
\end{point}

\reply{
}

\begin{point}{}
p 16, end of discussion: Does ‘Lévy stable distribution’ refer to the family of $\alpha$-stable
distributions? (I don’t think it has not been used before.)
\end{point}

\reply{
}

### 2023-11-15 reviews/Reviews.tex
reviews put in latex

\documentclass[11pt]{article}

\usepackage{fullpage}

\begin{document}

\section{Editor Comments}

The two reviewers find the results interesting and worth being eventually published in the special issue. However, they highlight a number of places in the text where the properties claimed or the precise questions addressed need serious clarification. As the readership of the special issue will be quite broad, it is important that the main messages are accessible to colleagues who do not necessarily know Fisher's infinitesimal model precisely. Therefore, although the mathematical results in themselves do not require deep modifications, I recommend that a major revision of the paper should be provided, which would take into account the different comments and suggestions of the reviewers.

\section{Reviewer Comments}

\subsection{Reviewer \#1:}

The paper explores inheritance of an additive trait influenced by a large number of loci with effect sizes drawn from a power law distribution, also fitting effect size distributions as estimated by GWAS (for several human disease traits) to estimate plausible power law exponents. The study is motivated by Barton et al's 2017 paper which shows how in the limit of a very large number of loci with effect sizes drawn from a distribution with finite variance, inheritance is described by the the infinitesimal model: in particular, that the distribution of trait values of the offspring of two parents is multivariate normal with a variance-covariance matrix that is independent of parental trait values. This paper explores whether some analogous statements can be made if the underlying effect size distribution does not have finite variance- whether, loosely speaking, some ``generalisation'' of the infinitesimal model might still apply. Below I summarise some of my concerns about the paper.

\begin{itemize}
\item Page 2, paragraph starting ``This suggests exploring whether stable distributions..''
The questions as stated: e.g., whether stable distributions can ``stand in'' for a Gaussian distribution in the infinitesimal model or whether the results of Barton et al ``carry over'' when the sum of effect sizes follows a stable law are quite vague and allow for multiple interpretations. Ideally, in order to address whether Barton et al's results (do not) carry over, one would like to at least show that with a power law distribution of effect sizes, the ``Mendelian noise term'', i.e., the difference between the offspring trait value and the midparent value, is (not) independent of the midparent value in the limit of a very large number of loci. And indeed this is what the authors address
somewhat non-rigorously (and for the case of a Cauchy distribution of effect sizes) in section 5 and using simulations : see e.g., figures 4C and 4f, figures 6C and 6F. Unsurprisingly, these simulations show that with e.g., Cauchy distribution of effect sizes, the non-independence between the segregation/noise term and the midparent trait value does not hold. However, a lot of the mathematical analysis in the paper (sections 3 and 4) explores when the distribution of trait values in a population is (not) Gaussian, which is a completely different question.

In the same paragraph: I don't think the claim that ``independence of offsprings' deviations and midparent values implies a Gaussian distribution'' (which is also stated in the abstract) is true? Surely, ``independence of offsprings' deviations and midparent values'' is just the basic statement of the infinitesimal model (say, in Barton et al 2017) and the infinitesimal model does not imply a Gaussian distribution of trait values? See also section 3.

\item Section 2 and figure 1: It is not obvious to me why one would study the ``frequency-weighted'' effect sizes. There is an assumption here I think that allele frequencies are uncorrelated with effect sizes: this will not be true if for example traits are under some kind of stabilising selection: see e.g., Simons et al (A Population Genetic Interpretation of GWAS Findings for Human Quantitative Traits) for a detailed exploration of this. Moreover, if SNPs are not truly causal but only tag causal alleles, then weighting by allele frequencies can be even more problematic?

As an alternative, would it make sense to look at what fraction of loci have effect size difference between alternative alleles to be $>2t$ without weighing by allele frequency?
More broadly, how do we interpret the frequency-weighted vs. unweighted distributions?

I was also wondering about the ``goodness of fit'' of observed distributions to power laws. I am not necessarily recommending a detailed quantitative analysis, but nevertheless it may be useful to comment on whether one has power to say reject alternative distributions (e.g., a mixture of exponentials) that are often considered. If not, then perhaps the exact exponents inferred are unimportant, and what the data points to more broadly, is a broad distribution of effect sizes.

\item Section 3, Sentence starting ``We know from Barton et al (2017)...'' : As far as I understand, the statement of Barton et al is that the joint distribution of trait values of offspring is multivaraite normal, conditioned on the two parents. I don't think the joint distribution of two randomly picked parents and their offspring is mutivariate normal, as this would require some kind of normality assumption about trait values in the population as a whole. It'd be helpful to distinguish more carefully between conditioned and unconditioned distributions, especially in this section.

In general, I find this section rather confused: the infinitesimal model basically implies that the ``Mendelian segregation term'' $R_{M}$ (which is the difference between an individual and the midparent value $\bar{Z}_{M}$) is normally distributed with mean zero and variance that depends on the relatedness between parents but is independent of the midparent value $\bar{Z}_{M}$. However, in proposition 1 (page 7), the authors claim that $\bar{Z}_{M}$ and $R_{M}$ are jointly Gaussian (which is equivalent to saying that the trait value of an individual and the mean of its parents' trait values is jointly Gaussian), which is in general not true.

I think it is also not correct to claim that ``Since models can be easily set up for which trait distributions are not Gaussian, the implication of this is that for such models, independence of the Mendelian sampling term is not likely a good assumption.'' (just below Proposition 1). As far as I understand, a non-Gaussian distribution of trait values can arise quite easily under the infinitesimal model (where the Mendelian segregation term is independent of the midparent value) e.g, due to strong or non-Gaussian selection or migration or some combination of the two (see, e.g., Figure 1 in Barton et al).

\item Section 4, first paragraph: I don't think the third assumption (that trait values in the population follow a Gaussian distribution) is a component of the ``infinitesimal model''. I think Turelli (2017) is also quite clear about this stating that a Gaussian distribution of trait values should only emerge for Gaussian or very weak selection on the population.

More generally, I found it a bit hard to follow the overall logical flow of section 4.1, especially once the authors launch into an exploration of the ``reproduction'' and ``noise'' terms (pages 9 and 10).
It would be useful to clarify what the biological/intuitive meaning of the noise and reproduction terms at the outset. It would also be useful to state (at the beginning) what the main goal of these explorations is: (is the goal to identify when both terms ``have well-posed limits independent of the other'' as stated at the end of page 8), and also to summarise (at the end) to what extent one can do this in all generality.

\item Last inequality on Page 13: Is this the distribution of the largest of the $M$ alleles carried by an individual, conditioned on the trait value $Z$ of the individual (and this is independent of $Z$?), or is it just the largest of $M$ iid draws from a Cauchy distribution? If the former, then it'd be good to say this explicitly. If the latter, then this does not quite answer the question posed at the beginning of the section about how much information knowing the trait value of an individual gives about underlying allelic effects.

Also, maybe worth specifying what the corresponding distribution for the largest of $M$ alleles looks like when the effect size distribution has finite variance, in order to highlight the contrast between ``well-behaved'' and heavy-tailed effect size distributions.
\end{itemize}

\subsubsection{Minor comments:}

\begin{itemize}
\item Introduction: Also worth citing Fisher, Bulmer etc. (the original references) when introducing the infinitesimal model?

\item Page 2, line 4: ``Perhaps the distribution of effect sizes within each gene is Normal'': I am confused by this. Do you mean that the distribution of effect sizes across all genes (with roughly the same ``proximity'' to the trait) is normal. If yes, then the above phrasing is a bit misleading. If what is meant that is really that the effect sizes of different alleles within a gene are normally distributed, then I am not sure if this is entirely plausible: see various papers by Turelli (e.g., ``Heritable Genetic Variation via Mutation-Selection Balance: Lerch's Zeta Meets the Abdominal Bristle.''). Presumably, a normal distribution of effect sizes for a sufficiently large genomic region is a good approximation but it is unclear what is large enough: smaller or larger than a typical gene...?

\item Page 2, sentence starting ``In fact, many traits claimed...'': This phrasing here is rather cryptic and I do not understand what is being said. Maybe rephrase? Also a reference would help...

\item Figure 2: Is there also a correlation between the number of cases and the number of SNPs? In other words, to what extent are (b) and (c) independent?

\item Page 12, sentence starting ``To see what knowing the parental trait value...'': In what sense is this the ``parental'' trait value??

\item Last inequality on Page 13: should the subscript be $j$ instead of $i$?

\item Simulations: Why not choose the same effect size distributions in the neutral case and the case with stabilizing selection? Alternatively, if different effect size distributions are used, it may be good to plot the distribution of (appropriately) scaled trait values. Right now the figures give the impression that more variation is maintained under stabilizing selection..

\item Figure 4 caption: I find the phrasing ``(a,b) chosen to span the 5\% of midparents centered on the 10\% quantile of midparent value'' a bit confusing. Does this mean a and b are respectively the 7.5\% and 12.5\% quantile of the distribution of midparent trait values?

\item Discussion: In the context of theory combining large and small-effect loci, the 2008 paper by Chevin and Hospital: ``Selective Sweep at a Quantitative Trait Locus in the Presence of Background Genetic Variation'' and the 1983 work by Lande: ``The response to selection on major and minor mutations affecting a metrical trait'' may also be somewhat relevant.
\end{itemize}

\end{document}

### 2023-11-15 writeup.tex
affils
sign error and comments
reviews put in latex

@@
 	\newtheorem{conj}{Conjecture}
+
+\newif\ifsubmission
+\submissiontrue
+% \submissionfalse
+\ifsubmission
+    % % stuff for submission
+    \usepackage{lineno}
+    % allow line numbers around math environments
+    % from https://tex.stackexchange.com/questions/43648/why-doesnt-lineno-number-a-paragraph-when-it-is-followed-by-an-align-equation
+        \newcommand*\patchAmsMathEnvironmentForLineno[1]{%
+          \expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
+          \expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
+          \renewenvironment{#1}%
+             {\linenomath\csname old#1\endcsname}%
+             {\csname oldend#1\endcsname\endlinenomath}}% 
+        \newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
+          \patchAmsMathEnvironmentForLineno{#1}%
+          \patchAmsMathEnvironmentForLineno{#1*}}%
+        \AtBeginDocument{%
+        \patchBothAmsMathEnvironmentsForLineno{equation}%
+        \patchBothAmsMathEnvironmentsForLineno{align}%
+        \patchBothAmsMathEnvironmentsForLineno{flalign}%
+        \patchBothAmsMathEnvironmentsForLineno{alignat}%
+        \patchBothAmsMathEnvironmentsForLineno{gather}%
+        \patchBothAmsMathEnvironmentsForLineno{multline}%
+        }
+\else
+    % for the nice version
+\fi
+
+% Responses to reviews
+\input{review-response-commands}
+% set this to show line numbers and include responses to reviews or not
+\newif\ifreviewresponses
+\reviewresponsestrue  % include them
+% \reviewresponsesfalse  % don't include them
+\newcommand{\responsefile}{review-responses}  % name of the review responses file
 
@@
 \author{ \begin{small}
-    Todd L. Parsons (LPSM, Sorbonne Universit\'e, CNRS UMR 8001, Paris, France, 75005) \&
-    Peter Ralph
+    Todd L. Parsons\footnote{LPSM, Sorbonne Universit\'e, CNRS UMR 8001, Paris, France, 75005}
+    ~\&~
+    Peter L. Ralph\footnote{Institute of Ecology \& Evolution, University of Oregon, Eugene, OR, USA, 97405}
 \end{small}}
@@
 This suggests a mixture model for the effect size distribution:
-perhaps the distribution of effect sizes within each gene is Normal,
+perhaps the distribution of effect sizes within each gene is Normal, \revpoint{1}{13}
 but with a standard deviation that varies with ``proximity'' to the trait;
@@
 This suggests exploring whether stable distributions might reasonably stand in for the Gaussian
-in the infinitesimal model.
+in the infinitesimal model. \revpoint{1}{1} \revpoint{1}{2}
 In this paper, we make preliminary explorations in that direction:
@@
 and interactions between alleles, between loci, and with the environment are often important in practice.
-In fact, many traits claimed to be genetically determined
+In fact, many traits claimed to be genetically determined \revpoint{1}{14}
 in fact only appear heritable through non-genetic means,
@@
 $$
-    \lim_{t \to \infty} \frac{ \log \P\{ |X| > 2 t \} }{ \log t } = \alpha ,
+    \lim_{t \to \infty} \frac{ - \log \P\{ |X| > 2 t \} }{ \log t } = \alpha ,
 $$
@@
 (which we take to be all values of $t$ for which the number of SNPs with absolute effect above $t$
-is between 20 and 10\% of all SNPs).
+is between 20 and 10\% of all SNPs). \revpoint{2}{3}
 Examples are shown in Figure~\ref{fig:example_snps}.
@@
 
-We know from \citet{barton2017infinitesimal} that if the $\eta^{[i]}_\ell$ are
+We know from \citet{barton2017infinitesimal} \revpoint{1}{5} \revpoint{2}{4}
+that if the $\eta^{[i]}_\ell$ are
 independent with mean zero and $\sum_{\ell=1}^M \var[\eta_\ell]/M \to \sigma^2$,
@@
 the implication of this is that for such models, independence of the Mendelian sampling term
-is not likely a good assumption.
+is not likely a good assumption. \revpoint{1}{7}
 Note that the proof does not depend on independence of the $X_\ell$:
 indeed, alleles may be inherited together,
-although in practice this would quickly create dependencies between nearby allelic effects.
+although in practice this would quickly create dependencies between nearby allelic effects. \revpoint{2}{5}
 Also note that it is \emph{not} usually true that the components --
@@
 
-As observed in \cite{turelli2017commentary}, there are at least three interpretations of what is called ``the infinitesimal model''.
+As observed in \cite{turelli2017commentary}, \revpoint{1}{8}
+there are at least three interpretations of what is called ``the infinitesimal model''.
 The maximalist of these, what \cite{turelli2017commentary} calls the Gaussian-population approximation, assumes a large number of small-effect loci,  parental trait-independent Gaussian Mendelian sampling terms  (as rigorously justified in \cite{barton2017infinitesimal,barton2022infinitesimal}), and, moreover, that  population trait values are in deterministic mutation-selection balance, with a Gaussian empirical distribution
@@
 whereas any fixed point must be the limit of an arbitrary characteristic function, $\varphi(\theta)$, under repeated applications of $F$,
-\ie
+\ie \revpoint{2}{7}
 \begin{align*}
@@
 We refer to the first term on the left hand side as the \emph{reproduction term} and the remaining sum as the \emph{noise term}, and only consider to the case when both have well-posed limits independent of the other.
+\llabel{limit_remark}
 
@@
 
-To see what knowing the parental trait value $Z^{[j]}_{M}$ tells us about the contribution of any given locus, $\eta^{[j]}_{\ell}$, we follow the arguments in  \citet{Fisher1918} and \citet{barton2017infinitesimal} using Bayes' Theorem.
+To see what knowing the parental trait value $Z^{[j]}_{M}$ \revpoint{1}{16}
+tells us about the contribution of any given locus, $\eta^{[j]}_{\ell}$, we follow the arguments in  \citet{Fisher1918} and \citet{barton2017infinitesimal} using Bayes' Theorem.
 Proceeding informally, with the probabilities indicating probability density functions, we have
@@
 
+% reviews here if toggled above
+% \ifsubmission\processdelayedfloats\fi
+
+\includereviews
+
 \end{document}

### 2023-11-27 writeup.tex
notes from meeting

@@
 \end{abstract}
+
+\comment{Make the point that another way to think about distibutions with non-finite variance
+    is that there's only a finite number of alleles above any fixed value;
+    and also the variance in any finite population will be finite.
+    (see reviewer comment, ``all the empirical trait distributions they find have mean and variance'')
+    Also discuss comparison to fat-tailed dispersal:
+    the question is, what distribution gives us a better model?
+    For dispersal, a fat-tailed distribution has in practice finite variance
+    but is a better model in that it describes noisiness better;
+    here, the analogy is something about polygenicity and noisiness of the process of adaptation.
+    If we're modeling non-polygenic (oligogenic) traits then
+    why stable as opposed to some other arbitrary distribution?
+    Well, the distribution of per-allele effect sizes might not be stable, but CLT tells us that maybe
+    stable gives a good approx to the process (as in Gaussian case).
+}
 

### 2023-11-28 review-responses.tex
many responses

@@
 \begin{point}{}
-    Section 2 and figure \ref{fig:example_snps}: It is not obvious to me why one would study the ``frequency-weighted'' effect sizes.
-There is an assumption here I think that allele frequencies are uncorrelated with effect sizes: this will not be true if for example traits are under some kind of stabilising selection: see e.g., Simons et al (A Population Genetic Interpretation of GWAS Findings for Human Quantitative Traits) for a detailed exploration of this.
+    Section 2 and Figure \ref{fig:example_snps}: It is not obvious to me why one would study the ``frequency-weighted'' effect sizes.
+There is an assumption here I think that allele frequencies are uncorrelated with effect sizes: this will not be true if for example traits are under some kind of stabilising selection: see e.g., \citet{simons2018population} for a detailed exploration of this.
 Moreover, if SNPs are not truly causal but only tag causal alleles, then weighting by allele frequencies can be even more problematic?
@@
 \reply{
+    Good question; this was somewhat obscure in the previous draft.
+    We've added a comment which hopefully explains why the frequency-weighted distribution
+    is the correct thing to consider for this calculation \revref,
+    and have added some more discussion about the point in general
+    to the Discussion \llname{discussion_DFEs}.
 }
@@
 \begin{point}{}
-    Section 3, Sentence starting ``We know from Barton et al (2017)...'' \revref: As far as I understand, the statement of Barton et al is that the joint distribution of trait values of offspring is multivaraite normal, conditioned on the two parents.
+    Section 3, Sentence starting ``We know from \citet{barton2017infinitesimal}, \ldots'' \revref: As far as I understand, the statement of \citet{barton2017infinitesimal} is that the joint distribution of trait values of offspring is multivaraite normal, conditioned on the two parents.
 I don't think the joint distribution of two randomly picked parents and their offspring is mutivariate normal, as this would require some kind of normality assumption about trait values in the population as a whole.
@@
 \reply{
+    Looking back at \citet{barton2017infinitesimal}, we see where the confusion comes from,
+    and have added some clarifying remarks at the end of this paragraph \revref.
+    % Quoting \citet{barton2017infinitesimal},
+    % ``Even though the trait distribution across the whole population may be far from Gaussian,
+    % the multivariate normal will play a central r\hat{o}le in our analysis''.
+    In more detail:
+    \citet{barton2017infinitesimal} (``main result'', equation 9)
+    shows that conditioned on the trait values in generation $t$,
+    the values of generation $t+1$ are also multivariate Gaussian.
+    This implies that the joint distribution of traits of any subset of individuals
+    is multivariate Gaussian (possibly conditioned on the initial generation)
+    -- as \citet{barton2017infinitesimal} says,
+    \textit{``equally we could have conditioned on the values of any subset of relatives
+    and the same would hold true: expected trait values would be
+    a linear functional (determined by the pedigree) of the values
+    on which we conditioned, but segregation variances would be unchanged.''}
+    (More generally, if $X$ is multivariate Gaussian, and $(Y|X)$ is also multivariate Gaussian
+    with mean but not covariance depending on $X$,
+    then $(X,Y)$ are jointly multivariate Gaussian.)
+    We now also discuss the point further in a subsequent paragraph \llname{pop_not_gaussian}.
 }
@@
 \reply{
+    If the reviewer had said ``confusing'' rather than ``confused'',
+    then we would heartily agree.
+    After some difficult thinking through of things,
+    we've provided a good bit of additional clarification
+    about what the Proposition means and doesn't mean:
+    see in particular \llname{pop_not_gaussian} and \revref.
 }
@@
 \reply{
+    Thanks -- we've changed this to ``since models can be easily set up for which the distribution
+    of Mendelian sampling terms is not Gaussian, \ldots''
+    and provided additional context \revref.
 }
@@
 
-\subsubsection{Minor comments:}
+\subsubsection*{Minor comments:}
 
@@
 \begin{point}{}
-Intro, p 2: Not too many readers will be familiar with $\alpha$-stable distributions. It may be
+Intro: Not too many readers will be familiar with $\alpha$-stable distributions. It may be
 useful to clearly point out the difference to the Gaussian case and what happens if $\alpha \ge 2$
@@
 \reply{
+    Good idea; we've added this \revref.
 }
@@
 \begin{point}{}
-p 3, top: State that M is the number of loci.
+p 3: State that M is the number of loci.
 \end{point}
@@
 \reply{
+    Whoops - done \revref.
 }
@@
 \reply{
+    No, but as written it was confusing, and we've clarified \revref.
+    We chose these values by empirical inspection:
+    the lower bound was 20 SNPs so that it was not unduly affected by a few large SNPs,
+    and the upper bound as 10\% because we wanted the ``tail'', not the ``bulk''.
+    (We also checked that changing these values somewhat didn't affect results.)
 }
@@
 \reply{
+    Good idea -- done \revref. (Also see \revreffull{2}{1}.)
 }
@@
 What means ‘nearby alleles’ ? Do you mean physically linked
-loci? Another 3 lines below: alleles are not polymorphic; loci can be polymorphic.
+    loci? Another \llname{loci_polymorphic}: alleles are not polymorphic; loci can be polymorphic.
 \end{point}
@@
 \reply{
+    Good points; both are fixed: \revref and \llname{loci_polymorphic}.
 }
@@
 
-\begin{point}{}
-p 16: This discussion is too vague. At least, adding a summary of the main findings
+\begin{point}{\revref}
+This discussion is too vague. At least, adding a summary of the main findings
 would be appropriate (always noting for which version of the infinitesimal model).
@@
 \reply{
+    We have added a good bit more discussion, including a summary \revref.
 }
 
-\begin{point}{}
-p 16, end of discussion: Does ‘Lévy stable distribution’ refer to the family of $\alpha$-stable
+\begin{point}{\revref}
+Does ‘Lévy stable distribution’ refer to the family of $\alpha$-stable
 distributions? (I don’t think it has not been used before.)
@@
 \reply{
+    That's right; changed to ``$\alpha$-stable distribution'' \revref.
 }

### 2023-11-28 writeup.tex
Update writeup.tex
many responses

@@
 
-\comment{Make the point that another way to think about distibutions with non-finite variance
+\comment{ADD TO DISCUSSION, WHICH IS SHORT:
+    Make the point that another way to think about distibutions with non-finite variance
     is that there's only a finite number of alleles above any fixed value;
@@
 The consensus in the field seems to be that traits exist on a spectrum between monogenic and polygenic,
-and that many - if not most - lie in the intermediate region,
+and that many -- if not most -- lie in the intermediate region,
 where heritable trait variation is due both to segregating large- and small-effect variants.
@@
 (and the approximation becomes exact, as $n \to \infty$).
+Oth the other hand, if $\alpha \ge 2$ then the distribution has finite variance,
+and so the sum of $n$, divided by $\sqrt{n}$, converges to a Gaussian distribution.
 In such ensembles, single large entries can be important:
@@
 so that e.g., for $\alpha=1$ a positive proportion of the sum of $n$ entries will be contributed by a single one.
+On the other hand, the largest of $n$ Gaussians is of order $\sqrt(\log n)$. \revpoint{2}{1}
 
@@
 $$
-where $\eta_\ell$ is the contribution of the allele carried at the $\ell^\text{th}$ locus.
+where $\eta_\ell$ is the contribution of the allele carried at the $\ell^\text{th}$ locus,
+and $M$ is the number of loci. \revpoint{2}{2}
 The effects of the alleles carried by the two parents
@@
 \section{Empirical observations}
+    \label{sec:empirical}
 
@@
 
-Since the mendelian sampling term is a sum of terms of the form
+Since the Mendelian sampling term is a sum of terms of the form
 $$
@@
 the appropriate central limit theorem for the sum is determined by
-the tail behavior of $|(X - 1/2) \eta| = |\eta|/2$,
+the tail behavior of $|(X - 1/2) \eta|$,
 where $X$ is Bernoulli(1/2) and $\eta$ is the effect
-of a uniformly chosen allele at a uniformly chosen locus.
-So, we'd like to know
+of a uniformly chosen one of the two alleles from a uniformly chosen parent
+at a uniformly chosen locus.
+Since $|X-1/2| = 1/2$, we'd like to know
 if $\P\{|X|>2t\}$ is proportional to $t^{-\alpha}$,
@@
 $$
+Note that the relevant effect size distribution here is weighted by frequency, \revpoint{1}{3}
+which is necessary because we're looking at the deviation of a typical offspring from their parents,
+and the chance a given allele is carried by a random parent is given by the frequency.
+This will make a difference as frequencies and effect sizes are often correlated in practice.
 So, to obtain a per-trait estimate of $\alpha$,
@@
 for a sequence of values of $t$,
-and fit a linear model to the tail
-(which we take to be all values of $t$ for which the number of SNPs with absolute effect above $t$
-is between 20 and 10\% of all SNPs). \revpoint{2}{3}
+and fit a linear model to the tail.
+(To do this,
+we take as the ``tail''
+all values of $t$ for which the number of SNPs with absolute effect above $t$
+is at least 20 and at most 10\% of all SNPs.) \revpoint{2}{3}
 Examples are shown in Figure~\ref{fig:example_snps}.
@@
 that if the $\eta^{[i]}_\ell$ are
-independent with mean zero and $\sum_{\ell=1}^M \var[\eta_\ell]/M \to \sigma^2$,
-then with $\alpha=2$,
+independent with mean zero and $\sum_{\ell=1}^M \var[\eta_\ell]/M \to \sigma^2$ as $M \to \infty$,
+then with $\alpha \ge 2$,
 the traits $(Z_M, Z_M^{[1]}, Z_M^{[2]})$ converge in distribution to a multivariate Gaussian distribution
 $(Z, Z^{[1]}, Z^{[2]})$ for which $Z - (Z^{[1]} + Z^{[2]})/2$ is independent of $Z^{[1]}$ and $Z^{[2]}$.
+(The main result in \citet{barton2017infinitesimal} is that the trait values in generation $t$
+are Gaussian conditioned on the traits in generation $t-1$,
+but as discussed in that paper,
+this implies that the traits of any set of individuals is multivariate Gaussian,
+with covariance matrix determined by the pedigree.
+Note, however, that this does \emph{not} imply that the distribution of trait values
+across the population is Gaussian, as a collection of correlated Gaussians is not itself Gaussian.)
 
@@
 in at least some situations,
-independence of the Mendelian sampling term implies a Gaussian distribution.
-To see this,
+independence of the Mendelian sampling term \emph{implies} a Gaussian distribution.
+To see this (and to see exactly what we mean by this),
 divide the parental loci into groups according to the results of inheritance,
@@
 
-% A similar statement holds for the offspring of two parents in a pedigree,
-% using the Skitovich-Darmois theorem \citet{kagan1973characterization},
-% except that the portion of the genome shared in common by the parents is unconstrained.
-
-Since models can be easily set up for which trait distributions are not Gaussian,
-the implication of this is that for such models, independence of the Mendelian sampling term
+We should now discuss what this counter-intuitive result says and does not say.
+The proposition says, roughly, that if the parental alleles are unlinked and independently chosen,
+and meotic segregation is fair,
+then independence of Mendelian sampling term from midparent value
+\emph{implies} a joint Gaussian distribution for the two.
+We have not assumed anything about the distribution of the trait in the population:
+rather, we have showed that any model of this sort
+that assumes independence of Mendelian sampling term and midparent value
+must also, by logical necessity, have a Gaussian distribution for the two.
+In other words,
+since models can be easily set up for which the distribution of Mendelian sampling terms
+is not Gaussian,
+the implication of the proposition
+is that for such models, independence of the Mendelian sampling term
 is not likely a good assumption. \revpoint{1}{7}
@@
 indeed, alleles may be inherited together,
-although in practice this would quickly create dependencies between nearby allelic effects. \revpoint{2}{5}
+although in practice this would quickly create dependencies between the effects
+of alleles at such linked loci. \revpoint{2}{5}
 Also note that it is \emph{not} usually true that the components --
@@
 are independent of each other.
+Finally, note that although they may not be independent,
+the Mendelian sampling term $R_M$ and the midparent value $\bar Z_M$
+\emph{are} always uncorrelated, since $Z_M$ and $V_M$ are exchangeable.
+
+Note that under the conditions of the proposition, \llabel{pop_not_gaussian}
+the parental traits, $Z^{[1]}_M$ and $Z^{[2]}_M$,
+are also (jointly) Gaussian.
+(This was not an assumption, however.)
+If parents are chosen uniformly, then
+the distribution of, say, $Z^{[1]}$ is just the population's trait distribution.
+This seems wrong, since it is well-known that forces such as selection or migration
+can make the population distribution non-Gaussian.
+However, there is no contradiction:
+we have stated the proposition only in terms of the \emph{first} generation
+of offspring of unrelated parents \citep[as in][]{barton2017infinitesimal}.
+Proposition~\ref{prop:parental_contribs} does not imply that
+the distribution of traits in the offspring generation is Gaussian:
+unless all offspring are unrelated,
+it will be a collection of correlated Gaussians,
+which is not itself Gaussian.
+(As an extreme example, suppose that half the offspring are produced by the two parents
+with largest trait values, and the other half by the parents with the smallest trait values;
+then although each offspring's distribution about the midparent is Gaussian,
+the offspring's population distribution is a mixture of two Gaussians
+about these two midparents.)
+What about further generations?
+A similar statement to the proposition
+holds for the offspring of two parents in an arbitrary pedigree,
+using the Skitovich-Darmois theorem \citet{kagan1973characterization}
+% except that the portion of the genome shared in common by the parents is unconstrained.
+-- however, there is nothing to be gained by the additional generality.
 
 As we discuss more later,
-the presence of a large effect polymorphic allele
+the presence of a large effect allele at a polymorphic locus \llabel{loci_polymorphic}
 makes $R$ and $\bar Z$ dependent:
@@
 then $R$ is close to $+5$ or $-5$ with equal probability.
+
+Finally, a possible point of confusion is that \revpoint{1}{6}
+the infinitesimal model is often phrased as a statement about offspring trait distributions
+\emph{given} parental trait values,
+while our statement above regards parental trait values as random.
+However, the notions are equivalent:
+the statement that
+``for each pair of parents the deviation of the offspring's trait from midparent
+has a distribution that does not depend on the parents' values''
+is essentially the definition of statistical independence
+of offspring's Mendelian sampling term and parental trait values.
 
@@
 \[
-	\Psi(\theta) = im\theta + c|\theta|^{\alpha} (1-i\beta \sgn(\theta)\omega_{\alpha}(\theta))
+	\Psi(\theta) = im\theta - c|\theta|^{\alpha} (1-i\beta \sgn(\theta)\omega_{\alpha}(\theta))
 \]
@@
 	 \sum_{k=1}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}
-	 &=  \sum_{k=1}^{n-1} 2^{(1-\alpha)k} c|\theta|^{\alpha} {\textstyle\left(1-i \beta\sgn(\theta)\left(\frac{\theta}{2^{k}}\right)
+	 &=  \sum_{k=1}^{n-1} -2^{(1-\alpha)k} c|\theta|^{\alpha} {\textstyle\left(1-i \beta\sgn\left(\frac{\theta}{2^{k}}\right)
 	 	 \omega_{\alpha}\left(\frac{\theta}{2^{k}}\right)\right)}\\
 	&= \begin{cases}
-		 c|\theta|^{\alpha}\frac{1-2^{(1-\alpha)n}}{1-2^{1-\alpha}}\left(1-i\beta \sgn(\theta) \tan{\frac{\pi\alpha}{2}}\right) & \text{if $\alpha \neq 1$, and}\\	
-         (n-1) c|\theta|^{\alpha}  \left(1+\frac{2i\beta}{\pi}\sgn(\theta)\left(\log{|\theta|} - \frac{n}{2} \log{2}\right)\right) & \text{if $\alpha = 1$,}
+		 -c|\theta|^{\alpha}\frac{1-2^{(1-\alpha)n}}{1-2^{1-\alpha}}\left(1-i\beta \sgn(\theta) \tan{\frac{\pi\alpha}{2}}\right) & \text{if $\alpha \neq 1$, and}\\	
+         -(n-1) c|\theta| \left(1+\frac{2i\beta}{\pi}\sgn(\theta)\left(\log{|\theta|} - \frac{n}{2} \log{2}\right)\right) & \text{if $\alpha = 1$,}
 	\end{cases}
@@
 \begin{equation}\label{eq:stablenoises}
-	 \frac{c}{1-2^{1-\alpha}}|\theta|^{\alpha}\left(1-i\beta \sgn(\theta) \tan{\frac{\pi\alpha}{2}}\right), 
+	 -\frac{c}{1-2^{1-\alpha}}|\theta|^{\alpha}\left(1-i\beta \sgn(\theta) \tan{\frac{\pi\alpha}{2}}\right), 
 \end{equation}
@@
 \section{Discussion}
+    \revpoint{2}{11}
+
+The infinitesimal model is a powerful and tractable model of trait inheritance,
+but is polygenic to a fault,
+excluding the possibility of loci of large effect
+(that are often seen in practice).
+In this paper,
+we have looked at the possibility of using a broader range of central limit theorems
+to extend the infinitesimal model to traits for which 
+effect size distributions fall in the domain of attraction of the $\alpha$-stable distributions
+(i.e., a larger proportion of variance
+is explained by a smaller number of loci).
+Indeed, we found evidence that this is true of many real traits,
+although our relatively uncritical use of GWAS estimates makes this evidence suggestive at best.
+Next, we show that in fact the infinitesimal model \emph{cannot} describe non-Gaussian distributions
+\comment{XXX TODO}.
+A recurring theme in the rest of the paper is
+that just as the infinitesimal model is useful precisely because it ignores the genetic details,
+it cannot apply once some genetic details are important
+(i.e., there are loci of large effect).
 
@@
 not because it seems most realistic
-(indeed data suggests the L\'evy stable distribution with $\alpha=3/2$ might be better)
+(indeed data suggests the $\alpha$-stable distribution with $\alpha=3/2$ might be better) \revpoint{2}{12}
 but for mathematical convenience.
+
+\subsection*{Selection and effect size distributions} \llabel{discussion_DFEs}
+A substantial body of theoretical work has used various models and assumptions
+to predict what distributions of effect sizes are expected based on biology.
+In comparing to this work, there are several distinctions that are important to make.
+For instance, selection makes the distribution of effect sizes identified by GWAS
+different than the distribution of effect sizes of \emph{de novo} mutations,
+and predictions are made about this relationship by \citet{simons2018population} % and \citet{simons2022simple}.
+(under stabilizing selection)
+and \citet{hayward2022polygenic}
+(under adaptation to a new environment).
+Our first question in this paper is orthogonal to this discussion:
+\emph{given} the available genetic variation,
+what model should we use to predict the short-term course of evolution?
+As we saw in Section~\ref{sec:empirical},
+the frequency-weighted effect size distribution (perhaps closer to that seen by GWAS)
+is more relevant for this question than the \emph{de novo} effect size distribution.
+In the latter part of this paper we look at long-term dynamics,
+but with major differences:
+most obviously, we only consider the distribution of the Mendelian noise term,
+rather than the distribution of individual allelic effects.
+These two are clearly related, but a more detailed analysis is beyond the scope of this paper.
 

### 2023-11-30 review-responses.tex
more responses

@@
 
-Thanks for the careful reading...
+Thanks for the careful reading and useful comments.
 
@@
 \reply{
+    We don't actually agree with this characterization --
+    sections 3 and 4 explore the consequences of different distributions
+    of the Mendelian segregation term (which derives from the sum of effects),
+    not the population distribution.
+    However, the assumptions and motivation were not nearly as clear as they could have been
+    in the previous draft, and this is hopefully more clear now --
+    see responses to points below.
 }
@@
 \reply{
+    We have modified the sentence to say
+    ``implies a Gaussian for the offsprings' deviations'',
+    to make it clear that this statement does not refer to the population distribution,
+    but rather to the Mendelian sampling term \revref.
+    See additional points below on this topic.
 }
@@
 \reply{
+    This is a good question, and we had in fact done some exploration of this question
+    but not included it in the paper.
+    We've still not done a detailed analysis,
+    but now provide some discussion and numerical exploration of our precision here:
+    see Figure~\ref{fig:power} and \revref.
 }
@@
 \reply{
+    We've rephrased this, and replaced ``gene'' with ``region'' --
+    really, we just want to motivate a mixture model,
+    as the "scale mixture of Gaussians" is a common way to get fat-tailed distributions,
+    and eminently biologically plausible (we think).
 }
@@
 \reply{
+    Yes; we've added this to Figure~\ref{fig:exponent_hist}.
 }
@@
 \reply{
+    These are good suggestions, however, we have encountered difficulties in implementing this,
+    so we've settled for a note alerting the reader not to read too much into the absolute scale
+    (see Figure~\ref{fig:sel_trait_distrns} caption).
 }
@@
 \reply{
+    That's right; we've rephrased this confusing explanation.
 }

### 2023-11-30 snp_effects/power.Rmd
more responses

Let's check that our estimation procedure for alpha works.

```{r fake_data}
n <- 1e3
gwas <- data.frame(minor_AF=runif(n),
                   beta=rcauchy(n)/100)

min_num <- 20
max_prop <- 0.1
```

Now let's let $(P, Z)$ be a random draw from this table, all $Z > 0$,
and $X = Z$ with probability $P$, and $X=0$ otherwise.
Here's the complementary cumulative distribution function of $X$,
i.e., if $N$ is the number of rows in the table,
$$\begin{aligned}
   \mathbb{P}\{X > x \}  
   &= \frac{1}{N} \sum_i P_i \mathbb{1}_{Z_i > x} \\
   &= \mathbb{P}\{X > 0\} - \frac{1}{N} \sum_i P_i \mathbb{1}_{Z_i \le x} \\
   &= \frac{1}{N}\sum_i P_i - \frac{1}{N} \sum_i P_i \mathbb{1}_{Z_i \le x} .
\end{aligned}$$

```{r plotit}
estim_alpha <- function (gwas, num_breaks=1001) {
    # For a sequence of value xi,
    # fits the linear model
    #  log(pxi) ~ log(xi)
    # where
    #  pxi = prob( a randomly chosen locus has effect above xi )
    # to only the 'tail', defined as those xi such that there
    # are at least min_num loci with effect above xi
    # and at most a proportion max_prop of loci with effect above xi.
    xx <- c(seq(0, 2 * quantile(abs(gwas$beta), 0.99), length.out=num_breaks), Inf)
    disc_beta <- cut(abs(gwas$beta), xx)
    num_points <- table(disc_beta)
    dpx <- tapply(gwas$minor_AF, disc_beta, sum)
    dpx[is.na(dpx)] <- 0
    p_zero <- mean(gwas$minor_AF)
    tail_prob <- p_zero - cumsum(dpx) / nrow(gwas)
    tail_nums <- nrow(gwas) - cumsum(num_points)
    x <- xx[-1]
    xlm <- lm(log(tail_prob) ~ log(x),
              subset=(is.finite(log(tail_prob))
                      & tail_nums > min_num
                      & tail_nums < max_prop * nrow(gwas)))
    return(xlm)
}

xlm <- estim_alpha(gwas)
logx <- xlm$model[[1]]
logy <- xlm$model[[2]]
plot(logx, logy, ylab='log(prob > x)', xlab='log(x)')
abline(coef(xlm))

summary(xlm)
```

## Power

Huh, how good of an estimator is this?
```{r boot}
cauchy_boots <- replicate(1000, {
    gwas <- data.frame(minor_AF=runif(n),
                       beta=rcauchy(n)/100)
    xlm <- estim_alpha(gwas)
    return(coef(xlm)) })
hist(cauchy_boots[2,], breaks=30)
abline(v=-1, col='red')
```
Looks ok.

## Gaussian

And, what if we give this Gaussian data?
```{r gboot}
gaussian_boots <- replicate(1000, {
    gwas <- data.frame(minor_AF=runif(n),
                       beta=rnorm(n)/100)
    xlm <- estim_alpha(gwas)
    return(coef(xlm)) })
hist(gaussian_boots[2,], breaks=30)
abline(v=-1, col='red')
```
Looks ok.

## t(df=1.5)

How about Student's t?
```{r tboot}
t_boots <- replicate(1000, {
    gwas <- data.frame(minor_AF=runif(n),
                       beta=rt(n, df=1.5)/100)
    xlm <- estim_alpha(gwas)
    return(coef(xlm)) })
hist(t_boots[2,], breaks=30)
abline(v=-1.5, col='red')
```
Looks ok.

## For the paper

```{r paper_fig}
pdf(file="power_demo.pdf", width=6, height=2.5, pointsize=10)
    layout(t(1:3))
    hist(-cauchy_boots[2,], breaks=30, xlab="estimated alpha", main='Cauchy')
    abline(v=1, col='red')
    hist(-t_boots[2,], breaks=30, xlab="estimated alpha", main="Student's t (df=1.5)")
    abline(v=1.5, col='red')
    hist(-gaussian_boots[2,], breaks=30, xlab="estimated alpha", main='Gaussian')
dev.off()
```

### 2023-11-30 writeup.tex
more responses

@@
 This suggests a mixture model for the effect size distribution:
-perhaps the distribution of effect sizes within each gene is Normal, \revpoint{1}{13}
-but with a standard deviation that varies with ``proximity'' to the trait;
-such mixtures can easily have large proportions of extreme values.
+perhaps a trait is affected by mutations in a number of genomic regions,
+and the distribution of effect sizes within each region is Normal, \revpoint{1}{13}
+but with a standard deviation that varies with ``proximity'' of that region to the trait
+(e.g., regions that are further ``upstream'' have higher standard deviations).
+Such mixtures can easily have large proportions of extreme values.
 
@@
 so that e.g., for $\alpha=1$ a positive proportion of the sum of $n$ entries will be contributed by a single one.
-On the other hand, the largest of $n$ Gaussians is of order $\sqrt(\log n)$. \revpoint{2}{1}
+On the other hand, the largest of $n$ Gaussians is of order $\sqrt{\log n}$. \revpoint{2}{1}
 
 This suggests exploring whether stable distributions might reasonably stand in for the Gaussian
-in the infinitesimal model. \revpoint{1}{1} \revpoint{1}{2}
+in the infinitesimal model. \revpoint{1}{1}
 In this paper, we make preliminary explorations in that direction:
@@
 The immediate answer to this will be ``no'' --
-in fact, we will see that independence of offsprings' deviations and midparent values \emph{implies} a Gaussian distribution in some sense.
+in fact, we will see that independence of offsprings' deviations and midparent values \emph{implies} a Gaussian distribution for the offsprings' deviations, in some sense. \revpoint{1}{2}
 This independence is a key reason why the infinitesimal model is tractable,
@@
         \textbf{(C)} number of cases.
+        \textbf{(D)} shows the number of SNPs and number of cases by phenotype.
         \label{fig:exponent_hist}
@@
 \end{figure}
+
+A possible concern is that tail exponents are difficult to estimate accurately, \revpoint{1}{4}
+but simulation-based estimates of accuracy (Supplemental Figure~\ref{fig:power})
+suggest that our estimates are unbiased but only accurate to within roughly $\pm 0.4$,
+so the distribution of Figure~\ref{fig:exponent_hist} should not be taken as
+a direct estimate of the underlying distribution of $\alpha$ values,
+but is robust evidence for exponents strictly between 1 and 2 (and probably around 1.5).
 
@@
         plotted against the midpoint of the corresponding quantile bin $(q_{k-1},q_k)$,
-        for $(a,b)$ chosen to span the 5\% of midparents centered on:
-        (red) the 10\% quantile,
-        (black) the median, and
-        (green) the 90\% quantile of midparent value.
+        for $(a,b)$ chosen to span the three segments of 5\% of
+        ``low'', ``medium'', and ``high'' midparents, i.e.:
+        midparents between (red) the 7.5\% and 12.5\% quantiles,
+        (black) the 47.5\% and 52.5\% quantiles, and
+        (green) the 87.5\% and 92.5\% quantiles of midparent value.
         \label{fig:seg_noise}
@@
         As in Figure~\ref{fig:trait_distrns},
-        except under a model of stabilizing selection (see text for details).
+        except under a model of stabilizing selection
+        (see text for details,
+        and note that the scale of the underlying mutations do not match
+        neutral simulations, so distributions are not directly comparable
+        to Figure~\ref{fig:trait_distrns}).
         \label{fig:sel_trait_distrns}
@@
     \begin{center}
+        \includegraphics{snp_effects/power_demo}
+    \end{center}
+    \caption{
+        Distributions of estimated tail exponents using the same estimator
+        as in Figure~\ref{fig:exponent_hist},
+        for 1000 independent draws from
+        \textbf{(left)} the Cauchy distribution (true $\alpha=1$),
+        \textbf{(center)} the Student's $t$ distribution with 1.5 degrees of freedom (true $\alpha=1.5$), and
+        \textbf{(right)} the Gaussian distribution (true $\alpha$ is undefined,
+        as the tail decays super-polynomially).
+        Note that the typical number of SNPs per trait (shown in Figure~\ref{fig:exponent_hist})
+        is around 1000.
+        \label{fig:power}
+    }
+\end{figure}
+
+\begin{figure}
+    \begin{center}
         \includegraphics{sims/selected_seg_noise_small}

### 2023-11-30 snp_effects/plot_results.R

@@ -90,15 +90,17 @@ with(results, {
 dev.off()
 
 pdf(file="results_10.pdf", width=6, height=4, pointsize=10)
-layout(matrix(c(1,2,1,3), nrow=2))
+layout(matrix(c(1,2,1,3,1,4), nrow=2))
 with(subset(results, !pruned), {
     par(mar=c(4, 4, 1, 1), mgp=c(2.1,1,0))
     hist(-slope, breaks=50, xlab='tail exponent', main='')
-mtext("(A)", 3, adj=-0.1, line=0)
+mtext("(A)", 3, adj=-0.07, line=0)
     plot(nsnps, -slope, pch=20, cex=0.5, col=adjustcolor('black', 0.5), log='x', xlab="number of SNPs", ylab='tail exponent')
 mtext("(B)", 3, adj=-0.25, line=0)
     plot(n_cases, -slope, pch=20, cex=0.5, col=adjustcolor('black', 0.5), log='x', xlab="number of cases", ylab='tail exponent')
 mtext("(C)", 3, adj=-0.25, line=0)
+    plot(n_cases, nsnps, pch=20, cex=0.5, col=adjustcolor('black', 0.5), log='xy', xlab="number of cases", ylab='number of SNPs')
+mtext("(D)", 3, adj=-0.25, line=0)
 })
 dev.off()
 

### 2023-12-01 review-responses.tex
minor add
Minor edits

@@
 
-\reply{
+\reply{This  is almost verbatim from the abstract of Turelli 2017:
+\begin{quote}
+This commentary distinguishes three nested approximations, referred to as “infinitesimal genetics,” “Gaussian descendants” and “Gaussian population,” each plausibly called “the infinitesimal model.” 
+\end{quote} 
 }
@@
 \reply{
+    Good idea; we have done this \revref.
 }
@@
 
-\begin{point}{}
+\begin{point}{\revref}
     Last inequality on Page 13: should the subscript be $j$ instead of $i$?
@@
 
-\reply{
+\reply{It should be $\ell$, and is now corrected \revref.  Thanks for noticing the discrepancy; is an artefact of an earlier version which used $i$ for an index on loci.
 }
@@
 
-\begin{point}{}
+\begin{point}{\revref}
 Third display equation in Sect 4.1 \revref: The index k in the product should start at k = 0.
@@
 
-\reply{
+\reply{Thank you for noticing this.  We have corrected this, and the subsequent indices \revref.
 }
 
-\begin{point}{}
+\begin{point}{\revref}
 p 8, last line: Delete ‘to’
@@
 
-\reply{
+\reply{Done \revref.
 }

### 2023-12-01 writeup.tex
minor add
More minor edits.

@@
 so that e.g., for $\alpha=1$ a positive proportion of the sum of $n$ entries will be contributed by a single one.
-On the other hand, the largest of $n$ Gaussians is of order $\sqrt{\log n}$. \revpoint{2}{1}
+On the other hand, the largest of $n$ Gaussians is of order $\sqrt{\log n}$. \revpoint{2}{1} \revpoint{1}{11}
 
@@
     &= \lim_{n \to \infty} F^{(n)}(\varphi)(\theta) \\
-    &= \lim_{n \to \infty} \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}\prod_{k=1}^{n-1}\Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}} .
+    &= \lim_{n \to \infty} \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}\prod_{k=0}^{n-1}\Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}} .
 \end{align*}
@@
 % \[
-% 	F^{(n)}(\varphi)(\theta) =  {\textstyle \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}}\prod_{k=1}^{n-1} {\textstyle \Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}},
+% 	F^{(n)}(\varphi)(\theta) =  {\textstyle \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}}\prod_{k=0}^{n-1} {\textstyle \Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}},
 % \]
@@
 % \[
-% 	\phi(\theta) = \lim_{n \to \infty} \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}\prod_{k=1}^{n-1}\Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}
+% 	\phi(\theta) = \lim_{n \to \infty} \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}\prod_{k=0}^{n-1}\Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}
 % \]
@@
 	\Log{F^{(n)}(\varphi)(\theta)} = {\textstyle 2^{n} \psi\left(\frac{\theta}{2^{n}}\right)} 
-		+ \sum_{k=1}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}.
+		+ \sum_{k=0}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}.
 \]
 We briefly explore what possible limiting values are valid characteristic functions.
-We refer to the first term on the left hand side as the \emph{reproduction term} and the remaining sum as the \emph{noise term}, and only consider to the case when both have well-posed limits independent of the other.
-\llabel{limit_remark}
+We refer to the first term on the left hand side as the \emph{reproduction term} and the remaining sum as the \emph{noise term}, and only consider the case \revpoint{2}{8} when both have well-posed limits independent of the other.
+\llabel{limit_remark} % We don't need to limit to this caee.
 
@@
 First, consider the noise terms.
-A necessary condition for $\sum_{k=1}^{\infty} 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)$ to converge is that $2^{k} \Psi\left(\frac{\theta}{2^{k}}\right) \to 0$; since $\Psi(0) = 0$, this implies that if it exists, $\Psi'(0+) = 0$.
+A necessary condition for $\sum_{k=0}^{\infty} 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)$ to converge is that $2^{k} \Psi\left(\frac{\theta}{2^{k}}\right) \to 0$; since $\Psi(0) = 0$, this implies that if it exists, $\Psi'(0+) = 0$.
 In particular, as we would expect, if $R$ has a mean, it must be 0.
@@
 \begin{align*}
-	 \sum_{k=1}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}
-	 &=  \sum_{k=1}^{n-1} 2^{k} {\textstyle \left(- \frac{\eta^{2}}{2}  \frac{\theta^{2}}{2^{2k}}
+	 \sum_{k=0}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}
+	 &=  \sum_{k=0}^{n-1} 2^{k} {\textstyle \left(- \frac{\eta^{2}}{2}  \frac{\theta^{2}}{2^{2k}}
 		+ r\left(\frac{\theta}{2^{k}}\right)\frac{\theta^{2}}{2^{2k}}\right)}\\
-	&=- \eta^{2} \theta^{2} (1 - 2^{-n+1})  + \sum_{k=1}^{n-1} {\textstyle r\left(\frac{\theta}{2^{k}}\right)\frac{\theta^{2}}{2^{k}}}
+	&=- \eta^{2} \theta^{2} (1 - 2^{-n+1})  + \sum_{k=0}^{n-1} {\textstyle r\left(\frac{\theta}{2^{k}}\right)\frac{\theta^{2}}{2^{k}}}
 \end{align*}
@@
 \begin{align*}
-	 \sum_{k=1}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}
-	 &=  \sum_{k=1}^{n-1} -2^{(1-\alpha)k} c|\theta|^{\alpha} {\textstyle\left(1-i \beta\sgn\left(\frac{\theta}{2^{k}}\right)
+	 \sum_{k=0}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}
+	 &=  \sum_{k=0}^{n-1} -2^{(1-\alpha)k} c|\theta|^{\alpha} {\textstyle\left(1-i \beta\sgn\left(\frac{\theta}{2^{k}}\right)
 	 	 \omega_{\alpha}\left(\frac{\theta}{2^{k}}\right)\right)}\\
@@
 \]
-If $\gamma > 0$, then we recognize $e^{\alpha(\theta)} = e^{im \theta -\gamma|\theta|}$ as the characteristic function of a Cauchy distribution with median 0 and median absolute deviation $\gamma$, whereas if $\gamma = 0$, then $e^{\alpha(\theta)} = e^{im \theta}$ is the Fourier transform of a Dirac mass at $m$, \ie the characteristic function for a monomorphic population with all individuals having trait value equal to $m$.
-(And, $\gamma < 0$ does not give a valid characteristic function.)
+If $\gamma > 0$, then we recognize $e^{\psi(\theta)} = e^{im \theta -\gamma|\theta|}$ as the characteristic function of a Cauchy distribution with median 0 and median absolute deviation $\gamma$, whereas if $\gamma = 0$, then $e^{\psi(\theta)} = e^{im \theta}$ is the Fourier transform of a Dirac mass at $m$, \ie the characteristic function for a monomorphic population with all individuals having trait value equal to $m$.
+
+% This might seem a bit roundabout, but the point is that we can prove that $e^{\psi(\theta)}$ is -- in this case -- a characteristic function, so we don't need to lose generality by assuming that it is one.
+Our goal is to ask when $\lim_{n \to \infty} \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}\prod_{k=0}^{n-1}\Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}$ is a valid characteristic function, and to see what it can be.
+To exclude the case $\gamma < 0$, we first observe that if, as we assume, $\varphi(\theta)$ is the characteristic function of some random variable, say $X$, then $\varphi_{n}(\theta) = \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}} = e^{2^{n}\psi\left(\frac{\theta}{2^{n}}\right)}$ is the characteristic function of $2^{-n} \sum_{i = 1}^{2^{n}} X_{i}$, where $X_{1},\ldots,X_{2^{n}}$ are i.i.d.
+copies of $X$; now, L\'evy's continuity theorem (see \eg Theorem 2.3.4 in \cite{Durrett2005}) tells us that if (i) $\phi_{n}(\theta)$ has a point-wise limit, say $\phi_{\infty}(\theta)$, and (ii) $\phi_{\infty}(\theta)$ is continuous at 0, then $\phi_{\infty}(\theta)$ is the characteristic function of \emph{some} random variable; in particular, if $\alpha(\theta)$ exists and is bounded (\eg if the left and right-hand side derivatives exist, as above), then $\psi(\theta) = \alpha(\theta)\theta$ is continuous at 0, and hence so is $e^{\psi(\theta)}$.
+Thus, in this case $e^{\psi(\theta)}$ is itself a characteristic function.
+Moreover, being a characteristic function, $e^{\psi(\theta)}$ must satisfy $|e^{\psi(\theta)}| = e^{\Re \psi(\theta)} \leq 1$, whence $\gamma \geq 0$.
+
 % % I think we've shown this characteristic function has to be exp(im\theta - \gamma|\theta|);
@@
 converges weakly as $M \to \infty$ to a Poisson random variable with mean $2/(\pi \epsilon)$ (see \eg Theorem 2.6.1 in \cite{Durrett2005}).
-In particular, for large values of $M$ and fixed $\epsilon > 0$,
+In particular, for large values of $M$ and fixed $\epsilon > 0$,\revpoint{1}{17}
 \begin{align*}
-    \P\left\{\max_{i=1}^{M} \frac{\eta^{[j]}_{\ell}}{\bar{\gamma} M} > \epsilon\right\}  \sim 1-e^{-\frac{2}{\pi \bar{\gamma} \epsilon}},
+    \P\left\{\max_{\ell=1}^{M} \frac{\eta^{[j]}_{\ell}}{\bar{\gamma} M} > \epsilon\right\}  \sim 1-e^{-\frac{2}{\pi \bar{\gamma} \epsilon}},
 \end{align*}

### 2023-12-02 review-responses.tex
peter edits

@@
     sections 3 and 4 explore the consequences of different distributions
-    of the Mendelian segregation term (which derives from the sum of effects),
+    of the Mendelian sampling term (which derives from the sum of effects),
     not the population distribution.
@@
 
-\reply{This  is almost verbatim from the abstract of Turelli 2017:
+\reply{
+    We agree that \citet{turelli2017commentary} is clear about when the Gaussian emerges.
+    It also seems quite clear to us that what we say in this paragraph agrees
+    with \citet{turelli2017commentary}, who says in his Abstract that
     \begin{quote}
@@
     \end{quote} 
+    We think that the underlying issue here may be about motivation for the section.
+    To make the motivation, and the point about the Gaussian population assumption more clear,
+    we've added some explanatory text \revref.
+    Also note that this section does \emph{not} make a Gaussian-population (or, non-Gaussian population) assumption;
+    to make this more clear we've changed the phrase later in this paragraph
+    from ``non-Gaussian model'' to ``model with non-Gaussian Mendelian sampling terms''.
 }
@@
 \reply{
+    Good suggestion; we've added more clarifying points;
+    see \llname{clarification1}
 }

### 2023-12-02 writeup.tex
peter edits

@@
 
-As observed in \cite{turelli2017commentary}, \revpoint{1}{8}
+As observed in \cite{turelli2017commentary},
 there are at least three interpretations of what is called ``the infinitesimal model''. 
-The maximalist of these, what \cite{turelli2017commentary} calls the Gaussian-population approximation, assumes a large number of small-effect loci,  parental trait-independent Gaussian Mendelian sampling terms  (as rigorously justified in \cite{barton2017infinitesimal,barton2022infinitesimal}), and, moreover, that  population trait values are in deterministic mutation-selection balance, with a Gaussian empirical distribution
+The maximalist of these, what \cite{turelli2017commentary} calls the Gaussian-population approximation,
+assumes parental trait-independent Gaussian Mendelian sampling terms,
+and, moreover, that  population trait values are in deterministic mutation-selection balance, with a Gaussian empirical distribution
 (\ie the \emph{fraction} of the population with trait value in $[z,z+dz)$
@@
 for suitable choices of the empirical mean and variance $m$ and $\sigma^{2}$).
+As also discussed in \citet{turelli2017commentary}, \revpoint{1}{8}
+this assumption is rather fragile --
+a Gaussian distribution for the population only emerges under weak or Gaussian selection.
+However, the model has been used in influential work.
+In principle, another factor that could produce a non-Gaussian population distribution
+is if the Mendelian sampling terms are not Gaussian, as we consider here --
+so, it is interesting to look for other situations where the population distribution is analytically tractable.
+
 Leaving aside 
@@
 the question of whether or not some combination of effect sizes, dominance, and epistasis can justify non-Gaussian, trait-independent,  Mendelian sampling terms, we might reasonably ask when (or if)
-deterministic mutation-selection balance is possible in a non-Gaussian model
-that retains independence of the Mendelian sampling terms.
+deterministic mutation-selection balance is possible in a model with non-Gaussian Mendelian sampling terms
+that retains independence of these terms.
 (Indeed, above we have put substantial restrictions on what models of inheritance
@@
 $ \Phi(\theta) = \mathbb{E}\left[e^{i\theta R}\right] $
-are the characteristic functions of $Z^{j}$ and $R$, respectively.  In what follows, we will explore some non-Gaussian solutions $\phi(\theta)$ to \eqref{eqn:MSB_CF}.
+are the characteristic functions of $Z^{j}$ and $R$, respectively. 
+In what follows, we will explore some non-Gaussian solutions $\phi(\theta)$ to equation \eqref{eqn:MSB_CF}:
 
@@
 First, consider the case without selection, so $\mu$ is a constant
-which we may take equal to 1.  Then, our condition on the characteristic function simplifies to
+which we may take equal to 1. 
+(To remind the reader, our goal is therefore to describe \llabel{clarification1}
+what forms the population distribution might take in a neutral model
+if we assume only that the Mendelian sampling terms are independent of parental values,
+and nothing else about the distribution of these terms or the initial population distribution.) 
+Then, our condition on the characteristic function simplifies to
 \begin{equation}\label{eq:CF_neutral}
@@
 
-The crucial step in the derivation of the infinitesimal model in \cite{barton2017infinitesimal} is to show that, following the arguments in \cite{Fisher1918}, knowing an individual's trait value gives vanishingly little information about the contribution of any given locus as $M \to \infty$, which in turn allows one to show that the within-family Mendelian segregation variance is independent of the parental traits.
+The crucial step in the derivation of the infinitesimal model in \cite{barton2017infinitesimal} is to show that, following the arguments in \cite{Fisher1918}, knowing an individual's trait value gives vanishingly little information about the contribution of any given locus as $M \to \infty$, which in turn allows one to show that the within-family Mendelian sampling term is independent of the parental traits.
 Here, we will consider how conditioning an individual's trait value affects the law of a single allelic contribution in the Cauchy setting described above.
@@
         \textbf{(B,E)} Histograms of the Mendelian sampling term.
-        \textbf{(B,E)} Enrichment in quantiles of the Mendelian sampling term conditional on midparent value.
+        \textbf{(C,F)} Enrichment in quantiles of the Mendelian sampling term conditional on midparent value.
         To compute the latter,
@@
 (i.e., there are loci of large effect).
+Also, we show that removing big effects and using the infinitesimal model for the remainder is good \comment{todo incorporate this}.
 

### 2023-12-02 review-responses.tex
done with rev 1 points

@@
     see \llname{clarification1}
+    and \llname{clarification2}.
 }
@@
 \begin{point}{}
-    Last inequality on Page 13: Is this the distribution of the largest of the $M$ alleles carried by an individual, conditioned on the trait value $Z$ of the individual (and this is independent of $Z$?), or is it just the largest of $M$ iid draws from a Cauchy distribution?
+    Last inequality above \revref: Is this the distribution of the largest of the $M$ alleles carried by an individual, conditioned on the trait value $Z$ of the individual (and this is independent of $Z$?), or is it just the largest of $M$ iid draws from a Cauchy distribution?
 If the former, then it'd be good to say this explicitly.
@@
 \reply{
+    It is both --
+    the calculation is, as the reviewer points out, about the largest of $M$ iid Cauchys,
+    but those are the alleles carried by a parent in the initial generation.
+    Furthermore, as explained at the beginning of this section \llname{crucial_step},
+    this calculation about iid effects is indeed the key step
+    (in \citet{Fisher1918} and \cite{barton2017infinitesimal})
+    in describing the distribution of trait values in offspring.
+    We've added an sentence that we think should remind the reader
+    why we have in fact answered the question \revref.
 }
@@
 \reply{
+    Good idea, we have added some citations \revref.
+    However, note that what exactly are the ``original references'' is not so clear:
+    ``\textit{Indeed, its origins are lost in the mists of time (M. Bulmer, W.G. Hill, pers.  comm.).}''
+    \citep[from][]{barton2017infinitesimal}
 }
@@
 \reply{
+    Perhaps it is better now? \revref
 }
@@
 \begin{point}{}
-    Discussion: In the context of theory combining large and small-effect loci, the 2008 paper by Chevin and Hospital: ``Selective Sweep at a Quantitative Trait Locus in the Presence of Background Genetic Variation'' and the 1983 work by Lande: ``The response to selection on major and minor mutations affecting a metrical trait'' may also be somewhat relevant.
+    Discussion: In the context of theory combining large and small-effect loci, \citet{chevin2008selective}: ``Selective Sweep at a Quantitative Trait Locus in the Presence of Background Genetic Variation'' and \citet{lande1983response}: ``The response to selection on major and minor mutations affecting a metrical trait'' may also be somewhat relevant.
 \end{point}
@@
 \reply{
+    Thanks! We've incorporated these \revref.
 }

### 2023-12-02 writeup.tex
done with rev 1 points

@@
 and a covariance independent of the parental trait values (but that depends on how they are related).
-\citet{barton2017infinitesimal} give a history of the model
-and a formal proof of conditions under which this would be a good approximation to reality,
+\citet{barton2017infinitesimal} give an overview of the complex history of the model, \revpoint{1}{12}
+which is used in or adjacent to a good deal of theoretical work on quantitative genetics
+\citep[e.g.,][]{fisher1918correlation,bulmer1971effect,lande1975maintenance}.
+\citet{barton2017infinitesimal} focus on giving a formal proof of conditions
+under which this would be a good approximation to reality,
 which they extended to include dominance in \citet{barton2022infinitesimal}.
@@
 and interactions between alleles, between loci, and with the environment are often important in practice.
-In fact, many traits claimed to be genetically determined \revpoint{1}{14}
-in fact only appear heritable through non-genetic means,
-e.g., systematic inequalities in resource distribution partitioned by class or race.
-% \citep[e.g.,][]{davenport1916hereditary}.
+In fact, many traits that have been claimed to be mostly genetically determined \revpoint{1}{14}
+only appear heritable through non-genetic means,
+e.g., systematic inequalities in resource distribution partitioned by class or race
+\citep[e.g.,][]{davenport1916hereditary}.
 In this paper, we follow the literature in working with the relatively simple purely additive case
@@
 We briefly explore what possible limiting values are valid characteristic functions. 
-We refer to the first term on the left hand side as the \emph{reproduction term} and the remaining sum as the \emph{noise term}, and only consider the case \revpoint{2}{8} when both have well-posed limits independent of the other.
-\llabel{limit_remark} % We don't need to limit to this caee.
+We refer to the first term on the left hand side as the \emph{reproduction term} and the remaining sum as the \emph{noise term}, and only consider the case \revpoint{2}{8}
+when both have well-posed limits independent of the other. \llabel{limit_remark} % We don't need to limit to this caee.
+Biologically,
+the ``noise term'' is the contribution to variation in the population
+generated by the Mendelian sampling terms (\ie, by recombination and mutation),
+while the ``reproduction term'' is the contribution from variation in the initial generation.
+\llabel{clarification2}
 
@@
 
-Unfortunately, once selection is included, \eqref{eqn:MSB_CF} is less amenable to the sort of simple arguments above.
+Unfortunately, once selection is included, equation \eqref{eqn:MSB_CF} is less amenable to the sort of simple arguments above.
 We will content ourselves with presenting two examples of stabilizing selection, one already familiar to quantitative genetics, the other a Cauchy solution obtained by trial-and-error, which illustrates how selection can stabilize an otherwise unviable solution.
@@
 
-The crucial step in the derivation of the infinitesimal model in \cite{barton2017infinitesimal} is to show that, following the arguments in \cite{Fisher1918}, knowing an individual's trait value gives vanishingly little information about the contribution of any given locus as $M \to \infty$, which in turn allows one to show that the within-family Mendelian sampling term is independent of the parental traits.
+The crucial step in the derivation of the infinitesimal model in \cite{barton2017infinitesimal} is to show that, \llabel{crucial_step}
+following the arguments in \citet{Fisher1918}, knowing an individual's trait value gives vanishingly little information about the contribution of any given locus as $M \to \infty$, which in turn allows one to show that the within-family Mendelian sampling term is independent of the parental traits.
 Here, we will consider how conditioning an individual's trait value affects the law of a single allelic contribution in the Cauchy setting described above.
@@
 
+To summarize, the key argument in \citet{Fisher1918} and \cite{barton2017infinitesimal}  \revpoint{1}{10}
+requires that knowing the sum of parental effects does not give much information 
+about any particular genotype;
+we have shown here that this does not hold for Cauchy effect sizes
+(at least when these are independent).
 In particular, there is at least one locus at which knowing the parental trait gives one information about the effect size (and there are on the order of $\Oh\left(\frac{1}{\epsilon}\right)$ that give ``an $\epsilon$ of information''), and these large effect loci in turn give rise to correlations between
@@
 that include contributions of both ``large'' and ``small'' effect loci.
+Various authors have taken the approach of separating contributions of large- and small-effect loci
+(e.g., \citet{lande1983response} and \citet{chevin2008selective}). \revpoint{1}{20}
 In practice, breeders have long embellished the ``animal model'' (essentially, the infinitesimal model)

### 2023-12-02 review-responses.tex
peter edited through

@@
 We are writing to submit a revision of our manuscript, ETCETERA
+
+In particular, please note that
+we provide a copy of the paper with both highlighted changes
+and hyperlinked reviewer comments.
 
@@
     see \llname{clarification1}
-    and \llname{clarification2}.
+    and \llname{clarification2}
+    (and \llname{clarification3}).
 }
@@
 \begin{quote}
-The authors investigate the possibility of extending results of Barton et al (2017) on the
+The authors investigate the possibility of extending results of \citet{barton2017infinitesimal} on et al (2017) on the
 justification of the infinitesimal model to distributions of effect sizes with a large tail (i.e.,
@@
 
-We now discuss selection XXX WHERE
+Thanks for the close reading.
+We didn't discuss selection here because the question of whether selection contributes
+to the distribution of segregating allelic effects
+is somewhat orthogonal to the question at hand here:
+in other words, the infinitesimal model describes (an approximation for)
+the \emph{results of} random reassortment (etcetera) of whatever ancestral variation is present,
+and is therefore relatively agnostic to how that ancestral variation got there.
+This works because the infinitesimal model describes trait variation after reproduction
+but before selection.
+Note that, for instance, there is no obstacle to there being a correlation between effect sizes and allele frequencies,
+as one might expect under selection.
+However, we have added some clarification about this --
+see \revreffull{1}{3} and \llname{discussion_DFEs}.
 
 \begin{quote}
-In Section 3, the authors show a kind of negative result concerning the possibility of $\alpha$-
-stable distributions with $\alpha < 2$ (Proposition 1): If the deviations of the offspring trait
+In Section 3, the authors show a kind of negative result concerning the possibility of $\alpha$-stable 
+distributions with $\alpha < 2$ (Proposition 1): If the deviations of the offspring trait
 from the midparent value (here called the Mendelian sampling term, $R$) are independent
-of the midparent value $\bar Z$ (part of the core result in Barton et al (2017)), then the joint
-distribution of $R$ and $\bar Z$ is Gaussian. Thus, this independence is not compatible with ‘$\alpha$-
-stable laws’. In fact, as they show (by example) and discuss more generally in Sect. 4, a
+of the midparent value $\bar Z$ (part of the core result in \citet{barton2017infinitesimal}), then the joint
+distribution of $R$ and $\bar Z$ is Gaussian. Thus, this independence is not compatible with ‘$\alpha$-stable 
+laws’. In fact, as they show (by example) and discuss more generally in Sect. 4, a
 single polymorphic large-effect locus causes dependence of $R$ and $\bar Z$.
+\end{quote}
 
+We'd like to remind the reviewer of the caveat that this result
+also depends on a bit more -- independence of inherited and non-inherited contributions
+(e.g., as under an additive model of unlinked loci).
+
+\begin{quote}
 In Section 4, a version of the infinitesimal model is investigated that proposes a much
@@
 
-We don't actually assume a Gaussian distribution....
+We're glad the reviewer liked this section,
+but perhaps our introduction to the section was misleading --
+in fact we do \emph{not} assume a Gaussian population distribution,
+or any population distribution.
+Instead, we \emph{ask} which distributions might be population distributions
+under this generalized model,
+using arguments similar to those found in the literature.
+We've tried to make this more clear -- see additional text at \revreffull{1}{8}
+and \llname{clarification3}
+(and our response to Reviewer 1, point 8).
 
 \begin{quote}
-The authors use a simplistic Moran model to investigate these questions. However, mu-
-tation doesn’t seem to be modeled explicitly. Is it mutation if a parent is replaced by an
-offspring of the type described in the display equation on p 7? Maybe yes, but not neces-
-sarily because of segregation and recombination contribute substantially to $R$. Doesn’t the
-present assumption imply that the mutation rate is one and that mutation is modeled as a
-kind of diffusion? In the latter case, a Gaussian distribution can result. With explicit mu-
-tation models, where mutations occur per locus and a finite number of loci determines the
+The authors use a simplistic Moran model to investigate these questions. However, mutation 
+doesn’t seem to be modeled explicitly. Is it mutation if a parent is replaced by an
+offspring of the type described in the display equation at \llname{offspring_eqn}?
+Maybe yes, but not necessarily because of segregation and recombination contribute substantially to $R$.
+Doesn’t the present assumption imply that the mutation rate is one and that mutation is modeled as a
+kind of diffusion? In the latter case, a Gaussian distribution can result. With explicit
+mutation models, where mutations occur per locus and a finite number of loci determines the
 trait additively, a Gaussian distribution never results (and in no known limit; see above).
-How does population size enter? Is this a deterministic Moran model? Anyway, the results
-in Sect 4 are of interest because they strongly constrain the possibilities to justify such
-‘Gaussian-population approximation’ by any infinitesimal-like model (essentially to mix-
-tures of Cauchy distributions. The noted relation to blending inheritance is intriguing!).
-My main concern is that especially this section needs are clearer description of the model,
-and clear distinction between this ‘maximalist’ model (which has little theoretical sup-
-port) and the model investigated by Barton et al (2017). Also a much clearer summary
-and discussion of the implications of the results shown here are needed. It should not be
-assumed that readers have read Turelli (2017) and even less that they have studied Barton
-et al (2017) in considerable detail!
+How does population size enter? Is this a deterministic Moran model?
 \end{quote}
 
-We now explain stuff more...
+Mutation would enter into $R$, a point we now make \llname{mutation_note},
+but without detail, as we are aiming here for a high-level and quick calculation,
+not a detailed investigation.
+In fact, we thought a bit about how to set up an explicit genotype-to-phenotype map
+in which the Mendelian sampling terms are non-Gaussian and independent of parental traits,
+but did not arrive at a concrete solution --
+we think this is probably possible,
+but as we know from Proposition~\ref{prop:parental_contribs},
+cannot be a simple additive model.
+Investigating how exactly a concrete model of mutation would affect things is beyond the scope of this paper.
+
+\begin{quote}
+Anyway, the results
+in Sect 4 are of interest because they strongly constrain the possibilities to justify such
+‘Gaussian-population approximation’ by any infinitesimal-like model (essentially to
+mixtures of Cauchy distributions. The noted relation to blending inheritance is intriguing!).
+My main concern is that especially this section needs are clearer description of the model,
+and clear distinction between this ‘maximalist’ model (which has little theoretical
+support) and the model investigated by Barton et al (2017). Also a much clearer summary
+and discussion of the implications of the results shown here are needed. It should not be
+assumed that readers have read Turelli (2017) and even less that they have studied
+\citet{barton2017infinitesimal} in considerable detail!
+\end{quote}
+
+Thanks for the suggestions.
+We also thought the section was useful (and also intruiging!),
+which is why we left the section in.
+(Also note it motivates the form of selection in our simulations.)
+We share the reviewer's concern about context,
+and have added some additional text
+while trying not to get into the weeds too much
+-- in particular, see the new text around \llname{clarification3}.
 
@@
 noise’ (and presumably for any other limit law with $\alpha < 2$). In the absence of giving the
-estimate in App D in Barton et al, the explanation at the bottom of p 12 is difficult to
+estimate in App D in Barton et al, the explanation around \llname{approx_eqn} is difficult to
 follow. Please provide their estimate for easier comparison (do not assume that readers
@@
 \end{quote}
+
+\comment{TODO}
 
@@
 Please clarify the model. It also needs to be pointed out clearly that such results seem to
-be at variance with all (?) explicit multilocus models that have been studied so far
+be at variance with all (?) explicit multilocus models that have been studied so far.
+\end{quote}
 
+Honestly, we're confused by this comment.
+The mutation rate is $10^{-9}$ per bp on a genome of length $10^8$,
+as stated at \llname{mutation_rate} -- perhaps the reviewer missed that paragraph?
+We're also unclear what exactly is meant by ``at variance with all \ldots models'' --
+it's true that no-one else has (probably?) simulated this exact model,
+but when it comes to simulations, this is often true?
+We're not sure which specific aspects of difference the reviewer has in mind pointing out.
+
+\begin{quote}
 In summary, this ms contains several interesting aspects and will deserve publication after
@@
 \end{quote}
+
+Thanks! We hope our improvements
+have made everything at least somewhat clearer.
 
@@
 \begin{point}{\revref}
-What means ‘nearby alleles’ ? Do you mean physically linked
+What means `nearby alleles'? Do you mean physically linked
     loci? Another \llname{loci_polymorphic}: alleles are not polymorphic; loci can be polymorphic.
@@
 \reply{
-    Good points; both are fixed: \revref and \llname{loci_polymorphic}.
+    Good points; both are fixed: \revref{} and \llname{loci_polymorphic}.
 }
@@
 \begin{point}{}
-p 7, Sect 4: As noted above, please describe this model in more detail to distinguish it
-from that in Barton et al (2017), which makes much weaker assumptions. Note that it is
+Sect \ref{sec:abnormal}: As noted above, please describe this model in more detail to distinguish it
+from that in \citet{barton2017infinitesimal}, which makes much weaker assumptions. Note that it is
 not justifiable except under quite extreme assumptions.
@@
 \reply{
+    We've added more text to hopefully
+    put the reader in the right frame of mind:
+    see especially \llname{clarification3}.
 }
@@
 \begin{point}{}
-p 10: It might be useful to mention earlier that the Cauchy distribution is an archetypical
+Around \llname{cauchy_note}: It might be useful to mention earlier that the Cauchy distribution is an archetypical
 example for an $\alpha$-stable distribution. The enthusiasm among biologists may be limited
@@
 don’t know if there are examples in which the variance grows (significantly) as the sample
-size increases. In the last para of Sect 4.1, please specify ‘one will get convergence’. I
+size increases. In the last para of Sect 4.1 \revref, please specify ‘one will get convergence’. I
 suppose this refers again to $M \to \infty$.
@@
 \reply{
+    Good suggestion -- we've added the note about the Cauchy \llname{cauchy_note}.
+    (We haven't tried to make biologists more enthusiastic, though,
+    since we don't know of a situation in which strict blending inheritance seems terribly likely.)
+    As for the note about ``one will get convergence'' -- it is as the number of generations increases,
+    not as $M \to \infty$ (in fact, $M$ does not enter into this section at all);
+    we've rewritten this statement to remind the reader
+    about what's going on.
 }
@@
 \reply{
+    \comment{TODO}
 }
@@
 \begin{point}{\revref}
-Does ‘Lévy stable distribution’ refer to the family of $\alpha$-stable
+Does `Lévy stable distribution’ refer to the family of $\alpha$-stable
 distributions? (I don’t think it has not been used before.)

### 2023-12-02 writeup.tex
peter edited through

@@
 \end{abstract}
-
-\comment{ADD TO DISCUSSION, WHICH IS SHORT:
-    Make the point that another way to think about distibutions with non-finite variance
-    is that there's only a finite number of alleles above any fixed value;
-    and also the variance in any finite population will be finite.
-    (see reviewer comment, ``all the empirical trait distributions they find have mean and variance'')
-    Also discuss comparison to fat-tailed dispersal:
-    the question is, what distribution gives us a better model?
-    For dispersal, a fat-tailed distribution has in practice finite variance
-    but is a better model in that it describes noisiness better;
-    here, the analogy is something about polygenicity and noisiness of the process of adaptation.
-    If we're modeling non-polygenic (oligogenic) traits then
-    why stable as opposed to some other arbitrary distribution?
-    Well, the distribution of per-allele effect sizes might not be stable, but CLT tells us that maybe
-    stable gives a good approx to the process (as in Gaussian case).
-}
 
@@
 The proposition says, roughly, that if the parental alleles are unlinked and independently chosen,
-and meotic segregation is fair,
+and meiotic segregation is fair,
 then independence of Mendelian sampling term from midparent value
@@
 \section{Abnormal quantitative genetics?}
+    \label{sec:abnormal}
 
@@
 
+To summarize, in this section we will uncritically assume that \llabel{clarification3}
+the independence assumption of the infinitesimal model holds,
+but with an arbitrary distribution for the Mendelian sampling term,
+and ask which population trait distributions are possible.
+To do this, we work at a higher level,
+ignoring the genetic basis of the trait
+\citep[and certainly not using the specific set-up of][]{barton2017infinitesimal}.
+
 Concretely, consider a Moran process where an individual whose trait value is $Z$ dies at rate $\mu(Z)$
 and is replaced by the offspring of two randomly chosen individuals,
-i.e., a new individual with trait
+i.e., a new individual with trait \llabel{offspring_eqn}
 \begin{align*}
@@
 (independent of $Z^{1}$ and $Z^{2}$, with distribution to be specified). 
+(Note that mutation might enter into $R$, but as the underlying genetic basis is not specified,
+we do not attempt to disambiguate the various possible contributions.) \llabel{mutation_note}
 Mutation-selection balance demands that that
@@
 % Since the set of functions $\mathcal{S} = \{e^{i \theta x} : \theta \in \R\}$ is separating (see \eg \S 3.4 in \cite{Ethier+Kurtz1986}), it suffices to show that \eqref{eqn:MSB_condition} holds for $f \in \mathcal{S}$, \ie
+If the population is large enough,
+we can treat the trait distribution as changing deterministically.
 Taking the characteristic function of both these distributions,
@@
 \]
-If $\gamma > 0$, then we recognize $e^{\psi(\theta)} = e^{im \theta -\gamma|\theta|}$ as the characteristic function of a Cauchy distribution with median 0 and median absolute deviation $\gamma$, whereas if $\gamma = 0$, then $e^{\psi(\theta)} = e^{im \theta}$ is the Fourier transform of a Dirac mass at $m$, \ie the characteristic function for a monomorphic population with all individuals having trait value equal to $m$.
+If $\gamma > 0$, then we recognize $e^{\psi(\theta)} = e^{im \theta -\gamma|\theta|}$
+as the characteristic function of a Cauchy distribution with median 0 and median absolute deviation $\gamma$
+(\ie, the canonical $\alpha$-stable distribution with $\alpha=1$), \llabel{cauchy_note}
+whereas if $\gamma = 0$, then $e^{\psi(\theta)} = e^{im \theta}$ is the Fourier transform of a Dirac mass at $m$, \ie the characteristic function for a monomorphic population with all individuals having trait value equal to $m$.
 
@@
 Our goal is to ask when $\lim_{n \to \infty} \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}\prod_{k=0}^{n-1}\Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}$ is a valid characteristic function, and to see what it can be.
-To exclude the case $\gamma < 0$, we first observe that if, as we assume, $\varphi(\theta)$ is the characteristic function of some random variable, say $X$, then $\varphi_{n}(\theta) = \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}} = e^{2^{n}\psi\left(\frac{\theta}{2^{n}}\right)}$ is the characteristic function of $2^{-n} \sum_{i = 1}^{2^{n}} X_{i}$, where $X_{1},\ldots,X_{2^{n}}$ are i.i.d.
-copies of $X$; now, L\'evy's continuity theorem (see \eg Theorem 2.3.4 in \cite{Durrett2005}) tells us that if (i) $\phi_{n}(\theta)$ has a point-wise limit, say $\phi_{\infty}(\theta)$, and (ii) $\phi_{\infty}(\theta)$ is continuous at 0, then $\phi_{\infty}(\theta)$ is the characteristic function of \emph{some} random variable; in particular, if $\alpha(\theta)$ exists and is bounded (\eg if the left and right-hand side derivatives exist, as above), then $\psi(\theta) = \alpha(\theta)\theta$ is continuous at 0, and hence so is $e^{\psi(\theta)}$.
+To exclude the case $\gamma < 0$, we first observe that if, as we assume, $\varphi(\theta)$ is the characteristic function of some random variable, say $X$, then $\varphi_{n}(\theta) = \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}} = e^{2^{n}\psi\left(\frac{\theta}{2^{n}}\right)}$ is the characteristic function of $2^{-n} \sum_{i = 1}^{2^{n}} X_{i}$, where $X_{1},\ldots,X_{2^{n}}$ are i.i.d.\ copies of $X$; now, L\'evy's continuity theorem (see \eg Theorem 2.3.4 in \cite{Durrett2005}) tells us that if (i) $\phi_{n}(\theta)$ has a point-wise limit, say $\phi_{\infty}(\theta)$, and (ii) $\phi_{\infty}(\theta)$ is continuous at 0, then $\phi_{\infty}(\theta)$ is the characteristic function of \emph{some} random variable; in particular, if $\alpha(\theta)$ exists and is bounded (\eg if the left and right-hand side derivatives exist, as above), then $\psi(\theta) = \alpha(\theta)\theta$ is continuous at 0, and hence so is $e^{\psi(\theta)}$.
 Thus, in this case $e^{\psi(\theta)}$ is itself a characteristic function.
@@
 
-More generally, one will get convergence if the noise is a mixture of components with finite variance and stable laws as above, whereas the population distribution will be a mixture of a (possibly degenerate) Cauchy distribution and such noise terms.
+More generally, \revpoint{2}{9}
+the trait distribution in the population will converge as time proceeds
+if the noise is a mixture of components with finite variance and stable laws as above,
+whereas the population distribution will be a mixture of a (possibly degenerate) Cauchy distribution and such noise terms.
 Note that in the presence of noise, a pure Cauchy-distributed solution is \emph{not} possible in the neutral case.
@@
 \section{The difficulty with an abnormal quantitative genetics}
+    \label{sec:difficulty}
 
@@
 % Unlike the case of bounded allelic effects, the bound is independent of the trait value $Z^{[j]}$. 
+\llabel{approx_eqn}
 In the case of bounded allelic effects, the error introduced by approximating the effect-size distribution conditional on the trait value by the unconditioned distribution is proportional to $|Z^{[j]}|$, so that extreme trait values are most informative of the size of allelic effects,
@@
 \section{Simulations}
+    \label{sec:simulation}
 
@@
 the genome is of length $10^8$bp with a uniform recombination rate of $10^{-8}$ crossovers per bp;
-mutations occur at a rate of $10^{-9}$ per bp.
+mutations occur at a rate of $10^{-9}$ per bp. \llabel{mutation_rate}
 Each new mutation is independently assigned an ``effect''
@@
 In this paper,
-we have looked at the possibility of using a broader range of central limit theorems
+we have looked at the possibility of using a broader range of Central Limit Theorems
 to extend the infinitesimal model to traits for which 
 effect size distributions fall in the domain of attraction of the $\alpha$-stable distributions
-(i.e., a larger proportion of variance
-is explained by a smaller number of loci).
+(for which a larger proportion of variance
+would be explained by a smaller number of loci).
 Indeed, we found evidence that this is true of many real traits,
@@
 Next, we show that in fact the infinitesimal model \emph{cannot} describe non-Gaussian distributions
-\comment{XXX TODO}.
-A recurring theme in the rest of the paper is
-that just as the infinitesimal model is useful precisely because it ignores the genetic details,
-it cannot apply once some genetic details are important
-(i.e., there are loci of large effect).
-Also, we show that removing big effects and using the infinitesimal model for the remainder is good \comment{todo incorporate this}.
+(\ie, non-Gaussian deviation of offspring trait from midparent value)
+if the traits are determined additively from independent loci
+(for a precise statement, see Proposition~\ref{prop:parental_contribs}).
+The problem is large-effect alleles:
+as described in Section~\ref{sec:difficulty},
+if large-effect alleles are present, then naturally parental trait values give information about parental genotypes,
+and hence about the distribution of offspring trait deviations.
+This also suggests that the infinitesimal model might be a good approximation
+once the largest alleles are removed,
+an idea we find support for in theory (Section~\ref{sec:difficulty}) and simulation (Section~\ref{sec:simulation}).
+Despite having spent a fair bit of time describing why
+the infinitesimal model cannot strictly apply 
+to an additive model with non-Gaussian effect size distributions,
+we still found it useful to work out some of the consequences if we took it seriously:
+the calculations in Section~\ref{sec:abnormal}
+suggested a tractable form of stabilizing selection that we used in simulations,
+and revealed an intriguing (if not strongly biologically motivated)
+model in which blending inheritance retains heritable variance.
+
+For simplicity, we have focused on the $\alpha=1$ case (which includes the Cauchy distribution).
+However, examination of Figure~\ref{fig:exponent_hist} suggests that $\alpha=3/2$
+would be a more natural choice for an effect size distribution.
+(One could use the Holtsmark distribution,
+or Student's $t$ with 1.5 degrees of freedom,
+which has the same tail behavior.)
+This would be particularly interesting to investigate further,
+because such distributions are not so extreme as the Cauchy,
+as the largest out of $M$ observations grows only as $M^{2/3}$ --
+so, in a trait determined by $M$ loci,
+the largest-effect allele will contribute only about $1/M^{1/3}$ of the typical trait value
+(and hence the largest effects are less important for more polygenic traits).
 
@@
 
+\subsection*{Finite variance}
+
+A finite sample from any distribution will have finite variance.
+However, if the distribution itself has infinite variance,
+then the sample variance will increase without bound as one takes a larger sample.
+This cannot be literally true in a biological model, since traits are bounded
+by physical constraints.
+However, effect size distributions with infinite variance
+might still produce predictions that better match reality,
+if the shape of the distributions match reality over the relevant scale.
+A similar situation is found in the discussions around models of animal movement \citep{benhamou2007animals}
+-- clearly, the amount by which an albatross can move in a day is bounded,
+but random walks driven by $\alpha$-stable noise
+might in principle still better describe statistical properties of an albatross's typical path.
+However, deciding \emph{which} statistical properties of the path to consider
+is subject to interpretation,
+and encounters the statistically difficult problem of estimating tail exponents.
+
 \subsection*{Selection and effect size distributions} \llabel{discussion_DFEs}
+
 A substantial body of theoretical work has used various models and assumptions

### 2023-12-03 review-responses.tex
edit through

@@
 
-We are writing to submit a revision of our manuscript, ETCETERA
+We are writing to submit a revised version of our manuscript,
+``Large effects and the infinitesimal model''.
+In response to the reviewer's comments
+we have made quite a few additions to the text
+which we hope will make the paper more accessible to a wider audience.
+We hope you agree!
 
-In particular, please note that
-we provide a copy of the paper with both highlighted changes
+Along with our revised paper and the responses to reviewers,
+we are also submitting a PDF that has both changes to the paper since the first submission highlighted,
 and hyperlinked reviewer comments.
+(Line numbers in the standalone responses to reviewers
+refer to the PDF of the paper without changes highlighted,
+while line numbers in the merged document with changes highlighted are self-consistent.)
+
+\vskip 2em
 
@@
 
-XXX
+Thanks very much to the reviewers for their comments.
+We agree that our initial version of the manuscript didn't have as broad an envisioned audience
+as it should have.
+We've tried to improve this, by adding additional context and discussion --
+in particular, in Sections 3 and 4, and to the Discussion.
 
@@
 Thanks for the careful reading and useful comments.
+We hope you like our revisions.
 
@@
 \reply{
-    We don't actually agree with this characterization --
-    sections 3 and 4 explore the consequences of different distributions
-    of the Mendelian sampling term (which derives from the sum of effects),
-    not the population distribution.
-    However, the assumptions and motivation were not nearly as clear as they could have been
+    We agree with most of this statement --
+    the best answer to the admittedly vague question we pose here
+    does have to do with whether the Mendelian noise term can be taken to be independent of parental values,
+    and we do explore some other -- we think related and interesting -- questions along the way.
+    However, we think there is some confusion here about both Sections 3 and \ref{sec:abnormal},
+    which we discuss more below: in particular, the focus of Section 3
+    is not on the population distribution,
+    but instead on the very question brought up here
+    (whether the Mendelian term can be independent of the parental values).
+    However, the assumptions and motivations were not as clear as they could have been
     in the previous draft, and this is hopefully more clear now --
@@
     We have modified the sentence to say
-    ``implies a Gaussian for the offsprings' deviations'',
+    ``implies a Gaussian distribution for the offsprings' deviations'',
     to make it clear that this statement does not refer to the population distribution,
@@
     This implies that the joint distribution of traits of any subset of individuals
-    is multivariate Gaussian (possibly conditioned on the initial generation)
+    is multivariate Gaussian (since the initial generation is Gaussian)
     -- as \citet{barton2017infinitesimal} says,
@@
     on which we conditioned, but segregation variances would be unchanged.''}
-    (More generally, if $X$ is multivariate Gaussian, and $(Y|X)$ is also multivariate Gaussian
+    (More generally, if $X$ is multivariate Gaussian, and $Y$ conditioned on $X$ is also multivariate Gaussian
     with mean but not covariance depending on $X$,
@@
     about what the Proposition means and doesn't mean:
-    see in particular \llname{pop_not_gaussian} and \revref.
+    see in particular \llname{pop_not_gaussian}.
 }
@@
 \reply{
-    Thanks -- we've changed this to ``since models can be easily set up for which the distribution
+    Thanks for the catch -- this was an unfortunate mistake that might have led to some of the previous confusion.
+    We've changed this to ``since models can be easily set up for which the distribution
     of Mendelian sampling terms is not Gaussian, \ldots''
@@
 \reply{
-    We've rephrased this, and replaced ``gene'' with ``region'' --
+    We've rephrased this, and replaced ``gene'' with ``region'' \revref --
     really, we just want to motivate a mixture model,
     as the "scale mixture of Gaussians" is a common way to get fat-tailed distributions,
-    and eminently biologically plausible (we think).
+    and at least roughly biologically plausible.
+    On the other hand, we don't want to get into the weeds with more detail or discussion here.
 }
@@
 \reply{
+    They are a member of the (first) ancestral generation, and therefore a (potential) parent.
+    We've clarified this \revref.
 }
@@
 \begin{quote}
-The authors investigate the possibility of extending results of \citet{barton2017infinitesimal} on et al (2017) on the
+The authors investigate the possibility of extending results of \citet{barton2017infinitesimal} on the
 justification of the infinitesimal model to distributions of effect sizes with a large tail (i.e.,
@@
 cannot be a simple additive model.
-Investigating how exactly a concrete model of mutation would affect things is beyond the scope of this paper.
+Investigating how exactly a concrete model of mutation would affect things is beyond the scope of this paper,
+and we hope the interesting bits of this section
+outweigh the lack of an explicit mutationi model.
+We have added a note that hopefully clarifies the questions
+about population size \llname{deterministic_point}.
 
@@
 \reply{
-    Good idea -- done \revref. (Also see \revreffull{2}{1}.)
+    Good idea -- done \revref. (Also see \revreffull{2}{1} and \revreffull{1}{11}.)
 }
@@
 \begin{point}{\revref}
-p 8, last line: Delete ‘to’
+Delete ‘to’
 \end{point}

### 2023-12-03 writeup.tex
edit through

@@
     is Normally distributed and independent of the parents' values.
-    Here, we investigate how the assumption of similar effect sizes
-    affects the model:
-    in particular, if the tail of the effect size distribution is polynomial
-    with exponent $\alpha < 2$, then sums of effects
-    should be well-approximated by a ``stable distribution'',
-    and find tail exponents between 1 and 2 in 
+    Here, we investigate how the assumption of similar effect sizes affects the model:
+    if, alternatively, the tail of the effect size distribution is polynomial
+    with exponent $\alpha < 2$, then a different Central Limit Theorem implies that
+    sums of effects should be well-approximated by a ``stable distribution'',
+    for which single large effects are often still important.
+    Empirically, we first find tail exponents between 1 and 2 in 
     effect sizes estimated by genome-wide association studies
     of many human disease-related traits.
-    We show that the independence of offspring trait deviations from parental averages
-    in many cases implies a Gaussian distribution,
+    We then show that the independence of offspring trait deviations from parental averages
+    in many cases implies the Gaussian aspect of the infinitesimal model,
     suggesting that non-Gaussian models of trait evolution must explicitly
@@
 (and the approximation becomes exact, as $n \to \infty$).
-Oth the other hand, if $\alpha \ge 2$ then the distribution has finite variance,
+Oth the other hand, if $\alpha \ge 2$ then the distribution has finite variance, \revpoint{2}{1} 
 and so the sum of $n$, divided by $\sqrt{n}$, converges to a Gaussian distribution.
@@
 so that e.g., for $\alpha=1$ a positive proportion of the sum of $n$ entries will be contributed by a single one.
-On the other hand, the largest of $n$ Gaussians is of order $\sqrt{\log n}$. \revpoint{2}{1} \revpoint{1}{11}
+On the other hand, the largest of $n$ Gaussians is of order $\sqrt{\log n}$. \revpoint{1}{11}
 
@@
 The immediate answer to this will be ``no'' --
-in fact, we will see that independence of offsprings' deviations and midparent values \emph{implies} a Gaussian distribution for the offsprings' deviations, in some sense. \revpoint{1}{2}
+in fact, we will see that independence of offsprings' deviations and midparent values \revpoint{1}{2}
+\emph{implies} a Gaussian distribution for the offsprings' deviations, in some sense.
 This independence is a key reason why the infinitesimal model is tractable,
@@
 and interactions between alleles, between loci, and with the environment are often important in practice.
-In fact, many traits that have been claimed to be mostly genetically determined \revpoint{1}{14}
+In fact, some important traits that have been claimed to be mostly genetically determined \revpoint{1}{14}
 only appear heritable through non-genetic means,
@@
 and the chance a given allele is carried by a random parent is given by the frequency.
-This will make a difference as frequencies and effect sizes are often correlated in practice.
+(This weighting will make a difference, as frequencies and effect sizes are often correlated in practice.)
 So, to obtain a per-trait estimate of $\alpha$,
@@
 but simulation-based estimates of accuracy (Supplemental Figure~\ref{fig:power})
-suggest that our estimates are unbiased but only accurate to within roughly $\pm 0.4$,
-so the distribution of Figure~\ref{fig:exponent_hist} should not be taken as
+show that our estimates are unbiased but only accurate to within roughly $\pm 0.4$.
+So, the distribution of Figure~\ref{fig:exponent_hist} should not be taken as
 a direct estimate of the underlying distribution of $\alpha$ values,
@@
 (\citet{kac1939characterization,bernstein1941property},
-and see \citet{kagan1973characterization} for many much more general results).
+and see \citep{kagan1973characterization} for many much more general results).
 Since $R_M$ is (half) the difference and $\bar Z_M$ is (half) the sum
@@
 must also, by logical necessity, have a Gaussian distribution for the two.
-In other words,
-since models can be easily set up for which the distribution of Mendelian sampling terms
-is not Gaussian,
+In other words, \revpoint{1}{7}
+since models can be easily set up for which offspring minus midparent traits
+are not Gaussian,
 the implication of the proposition
 is that for such models, independence of the Mendelian sampling term
-is not likely a good assumption. \revpoint{1}{7}
+is not likely a good assumption.
 Note that the proof does not depend on independence of the $X_\ell$:
@@
 
-Finally, a possible point of confusion is that \revpoint{1}{6}
-the infinitesimal model is often phrased as a statement about offspring trait distributions
-\emph{given} parental trait values,
-while our statement above regards parental trait values as random.
-However, the notions are equivalent:
-the statement that
-``for each pair of parents the deviation of the offspring's trait from midparent
-has a distribution that does not depend on the parents' values''
-is essentially the definition of statistical independence
-of offspring's Mendelian sampling term and parental trait values.
+% NOT HELPFUL
+% Finally, a possible point of confusion is that \revpoint{1}{6}
+% the infinitesimal model is often phrased as a statement about offspring trait distributions
+% \emph{given} parental trait values,
+% while our statement above regards parental trait values as random.
+% However, the notions are equivalent:
+% the statement that
+% ``for each pair of parents the deviation of the offspring's trait from midparent
+% has a distribution that does not depend on the parents' values''
+% is essentially the definition of statistical independence
+% of offspring's Mendelian sampling term and parental trait values.
 
@@
 a Gaussian distribution for the population only emerges under weak or Gaussian selection.
-However, the model has been used in influential work.
+% However, the model has been used in influential work.
 In principle, another factor that could produce a non-Gaussian population distribution
@@
 % Since the set of functions $\mathcal{S} = \{e^{i \theta x} : \theta \in \R\}$ is separating (see \eg \S 3.4 in \cite{Ethier+Kurtz1986}), it suffices to show that \eqref{eqn:MSB_condition} holds for $f \in \mathcal{S}$, \ie
-If the population is large enough,
+If the population is large enough, \llabel{deterministic_point}
 we can treat the trait distribution as changing deterministically.
@@
 when both have well-posed limits independent of the other. \llabel{limit_remark} % We don't need to limit to this caee.
-Biologically,
+Biologically, \llabel{clarification2}
 the ``noise term'' is the contribution to variation in the population
@@
 while the ``reproduction term'' is the contribution from variation in the initial generation.
-\llabel{clarification2}
 
@@
 
-To see what knowing the parental trait value $Z^{[j]}_{M}$ \revpoint{1}{16}
-tells us about the contribution of any given locus, $\eta^{[j]}_{\ell}$, we follow the arguments in  \citet{Fisher1918} and \citet{barton2017infinitesimal} using Bayes' Theorem.
+To see what knowing the trait value $Z^{[j]}_{M}$ \revpoint{1}{16}
+of a particular parent in the ancestral generation
+tells us about the contribution of any given locus, $\eta^{[j]}_{\ell}$, to their trait value,
+we follow the arguments in  \citet{Fisher1918} and \citet{barton2017infinitesimal} using Bayes' Theorem.
 Proceeding informally, with the probabilities indicating probability density functions, we have 
@@
 and to Gregor Gorjanc for insight into modern uses
-of the animal model and genomic selection. 
+of the animal model and genomic selection,
+as well as to two anonymous reviewers for helpful comments.
 This project was started on a visit by TLP to the University of Oregon that was partially supported by the CNRS PEPS grant ``Jeunes Chercheuses et Jeunes Chercheurs''.

### 2023-12-03 review-responses.tex
TO DOs done, typos corrected.
Still need to read through and have improvements to make to 4 & 5.

@@
 
-\comment{TODO}
+\comment{We deliberately avoided giving the analogous estimate from Barton et al.
+(2017), as the expression is complicated and far from transparent (its definition spans several paragraphs).
+As such, we instead chose to focus on its salient features, namely that scales as $\frac{1}{\sqrt{M}}$ and grows (approximately) proportionally to the magnitude of the trait.
+We have added a reference to the relevant appendix in Barton et al.
+(2017) for those sufficiently motivated, and have attempted to give some more intuition for the consequences of the difference in the error bounds.}
 
@@
 
-\begin{point}{}
+\begin{point}{\revref}
 p 12, last display equation: Please relate this explicitly to the corresponding result in
@@
 \reply{
-    \comment{TODO}
+There isn't an analogous result in Barton et al., as they have assumed that all allele effects are bounded random variables rescaled by a factor of $\frac{1}{\sqrt{M}}$.
+Thus, no matter how many alleles that one considers, none will have the large effects necessary to give information about that trait value.
+We've added some text to clarify this difference at \revref.
 }

### 2023-12-03 writeup.tex
TO DOs done, typos corrected.
Still need to read through and have improvements to make to 4 & 5.

@@
         +
-        \frac{\gamma_{\ell}(\bar{\gamma}^{2}-z^{2})+2zy}{\gamma_{\ell}(\bar{\gamma}^{2}+z^{2})}
+        \frac{\gamma_{\ell}(\bar{\gamma}^{2}-z^{2})+2\bar{\gamma} zy}{\bar{\gamma}(\bar{\gamma}^{2}+z^{2})}
         \frac{1}{M}
@@
         \left|
-        \frac{\gamma_{\ell}(\bar{\gamma}^{2}-z^{2})+2zy}{\gamma_{\ell}(\bar{\gamma}^{2}+z^{2})}
+        \frac{\gamma_{\ell}(\bar{\gamma}^{2}-z^{2})+2\bar{\gamma}zy}{\bar{\gamma}(\bar{\gamma}^{2}+z^{2})}
         \right|
-        \le \frac{|y|}{\gamma_\ell \bar{\gamma}} ,
+        \le \frac{\gamma_{\ell}}{\bar{\gamma}} + \frac{|y|}{\bar{\gamma}},
 \end{align*}
@@
     \right|
-    \le \left(\left(1 + \frac{|y|}{\gamma_\ell \bar{\gamma}}\right) \frac{1}{M} + \Oh\left(\frac{1}{M^2}\right)\right)  \P\{\eta^{[j]}_{\ell} = y\}.
+    \le \left(\left(\frac{\gamma_{\ell}}{\bar{\gamma}} + \frac{|y|}{\bar{\gamma}}\right) \frac{1}{M} + \Oh\left(\frac{1}{M^2}\right)\right)  \P\{\eta^{[j]}_{\ell} = y\}.
 \]
@@
 \llabel{approx_eqn}
-In the case of bounded allelic effects, the error introduced by approximating the effect-size distribution conditional on the trait value by the unconditioned distribution is proportional to $|Z^{[j]}|$, so that extreme trait values are most informative of the size of allelic effects,
-whereas here,
-the bound is independent of the trait value,
-and the error is maximized when the trait value is approximately equal to
-the mean absolute deviation of its distribution, \ie the median value of $|Z^{[j]}|$.
-
-The argument above shows for a typical locus, the effect of conditioning on the trait gives negligible information about the contribution from that site.   Recalling that
-\[
-    \P\left\{|\eta^{[j]}_{\ell}| > y\right\} =  1-\frac{2}{\pi} \arctan\left(\frac{y}{\gamma_{\ell}}\right) \sim \frac{2\gamma_{\ell}}{\pi y},
+
+When allelic effects have magnitude bounded by  $\frac{B}{\sqrt{M}}$, (see Appendix C in \citet{barton2017infinitesimal}), the error introduced by approximating the effect-size distribution conditional on the trait value by the unconditioned distribution takes the form
+\[
+	\left|\P\{\eta^{[j]}_{\ell} = y \vert Z^{[j]} = z\} - \P\{\eta^{[j]}_{\ell} = y\} \right| \leq \frac{1}{\sqrt{M}} C(\hat{\sigma}_{M}^{2},z) \P\{\eta^{[j]}_{\ell} = y\},
+\]
+where $\hat{\sigma}_{M}$ is the variance in the trait value $Z^{[j]}$ and $C(\hat{\sigma}_{M}^{2},z)$ is a somewhat complicated expression that is approximately proportional to $z$, but independent of the allelic effect size $y$ (Equation (12), \textit{loc.
+cit.}).
+Thus, the trait value $Z^{[j]}$ is most informative of the size of the allelic effects at any given locus when it differs significantly from its mean: intuitively, when allele effects are uniformly bounded, a large trait value is evidence for many atypically large allelic effects.
+Interestingly, for the Cauchy distribution, we have the opposite: the error bound is independent of the trait value, $Z^{[j]} = z$, but is proportional to the allele effect size, $\eta^{[j]}_{\ell} = y$, at the locus under consideration (indeed, the error is maximized when the trait value is approximately equal to the mean absolute deviation of its distribution, \ie the median value of $|Z^{[j]}|$, rather than at extreme trait values).
+This suggests that with Cauchy-distributed allelic effects, knowing the trait value gives evidence for large allelic effects at individual loci.
+
+To elaborate, the argument above shows for a typical locus, the effect of conditioning on the trait gives negligible information about the contribution from that site.   Recalling that
+\[
+    \P\left\{|\eta^{[j]}_{\ell}| > y\right\} =  1-\frac{2}{\pi} \arctan\left(\frac{y}{\gamma_{\ell}}\right) \sim \frac{2\gamma_{\ell}}{\pi y} \quad \text{for large values of $y$},
 \]
@@
 \[
-    \P\left\{\frac{\eta^{[j]}_{\ell}}{\gamma_\ell \bar{\gamma} M} > \epsilon\right\} \sim \frac{2\gamma_{\ell}}{\pi \gamma_\ell \bar{\gamma} M \epsilon}.
+    \P\left\{\frac{|\eta^{[j]}_{\ell}|}{\bar{\gamma} M} > \epsilon\right\} \sim \frac{2\gamma_{\ell}}{\pi \bar{\gamma} M \epsilon}.
 \]
@@
 	\1^{M}_{\ell} = \begin{cases}
-        1 & \text{if $\frac{\eta^{[j]}_{\ell}}{\bar{\gamma} M} > \epsilon$, and}\\
+        1 & \text{if $\frac{|\eta^{[j]}_{\ell}|}{\bar{\gamma} M} > \epsilon$, and}\\
 		0 & \text{otherwise.}
@@
 \[
-	\lim_{M \to \infty} \sum_{\ell = 1}^{M} \frac{2\gamma_{\ell}}{\pi \bar{\gamma} M \epsilon} = \frac{2}{\pi \epsilon},
-\]
-the number of loci at which the magnitude of the effect size exceeds $M \epsilon$, which is 
+	\lim_{M \to \infty} \sum_{\ell = 1}^{M} \frac{2 \gamma_{\ell}}{\pi \bar{\gamma} M \epsilon} = \frac{2}{\pi \epsilon},
+\]
+the number of loci at which the magnitude of the scaled effect size exceeds $\bar{\gamma} \epsilon$, which is 
 $\sum_{\ell = 1}^{M} \1^{M}_{\ell}$,
@@
 \begin{align*}
-    \P\left\{\max_{\ell=1}^{M} \frac{\eta^{[j]}_{\ell}}{\bar{\gamma} M} > \epsilon\right\}  \sim 1-e^{-\frac{2}{\pi \bar{\gamma} \epsilon}},
+    \P\left\{\max_{\ell=1}^{M} \frac{|\eta^{[j]}_{\ell}|}{\bar{\gamma} M} > \epsilon\right\}  \sim 1-e^{-\frac{2}{\pi \epsilon}},
 \end{align*}
 which approaches 1 as $\epsilon \to 0$, while the expected number of $M$ alleles such that $\frac{\eta^{[j]}_{\ell}}{\bar{\gamma} M} > \epsilon$ is asymptotically $\frac{2}{\pi\epsilon}$. 
+By contrast, in \citet{Fisher1918} and \citet{barton2017infinitesimal} \revpoint{2}{10}, the scaled allelic effect at any locus $\ell$ has magnitude bounded by $\frac{B}{\sqrt{M}}$, and thus, no matter how many alleles one considers, there will be no large effect alleles that are informative of the trait value.
 

### 2023-12-04 writeup.tex
Working on Section 4.

@@
 (and the approximation becomes exact, as $n \to \infty$).
-Oth the other hand, if $\alpha \ge 2$ then the distribution has finite variance, \revpoint{2}{1} 
+On the other hand, if $\alpha \ge 2$ then the distribution has finite variance, \revpoint{2}{1} 
 and so the sum of $n$, divided by $\sqrt{n}$, converges to a Gaussian distribution.
@@
 Since $|X-1/2| = 1/2$, we'd like to know
-if $\P\{|X|>2t\}$ is proportional to $t^{-\alpha}$,
+if $\P\{|\eta|>2t\}$ is proportional to $t^{-\alpha}$,
 i.e., the value of
 $$
-    \lim_{t \to \infty} \frac{ - \log \P\{ |X| > 2 t \} }{ \log t } = \alpha ,
+    \lim_{t \to \infty} \frac{ - \log \P\{ |\eta| > 2 t \} }{ \log t } = \alpha ,
 $$
@@
 $$
-    \P\{ |\eta| > 2t \} \approx \frac{1}{M} \sum_{\ell : |e_\ell| > 2t} p_\ell .
+    \P\{ |\eta_{\ell}| > 2t \} \approx \frac{1}{M} \sum_{\ell : |e_\ell| > 2t} p_\ell .
 $$
@@
 are the characteristic functions of $Z^{j}$ and $R$, respectively. 
-In what follows, we will explore some non-Gaussian solutions $\phi(\theta)$ to equation \eqref{eqn:MSB_CF}:
+In what follows, we will explore some non-Gaussian solutions $\phi(\theta)$ to equation \eqref{eqn:MSB_CF}.
 
@@
 \]
-whereas any fixed point must be the limit of an arbitrary characteristic function, $\varphi(\theta)$, under repeated applications of $F$,
+whereas any fixed point must be the limit of an arbitrary function, $\varphi(\theta)$, under repeated applications of $F$,
 \ie \revpoint{2}{7}
-\begin{align*}
+\begin{equation}\label{eq:fpeq}
+\begin{aligned}
 	\phi(\theta) 
@@
     &= \lim_{n \to \infty} \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}\prod_{k=0}^{n-1}\Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}} .
-\end{align*}
+\end{aligned}
+\end{equation}
 % Iterating $F$ $n$ times, we get
@@
 % \]
-is a fixed point whenever this limit exists.
-
-Since $\varphi$ and $\Phi$ are characteristic functions, they are continuous and $\phi(0) = \Phi(0) = 1$, and thus both are non-zero in a neighbourhood of 0.
+is a fixed point whenever this limit exists, whereas any fixed point $\phi$ of $F$ satisfies  $\phi(\theta) = \lim_{n \to \infty} F^{(n)}(\phi)(\theta)$.
+As we are interested in identifying those fixed points that are characteristic functions, we henceforth assume, without loss of generality, that the $\varphi$ is itself a characteristic function (if $\phi$ exists then we can always take $\varphi = \phi$) and that the limit $\lim_{n \to \infty} F^{(n)}(\varphi)(\theta)$ exists.
+
+Since $\varphi$ and $\Phi$ are characteristic functions, say for real-valued random variables $Z$ and $R$, respectively, then $\varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}$, $\Phi_{n}(\theta) = \prod_{k=0}^{n-1}\Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}$, and $F^{(n)}(\varphi)(\theta)$ are all characteristic functions, of
+\[
+	 2^{-n} \sum_{i = 1}^{2^{n}} Z_{i}, \quad \sum_{k = 0}^{n-1} 2^{-k} \sum_{j = 1}^{2^{k}} R_{j}, 
+	 \quad \text{and} \quad 
+	 2^{-n} \sum_{i = 1}^{2^{n}} Z_{i} + \sum_{k = 0}^{n-1} 2^{-k} \sum_{j = 1}^{2^{k}} R_{k,j},
+\]
+respectively, where the $Z_{i}$ and $R_{k,j}$ are i.i.d.
+copies of $Z$ and $R$.
+L\'evy's continuity theorem (see \eg Theorem 2.3.4 in \cite{Durrett2005}) tells us that if $\varphi_{n}(\theta)$ has a point-wise limit, say $\varphi_{\infty}(\theta)$, then $\varphi_{\infty}(\theta)$ is the characteristic function of some random variable if and only if $\varphi_{\infty}(\theta)$ is continuous at 0 (in which case it is continuous for all $\theta \in \mathbb{R}$), and similarly for $\Phi_{n}(\theta)$ and $F^{(n)}(\varphi)(\theta)$.
+Note that, since we are assuming that $F^{(n)}(\varphi)(\theta)$ has a point-wise limit, then either both $\varphi_{n}$ and $\Phi_{n}$ have point-wise limits, or neither does.
+
+Now, as characteristic functions, $\varphi$ and $\Phi$ are continuous and $\varphi(0) = \Phi(0) = 1$, and thus both are non-zero in some neighbourhood of 0.
 In particular, $\psi(\theta) = \Log{\varphi}(\theta)$ and $\Psi(\theta) = \Log{\Phi(\theta)}$ exist in some neighbourhood of 0 as well\footnote{We use $\Log$ to denote the principal value of the complex logarithm, for which the imaginary part always lies in the interval $(-\pi,\pi]$.}.
@@
 We briefly explore what possible limiting values are valid characteristic functions. 
-We refer to the first term on the left hand side as the \emph{reproduction term} and the remaining sum as the \emph{noise term}, and only consider the case \revpoint{2}{8}
+We refer to the first term on the left hand side, $\psi_{n}(\theta) = 2^{n} \psi\left(\frac{\theta}{2^{n}}\right)$,  as the \emph{reproduction term} and the remaining sum, $\Psi_{n}(\theta) = \sum_{k=0}^{n-1} 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)$, as the \emph{noise term}, and only consider the case \revpoint{2}{8}
 when both have well-posed limits independent of the other. \llabel{limit_remark} % We don't need to limit to this caee.
@@
 First, consider the noise terms.
-A necessary condition for $\sum_{k=0}^{\infty} 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)$ to converge is that $2^{k} \Psi\left(\frac{\theta}{2^{k}}\right) \to 0$; since $\Psi(0) = 0$, this implies that if it exists, $\Psi'(0+) = 0$.
+A necessary condition for $\Psi_{n}(\theta)$ to converge is that $2^{k} \Psi\left(\frac{\theta}{2^{k}}\right) \to 0$; since $\Psi(0) = 0$, this implies that if it exists, $\Psi'(0+) = 0$.
 In particular, as we would expect, if $R$ has a mean, it must be 0.
@@
 \paragraph{Reproductive Terms:}
-We now turn our attention to the limit
-\[
-		\alpha(\theta) = \lim_{n \to \infty} 2^{n}\psi\left(\frac{\theta}{2^{n}}\right)
-\]
-If it exists, then for all integers $k$, 
-\[
-	\alpha(2^{k} \theta) = \lim_{n \to \infty} 2^{n}\psi\left(\frac{2^{k} \theta}{2^{n}}\right) =  \lim_{n \to \infty} 2^{k} 2^{n-k} \psi\left(\frac{\theta}{2^{n-k}}\right) = 2^{k} \alpha(\theta).
-\]
-and, since $\alpha(0) = 0$,
-\[
-	\frac{\alpha(\theta)}{\theta} = \frac{\alpha(2^{k} \theta)}{2^{k}\theta} 
-	\to \begin{cases} 
-		\alpha'(0+) & \text{if $\theta > 0$, and}\\
-		\alpha'(0-) & \text{if $\theta < 0$,}
-	\end{cases}
-\]
-as $k \to - \infty$ if the limits exist (\ie there are unique left-and right-hand limits at zero; this is the case if \eg we require $\alpha(\theta)$, and thus $\phi(\theta)$, to be convex functions).
-We henceforth assume this.
-
-Since $\varphi(\theta)$ is a characteristic function, it has the Hermitian property: $\phi(-\theta) = \overline{\phi(\theta)}$, and thus $\psi(-\theta) = \overline{\psi(\theta)}$ and $\alpha(-\theta) = \overline{\alpha(\theta)}$.
-In particular, if one (and by extension both) limits exist, we must have $\alpha'(0-) = \overline{\alpha'(0+)}$.
-In this case, setting $-\gamma + i m = \alpha'(0+)$, we have
-\[
-	\alpha(\theta) = \begin{cases} 
+We now turn our attention to the (point-wise) limit
+\[
+		\psi_{\infty}(\theta) = \lim_{n \to \infty} \psi_{n}(\theta).
+\]
+Since $\varphi(\theta)$ is a characteristic function, it has the Hermitian property: $\varphi(-\theta) = \overline{\varphi(\theta)}$, and thus $\psi_{n}(-\theta) = \overline{\psi_{n}(\theta)}$ for all $n$ and $\psi_{\infty}(-\theta) = \overline{\psi_{\infty}(\theta)}$ for all $\theta$ at which the limit exists.
+We can thus, without loss, assume that $\theta > 0$.
+
+Suppose $\psi_{\infty}(\theta)$ exists for some $\theta > 0$.  Then for all integers $k$, 
+\[
+	\psi_{\infty}(2^{k} \theta) = \lim_{n \to \infty} 2^{n}\psi\left(\frac{2^{k} \theta}{2^{n}}\right) =  \lim_{n \to \infty} 2^{k} 2^{n-k} \psi\left(\frac{\theta}{2^{n-k}}\right) 
+	= 2^{k} \psi_{\infty}(\theta).
+\]
+In particular, since $\psi_{\infty}(0) = 0$, 
+\[
+	\frac{\psi_{\infty}(\theta)}{\theta} = \frac{\psi_{\infty}(2^{k} \theta)}{2^{k}\theta} = \frac{\psi_{\infty}(2^{k} \theta)-\psi_{\infty}(0)}{2^{k}\theta - 0}, 
+\]
+so if $\psi_{\infty}$ has a right-hand side derivative at zero, then
+\[
+	\frac{\psi_{\infty}(\theta)}{\theta} = \lim_{k \to -\infty} \frac{\psi_{\infty}(2^{k} \theta)}{2^{k}\theta} = \psi_{\infty}'(0+).
+\]
+
+Suppose that this derivative exists, say $\psi_{\infty}'(0+) = -\gamma + i m$ for some $m, \gamma \in \mathbb{R}$.
+Then, by the Hermitian property,  $\psi_{\infty}'(0-) = \overline{\psi_{\infty}'(0+)}$ must exist as well, so
+\[
+	\psi_{\infty}(\theta) = \begin{cases} 
 		i m\theta - \gamma \theta & \text{if $\theta \geq 0$, and}\\
@@
 \]
-If $\gamma > 0$, then we recognize $e^{\psi(\theta)} = e^{im \theta -\gamma|\theta|}$
-as the characteristic function of a Cauchy distribution with median 0 and median absolute deviation $\gamma$
-(\ie, the canonical $\alpha$-stable distribution with $\alpha=1$), \llabel{cauchy_note}
-whereas if $\gamma = 0$, then $e^{\psi(\theta)} = e^{im \theta}$ is the Fourier transform of a Dirac mass at $m$, \ie the characteristic function for a monomorphic population with all individuals having trait value equal to $m$.
-
-% This might seem a bit roundabout, but the point is that we can prove that $e^{\psi(\theta)}$ is -- in this case -- a characteristic function, so we don't need to lose generality by assuming that it is one.
-Our goal is to ask when $\lim_{n \to \infty} \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}\prod_{k=0}^{n-1}\Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}$ is a valid characteristic function, and to see what it can be.
-To exclude the case $\gamma < 0$, we first observe that if, as we assume, $\varphi(\theta)$ is the characteristic function of some random variable, say $X$, then $\varphi_{n}(\theta) = \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}} = e^{2^{n}\psi\left(\frac{\theta}{2^{n}}\right)}$ is the characteristic function of $2^{-n} \sum_{i = 1}^{2^{n}} X_{i}$, where $X_{1},\ldots,X_{2^{n}}$ are i.i.d.\ copies of $X$; now, L\'evy's continuity theorem (see \eg Theorem 2.3.4 in \cite{Durrett2005}) tells us that if (i) $\phi_{n}(\theta)$ has a point-wise limit, say $\phi_{\infty}(\theta)$, and (ii) $\phi_{\infty}(\theta)$ is continuous at 0, then $\phi_{\infty}(\theta)$ is the characteristic function of \emph{some} random variable; in particular, if $\alpha(\theta)$ exists and is bounded (\eg if the left and right-hand side derivatives exist, as above), then $\psi(\theta) = \alpha(\theta)\theta$ is continuous at 0, and hence so is $e^{\psi(\theta)}$.
-Thus, in this case $e^{\psi(\theta)}$ is itself a characteristic function.
-Moreover, being a characteristic function, $e^{\psi(\theta)}$ must satisfy $|e^{\psi(\theta)}| = e^{\Re \psi(\theta)} \leq 1$, whence $\gamma \geq 0$.
+Moreover,  $\psi_{\infty}(\theta) = i m \theta - \gamma |\theta|$ is continuous at 0, and hence so is $\varphi_{\infty}(\theta) = e^{\psi_{\infty}(\theta)}$.
+As we observed above, this implies that $e^{\psi_{\infty}(\theta)}$ is itself a characteristic function.
+Thus, $e^{\psi_{\infty}(\theta)}$ must satisfy $|e^{\psi_{\infty}(\theta)}| = e^{\Re \psi_{\infty}(\theta)} \leq 1$, whence $\Re \psi_{\infty}(\theta) \leq 0$ and $\gamma \geq 0$.
+In particular, if $\gamma > 0$, then we recognize $e^{\psi_{\infty}(\theta)} = e^{im \theta -\gamma|\theta|}$ as the characteristic function of a Cauchy distribution with median m and median absolute deviation $\gamma$ (\ie, the canonical $\alpha$-stable distribution with $\alpha=1$), \llabel{cauchy_note} whereas if $\gamma = 0$, then $e^{\psi_{\infty}(\theta)} = e^{im \theta}$ is the Fourier transform of a Dirac mass at $m$, \ie the characteristic function for a monomorphic population with all individuals having trait value equal to $m$.
 
@@
 which we recognize as being equal to $\phi\left(\frac{\theta}{2}\right)^{2} \Phi(\theta)$, where $\phi(\theta) = e^{i\gamma|\theta|}$ and $\Phi(\theta) = e^{i(\delta - \gamma)|\theta|}$ are the characteristic functions of the population density and the Mendelian sampling term, respectively,
-giving a.non-Gaussian solution to \eqref{eqn:MSB_CF} with both noise and selection.
+giving a non-Gaussian solution to \eqref{eqn:MSB_CF} with both noise and selection.
 

### 2023-12-04 writeup.tex
Fixing typos.

@@
 \]
-Since $\varphi(\theta)$ is a characteristic function, it has the Hermitian property: $\varphi(-\theta) = \overline{\varphi(\theta)}$, and thus $\psi_{n}(-\theta) = \overline{\psi_{n}(\theta)}$ for all $n$ and $\psi_{\infty}(-\theta) = \overline{\psi_{\infty}(\theta)}$ for all $\theta$ at which the limit exists.
-We can thus, without loss, assume that $\theta > 0$.
-
-Suppose $\psi_{\infty}(\theta)$ exists for some $\theta > 0$.  Then for all integers $k$, 
+
+Suppose $\psi_{\infty}(\theta)$ exists for some $\theta \in \mathbb{R}$.  Then for all integers $k$, 
 \[
@@
 Suppose that this derivative exists, say $\psi_{\infty}'(0+) = -\gamma + i m$ for some $m, \gamma \in \mathbb{R}$.
-Then, by the Hermitian property,  $\psi_{\infty}'(0-) = \overline{\psi_{\infty}'(0+)}$ must exist as well, so
+Since $\varphi(\theta)$ is a characteristic function, it has the Hermitian property: $\varphi(-\theta) = \overline{\varphi(\theta)}$, and thus $\psi_{n}(-\theta) = \overline{\psi_{n}(\theta)}$ for all $n$ and $\psi_{\infty}(-\theta) = \overline{\psi_{\infty}(\theta)}$ for all $\theta$ at which the limit exists.
+In particular, if the right-hand side derivative, then $\psi_{\infty}'(0-) = -\overline{\psi_{\infty}'(0+)}$ must exist as well, so
 \[
@@
 \]
+(\nb we could have equally well assumed the left-hand side derivative existed and derived the right-hand side derivative).
 Moreover,  $\psi_{\infty}(\theta) = i m \theta - \gamma |\theta|$ is continuous at 0, and hence so is $\varphi_{\infty}(\theta) = e^{\psi_{\infty}(\theta)}$.

### 2023-12-04 writeup.tex
Update writeup.tex

@@
 \]
-
 Suppose $\psi_{\infty}(\theta)$ exists for some $\theta \in \mathbb{R}$.  Then for all integers $k$, 
@@
 \]
-so if $\psi_{\infty}$ has a right-hand side derivative at zero, then
+so if $\psi_{\infty}$ has a right-hand side derivative at zero (\eg if $\psi_{\infty}$ is convex), then
 \[
@@
 \]
-(\nb we could have equally well assumed the left-hand side derivative existed and derived the right-hand side derivative).
+(\nb we could have equally well assumed the left-hand side derivative existed and derived the right-hand side derivative from the Hermitian property).
+
 Moreover,  $\psi_{\infty}(\theta) = i m \theta - \gamma |\theta|$ is continuous at 0, and hence so is $\varphi_{\infty}(\theta) = e^{\psi_{\infty}(\theta)}$.

### 2023-12-05 review-responses.tex
final? edits

@@
 ``Large effects and the infinitesimal model''.
-In response to the reviewer's comments
+In response to the reviewers' comments
 we have made quite a few additions to the text
@@
     this calculation about iid effects is indeed the key step
-    (in \citet{Fisher1918} and \cite{barton2017infinitesimal})
+    (in \citet{fisher1918correlation} and \cite{barton2017infinitesimal})
     in describing the distribution of trait values in offspring.
@@
 
-\comment{We deliberately avoided giving the analogous estimate from Barton et al.
+We deliberately avoided giving the analogous estimate from Barton et al.
 (2017), as the expression is complicated and far from transparent (its definition spans several paragraphs).
@@
 We have added a reference to the relevant appendix in Barton et al.
-(2017) for those sufficiently motivated, and have attempted to give some more intuition for the consequences of the difference in the error bounds.}
+(2017) for those sufficiently motivated, and have attempted to give some more intuition for the consequences of the difference in the error bounds.
 
@@
 \reply{
-    That's right; changed to ``$\alpha$-stable distribution'' \revref.
+    That's right; we've edited this bit, which no longer says ``Lévy''. \revref
 }

### 2023-12-05 writeup.tex
final? edits

@@
 under which this would be a good approximation to reality,
-which they extended to include dominance in \citet{barton2022infinitesimal}.
+a proof they later extended to include dominance in \citet{barton2022infinitesimal}.
 
 As suggested by its name,
-under the infinitesimal model the effects of any given genetic variant on the trait
+under the infinitesimal model the effect of any given genetic variant on the trait
 is small (indeed, vanishing),
 and selection-driven trait shifts are mediated by small changes in frequencies of many alleles of small effect.
-However, empirical understanding of the genetic basis of trait variation in real-world species
+However, empirical studies of the genetic basis of trait variation in real-world species
 often finds a significant portion of heritable variation explained by variants at only a few loci.
@@
 and the distribution of effect sizes within each region is Normal, \revpoint{1}{13}
-but with a standard deviation that varies with ``proximity'' of that region to the trait
+but with a standard deviation specific to the region that varies with ``proximity'' of that region to the trait
 (e.g., regions that are further ``upstream'' have higher standard deviations).
@@
 On the other hand, if $\alpha \ge 2$ then the distribution has finite variance, \revpoint{2}{1} 
-and so the sum of $n$, divided by $\sqrt{n}$, converges to a Gaussian distribution.
+and so the sum of $n$ copies, divided by $\sqrt{n}$, converges to a Gaussian distribution.
 In such ensembles, single large entries can be important:
@@
 so that e.g., for $\alpha=1$ a positive proportion of the sum of $n$ entries will be contributed by a single one.
-On the other hand, the largest of $n$ Gaussians is of order $\sqrt{\log n}$. \revpoint{1}{11}
+(On the other hand, the largest of $n$ Gaussians is of order $\sqrt{\log n}$.) \revpoint{1}{11}
 
@@
 In fact, some important traits that have been claimed to be mostly genetically determined \revpoint{1}{14}
-only appear heritable through non-genetic means,
-e.g., systematic inequalities in resource distribution partitioned by class or race
+only appear heritable through non-genetic means
+such as systematic inequalities in resource distribution partitioned by class or race
 \citep[e.g.,][]{davenport1916hereditary}.
@@
 Note in particular that
-the contribution of locus $\ell$ to $R_M$ is $\pm |\eta_\ell^{[1]} - \eta_\ell^{[2]}|$.
+the contribution of locus $\ell$ to $R_M$ is $\pm |\eta_\ell^{[1]} - \eta_\ell^{[2]}|/2$.
 
@@
 a direct estimate of the underlying distribution of $\alpha$ values,
-but is robust evidence for exponents strictly between 1 and 2 (and probably around 1.5).
+but it is robust evidence for exponents strictly between 1 and 2 (and probably around 1.5).
 
@@
 
+Now we return to the theory.
 We know from \citet{barton2017infinitesimal} \revpoint{1}{5} \revpoint{2}{4}
@@
 \end{align*}
-so that 
+so that the parental traits are
 $Z_M^{[1]} = U_M^{[1]} + V_M^{[1]}$ and
@@
 (\citet{kac1939characterization,bernstein1941property},
-and see \citep{kagan1973characterization} for many much more general results).
+and see \citet{kagan1973characterization} for many much more general results).
 Since $R_M$ is (half) the difference and $\bar Z_M$ is (half) the sum
@@
 
-Note that under the conditions of the proposition, \llabel{pop_not_gaussian}
+Under the conditions of the proposition, \llabel{pop_not_gaussian}
 the parental traits, $Z^{[1]}_M$ and $Z^{[2]}_M$,
@@
 
+Although an infinitesimal model with non-Gaussian noise is incompatible with the usual additivity assumptions,
+such models still exist (and might be useful);
+next, we describe some of their properties.
 As observed in \cite{turelli2017commentary},
@@
 \begin{align*}
-  \frac{Z^{1} +Z^{2}}{2} + R ,
+    \frac{Z^{[1]} +Z^{[2]}}{2} + R ,
 \end{align*}
-where $Z^{1}$ and $Z^{2}$ are independent draws from the empirical population distribution
+where $Z^{[1]}$ and $Z^{[2]}$ are independent draws from the empirical population distribution
 and $R$ is an independent draw from the Mendelian sampling distribution
-(independent of $Z^{1}$ and $Z^{2}$, with distribution to be specified). 
+(independent of $Z^{[1]}$ and $Z^{[2]}$, with distribution to be specified). 
 (Note that mutation might enter into $R$, but as the underlying genetic basis is not specified,
@@
 and the distribution of individuals added by birth are equal,
-\ie that $(Z^{1} +Z^{2})/2 + R$ is equal in distribution
+\ie that $(Z^{[1]} +Z^{[2]})/2 + R$ is equal in distribution
 to a $\mu$-weighted choice of $Z$.
@@
 % \begin{equation} \label{eqn:MSB_condition}
-%     \E\left[ f\left(\frac{Z^{1} +Z^{2}}{2} + R \right) \right] = \frac{\E\left[ f(Z) \mu(Z) \right]}{\E[\mu(Z)]},
+%     \E\left[ f\left(\frac{Z^{[1]} +Z^{[2]}}{2} + R \right) \right] = \frac{\E\left[ f(Z) \mu(Z) \right]}{\E[\mu(Z)]},
 % \end{equation}
-% where the expectation is with respect to the joint law of the population empirical distribution (the law of $Z$, $Z^{1}$ and $Z^{2}$), and $R$.
+% where the expectation is with respect to the joint law of the population empirical distribution (the law of $Z$, $Z^{[1]}$ and $Z^{[2]}$), and $R$.
 % 
@@
     =
-    \E\left[ e^{i\theta\left(\frac{Z^{1} +Z^{2}}{2} + R \right)}\right] .
+    \E\left[ e^{i\theta\left(\frac{Z^{[1]} + Z^{[2]}}{2} + R \right)}\right] .
 \end{align}
-Now, we assumed that $Z^{1}$, $Z^{2}$, and $R$ are independent, so we can write \eqref{eqn:MSB_CF} as 
+Now, we assumed that $Z^{[1]}$, $Z^{[2]}$, and $R$ are independent, so we can write \eqref{eqn:MSB_CF} as 
 \begin{align*}
-    \frac{\E\left[ e^{i\theta Z} \mu(Z) \right]}{\E[\mu(Z)]} &= \E\left[ e^{i\theta\left(\frac{Z^{1} +Z^{2}}{2} + R \right)}\right]\\
+    \frac{\E\left[ e^{i\theta Z} \mu(Z) \right]}{\E[\mu(Z)]} &= \E\left[ e^{i\theta\left(\frac{Z^{[1]} +Z^{[2]}}{2} + R \right)}\right]\\
 	% &= \E\left[ e^{i\frac{\theta}{2} Z^{1}}e^{i\frac{\theta}{2} Z^{2}}e^{\i\theta R}\right]\\
-	&= \E\left[e^{i\frac{\theta}{2} Z^{1}}\right]\E\left[e^{i\frac{\theta}{2} Z^{2}}\right]\E\left[e^{i\theta R}\right]\\
+    &= \E\left[e^{i\frac{\theta}{2} Z^{[1]}}\right]\E\left[e^{i\frac{\theta}{2} Z^{[2]}}\right]\E\left[e^{i\theta R}\right]\\
 	&= \phi\left(\frac{\theta}{2}\right)^{2} \Phi(\theta),
@@
 where
-$ \phi(\theta) = \mathbb{E}\left[e^{i\theta Z^{1}}\right] = \mathbb{E}\left[e^{i\theta Z^{2}}\right] $
+$ \phi(\theta) = \mathbb{E}\left[e^{i\theta Z^{[1]}}\right] = \mathbb{E}\left[e^{i\theta Z^{[2]}}\right] $
 and
 $ \Phi(\theta) = \mathbb{E}\left[e^{i\theta R}\right] $
-are the characteristic functions of $Z^{j}$ and $R$, respectively. 
+are the characteristic functions of $Z^{[j]}$ and $R$, respectively. 
 In what follows, we will explore some non-Gaussian solutions $\phi(\theta)$ to equation \eqref{eqn:MSB_CF}.
@@
  \[
-	F(\phi)(\theta) = {\textstyle \phi\left(\frac{\theta}{2}\right)^{2}} \Phi(\theta),
-\]
-whereas any fixed point must be the limit of an arbitrary function, $\varphi(\theta)$, under repeated applications of $F$,
+	F(\phi)(\theta) = {\textstyle \phi\left(\frac{\theta}{2}\right)^{2}} \Phi(\theta) .
+\]
+Any fixed point $\phi(\theta)$ must be the limit of an arbitrary function, $\varphi(\theta)$, under repeated applications of $F$,
 \ie \revpoint{2}{7}
@@
 % \]
-is a fixed point whenever this limit exists, whereas any fixed point $\phi$ of $F$ satisfies  $\phi(\theta) = \lim_{n \to \infty} F^{(n)}(\phi)(\theta)$.
-As we are interested in identifying those fixed points that are characteristic functions, we henceforth assume, without loss of generality, that the $\varphi$ is itself a characteristic function (if $\phi$ exists then we can always take $\varphi = \phi$) and that the limit $\lim_{n \to \infty} F^{(n)}(\varphi)(\theta)$ exists.
-
-Since $\varphi$ and $\Phi$ are characteristic functions, say for real-valued random variables $Z$ and $R$, respectively, then $\varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}$, $\Phi_{n}(\theta) = \prod_{k=0}^{n-1}\Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}$, and $F^{(n)}(\varphi)(\theta)$ are all characteristic functions, of
+is a fixed point whenever this limit exists,
+whereas any fixed point $\phi$ of $F$ satisfies  $\phi(\theta) = \lim_{n \to \infty} F^{(n)}(\varphi)(\theta)$
+for some function $\varphi(\theta)$.
+As we are interested in identifying those fixed points that are characteristic functions, we henceforth assume, without loss of generality, that $\varphi$ is itself a characteristic function (if $\phi$ exists then we can always take $\varphi = \phi$) and that the limit $\lim_{n \to \infty} F^{(n)}(\varphi)(\theta)$ exists.
+
+Since $\varphi$ and $\Phi$ are characteristic functions, say for real-valued random variables $Z$ and $R$, respectively, then $\varphi_n(\theta) = \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}$, $\Phi_{n}(\theta) = \prod_{k=0}^{n-1}\Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}$, and $F^{(n)}(\varphi)(\theta)$ are all characteristic functions, of
 \[
@@
 the ``noise term'' is the contribution to variation in the population
-generated by the Mendelian sampling terms (\ie, by recombination and mutation),
+generated by the Mendelian sampling terms (\ie by recombination and mutation),
 while the ``reproduction term'' is the contribution from variation in the initial generation.
@@
 \]
-for $\alpha \in (0,2]$, $\beta \in [-1,1]$, $c \geq 0$, and $m \in \mathbb{R}$ and
+for some $\alpha \in (0,2]$, $\beta \in [-1,1]$, $c \geq 0$, and $m \in \mathbb{R}$ and
 \[
@@
 which corresponds to a $\left(\alpha,\beta, \frac{c}{1-2^{1-\alpha}},0\right)$-stable law.
+In other words, if the Mendelian sampling noise is stable with some $\alpha > 1$,
+then the cumulative effect across generations of this noise has a stable distribution of the same form,
+but with a larger scale -- analogous to the Gaussian case.
 
@@
 In particular, if $\gamma > 0$, then we recognize $e^{\psi_{\infty}(\theta)} = e^{im \theta -\gamma|\theta|}$ as the characteristic function of a Cauchy distribution with median m and median absolute deviation $\gamma$ (\ie, the canonical $\alpha$-stable distribution with $\alpha=1$), \llabel{cauchy_note} whereas if $\gamma = 0$, then $e^{\psi_{\infty}(\theta)} = e^{im \theta}$ is the Fourier transform of a Dirac mass at $m$, \ie the characteristic function for a monomorphic population with all individuals having trait value equal to $m$.
+
+In summary, the limiting effect of the initial generation can take one of only forms:
+either a constant (i.e., the initial generation only determines the mean),
+or a Cauchy distribution.
 
@@
 $ \phi(\theta) = \exp(i m \theta - \frac{1}{2} \sigma^{2} \theta^{2}).  $
-Moreover, provided $a < \frac{1}{2\sigma^{2}}$, we can explicitly compute $\frac{\mathbb{E}\left[e^{i \theta Z}\mu(Z)\right]}{\mathbb{E}[\mu(Z)]}$: %  using the Gaussian probability density function to conclude that
+Moreover, provided $a < 1 / 2\sigma^{2}$, we can explicitly compute
+$\mathbb{E}\left[e^{i \theta Z}\mu(Z)\right] / \mathbb{E}[\mu(Z)]$:
+%  using the Gaussian probability density function to conclude that
 if $R$ is Gaussian with mean zero and variance $\eta^2$, then equation~\eqref{eqn:MSB_CF} becomes
@@
 
+%%%%%%%%%%%%%%%%%%%%
 \section{The difficulty with an abnormal quantitative genetics}
@@
 The crucial step in the derivation of the infinitesimal model in \cite{barton2017infinitesimal} is to show that, \llabel{crucial_step}
-following the arguments in \citet{Fisher1918}, knowing an individual's trait value gives vanishingly little information about the contribution of any given locus as $M \to \infty$, which in turn allows one to show that the within-family Mendelian sampling term is independent of the parental traits.
+following the arguments in \citet{fisher1918correlation}, knowing an individual's trait value gives vanishingly little information about the contribution of any given locus as $M \to \infty$, which in turn allows one to show that the within-family Mendelian sampling term is independent of the parental traits.
 Here, we will consider how conditioning an individual's trait value affects the law of a single allelic contribution in the Cauchy setting described above.
@@
 tells us about the contribution of any given locus, $\eta^{[j]}_{\ell}$, to their trait value,
-we follow the arguments in  \citet{Fisher1918} and \citet{barton2017infinitesimal} using Bayes' Theorem.
+we follow the arguments in  \citet{fisher1918correlation} and \citet{barton2017infinitesimal} using Bayes' Theorem.
 Proceeding informally, with the probabilities indicating probability density functions, we have 
@@
 which approaches 1 as $\epsilon \to 0$, while the expected number of $M$ alleles such that $\frac{\eta^{[j]}_{\ell}}{\bar{\gamma} M} > \epsilon$ is asymptotically $\frac{2}{\pi\epsilon}$. 
-By contrast, in \citet{Fisher1918} and \citet{barton2017infinitesimal} \revpoint{2}{10}, the scaled allelic effect at any locus $\ell$ has magnitude bounded by $\frac{B}{\sqrt{M}}$, and thus, no matter how many alleles one considers, there will be no large effect alleles that are informative of the trait value.
-
-To summarize, the key argument in \citet{Fisher1918} and \cite{barton2017infinitesimal}  \revpoint{1}{10}
+By contrast, in \citet{fisher1918correlation} and \citet{barton2017infinitesimal} \revpoint{2}{10}, the scaled allelic effect at any locus $\ell$ has magnitude bounded by $\frac{B}{\sqrt{M}}$, and thus, no matter how many alleles one considers, there will be no large effect alleles that are informative of the trait value.
+
+To summarize, the key argument in \citet{fisher1918correlation} and \citet{barton2017infinitesimal}  \revpoint{1}{10}
 requires that knowing the sum of parental effects does not give much information 
@@
 an idea we find support for in theory (Section~\ref{sec:difficulty}) and simulation (Section~\ref{sec:simulation}).
-Despite having spent a fair bit of time describing why
+Despite having described why
 the infinitesimal model cannot strictly apply 
@@
 For simplicity, we have focused on the $\alpha=1$ case (which includes the Cauchy distribution).
-However, examination of Figure~\ref{fig:exponent_hist} suggests that $\alpha=3/2$
+However, examination of Figure~\ref{fig:exponent_hist} suggests that $\alpha=3/2$ \revpoint{2}{12}
 would be a more natural choice for an effect size distribution.
@@
 
-Finally, we note that above we've most often used the Cauchy distribution
-not because it seems most realistic
-(indeed data suggests the $\alpha$-stable distribution with $\alpha=3/2$ might be better) \revpoint{2}{12}
-but for mathematical convenience.
-
-\subsection*{Finite variance}
-
+\paragraph{Finite variance}
 A finite sample from any distribution will have finite variance.
@@
 
-\subsection*{Selection and effect size distributions} \llabel{discussion_DFEs}
-
+\paragraph{Selection and effect size distributions} \llabel{discussion_DFEs}
 A substantial body of theoretical work has used various models and assumptions

### 2023-12-05 writeup.tex
minor fixen

@@
 
-In summary, the limiting effect of the initial generation can take one of only forms:
+In summary, if the left- or right-hand derivative of $\psi_\infty$ exists,
+then the limiting effect of the initial generation can take one of only two forms:
 either a constant (i.e., the initial generation only determines the mean),
 or a Cauchy distribution.
+However, it is unclear if there are cases where these derivatives do not exist
+and these effects converge to a meaningful limit.
 

### 2024-01-23 review-responses.tex
new reviewer responses

--- 2023-12-03 review-responses.tex
@@
 ``Large effects and the infinitesimal model''.
-In response to the reviewer's comments
-we have made quite a few additions to the text
-which we hope will make the paper more accessible to a wider audience.
-We hope you agree!
+We have addressed the remaining points of the reviwers,
+hopefully to everyone's satisfaction.
 
@@
 \begin{quote}
-The two reviewers find the results interesting and worth being eventually published in the special issue.
-However, they highlight a number of places in the text where the properties claimed or the precise questions addressed need serious clarification.
-As the readership of the special issue will be quite broad, it is important that the main messages are accessible to colleagues who do not necessarily know Fisher's infinitesimal model precisely.
-Therefore, although the mathematical results in themselves do not require deep modifications, I recommend that a major revision of the paper should be provided, which would take into account the different comments and suggestions of the reviewers.
+As you will read from the reviewers' comments below, both of them find that the quality and clarity of the exposition has improved a lot, and they only have a list of minor (nonetheless important) comments that should be addressed in a minor revision.
+Provided the points raised are suitably addressed, the paper will then be accepted for publication.
+And though this has nothing to do with the paper, happy New Year to the authors ;-).
 \end{quote}
 
-Thanks very much to the reviewers for their comments.
-We agree that our initial version of the manuscript didn't have as broad an envisioned audience
-as it should have.
-We've tried to improve this, by adding additional context and discussion --
-in particular, in Sections 3 and 4, and to the Discussion.
+Happy New Year!
 
@@
 
-\begin{quote}
-The paper explores inheritance of an additive trait influenced by a large number of loci with effect sizes drawn from a power law distribution, also fitting effect size distributions as estimated by GWAS (for several human disease traits) to estimate plausible power law exponents.
-The study is motivated by Barton et al's 2017 paper which shows how in the limit of a very large number of loci with effect sizes drawn from a distribution with finite variance, inheritance is described by the the infinitesimal model: in particular, that the distribution of trait values of the offspring of two parents is multivariate normal with a variance-covariance matrix that is independent of parental trait values.
-This paper explores whether some analogous statements can be made if the underlying effect size distribution does not have finite variance- whether, loosely speaking, some ``generalisation'' of the infinitesimal model might still apply.
-Below I summarise some of my concerns about the paper.
-\end{quote}
-
-Thanks for the careful reading and useful comments.
-We hope you like our revisions.
-
 \begin{point}{}
-    Paragraph starting ``This suggests exploring whether stable distributions..'' \revref:
-The questions as stated: e.g., whether stable distributions can ``stand in'' for a Gaussian distribution in the infinitesimal model or whether the results of Barton et al ``carry over'' when the sum of effect sizes follows a stable law are quite vague and allow for multiple interpretations.
-Ideally, in order to address whether Barton et al's results (do not) carry over, one would like to at least show that with a power law distribution of effect sizes, the ``Mendelian noise term'', i.e., the difference between the offspring trait value and the midparent value, is (not) independent of the midparent value in the limit of a very large number of loci.
-And indeed this is what the authors address
-somewhat non-rigorously (and for the case of a Cauchy distribution of effect sizes) in section 5 and using simulations : see e.g., figures 4C and 4f, figures 6C and 6F.
-Unsurprisingly, these simulations show that with e.g., Cauchy distribution of effect sizes, the non-independence between the segregation/noise term and the midparent trait value does not hold.
-However, a lot of the mathematical analysis in the paper (sections 3 and 4) explores when the distribution of trait values in a population is (not) Gaussian, which is a completely different question.
+     As far as I understand, the crucial assumption is that the effect sizes $\eta_{l}^{i}$
+     of the alleles carried by an individual $i$ are independent
+     (as in the first line of Proposition \ref{prop:parental_contribs}),
+     which basically amounts to assuming linkage equilibrium in the population.
+     In that case, of course, all trait values across any group of individuals are multi-variate Gaussian
+     (provided effect size distributions are not heavy-tailed),
+     without needing to condition on any ancestors.
+     Moreover, under these conditions,
+     even the trait value distribution in the population would be Gaussian
+     (provided effect size distributions is not heavy-tailed).
+     As far as I understand, \citet{barton2017infinitesimal} is more general
+     in that it does not require $\eta_{l}^{i}$ to be independent,
+     except in the founder $t=0$ generation
+     (and I am not sure how crucial even that assumption is);
+     in later generations, there can be an arbitrary amount of LD.
+     This is a crucial point, since in general, under the infinitesimal model,
+     non-Gaussian population distributions only arise because of LD, in which case,
+     the $\eta_{l}^{i}$ are not independent.
+     Thus, it is worth highlighting that assuming $\eta_{l}^{i}$ to be independent
+     basically implies linkage equilibrium, and that this is in general,
+     not a requirement for the infinitesimal model
+     or for the derivations of \citet{barton2017infinitesimal},
+     whereas it is a key requirement here.
+     I think this had been a big part of my earlier confusion.
 \end{point}
@@
 \reply{
-    We agree with most of this statement --
-    the best answer to the admittedly vague question we pose here
-    does have to do with whether the Mendelian noise term can be taken to be independent of parental values,
-    and we do explore some other -- we think related and interesting -- questions along the way.
-    However, we think there is some confusion here about both Sections 3 and \ref{sec:abnormal},
-    which we discuss more below: in particular, the focus of Section 3
-    is not on the population distribution,
-    but instead on the very question brought up here
-    (whether the Mendelian term can be independent of the parental values).
-    However, the assumptions and motivations were not as clear as they could have been
-    in the previous draft, and this is hopefully more clear now --
-    see responses to points below.
+    It is true that if the $\eta$ are not independent then $Z_M$ and $V_M$ are not independent,
+    and so the Kac-Bernstein theorem does not apply.
+    However -- returning to the goal of this section here -- it seems hard and maybe impossible
+    to use LD to obtain a inifinitesimal-like, non-Gaussian model.
+    But, we also don't see why it's impossible, and have added a note to this effect \revref.
+    (Also we'd like to point out that the independence only applies to the \emph{parental} generation;
+    there is in fact LD in the offspring's generation, which is the generation Proposition \ref{prop:parental_contribs}
+    refers to. Our results could be extended to further generations, as we remark just before \revref,
+    but since our point is about non-existence of other models, this seemed superfluous.
+    Note also that although we assume independence, we don't assume the $\eta$ are identically distributed,
+    so that a Gaussian or other stable distribution is not guaranteed.)
 }
 
-\begin{point}{\revref}
-In the same paragraph: I don't think the claim that ``independence of offsprings' deviations and midparent values implies a Gaussian distribution'' (which is also stated in the abstract) is true?
-Surely, ``independence of offsprings' deviations and midparent values'' is just the basic statement of the infinitesimal model (say, in Barton et al 2017) and the infinitesimal model does not imply a Gaussian distribution of trait values?
-See also section 3.
+\begin{point}{}
+     I think this also needs to be emphasised in later sections,
+     e.g., section 4, where the statement that no assumption is made about the initial distribution of the population
+     doesn't seem to be quite right.
+     I think the implicit assumption of Linkage equilibrium does come in
+     once one assumes that all the $\eta_{l}^{i}$ are independent.
 \end{point}
@@
 \reply{
-    We have modified the sentence to say
-    ``implies a Gaussian distribution for the offsprings' deviations'',
-    to make it clear that this statement does not refer to the population distribution,
-    but rather to the Mendelian sampling term \revref.
-    See additional points below on this topic.
-}
-
-\begin{point}{}
-    Section 2 and Figure \ref{fig:example_snps}: It is not obvious to me why one would study the ``frequency-weighted'' effect sizes.
-There is an assumption here I think that allele frequencies are uncorrelated with effect sizes: this will not be true if for example traits are under some kind of stabilising selection: see e.g., \citet{simons2018population} for a detailed exploration of this.
-Moreover, if SNPs are not truly causal but only tag causal alleles, then weighting by allele frequencies can be even more problematic?
-
-As an alternative, would it make sense to look at what fraction of loci have effect size difference between alternative alleles to be $>2t$ without weighing by allele frequency?
-More broadly, how do we interpret the frequency-weighted vs. unweighted distributions?
-\end{point}
-
-\reply{
-    Good question; this was somewhat obscure in the previous draft.
-    We've added a comment which hopefully explains why the frequency-weighted distribution
-    is the correct thing to consider for this calculation \revref,
-    and have added some more discussion about the point in general
-    to the Discussion \llname{discussion_DFEs}.
-}
-
-\begin{point}{}
-I was also wondering about the ``goodness of fit'' of observed distributions to power laws.
-I am not necessarily recommending a detailed quantitative analysis, but nevertheless it may be useful to comment on whether one has power to say reject alternative distributions (e.g., a mixture of exponentials) that are often considered.
-If not, then perhaps the exact exponents inferred are unimportant, and what the data points to more broadly, is a broad distribution of effect sizes.
-\end{point}
-
-\reply{
-    This is a good question, and we had in fact done some exploration of this question
-    but not included it in the paper.
-    We've still not done a detailed analysis,
-    but now provide some discussion and numerical exploration of our precision here:
-    see Figure~\ref{fig:power} and \revref.
-}
-
-\begin{point}{}
-    Section 3, Sentence starting ``We know from \citet{barton2017infinitesimal}, \ldots'' \revref: As far as I understand, the statement of \citet{barton2017infinitesimal} is that the joint distribution of trait values of offspring is multivaraite normal, conditioned on the two parents.
-I don't think the joint distribution of two randomly picked parents and their offspring is mutivariate normal, as this would require some kind of normality assumption about trait values in the population as a whole.
-It'd be helpful to distinguish more carefully between conditioned and unconditioned distributions, especially in this section.
-\end{point}
-
-\reply{
-    Looking back at \citet{barton2017infinitesimal}, we see where the confusion comes from,
-    and have added some clarifying remarks at the end of this paragraph \revref.
-    % Quoting \citet{barton2017infinitesimal},
-    % ``Even though the trait distribution across the whole population may be far from Gaussian,
-    % the multivariate normal will play a central r\hat{o}le in our analysis''.
-    In more detail:
-    \citet{barton2017infinitesimal} (``main result'', equation 9)
-    shows that conditioned on the trait values in generation $t$,
-    the values of generation $t+1$ are also multivariate Gaussian.
-    This implies that the joint distribution of traits of any subset of individuals
-    is multivariate Gaussian (since the initial generation is Gaussian)
-    -- as \citet{barton2017infinitesimal} says,
-    \textit{``equally we could have conditioned on the values of any subset of relatives
-    and the same would hold true: expected trait values would be
-    a linear functional (determined by the pedigree) of the values
-    on which we conditioned, but segregation variances would be unchanged.''}
-    (More generally, if $X$ is multivariate Gaussian, and $Y$ conditioned on $X$ is also multivariate Gaussian
-    with mean but not covariance depending on $X$,
-    then $(X,Y)$ are jointly multivariate Gaussian.)
-    We now also discuss the point further in a subsequent paragraph \llname{pop_not_gaussian}.
-}
-
-\begin{point}{}
-    In general, I find this section rather confused: the infinitesimal model basically implies that the ``Mendelian segregation term'' $R_{M}$ (which is the difference between an individual and the midparent value $\bar{Z}_{M}$) is normally distributed with mean zero and variance that depends on the relatedness between parents but is independent of the midparent value $\bar{Z}_{M}$.
-However, in Proposition \ref{prop:parental_contribs}, the authors claim that $\bar{Z}_{M}$ and $R_{M}$ are jointly Gaussian (which is equivalent to saying that the trait value of an individual and the mean of its parents' trait values is jointly Gaussian), which is in general not true.
-\end{point}
-
-\reply{
-    If the reviewer had said ``confusing'' rather than ``confused'',
-    then we would heartily agree.
-    After some difficult thinking through of things,
-    we've provided a good bit of additional clarification
-    about what the Proposition means and doesn't mean:
-    see in particular \llname{pop_not_gaussian}.
-}
-
-\begin{point}{}
-    I think it is also not correct to claim that ``Since models can be easily set up for which trait distributions are not Gaussian, the implication of this is that for such models, independence of the Mendelian sampling term is not likely a good assumption.'' \revref.
-As far as I understand, a non-Gaussian distribution of trait values can arise quite easily under the infinitesimal model (where the Mendelian segregation term is independent of the midparent value) e.g, due to strong or non-Gaussian selection or migration or some combination of the two (see, e.g., Figure 1 in Barton et al).
-\end{point}
-
-\reply{
-    Thanks for the catch -- this was an unfortunate mistake that might have led to some of the previous confusion.
-    We've changed this to ``since models can be easily set up for which the distribution
-    of Mendelian sampling terms is not Gaussian, \ldots''
-    and provided additional context \revref.
-}
-
-\begin{point}{}
-    Section 4, first paragraph \revref: I don't think the third assumption (that trait values in the population follow a Gaussian distribution) is a component of the ``infinitesimal model''.
-I think Turelli (2017) is also quite clear about this stating that a Gaussian distribution of trait values should only emerge for Gaussian or very weak selection on the population.
-\end{point}
-
-\reply{
-    We agree that \citet{turelli2017commentary} is clear about when the Gaussian emerges.
-    It also seems quite clear to us that what we say in this paragraph agrees
-    with \citet{turelli2017commentary}, who says in his Abstract that
-    \begin{quote}
-    This commentary distinguishes three nested approximations, referred to as “infinitesimal genetics,” “Gaussian descendants” and “Gaussian population,” each plausibly called “the infinitesimal model.”
-    \end{quote} 
-    We think that the underlying issue here may be about motivation for the section.
-    To make the motivation, and the point about the Gaussian population assumption more clear,
-    we've added some explanatory text \revref.
-    Also note that this section does \emph{not} make a Gaussian-population (or, non-Gaussian population) assumption;
-    to make this more clear we've changed the phrase later in this paragraph
-    from ``non-Gaussian model'' to ``model with non-Gaussian Mendelian sampling terms''.
-}
-
-\begin{point}{}
-More generally, I found it a bit hard to follow the overall logical flow of section 4.1, especially once the authors launch into an exploration of the ``reproduction'' and ``noise'' terms (pages 9 and 10).
-    It would be useful to clarify what the biological/intuitive meaning of the noise and reproduction terms at the outset.
-It would also be useful to state (at the beginning) what the main goal of these explorations is: (is the goal to identify when both terms ``have well-posed limits independent of the other'' as stated at \llname{limit_remark}), and also to summarise (at the end) to what extent one can do this in all generality.
-\end{point}
-
-\reply{
-    Good suggestion; we've added more clarifying points;
-    see \llname{clarification1}
-    and \llname{clarification2}
-    (and \llname{clarification3}).
-}
-
-\begin{point}{}
-    Last inequality above \revref: Is this the distribution of the largest of the $M$ alleles carried by an individual, conditioned on the trait value $Z$ of the individual (and this is independent of $Z$?), or is it just the largest of $M$ iid draws from a Cauchy distribution?
-If the former, then it'd be good to say this explicitly.
-If the latter, then this does not quite answer the question posed at the beginning of the section about how much information knowing the trait value of an individual gives about underlying allelic effects.
-\end{point}
-
-\reply{
-    It is both --
-    the calculation is, as the reviewer points out, about the largest of $M$ iid Cauchys,
-    but those are the alleles carried by a parent in the initial generation.
-    Furthermore, as explained at the beginning of this section \llname{crucial_step},
-    this calculation about iid effects is indeed the key step
-    (in \citet{Fisher1918} and \cite{barton2017infinitesimal})
-    in describing the distribution of trait values in offspring.
-    We've added an sentence that we think should remind the reader
-    why we have in fact answered the question \revref.
-}
-
-\begin{point}{}
-Also, maybe worth specifying what the corresponding distribution for the largest of $M$ alleles looks like when the effect size distribution has finite variance, in order to highlight the contrast between ``well-behaved'' and heavy-tailed effect size distributions.
-\end{point}
-
-\reply{
-    Good idea; we have done this \revref.
-}
-
-\subsubsection*{Minor comments:}
-
-\begin{point}{}
-    Introduction: Also worth citing Fisher, Bulmer etc. (the original references) when introducing the infinitesimal model?
-\end{point}
-
-\reply{
-    Good idea, we have added some citations \revref.
-    However, note that what exactly are the ``original references'' is not so clear:
-    ``\textit{Indeed, its origins are lost in the mists of time (M. Bulmer, W.G. Hill, pers.  comm.).}''
-    \citep[from][]{barton2017infinitesimal}
-}
-
-\begin{point}{\revref}
-    ``Perhaps the distribution of effect sizes within each gene is Normal'': I am confused by this.
-Do you mean that the distribution of effect sizes across all genes (with roughly the same ``proximity'' to the trait) is normal.
-If yes, then the above phrasing is a bit misleading.
-If what is meant that is really that the effect sizes of different alleles within a gene are normally distributed, then I am not sure if this is entirely plausible: see various papers by Turelli (e.g., ``Heritable Genetic Variation via Mutation-Selection Balance: Lerch's Zeta Meets the Abdominal Bristle.'').
-Presumably, a normal distribution of effect sizes for a sufficiently large genomic region is a good approximation but it is unclear what is large enough: smaller or larger than a typical gene...?
-\end{point}
-
-\reply{
-    We've rephrased this, and replaced ``gene'' with ``region'' \revref --
-    really, we just want to motivate a mixture model,
-    as the "scale mixture of Gaussians" is a common way to get fat-tailed distributions,
-    and at least roughly biologically plausible.
-    On the other hand, we don't want to get into the weeds with more detail or discussion here.
-}
-
-\begin{point}{\revref}
-    Sentence starting ``In fact, many traits claimed...'': This phrasing here is rather cryptic and I do not understand what is being said. Maybe rephrase? Also a reference would help...
-\end{point}
-
-\reply{
-    Perhaps it is better now? \revref
-}
-
-\begin{point}{}
-    Figure \ref{fig:exponent_hist}: Is there also a correlation between the number of cases and the number of SNPs? In other words, to what extent are (b) and (c) independent?
-\end{point}
-
-\reply{
-    Yes; we've added this to Figure~\ref{fig:exponent_hist}.
-}
-
-\begin{point}{\revref}
-    Sentence starting ``To see what knowing the parental trait value...'': In what sense is this the ``parental'' trait value??
-\end{point}
-
-\reply{
-    They are a member of the (first) ancestral generation, and therefore a (potential) parent.
-    We've clarified this \revref.
-}
-
-\begin{point}{\revref}
-    Last inequality on Page 13: should the subscript be $j$ instead of $i$?
-\end{point}
-
-\reply{It should be $\ell$, and is now corrected \revref.  Thanks for noticing the discrepancy; is an artefact of an earlier version which used $i$ for an index on loci.
-}
-
-\begin{point}{}
-    Simulations: Why not choose the same effect size distributions in the neutral case and the case with stabilizing selection?
-Alternatively, if different effect size distributions are used, it may be good to plot the distribution of (appropriately) scaled trait values.
-Right now the figures give the impression that more variation is maintained under stabilizing selection..
-\end{point}
-
-\reply{
-    These are good suggestions, however, we have encountered difficulties in implementing this,
-    so we've settled for a note alerting the reader not to read too much into the absolute scale
-    (see Figure~\ref{fig:sel_trait_distrns} caption).
-}
-
-\begin{point}{}
-    Figure \ref{fig:seg_noise} caption: I find the phrasing ``(a,b) chosen to span the 5\% of midparents centered on the 10\% quantile of midparent value'' a bit confusing.
-Does this mean a and b are respectively the 7.5\% and 12.5\% quantile of the distribution of midparent trait values?
-\end{point}
-
-\reply{
-    That's right; we've rephrased this confusing explanation.
-}
-
-\begin{point}{}
-    Discussion: In the context of theory combining large and small-effect loci, \citet{chevin2008selective}: ``Selective Sweep at a Quantitative Trait Locus in the Presence of Background Genetic Variation'' and \citet{lande1983response}: ``The response to selection on major and minor mutations affecting a metrical trait'' may also be somewhat relevant.
-\end{point}
-
-\reply{
-    Thanks! We've incorporated these \revref.
+    We don't think this applies, as
+    Section 4 is explicitly not working with the model of Section 3 --
+    there are no loci explicitly modeled at all in Section 4.
+    We have added a reminder to emphasize this fact \revref.
 }
@@
 
-\begin{quote}
-The authors investigate the possibility of extending results of \citet{barton2017infinitesimal} on the
-justification of the infinitesimal model to distributions of effect sizes with a large tail (i.e.,
-polynomial decay with exponent $\alpha < 2$). This is biologically interesting and important
-because there are clear indications that variation in some traits is caused by a mixture
-of polygenic (infinitesimal) contributions and one or few loci of large effect. It also a
-mathematically difficult topic because it leads into the realm of generalized Central Limit
-Theorems, namely stable laws.
-
-After a brief, partly rough (see detailed comments below) introduction the ms starts with
-a statistical analysis of the effect size distributions of a large number of traits (all human
-disease related) for which GWAS data are available. The evaluation shows that almost
-all estimated tail exponents are between 1 and 2.5 with a quite flat distribution between
-about 1.3 and 2 (some potential statistical artifacts are noted). However, the authors do
-not discuss the influence of selection, which could be essential because the traits are disease
-related.
-\end{quote}
-
-Thanks for the close reading.
-We didn't discuss selection here because the question of whether selection contributes
-to the distribution of segregating allelic effects
-is somewhat orthogonal to the question at hand here:
-in other words, the infinitesimal model describes (an approximation for)
-the \emph{results of} random reassortment (etcetera) of whatever ancestral variation is present,
-and is therefore relatively agnostic to how that ancestral variation got there.
-This works because the infinitesimal model describes trait variation after reproduction
-but before selection.
-Note that, for instance, there is no obstacle to there being a correlation between effect sizes and allele frequencies,
-as one might expect under selection.
-However, we have added some clarification about this --
-see \revreffull{1}{3} and \llname{discussion_DFEs}.
-
-\begin{quote}
-In Section 3, the authors show a kind of negative result concerning the possibility of $\alpha$-stable 
-distributions with $\alpha < 2$ (Proposition 1): If the deviations of the offspring trait
-from the midparent value (here called the Mendelian sampling term, $R$) are independent
-of the midparent value $\bar Z$ (part of the core result in \citet{barton2017infinitesimal}), then the joint
-distribution of $R$ and $\bar Z$ is Gaussian. Thus, this independence is not compatible with ‘$\alpha$-stable 
-laws’. In fact, as they show (by example) and discuss more generally in Sect. 4, a
-single polymorphic large-effect locus causes dependence of $R$ and $\bar Z$.
-\end{quote}
-
-We'd like to remind the reviewer of the caveat that this result
-also depends on a bit more -- independence of inherited and non-inherited contributions
-(e.g., as under an additive model of unlinked loci).
-
-\begin{quote}
-In Section 4, a version of the infinitesimal model is investigated that proposes a much
-stronger result than a Gaussian Mendelian sampling term independent of the parental
-trait distribution, namely a Gaussian distribution of the trait (within the population)
-under mutation-selection balance. Although such a model has occasionally been assumed
-in the literature, it has never been justified rigorously and, presumably, is not justifiable
-(on the basis of realistic assumptions – as clearly noted by Turelli’s brief review and
-previous extensive work by Turelli, Barton, and others). Nevertheless, the analysis of the
-authors is interesting from a theoretical point of view because they investigate whether
-such deterministic mutation-selection balance is possible in a non-Gaussian model that
-retains independence of the Mendelian sampling terms, and if such a model might yield a
-reasonable approximation to an additive model with non-Gaussian effects.
-\end{quote}
-
-We're glad the reviewer liked this section,
-but perhaps our introduction to the section was misleading --
-in fact we do \emph{not} assume a Gaussian population distribution,
-or any population distribution.
-Instead, we \emph{ask} which distributions might be population distributions
-under this generalized model,
-using arguments similar to those found in the literature.
-We've tried to make this more clear -- see additional text at \revreffull{1}{8}
-and \llname{clarification3}
-(and our response to Reviewer 1, point 8).
-
-\begin{quote}
-The authors use a simplistic Moran model to investigate these questions. However, mutation 
-doesn’t seem to be modeled explicitly. Is it mutation if a parent is replaced by an
-offspring of the type described in the display equation at \llname{offspring_eqn}?
-Maybe yes, but not necessarily because of segregation and recombination contribute substantially to $R$.
-Doesn’t the present assumption imply that the mutation rate is one and that mutation is modeled as a
-kind of diffusion? In the latter case, a Gaussian distribution can result. With explicit
-mutation models, where mutations occur per locus and a finite number of loci determines the
-trait additively, a Gaussian distribution never results (and in no known limit; see above).
-How does population size enter? Is this a deterministic Moran model?
-\end{quote}
-
-Mutation would enter into $R$, a point we now make \llname{mutation_note},
-but without detail, as we are aiming here for a high-level and quick calculation,
-not a detailed investigation.
-In fact, we thought a bit about how to set up an explicit genotype-to-phenotype map
-in which the Mendelian sampling terms are non-Gaussian and independent of parental traits,
-but did not arrive at a concrete solution --
-we think this is probably possible,
-but as we know from Proposition~\ref{prop:parental_contribs},
-cannot be a simple additive model.
-Investigating how exactly a concrete model of mutation would affect things is beyond the scope of this paper,
-and we hope the interesting bits of this section
-outweigh the lack of an explicit mutationi model.
-We have added a note that hopefully clarifies the questions
-about population size \llname{deterministic_point}.
-
-\begin{quote}
-Anyway, the results
-in Sect 4 are of interest because they strongly constrain the possibilities to justify such
-‘Gaussian-population approximation’ by any infinitesimal-like model (essentially to
-mixtures of Cauchy distributions. The noted relation to blending inheritance is intriguing!).
-My main concern is that especially this section needs are clearer description of the model,
-and clear distinction between this ‘maximalist’ model (which has little theoretical
-support) and the model investigated by Barton et al (2017). Also a much clearer summary
-and discussion of the implications of the results shown here are needed. It should not be
-assumed that readers have read Turelli (2017) and even less that they have studied
-\citet{barton2017infinitesimal} in considerable detail!
-\end{quote}
-
-Thanks for the suggestions.
-We also thought the section was useful (and also intruiging!),
-which is why we left the section in.
-(Also note it motivates the form of selection in our simulations.)
-We share the reviewer's concern about context,
-and have added some additional text
-while trying not to get into the weeds too much
--- in particular, see the new text around \llname{clarification3}.
-
-\begin{quote}
-In Section 5, it is pointed out why the Barton et al. (2017) fails in the case of ‘Cauchy
-noise’ (and presumably for any other limit law with $\alpha < 2$). In the absence of giving the
-estimate in App D in Barton et al, the explanation around \llname{approx_eqn} is difficult to
-follow. Please provide their estimate for easier comparison (do not assume that readers
-studied Barton et al; assume that they scanned (parts) of it).
-\end{quote}
-
-\comment{TODO}
-
-\begin{quote}
-Finally, Section 6 presents an insightful comparison of the distributions of trait values
-resulting from either Gaussian or Cauchy mutation-effect distributions. This seems to be
-an explicit genetic model, but again, what is the mutation rate? Clearly, the mutation
-rate determines the variance (or more generally variability) of the population distribution.
-Please clarify the model. It also needs to be pointed out clearly that such results seem to
-be at variance with all (?) explicit multilocus models that have been studied so far.
-\end{quote}
-
-Honestly, we're confused by this comment.
-The mutation rate is $10^{-9}$ per bp on a genome of length $10^8$,
-as stated at \llname{mutation_rate} -- perhaps the reviewer missed that paragraph?
-We're also unclear what exactly is meant by ``at variance with all \ldots models'' --
-it's true that no-one else has (probably?) simulated this exact model,
-but when it comes to simulations, this is often true?
-We're not sure which specific aspects of difference the reviewer has in mind pointing out.
-
-\begin{quote}
-In summary, this ms contains several interesting aspects and will deserve publication after
-improving the presentation and clarifying a number of issues.
-\end{quote}
-
-Thanks! We hope our improvements
-have made everything at least somewhat clearer.
-
-\begin{point}{}
-Intro: Not too many readers will be familiar with $\alpha$-stable distributions. It may be
-useful to clearly point out the difference to the Gaussian case and what happens if $\alpha \ge 2$
-(obviously, this is of relevance).
+\begin{point}{\revref}
+     p.4, line 3 (from top): garbled sentence
 \end{point}
@@
 \reply{
-    Good idea; we've added this \revref.
+    We've simplified the sentence.
+}
+
+\begin{point}{\revref}
+     delete one `that'
+\end{point}
+
+\reply{
+    Done.
 }
@@
 \begin{point}{}
-p 3: State that M is the number of loci.
+     Section 6: it is unfortunate to use $\mu$ for the mutation probability and $\mu(x)$ for the death rate. Maybe use $u$ for the mutation prob.
 \end{point}
@@
 \reply{
-    Whoops - done \revref.
+    Good call; done.
 }
@@
 \begin{point}{\revref}
-Should this read ‘between 20\% and 10\%’? Why this choice?
+     p.18, line 460: delete one comma
 \end{point}
@@
 \reply{
-    No, but as written it was confusing, and we've clarified \revref.
-    We chose these values by empirical inspection:
-    the lower bound was 20 SNPs so that it was not unduly affected by a few large SNPs,
-    and the upper bound as 10\% because we wanted the ``tail'', not the ``bulk''.
-    (We also checked that changing these values somewhat didn't affect results.)
+    Whoops! Done.
 }
-
-\begin{point}{\revref}
-First line in Sect 3: add ‘as $M \to \infty$’. In the next line, this is first time that $\alpha = 2$ is
-mentioned. Clarify earlier that this case is very different from $\alpha < 2$ (maybe already below
-the last display eq on p3).
-\end{point}
-
-\reply{
-    Good idea -- done \revref. (Also see \revreffull{2}{1} and \revreffull{1}{11}.)
-}
-
-\begin{point}{\revref}
-What means `nearby alleles'? Do you mean physically linked
-    loci? Another \llname{loci_polymorphic}: alleles are not polymorphic; loci can be polymorphic.
-\end{point}
-
-\reply{
-    Good points; both are fixed: \revref{} and \llname{loci_polymorphic}.
-}
-
-\begin{point}{}
-Sect \ref{sec:abnormal}: As noted above, please describe this model in more detail to distinguish it
-from that in \citet{barton2017infinitesimal}, which makes much weaker assumptions. Note that it is
-not justifiable except under quite extreme assumptions.
-\end{point}
-
-\reply{
-    We've added more text to hopefully
-    put the reader in the right frame of mind:
-    see especially \llname{clarification3}.
-}
-
-\begin{point}{\revref}
-Third display equation in Sect 4.1 \revref: The index k in the product should start at k = 0.
-\end{point}
-
-\reply{Thank you for noticing this.  We have corrected this, and the subsequent indices \revref.
-}
-
-\begin{point}{\revref}
-Delete ‘to’
-\end{point}
-
-\reply{Done \revref.
-}
-
-\begin{point}{}
-Around \llname{cauchy_note}: It might be useful to mention earlier that the Cauchy distribution is an archetypical
-example for an $\alpha$-stable distribution. The enthusiasm among biologists may be limited
-though, because all the empirical trait distributions they find have mean and variance. I
-don’t know if there are examples in which the variance grows (significantly) as the sample
-size increases. In the last para of Sect 4.1 \revref, please specify ‘one will get convergence’. I
-suppose this refers again to $M \to \infty$.
-\end{point}
-
-\reply{
-    Good suggestion -- we've added the note about the Cauchy \llname{cauchy_note}.
-    (We haven't tried to make biologists more enthusiastic, though,
-    since we don't know of a situation in which strict blending inheritance seems terribly likely.)
-    As for the note about ``one will get convergence'' -- it is as the number of generations increases,
-    not as $M \to \infty$ (in fact, $M$ does not enter into this section at all);
-    we've rewritten this statement to remind the reader
-    about what's going on.
-}
-
-\begin{point}{}
-p 12, last display equation: Please relate this explicitly to the corresponding result in
-Barton et al. (e.g., by stating their estimate).
-\end{point}
-
-\reply{
-    \comment{TODO}
-}
-
-\begin{point}{\revref}
-This discussion is too vague. At least, adding a summary of the main findings
-would be appropriate (always noting for which version of the infinitesimal model).
-\end{point}
-
-\reply{
-    We have added a good bit more discussion, including a summary \revref.
-}
-
-\begin{point}{\revref}
-Does `Lévy stable distribution’ refer to the family of $\alpha$-stable
-distributions? (I don’t think it has not been used before.)
-\end{point}
-
-\reply{
-    That's right; changed to ``$\alpha$-stable distribution'' \revref.
-}

### 2024-01-23 writeup.tex
bob note
new reviewer responses

@@
 and a covariance independent of the parental trait values (but that depends on how they are related).
-\citet{barton2017infinitesimal} give an overview of the complex history of the model, \revpoint{1}{12}
+\citet{barton2017infinitesimal} give an overview of the complex history of the model,
 which is used in or adjacent to a good deal of theoretical work on quantitative genetics
@@
 perhaps a trait is affected by mutations in a number of genomic regions,
-and the distribution of effect sizes within each region is Normal, \revpoint{1}{13}
+and the distribution of effect sizes within each region is Normal,
 but with a standard deviation specific to the region that varies with ``proximity'' of that region to the trait
@@
 (and the approximation becomes exact, as $n \to \infty$).
-On the other hand, if $\alpha \ge 2$ then the distribution has finite variance, \revpoint{2}{1} 
+On the other hand, if $\alpha \ge 2$ then the distribution has finite variance,
 and so the sum of $n$ copies, divided by $\sqrt{n}$, converges to a Gaussian distribution.
@@
 so that e.g., for $\alpha=1$ a positive proportion of the sum of $n$ entries will be contributed by a single one.
-(On the other hand, the largest of $n$ Gaussians is of order $\sqrt{\log n}$.) \revpoint{1}{11}
+(On the other hand, the largest of $n$ Gaussians is of order $\sqrt{\log n}$.)
 
 This suggests exploring whether stable distributions might reasonably stand in for the Gaussian
-in the infinitesimal model. \revpoint{1}{1}
+in the infinitesimal model.
 In this paper, we make preliminary explorations in that direction:
@@
 The immediate answer to this will be ``no'' --
-in fact, we will see that independence of offsprings' deviations and midparent values \revpoint{1}{2}
+in fact, we will see that independence of offsprings' deviations and midparent values
 \emph{implies} a Gaussian distribution for the offsprings' deviations, in some sense.
@@
 and interactions between alleles, between loci, and with the environment are often important in practice.
-In fact, some important traits that have been claimed to be mostly genetically determined \revpoint{1}{14}
+In fact, some important traits that have been claimed to be mostly genetically determined
 only appear heritable through non-genetic means
@@
 where $\eta_\ell$ is the contribution of the allele carried at the $\ell^\text{th}$ locus,
-and $M$ is the number of loci. \revpoint{2}{2}
+and $M$ is the number of loci.
 The effects of the alleles carried by the two parents
@@
 
+We do not consider the effects of new mutations.
+
 %%%%%%%%%%%%%%%%%%%%%%
@@
 $$
-where $\eta_\ell^{[i]}$ is the effect of the allele in parent $i$ at locus $\ell$,
 the appropriate central limit theorem for the sum is determined by
-the tail behavior of $|(X - 1/2) \eta|$,
+the tail behavior of $|(X - 1/2) \eta|$,  \revpoint{2}{1}
 where $X$ is Bernoulli(1/2) and $\eta$ is the effect
@@
 at a uniformly chosen locus.
-Since $|X-1/2| = 1/2$, we'd like to know
+% $\eta_\ell^{[i]}$ is the effect of the allele in parent $i$ at locus $\ell$,
+Since $|X-1/2| = 1/2$, we'd therefore like to know
 if $\P\{|\eta|>2t\}$ is proportional to $t^{-\alpha}$,
@@
 $$
-Note that the relevant effect size distribution here is weighted by frequency, \revpoint{1}{3}
+Note that the relevant effect size distribution here is weighted by frequency,
 which is necessary because we're looking at the deviation of a typical offspring from their parents,
@@
 all values of $t$ for which the number of SNPs with absolute effect above $t$
-is at least 20 and at most 10\% of all SNPs.) \revpoint{2}{3}
+is at least 20 and at most 10\% of all SNPs.)
 Examples are shown in Figure~\ref{fig:example_snps}.
@@
 
-A possible concern is that tail exponents are difficult to estimate accurately, \revpoint{1}{4}
+A possible concern is that tail exponents are difficult to estimate accurately,
 but simulation-based estimates of accuracy (Supplemental Figure~\ref{fig:power})
@@
 Now we return to the theory.
-We know from \citet{barton2017infinitesimal} \revpoint{1}{5} \revpoint{2}{4}
+We know from \citet{barton2017infinitesimal}
 that if the $\eta^{[i]}_\ell$ are
@@
 must also, by logical necessity, have a Gaussian distribution for the two.
-In other words, \revpoint{1}{7}
+In other words,
 since models can be easily set up for which offspring minus midparent traits
@@
 although in practice this would quickly create dependencies between the effects
-of alleles at such linked loci. \revpoint{2}{5}
+of alleles at such linked loci.
 Also note that it is \emph{not} usually true that the components --
@@
 
+Finally, note that we follow \citet{barton2017infinitesimal} by assuming that loci are independent
+(\ie no linkage disequilibrium) \revpoint{1}{1}
+in the parental generation.
+% (They are not necessarily identically distributed, however,
+% so the Central Limit Theorem need not apply \emph{a priori}.)
+One might then wonder if by relaxing this assumption
+one could find a model in which the Mendelian sampling term was independent of midparent value
+and yet had a non-Gaussian distribution.
+To be interesting, such a model would need to maintain the nonindependence structure
+that guarantees these properties,
+which seems hard to arrange.
+
 As we discuss more later,
-the presence of a large effect allele at a polymorphic locus \llabel{loci_polymorphic}
+the presence of a large effect allele at a polymorphic locus
 makes $R$ and $\bar Z$ dependent:
@@
 % NOT HELPFUL
-% Finally, a possible point of confusion is that \revpoint{1}{6}
+% Finally, a possible point of confusion is that
 % the infinitesimal model is often phrased as a statement about offspring trait distributions
@@
 for suitable choices of the empirical mean and variance $m$ and $\sigma^{2}$).
-As also discussed in \citet{turelli2017commentary}, \revpoint{1}{8}
+As also discussed in \citet{turelli2017commentary},
 this assumption is rather fragile --
@@
 (Indeed, above we have put substantial restrictions on what models of inheritance
-might lead to such models.)
+might lead to such models;
+in particular,
+we are \emph{not} working with the explicit sum-of-$\eta_\ell$ model of the previous sections.) \revpoint{1}{2}
 We might also ask whether such a model
@@
 we do not attempt to disambiguate the various possible contributions.) \llabel{mutation_note}
-Mutation-selection balance demands that that
+Mutation-selection balance demands that \revpoint{2}{2}
 the distribution of individuals removed by death
@@
 Any fixed point $\phi(\theta)$ must be the limit of an arbitrary function, $\varphi(\theta)$, under repeated applications of $F$,
-\ie \revpoint{2}{7}
+\ie
 \begin{equation}\label{eq:fpeq}
@@
 We briefly explore what possible limiting values are valid characteristic functions. 
-We refer to the first term on the left hand side, $\psi_{n}(\theta) = 2^{n} \psi\left(\frac{\theta}{2^{n}}\right)$,  as the \emph{reproduction term} and the remaining sum, $\Psi_{n}(\theta) = \sum_{k=0}^{n-1} 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)$, as the \emph{noise term}, and only consider the case \revpoint{2}{8}
+We refer to the first term on the left hand side, $\psi_{n}(\theta) = 2^{n} \psi\left(\frac{\theta}{2^{n}}\right)$,  as the \emph{reproduction term} and the remaining sum, $\Psi_{n}(\theta) = \sum_{k=0}^{n-1} 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)$, as the \emph{noise term}, and only consider the case
 when both have well-posed limits independent of the other. \llabel{limit_remark} % We don't need to limit to this caee.
@@
 Thus, $e^{\psi_{\infty}(\theta)}$ must satisfy $|e^{\psi_{\infty}(\theta)}| = e^{\Re \psi_{\infty}(\theta)} \leq 1$, whence $\Re \psi_{\infty}(\theta) \leq 0$ and $\gamma \geq 0$.
-In particular, if $\gamma > 0$, then we recognize $e^{\psi_{\infty}(\theta)} = e^{im \theta -\gamma|\theta|}$ as the characteristic function of a Cauchy distribution with median m and median absolute deviation $\gamma$ (\ie, the canonical $\alpha$-stable distribution with $\alpha=1$), \llabel{cauchy_note} whereas if $\gamma = 0$, then $e^{\psi_{\infty}(\theta)} = e^{im \theta}$ is the Fourier transform of a Dirac mass at $m$, \ie the characteristic function for a monomorphic population with all individuals having trait value equal to $m$.
+In particular, if $\gamma > 0$, then we recognize $e^{\psi_{\infty}(\theta)} = e^{im \theta -\gamma|\theta|}$ as the characteristic function of a Cauchy distribution with median m and median absolute deviation $\gamma$ (\ie the canonical $\alpha$-stable distribution with $\alpha=1$), \llabel{cauchy_note} whereas if $\gamma = 0$, then $e^{\psi_{\infty}(\theta)} = e^{im \theta}$ is the Fourier transform of a Dirac mass at $m$, \ie the characteristic function for a monomorphic population with all individuals having trait value equal to $m$.
 
@@
 
-More generally, \revpoint{2}{9}
+More generally,
 the trait distribution in the population will converge as time proceeds
@@
 
-To see what knowing the trait value $Z^{[j]}_{M}$ \revpoint{1}{16}
+To see what knowing the trait value $Z^{[j]}_{M}$
 of a particular parent in the ancestral generation
@@
 converges weakly as $M \to \infty$ to a Poisson random variable with mean $2/(\pi \epsilon)$ (see \eg Theorem 2.6.1 in \cite{Durrett2005}).
-In particular, for large values of $M$ and fixed $\epsilon > 0$,\revpoint{1}{17}
+In particular, for large values of $M$ and fixed $\epsilon > 0$,
 \begin{align*}
@@
 which approaches 1 as $\epsilon \to 0$, while the expected number of $M$ alleles such that $\frac{\eta^{[j]}_{\ell}}{\bar{\gamma} M} > \epsilon$ is asymptotically $\frac{2}{\pi\epsilon}$. 
-By contrast, in \citet{fisher1918correlation} and \citet{barton2017infinitesimal} \revpoint{2}{10}, the scaled allelic effect at any locus $\ell$ has magnitude bounded by $\frac{B}{\sqrt{M}}$, and thus, no matter how many alleles one considers, there will be no large effect alleles that are informative of the trait value.
-
-To summarize, the key argument in \citet{fisher1918correlation} and \citet{barton2017infinitesimal}  \revpoint{1}{10}
+By contrast, in \citet{fisher1918correlation} and \citet{barton2017infinitesimal}, the scaled allelic effect at any locus $\ell$ has magnitude bounded by $\frac{B}{\sqrt{M}}$, and thus, no matter how many alleles one considers, there will be no large effect alleles that are informative of the trait value.
+
+To summarize, the key argument in \citet{fisher1918correlation} and \citet{barton2017infinitesimal}
 requires that knowing the sum of parental effects does not give much information 
@@
         had populations of size $N=1000$,
-        a genome of length $L=10^8$ bp and mutation rate of $\mu=10^{-9}$ per bp per generation;
+        a genome of length $L=10^8$ bp and mutation rate of $u=10^{-9}$ per bp per generation;
         the effect size distributions were
-        Normal with mean 0 and standard deviation $1/\sqrt{4NL\mu}$
-        and Cauchy with center 0 and scale $1/(4NL\mu)$, respectively.
+        Normal with mean 0 and standard deviation $1/\sqrt{4NLu}$
+        and Cauchy with center 0 and scale $1/(4NLu)$, respectively.
         \label{fig:trait_distrns}
@@
 \section{Discussion}
-    \revpoint{2}{11}
 
@@
 Next, we show that in fact the infinitesimal model \emph{cannot} describe non-Gaussian distributions
-(\ie, non-Gaussian deviation of offspring trait from midparent value)
+(\ie non-Gaussian deviation of offspring trait from midparent value) \revpoint{2}{4}
 if the traits are determined additively from independent loci
@@
 For simplicity, we have focused on the $\alpha=1$ case (which includes the Cauchy distribution).
-However, examination of Figure~\ref{fig:exponent_hist} suggests that $\alpha=3/2$ \revpoint{2}{12}
+However, examination of Figure~\ref{fig:exponent_hist} suggests that $\alpha=3/2$
 would be a more natural choice for an effect size distribution.
@@
 Various authors have taken the approach of separating contributions of large- and small-effect loci
-(e.g., \citet{lande1983response} and \citet{chevin2008selective}). \revpoint{1}{20}
+(e.g., \citet{lande1983response} and \citet{chevin2008selective}).
 In practice, breeders have long embellished the ``animal model'' (essentially, the infinitesimal model)
@@
 \subsection*{Acknowledgements}
-Thanks go to Nate Pope for useful discussion
+Thanks go to Nate Pope and Bob Week for useful discussion
 and to Gregor Gorjanc for insight into modern uses

### 2024-02-21 review-responses.tex
final response to reviewers

@@
 \begin{quote}
-As you will read from the reviewers' comments below, both of them find that the quality and clarity of the exposition has improved a lot, and they only have a list of minor (nonetheless important) comments that should be addressed in a minor revision.
-Provided the points raised are suitably addressed, the paper will then be accepted for publication.
-And though this has nothing to do with the paper, happy New Year to the authors ;-).
+It would be good to try to address the remaining concern of Reviewer 1 and to
+clarify the corresponding assumption on which your analysis seems to rest.
+After that, the paper will be accepted for publication without further
+reviewing.  
 \end{quote}
 
-Happy New Year!
+Thanks very much, and apologies such a minor change took some time.
+We've tried our best here.
 
@@
 \begin{point}{}
-     As far as I understand, the crucial assumption is that the effect sizes $\eta_{l}^{i}$
-     of the alleles carried by an individual $i$ are independent
-     (as in the first line of Proposition \ref{prop:parental_contribs}),
-     which basically amounts to assuming linkage equilibrium in the population.
-     In that case, of course, all trait values across any group of individuals are multi-variate Gaussian
-     (provided effect size distributions are not heavy-tailed),
-     without needing to condition on any ancestors.
-     Moreover, under these conditions,
-     even the trait value distribution in the population would be Gaussian
-     (provided effect size distributions is not heavy-tailed).
-     As far as I understand, \citet{barton2017infinitesimal} is more general
-     in that it does not require $\eta_{l}^{i}$ to be independent,
-     except in the founder $t=0$ generation
-     (and I am not sure how crucial even that assumption is);
-     in later generations, there can be an arbitrary amount of LD.
-     This is a crucial point, since in general, under the infinitesimal model,
-     non-Gaussian population distributions only arise because of LD, in which case,
-     the $\eta_{l}^{i}$ are not independent.
-     Thus, it is worth highlighting that assuming $\eta_{l}^{i}$ to be independent
-     basically implies linkage equilibrium, and that this is in general,
-     not a requirement for the infinitesimal model
-     or for the derivations of \citet{barton2017infinitesimal},
-     whereas it is a key requirement here.
-     I think this had been a big part of my earlier confusion.
+I don't think the authors quite address the point I had raised, and their
+response and also the change in the paper that they make \revref{} is
+rather cryptic and hard to understand. I think all that I was trying to say was
+that the independence of allelic states, i.e., the LE assumption, is a big part
+of their argument but not of the original infinitesimal model or of the TPB
+2017 paper. And I thus felt that it is important to highlight the rather
+limited nature of their proposition, which I don't think quite comes across.
 \end{point}
 
-\reply{
-    It is true that if the $\eta$ are not independent then $Z_M$ and $V_M$ are not independent,
-    and so the Kac-Bernstein theorem does not apply.
-    However -- returning to the goal of this section here -- it seems hard and maybe impossible
-    to use LD to obtain a inifinitesimal-like, non-Gaussian model.
-    But, we also don't see why it's impossible, and have added a note to this effect \revref.
-    (Also we'd like to point out that the independence only applies to the \emph{parental} generation;
-    there is in fact LD in the offspring's generation, which is the generation Proposition \ref{prop:parental_contribs}
-    refers to. Our results could be extended to further generations, as we remark just before \revref,
-    but since our point is about non-existence of other models, this seemed superfluous.
-    Note also that although we assume independence, we don't assume the $\eta$ are identically distributed,
-    so that a Gaussian or other stable distribution is not guaranteed.)
-}
+\paragraph{Reply:}
+Although what you're saying here seems very clear,
+we must be talking past each other somehow,
+since \citet{barton2017infinitesimal} unambiguously \emph{does} assume LE,
+for at least the main results on
+``the infinitesimal model as the limit of Mendelian inheritance''.
+In \citet{barton2017infinitesimal}, Assumption (4) (``Ancestral population'')
+on page 58 of the TPB PDF,
+a prerequisite to the main convergence results, says:
+``\emph{Although it is not strictly necessary,
+we assume that in generation zero, the individuals that
+found the pedigree are unrelated. They are sampled from
+an ancestral population in which all loci are assumed to be
+in linkage equilibrium.}''
 
-\begin{point}{}
-     I think this also needs to be emphasised in later sections,
-     e.g., section 4, where the statement that no assumption is made about the initial distribution of the population
-     doesn't seem to be quite right.
-     I think the implicit assumption of Linkage equilibrium does come in
-     once one assumes that all the $\eta_{l}^{i}$ are independent.
-\end{point}
+Of course, loci in \emph{subsequent} generations won't be in LE,
+and we assumed in our previous revision that this is what the reviewer was referring to.
+Our argument here is that if the infinitesimal model
+is not a good model for this simplest LE situation,
+then it won't be in subsequent generations also.
 
-\reply{
-    We don't think this applies, as
-    Section 4 is explicitly not working with the model of Section 3 --
-    there are no loci explicitly modeled at all in Section 4.
-    We have added a reminder to emphasize this fact \revref.
-}
+Perhaps by ``the original infinitesimal model or the TPB 2017 paper'',
+the reviewer means something different than these convergence results?
+We certainly agree that \emph{the infinitesimal model},
+stated as a phenomenological model for trait inheritance,
+does not make any assumptions whatsoever about the precise genetic basis
+(including linkage equilibrium or not).
+However, that's probably not what the reviewer means, since that's entirely
+beside the point (as in this section of the paper we're looking at
+consequences of the genetic basis)?
 
-%%%%%%%%%%%%%%
-\reviewersection{2}
-
-\begin{point}{\revref}
-     p.4, line 3 (from top): garbled sentence
-\end{point}
-
-\reply{
-    We've simplified the sentence.
-}
-
-\begin{point}{\revref}
-     delete one `that'
-\end{point}
-
-\reply{
-    Done.
-}
-
-\begin{point}{}
-     Section 6: it is unfortunate to use $\mu$ for the mutation probability and $\mu(x)$ for the death rate. Maybe use $u$ for the mutation prob.
-\end{point}
-
-\reply{
-    Good call; done.
-}
-
-\begin{point}{\revref}
-     p.18, line 460: delete one comma
-\end{point}
-
-\reply{
-    Whoops! Done.
-}
+In summary, we're probably missing something,
+but despite significant thinking, haven't figured out what it might be.
+We've tried (again) to clarify the general point;
+maybe it will help.

### 2024-02-21 writeup.tex
final response to reviewers

@@
 the appropriate central limit theorem for the sum is determined by
-the tail behavior of $|(X - 1/2) \eta|$,  \revpoint{2}{1}
+the tail behavior of $|(X - 1/2) \eta|$,
 where $X$ is Bernoulli(1/2) and $\eta$ is the effect
@@
 
-Under the conditions of the proposition, \llabel{pop_not_gaussian}
+Under the conditions of the proposition,
 the parental traits, $Z^{[1]}_M$ and $Z^{[2]}_M$,
@@
 holds for the offspring of two parents in an arbitrary pedigree,
-using the Skitovich-Darmois theorem \citet{kagan1973characterization}
-% except that the portion of the genome shared in common by the parents is unconstrained.
--- however, there is nothing to be gained by the additional generality.
-
-Finally, note that we follow \citet{barton2017infinitesimal} by assuming that loci are independent
-(\ie no linkage disequilibrium) \revpoint{1}{1}
-in the parental generation.
+using the Skitovich-Darmois theorem \citep{kagan1973characterization},
+except that the distribution of the portion of the genome shared in common by the parents
+is not constrained.
+However, there is little to be gained by the additional generality.
+
+Finally, note that we follow \citet{barton2017infinitesimal}
+by assuming that loci in the founding generation are independent
+(\ie no linkage disequilibrium). \revpoint{1}{1}
 % (They are not necessarily identically distributed, however,
 % so the Central Limit Theorem need not apply \emph{a priori}.)
-One might then wonder if by relaxing this assumption
+Of course, following the offspring foward through a pedigree
+(as \citet{barton2017infinitesimal} do),
+the loci will no longer be independent.
+One might then wonder if by relaxing the assumption of independence between loci
 one could find a model in which the Mendelian sampling term was independent of midparent value
@@
 in particular,
-we are \emph{not} working with the explicit sum-of-$\eta_\ell$ model of the previous sections.) \revpoint{1}{2}
+we are \emph{not} working with the explicit sum-of-$\eta_\ell$ model of the previous sections.)
 We might also ask whether such a model
@@
 
-To summarize, in this section we will uncritically assume that \llabel{clarification3}
+To summarize, in this section we will uncritically assume that
 the independence assumption of the infinitesimal model holds,
@@
 and is replaced by the offspring of two randomly chosen individuals,
-i.e., a new individual with trait \llabel{offspring_eqn}
+i.e., a new individual with trait
 \begin{align*}
@@
 (Note that mutation might enter into $R$, but as the underlying genetic basis is not specified,
-we do not attempt to disambiguate the various possible contributions.) \llabel{mutation_note}
-Mutation-selection balance demands that \revpoint{2}{2}
+we do not attempt to disambiguate the various possible contributions.)
+Mutation-selection balance demands that
 the distribution of individuals removed by death
@@
 % Since the set of functions $\mathcal{S} = \{e^{i \theta x} : \theta \in \R\}$ is separating (see \eg \S 3.4 in \cite{Ethier+Kurtz1986}), it suffices to show that \eqref{eqn:MSB_condition} holds for $f \in \mathcal{S}$, \ie
-If the population is large enough, \llabel{deterministic_point}
+If the population is large enough,
 we can treat the trait distribution as changing deterministically.
@@
 which we may take equal to 1. 
-(To remind the reader, our goal is therefore to describe \llabel{clarification1}
+(To remind the reader, our goal is therefore to describe
 what forms the population distribution might take in a neutral model
@@
 We refer to the first term on the left hand side, $\psi_{n}(\theta) = 2^{n} \psi\left(\frac{\theta}{2^{n}}\right)$,  as the \emph{reproduction term} and the remaining sum, $\Psi_{n}(\theta) = \sum_{k=0}^{n-1} 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)$, as the \emph{noise term}, and only consider the case
-when both have well-posed limits independent of the other. \llabel{limit_remark} % We don't need to limit to this caee.
-Biologically, \llabel{clarification2}
+when both have well-posed limits independent of the other.
+Biologically,
 the ``noise term'' is the contribution to variation in the population
@@
 Thus, $e^{\psi_{\infty}(\theta)}$ must satisfy $|e^{\psi_{\infty}(\theta)}| = e^{\Re \psi_{\infty}(\theta)} \leq 1$, whence $\Re \psi_{\infty}(\theta) \leq 0$ and $\gamma \geq 0$.
-In particular, if $\gamma > 0$, then we recognize $e^{\psi_{\infty}(\theta)} = e^{im \theta -\gamma|\theta|}$ as the characteristic function of a Cauchy distribution with median m and median absolute deviation $\gamma$ (\ie the canonical $\alpha$-stable distribution with $\alpha=1$), \llabel{cauchy_note} whereas if $\gamma = 0$, then $e^{\psi_{\infty}(\theta)} = e^{im \theta}$ is the Fourier transform of a Dirac mass at $m$, \ie the characteristic function for a monomorphic population with all individuals having trait value equal to $m$.
+In particular, if $\gamma > 0$, then we recognize $e^{\psi_{\infty}(\theta)} = e^{im \theta -\gamma|\theta|}$ as the characteristic function of a Cauchy distribution with median m and median absolute deviation $\gamma$ (\ie the canonical $\alpha$-stable distribution with $\alpha=1$), whereas if $\gamma = 0$, then $e^{\psi_{\infty}(\theta)} = e^{im \theta}$ is the Fourier transform of a Dirac mass at $m$, \ie the characteristic function for a monomorphic population with all individuals having trait value equal to $m$.
 
@@
 
-The crucial step in the derivation of the infinitesimal model in \cite{barton2017infinitesimal} is to show that, \llabel{crucial_step}
+The crucial step in the derivation of the infinitesimal model in \cite{barton2017infinitesimal} is to show that,
 following the arguments in \citet{fisher1918correlation}, knowing an individual's trait value gives vanishingly little information about the contribution of any given locus as $M \to \infty$, which in turn allows one to show that the within-family Mendelian sampling term is independent of the parental traits.
@@
 % Unlike the case of bounded allelic effects, the bound is independent of the trait value $Z^{[j]}$.
-\llabel{approx_eqn}
 
@@
 the genome is of length $10^8$bp with a uniform recombination rate of $10^{-8}$ crossovers per bp;
-mutations occur at a rate of $10^{-9}$ per bp. \llabel{mutation_rate}
+mutations occur at a rate of $10^{-9}$ per bp.
 Each new mutation is independently assigned an ``effect''
@@
 Next, we show that in fact the infinitesimal model \emph{cannot} describe non-Gaussian distributions
-(\ie non-Gaussian deviation of offspring trait from midparent value) \revpoint{2}{4}
+(\ie non-Gaussian deviation of offspring trait from midparent value)
 if the traits are determined additively from independent loci
@@
 
-\paragraph{Selection and effect size distributions} \llabel{discussion_DFEs}
+\paragraph{Selection and effect size distributions}
 A substantial body of theoretical work has used various models and assumptions
